### 2016-04-12 paper/paper.tex
Add latex template

\documentclass[10pt,numbers,preprint,nocopyrightspace]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.

\usepackage{amsmath}
\usepackage{color}
\usepackage{listings}
\usepackage{balance}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{theorem}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{mdwlist}
\usepackage{morefloats}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{mathptmx}
\usepackage[normalem]{ulem}
\usepackage{booktabs}
\usepackage{tablefootnote}
\usepackage{enumitem}

\setlist[itemize]{noitemsep,nolistsep}
\setlist[enumerate]{noitemsep,nolistsep}

\lstset{ %
language=Java,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbers=none,
stepnumber=1,
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
aboveskip=3pt,
belowskip=-1pt
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}
\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Title}

\authorinfo{}
           {}
           {}
%\authorinfo{Name}
%           {Affiliation2/3}
%           {Email2/3}

\maketitle

\begin{abstract}
TODO
\end{abstract}

\begin{comment}
\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2
\end{comment}

%===============================================================================
\input{sections/Introduction.tex}
\input{sections/Overview.tex}
\input{sections/CaseStudy.tex}
\input{sections/RelatedWork.tex}
%\acks

%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{paper}

% The bibliography should be embedded for final submission.

\end{document}

### 2016-04-12 paper/computePositions.rb

@@ -1,0 +1,68 @@
+require 'fileutils'
+
+# Example 1 - Read File and close
+def scan(fileName, beginpos, endpos, fileloc)
+  file = File.new(fileName, "r")
+  counter = 1
+  while (line = file.gets)
+    if line =~ /BEGIN_(\w+)/
+      throw "Symbol '#{$1}' already defined " + fileName if fileloc[$1]
+      fileloc[$1] = fileName
+      beginpos[$1] = counter+1
+      puts "BEGIN #{$1}: #{beginpos[$1]}"
+    end
+    if line =~ /END_(\w+)/
+      endpos[$1] = counter-1
+      puts "END #{$1}: #{endpos[$1]}"
+    end
+    counter = counter + 1
+  end
+  file.close
+end
+
+def process(file, out, beginpos, endpos, fileloc)
+  while (line = file.gets)
+    if line =~ /APPLY:\s?(\w+)/
+      label = $1
+      puts "Processing #{label}"
+      #puts "Line before: #{line}"
+      if fileloc.has_key?(label) then
+        line = line.sub( /linerange=([0-9-]*)/, "linerange=#{beginpos[label]}-#{endpos[label]}" )
+        #puts "Line is now 1: #{line}"
+        line = line.sub( /\{[^{]*\}/, "{#{fileloc[label]}}" )
+        #puts "Line is now 2: #{line}"
+      else
+        puts "Not found!"
+      end
+    end
+    out.write(line)
+  end
+  file.close
+  out.close
+end
+
+fileloc = {}
+beginpos = {}
+endpos = {}
+
+Dir['../src/*/*.java'].each do |file|
+  scan(file, beginpos, endpos, fileloc)
+end
+
+Dir['*.tex'].each do |file|
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+
+Dir['./sections/*.tex'].each do |file|
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end

### 2016-04-12 paper/run.sh

@@ -1,0 +1,27 @@
+#!/bin/bash
+
+if [ $1 ]
+then
+    case "$1" in
+    bib )       
+        pdflatex paper.tex
+        bibtex paper     
+    ;;
+    * )
+        echo "Usage (In terminal with zsh):"
+        echo "\t\t./run.sh"
+        echo "render and open a new pdf"
+        echo "\t\t./run.sh bib"
+        echo "render and open a new pdf with bib changes"
+    esac   
+fi
+
+ruby computePositions.rb
+pdflatex paper.tex
+open paper.pdf
+
+rm footempfile.txt
+# rm *.bbl
+# rm *.blg
+rm *.log
+rm *-old

### 2016-04-12 src/circuit/IntList.java
Add case studies

@@ -1,0 +1,12 @@
+package circuit;
+
+import java.util.List;
+
+import lombok.Obj;
+
+// wrapper for List<Integer>
+// the inner type (Integer) will be lost in the sub-family
+@Obj
+public interface IntList {
+    List<Integer> out();
+}

### 2016-04-12 src/circuit/IntPair.java

@@ -1,0 +1,18 @@
+package circuit;
+
+import java.util.function.IntUnaryOperator;
+
+import lombok.Obj;
+
+@Obj
+public interface IntPair {
+    int x();
+    int y();
+
+    default String show() {
+        return "(" + x() + "," + y() + ")";
+    }
+    default IntPair map(IntUnaryOperator f) {
+        return IntPair.of(f.applyAsInt(x()), f.applyAsInt(y()));
+    }
+}

### 2016-04-12 src/circuit/Layout.java

@@ -1,0 +1,20 @@
+package circuit;
+
+import static java.util.stream.Collectors.toList;
+
+import java.util.List;
+import java.util.function.IntUnaryOperator;
+
+import lombok.Obj;
+
+@Obj
+public interface Layout {
+    List<List<IntPair>> layout();
+    default Layout map(IntUnaryOperator f) {
+            return Layout.of(layout().stream()
+                    .map(xs -> xs.stream()
+                            .map(pr -> pr.map(f))
+                            .collect(toList()))
+                    .collect(toList()));
+    }
+}

### 2016-04-12 src/circuit/Test.java

@@ -1,0 +1,105 @@
+package circuit;
+
+import java.awt.Dimension;
+import java.awt.Graphics;
+import java.awt.Graphics2D;
+import java.util.Arrays;
+import java.util.List;
+
+import javax.swing.JFrame;
+import javax.swing.JPanel;
+import javax.swing.SwingUtilities;
+
+import circuit.TlayoutFeature.Above;
+import circuit.TlayoutFeature.Beside;
+import circuit.TlayoutFeature.Circuit;
+import circuit.TlayoutFeature.Fan;
+import circuit.TlayoutFeature.Identity;
+import circuit.TlayoutFeature.RStretch;
+import circuit.TlayoutFeature.Stretch;
+
+class DrawCircuit extends JPanel {
+    Layout layout;
+    static int L = 30;
+    static int W = 30;
+    static int R = 6;
+
+    public DrawCircuit(Layout layout) {
+        this.layout = layout;
+    }
+
+    public IntPair toCoord(int c, int r) {
+        return IntPair.of((c+1)*W, (r+1)*L);
+    }
+
+    public void drawDot(Graphics2D g, IntPair p) {
+        int x = p.x() - (R/2);
+        int y = p.y() - (R/2);
+        g.fillOval(x, y, R, R);
+
+    }
+
+    public void drawLine(Graphics2D g, IntPair from, IntPair to) {
+        g.drawLine(from.x(), from.y(), to.x(), to.y());
+    }
+
+    @Override
+    public void paintComponent(Graphics g) {
+        super.paintComponent(g);
+        Graphics2D g2d = (Graphics2D) g;
+        List<List<IntPair>> ll = layout.layout();
+        int col = ll.stream().flatMap(l -> l.stream()).map(pr -> pr.y()).reduce(Integer::max).orElse(0);
+
+        // draw vertical lines
+        for (int c = 0; c <= col; c++)
+            drawLine(g2d, toCoord(c, 0), toCoord(c, ll.size()));
+        for (int r = 0; r < ll.size(); r++) {
+            for (IntPair pr : ll.get(r)) {
+                IntPair from = toCoord(pr.x(), r);
+                IntPair to = toCoord(pr.y(), r+1);
+                drawDot(g2d, from);
+                drawDot(g2d, to);
+                drawLine(g2d, from, to);
+            }
+        }
+    }
+
+    @Override
+    public Dimension getPreferredSize() {
+        return new Dimension(300, 300);
+    }
+}
+public class Test {
+    public static void main(String[] args) {
+        // Figure 3
[…]

### 2016-04-12 src/circuit/TlayoutFeature.java

@@ -1,0 +1,250 @@
+package circuit;
+import static java.util.Collections.emptyList;
+import static java.util.Collections.singletonList;
+import static java.util.stream.Collectors.toList;
+
+import java.util.ArrayList;
+import java.util.List;
+import java.util.function.BinaryOperator;
+import java.util.function.IntUnaryOperator;
+import java.util.stream.IntStream;
+import java.util.stream.Stream;
+
+import lombok.Obj;
+
+@Obj
+interface AST {
+    interface Circuit {
+    }
+    interface Identity extends Circuit {
+        int n();
+    }
+    interface Fan extends Circuit {
+        int n();
+    }
+    interface Above extends Circuit {
+        Circuit c1();
+        Circuit c2();
+    }
+    interface Beside extends Circuit {
+        Circuit c1();
+        Circuit c2();
+    }
+    interface Stretch extends Circuit {
+        IntList ns();
+        Circuit c();
+    }
+}
+
+@Obj
+interface DepthFeature extends AST {
+    interface Circuit {
+        int _depth();
+    }
+    interface Identity{
+        default int _depth() {
+            return 0;
+        }
+    }
+    interface Fan {
+        default int _depth() {
+            return 1;
+        }
+    }
+    interface Above {
+        default int _depth() {
+            return c1()._depth() + c2()._depth();
+        }
+    }
+    interface Beside {
+        default int _depth() {
+            return Math.max(c1()._depth(), c2()._depth());
+        }
+    }
+    interface Stretch {
+        default int _depth() {
+            return c()._depth();
+        }
+    }
+}
+
+@Obj
+interface WidthFeature extends AST {
+    interface Circuit {
+        int _width();
+    }
+    interface Identity {
+        default int _width() {
+            return n();
+        }
+    }
+    interface Fan {
+        default int _width() {
+            return n();
+        }
+    }
+    interface Above {
+        default int _width() {
+            return c1()._width();
+        }
+    }
+    interface Beside {
+        default int _width() {
+            return c1()._width() + c2()._width();
+        }
+    }
+    interface Stretch {
+        default int _width() {
+            return ns().out().stream().reduce(0, (a, b) -> a + b); // bug: the inner type(Integer) is lost
+        }
+    }
+}
+
+
+@Obj
+interface WellSizeFeature extends WidthFeature, DepthFeature {
+    interface Circuit {
[…]

### 2016-04-12 src/diagrams/Attr.java

@@ -1,0 +1,12 @@
+package diagrams;
+
+import lombok.Obj;
+
+@Obj
+interface Attr {
+    String name();
+    String value();
+    default String show() {
+        return name() + "=\"" + value() + "\"";
+    }
+}

### 2016-04-12 src/diagrams/Diagrams.java

@@ -1,0 +1,251 @@
+package diagrams;
+
+import static java.util.Arrays.asList;
+import static java.util.Collections.emptyList;
+import static java.util.Collections.singletonList;
+import static java.util.stream.Collectors.joining;
+import static java.util.stream.Collectors.toList;
+
+import java.io.IOException;
+import java.nio.file.Files;
+import java.nio.file.Path;
+import java.nio.file.Paths;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.function.Function;
+import java.util.function.UnaryOperator;
+import java.util.stream.IntStream;
+import java.util.stream.Stream;
+
+import lombok.Obj;
+
+@Obj
+interface Family {
+    interface Shape {
+        Extent _toExtent();
+        XML _toXML(List<Attr> styleAttrs, Function<Pos, Pos> trans);
+        String _show();
+        default Drawing draw(StyleSheet style) {
+            return Drawing.of(singletonList(this), singletonList(style), singletonList(p -> p));
+        }
+    }
+
+    interface Rectangle extends Shape {
+        double width();
+        double height();
+        default Extent _toExtent() {
+            return Extent.of(Pos.of(width(), height()).resize(-0.5), Pos.of(width(), height()).resize(0.5));
+        }
+        default XML _toXML(List<Attr> styleAttrs, Function<Pos, Pos> trans) {
+            List<Attr> attrs = new ArrayList<>(trans.apply(Pos.of(width(), height()).resize(-0.5)).toAttrs("x", "y"));
+            attrs.addAll(Pos.of(width(), height()).toAttrs("width", "height"));
+            attrs.addAll(styleAttrs);
+            return XML.of(attrs, "rect", emptyList());
+        }
+        default String _show() {
+            return "Rectangle " + width() + " " + height();
+        }
+    }
+
+    interface Ellipse extends Shape {
+        double rx();
+        double ry();
+        @Override
+        default Extent _toExtent() {
+            return Extent.of(Pos.of(-rx(), -ry()), Pos.of(rx(), ry()));
+        }
+        default XML _toXML(List<Attr> styleAttrs, Function<Pos, Pos> trans) {
+            List<Attr> attrs = new ArrayList<>(trans.apply(Pos.of(0, 0)).toAttrs("cx", "cy"));
+            attrs.addAll(Pos.of(rx(), ry()).toAttrs("rx", "ry"));
+            attrs.addAll(styleAttrs);
[…]

### 2016-04-12 src/diagrams/Extent.java

@@ -1,0 +1,21 @@
+package diagrams;
+
+import java.util.function.Function;
+
+import lombok.Obj;
+
+@Obj
+interface Extent {
+    Pos p1();
+    Pos p2();
+    default Extent union(Extent e) {
+        return Extent.of(Pos.of(Math.min(p1().x(), e.p1().x()), Math.min(p1().y(), e.p1().y())),
+                Pos.of(Math.max(p2().x(), e.p2().x()), Math.max(p2().y(), e.p2().y())));
+    }
+    default Extent transform(Function<Pos, Pos> t) {
+        return Extent.of(t.apply(p1()), t.apply(p2()));
+    }
+    default String show() {
+        return "(" + p1().show() + "," + p2().show() + ")";
+    }
+}

### 2016-04-12 src/diagrams/Pos.java

@@ -1,0 +1,28 @@
+package diagrams;
+
+import static java.util.Arrays.asList;
+
+import java.util.List;
+
+import lombok.Obj;
+
+@Obj
+interface Pos {
+    double x();
+    double y();
+    default Pos add(Pos other) {
+        return Pos.of(x() + other.x(), y() + other.y());
+    }
+    default Pos resize(double scale) {
+        return Pos.of(scale*x(), scale*y());
+    }
+    default String show() {
+        return x() + "," + y();
+    }
+    default List<Attr> toAttrs(String x, String y) {
+        return asList(Attr.of(x, ""+x()), Attr.of(y, ""+y()));
+    }
+    default Pos conjugate() {
+        return Pos.of(x(), -y());
+    }
+}

### 2016-04-12 src/diagrams/XML.java

@@ -1,0 +1,22 @@
+package diagrams;
+
+import static java.util.stream.Collectors.joining;
+
+import java.util.List;
+
+import lombok.Obj;
+
+@Obj
+interface XML {
+    String tag();
+    List<Attr> attrs();
+    List<XML> xmls();
+
+    default String show() {
+        String attrs = " " + attrs().stream().map(Attr::show).collect(joining(" "));
+        if (xmls().size() == 0)
+            return "<" + tag() + attrs + "/>";
+        else
+            return "<" + tag() + attrs + ">\n" + xmls().stream().map(XML::show).collect(joining("\n")) + "\n</" + tag() + ">";
+    }
+}

### 2016-04-19 paper/sections/CaseStudy.tex
paper structure

\section{Case Study}

Talk about the 3 case studies

Talk about Pretty Printer case study last and discuss issues with 
extensible transformations.

### 2016-04-19 paper/sections/Overview.tex
paper structure

\section{Overview}

Go over jeremy's examples, maybe having only 3 diagram 
constructs instead of 5 for space reasons.

Use five constructs to show extensibility.

Think of how to introduce our tool? Using the Jeremy's examples? 
or introducing before with some other examples?

### 2016-04-19 paper/sections/RelatedWork.tex
paper structure

\section{Related Work}

\paragraph{Deep and Shallow Embeddings}

\paragraph{Modularity of DSLs?}
Data types a la Carte; Finally Tagless; 
Object Algebras

\paragraph{Family Polymorphism}

\paragraph{ThisType}

\paragraph{Multiple Inheritance}

### 2016-04-19 paper/sections/ShallowOO.tex
paper structure

\section{Shallow Object-Oriented Programming}

Argue that shallow embeddings and straightforward OO 
programs are essentially the same thing.

Start from a simple shallow DSL in Haskell, 
and iterate throught it until you reach a form 
that looks like an OO program.

Show how todo transformations in Shallow embeddings
using the insight of how to do transformations in OO
programs.

Show the correponding Java programs and the Java program 
with transformation that we can port back to Haskell.

\begin{verbatim}
{-
type Exp = Int

lit x = x

add e1 e2 = e1 + e2
-}

{-
newtype Exp = E Int

lit x = E x

add (E e1) (E e2) = E (e1 + e2)
-}

data Exp = E {
  eval :: Int,
  transform :: Exp
}

lit x = E {eval = x, transform = lit x}

add e1 e2 = E {
  eval = eval e1 + eval e2,
  transform = add (lit 0) e2
}

p = eval $ transform (add (lit 5) (lit 3))
\end{verbatim}

### 2016-04-19 paper/sections/Tool.tex
paper structure

\section{Code Generation}

How to generate code for family polymorphism!

### 2016-04-19 paper/sections/Discussion.tex
paper structure

\section{Discussion and Limitations}

- Transformations and Binary methods (but argue that 
these types of operations are not used in shallow DSLs 
any way!).

- Java syntax not great for DSLs. 
operator overloding; infix operators (but this is a Java limitation; 
not a limitation of our approach).

- Limitations of the tool: separate compiltions and ... 
generics not fully implemented.

- How to apply our approach to other OO language 
  - C\# has annotations, but can you do AST rewriting?
  - Scala has macros;
  - Other languages?

### 2016-04-19 paper/sections/ShallowOO.tex
Delegating tasks
Move Bruno's code into a separate file

@@
 \section{Shallow Object-Oriented Programming}
+
+Weixin writes this part.
 
@@
 
-\begin{verbatim}
-{-
-type Exp = Int
-
-lit x = x
-
-add e1 e2 = e1 + e2
--}
-
-{-
-newtype Exp = E Int
-
-lit x = E x
-
-add (E e1) (E e2) = E (e1 + e2)
--}
-
-data Exp = E {
-  eval :: Int,
-  transform :: Exp
-}
-
-lit x = E {eval = x, transform = lit x}
-
-add e1 e2 = E {
-  eval = eval e1 + eval e2,
-  transform = add (lit 0) e2
-}
-
-p = eval $ transform (add (lit 5) (lit 3))
-\end{verbatim}
+\lstinputlisting[language=haskell,linerange=2-5]{./code/shallow.hs}%APPLY:EXPR1
+\lstinputlisting[language=haskell,linerange=9-12]{./code/shallow.hs}%APPLY:EXPR2
+\lstinputlisting[language=haskell,linerange=16-18]{./code/shallow.hs}%APPLY:EXPR3
+\lstinputlisting[language=haskell,linerange=22-33]{./code/shallow.hs}%APPLY:EXPR4

### 2016-04-19 paper/code/shallow.hs

@@ -1,0 +1,34 @@
+--BEGIN_EXPR1
+type Expr1 = Int
+lit1 n = n
+add1 e1 e2 = e1 + e2
+eval1 = id
+--END_EXPR1
+
+--BEGIN_EXPR2
+newtype Expr2 = Expr2 Int
+lit2 n = Expr2 n
+add2 (Expr2 e1) (Expr2 e2) = Expr2 $ e1 + e2
+eval2 (Expr2 n) = n
+--END_EXPR2
+
+--BEGIN_EXPR3
+data Expr3 = Expr3 { eval3 :: Int }
+lit3 n = Expr3 { eval3 = n }
+add3 e1 e2 = Expr3 { eval3 = eval3 e1 + eval3 e2 }
+--END_EXPR3
+
+--BEGIN_EXPR4
+data Expr4 = Expr4 {
+  eval4 :: Int,
+  transform :: Expr4
+}
+lit4 n = Expr4 {
+  eval4 = n,
+  transform = lit4 n
+}
+add4 e1 e2 = Expr4 {
+  eval4 = eval4 e1 + eval4 e2,
+  transform = add4 (lit4 0) e2
+}
+--END_EXPR4

### 2016-04-19 paper/computePositions.rb

@@ -49,6 +49,10 @@ Dir['../src/*/*.java'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end
 
+Dir['./code/*'].each do |file|
+  scan(file, beginpos, endpos, fileloc)
+end
+
 Dir['*.tex'].each do |file|
   #temp = Tempfile.new('compute_positions')
   tempname = "footempfile.txt"

### 2016-04-23 paper/sections/Overview.tex
Delegating tasks
Some text for S3

@@
 \section{Overview}
+
+\begin{comment}
+Weixin writes this one.
 
@@
 or introducing before with some other examples?
+
+\end{comment}
+
+%As language evolves, the need of new syntax often arises.
+DSLs may evolve along the time with new syntax and semantics introduced.
+This is essentially the Expression Problem (EP)~\cite{wadler}, which requires
+the host language equipped with two dimensions of extensibility.
+To illustrate, we start with a subset of the DSL presented in Gibbons and Wu's
+paper with a single semantic function. And then we show how to extend this
+language with full language constructs and additional semantics both in Haskell
+and in Java. The key difference is that the extension on semantics for Java is
+\emph{modular} whereas the Haskell approach needs to modify existing code.
+
+\subsection{A DSL for Parallel Prefix Circuits}
+Initially, the language contains three language constructs: a primitive
+\emph{fan} and two combinators \emph{beside} and \emph{stretch}. Their types are
+given as follows:
+
+\lstinputlisting[language=haskell,linerange=2-5]{./code/shallowCircuit.hs}%APPLY:CTYPE
+\emph{fan n} contains \emph{n} vertical wires with the first wire connected to
+all the remaining wires from top to bottom; \emph{beside c1 c2} joins two circuits
+horizontally; \emph{stretch ns c} insertes additional wires into the circuit
+\emph{c} by summing up \emph{ns}.
+For example, a circuit constructed with these three operator \emph{stretch [2,2,2] (fan 3) `beside` fan 1} is visualized in Figure~\ref{}.
+The semantics of the language is given below:
+
+\lstinputlisting[language=haskell,linerange=9-12]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
+which calculates the width of a circuit, i.e. the number of vertical wires in a circuit.
+The corresponding Java implementation is shown in Figure~\ref{}.
+
+\lstinputlisting[linerange=9-9]{../src/paper/Circuit.java}%APPLY:CIRCUIT
+\lstinputlisting[linerange=13-16]{../src/paper/Circuit.java}%APPLY:FAN
+\lstinputlisting[linerange=20-26]{../src/paper/Circuit.java}%APPLY:STRETCH
+\lstinputlisting[linerange=30-36]{../src/paper/Circuit.java}%APPLY:BESIDE
+The translation is straightforward. The only difference is that classes are
+replaced with interfaces and fields become unimplemented getter methods. This
+change is vital for extensibility, which will be discussed later in this section.
+
+\subsection{Adding New Syntax}
+Shallow embedding makes it easy to \emph{modularly} add new syntax both in FP and OOP.
+Suppose that we would like to add a primitive \emph{identity n} which contains
+\emph{n} isolated wires and a combinator \emph{above c1 c2} which combines two
+circuits of the same width vertically.
+
+In Haskell, we can define two more cases for the semantic function \texttt{identity} and
+\texttt{above} like this:
+\lstinputlisting[linerange=51-52]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
+It is also trivial in Java. We just need to define two new interfaces
+\texttt{Identity} and \texttt{Above} that both extend \texttt{Circuit} and
+implement the \texttt{width} method:
+
+\lstinputlisting[linerange=40-43]{../src/paper/Circuit.java}%APPLY:ID
+\lstinputlisting[linerange=47-51]{../src/paper/Circuit.java}%APPLY:ABOVE
+
+\subsection{Adding New Semantics}
+Adding new semantics, however, becomes hard.
+For example, we would like to have an operation that checks
+whether the circuit is of correct size. We say a circuit is \emph{well-sized}
+only if: 1) for \emph{fan n}, \emph{n} is a positive number; 2) for
+\emph{beside c2}, both \emph{c1} and \emph{c2} are well-sized; 3) for
+\emph{stretch ns c} \emph{c} is well-sized and its width is the same as the
+length of \emph{ns}.
+
+Gibbons and Wu solved this problem similar to how we add \texttt{peval} to
+\texttt{Exp}, where they used a pair instead of a record to capture the new
+semantic domain, then defined the two semantic functions simultaneously for each
+case and split the definitions through projections on the tuple in the end.
+This solution, however, modifies existing code, breaking the requirement of EP.
+
+Conversely, the support of covariant type-refinements and inheritance for OO
+languages allows us to add new semantics \emph{modularly}:
+%Different from records or tuples, interfaces are extensible.
+\lstinputlisting[linerange=56-58]{../src/paper/Circuit.java}%APPLY:CIRCUITWS
+\lstinputlisting[linerange=62-64]{../src/paper/Circuit.java}%APPLY:FANWS
+\lstinputlisting[linerange=68-73]{../src/paper/Circuit.java}%APPLY:STRETCHWS
+\lstinputlisting[linerange=77-83]{../src/paper/Circuit.java}%APPLY:BESIDEWS
+
+Interface \texttt{CircuitWS} extends the original interface and declare a new
+semantic function \texttt{wellSized} inside.
+Then all existing cases should extend this new interface with the semantic
+function implemented. Note that multiple inheritance is used for composing
+existing semantic functions and also for methods whose return type is
+\texttt{Circuit} should be refined to \texttt{CircuitWS} so that we can call
+\texttt{wellSized} method on such fields.
+
+extensibility.
+for return types. This is \ref{EPtrivially}
+(that is why we implement language constructs as interfaces rather than classes)
+
+\subsection{\name's Support}
+Although manually . It has some drawbacks.
+It is cubersome to .
+To retain extensibility, all should be kept as interfaces, which means that
+additional classes should be created for each of them so that objects can be constructed.
+Once new semantics is introduced, all the
+Programmers would not get warned may forget to implement
+Also, it is cubersome to express the dependencies boilerplate and do
+type-refinements manually.
+
+\name addresses all these problems through code instrument.
+By using \name, we can rewrite the extensions in the way shown in Figure\ref{}.
+
+Figure~\ref{} shows the \name's implementation of Circuit with extensions.
+Note that \name generates \texttt{of} method for class-like interfaces for
+constructing objects. Inside \texttt{of}, an anonymous class is created with all
+field-like methods implemented and its instance is returned.
+Moreover, \name has a good support family polymorphism.
+Client users need only explicitly declared the dependencies on other families
+through (...). \name will re-declare all members from the base families in the extended family.
+And the dependencies between members of families are automatically deducted and
+implicitly expressed in member re-declaration.

### 2016-04-23 paper/code/shallow.hs

@@ -1,34 +1,53 @@
+--BEGIN_TYPE
+lit :: Int -> Exp
+add :: Exp -> Exp -> Exp
+--END_TYPE
+
 --BEGIN_EXPR1
-type Expr1 = Int
-lit1 n = n
-add1 e1 e2 = e1 + e2
-eval1 = id
+type Exp1 = Bool
+lit1 n = n == 0
+add1 _ _ = false
+is0 = id
 --END_EXPR1
 
 --BEGIN_EXPR2
-newtype Expr2 = Expr2 Int
-lit2 n = Expr2 n
-add2 (Expr2 e1) (Expr2 e2) = Expr2 $ e1 + e2
-eval2 (Expr2 n) = n
+newtype Exp2 = Exp2 Bool
+lit2 n = Exp2 n
+add2 (Exp2 e1) (Exp2 e2) = Exp2 $ e1 + e2
+is0 (Exp2 e) = e
 --END_EXPR2
 
 --BEGIN_EXPR3
-data Expr3 = Expr3 { eval3 :: Int }
-lit3 n = Expr3 { eval3 = n }
-add3 e1 e2 = Expr3 { eval3 = eval3 e1 + eval3 e2 }
+newtype Exp3 = Exp3 { eval3 :: Int }
+lit3 n = Exp3 { eval3 = n }
+add3 e1 e2 = Exp3 { eval3 = eval3 e1 + eval3 e2 }
 --END_EXPR3
 
 --BEGIN_EXPR4
-data Expr4 = Expr4 {
+data Exp4 = Exp4 {
   eval4 :: Int,
-  transform :: Expr4
+  transform :: Exp4
 }
-lit4 n = Expr4 {
+lit4 n = Exp4 {
   eval4 = n,
   transform = lit4 n
 }
-add4 e1 e2 = Expr4 {
+add4 e1 e2 = Exp4 {
   eval4 = eval4 e1 + eval4 e2,
   transform = add4 (lit4 0) e2
+}
+
+data Exp = Exp {
+  is0 :: Bool,
+  peval :: Exp
+}
+lit n = Exp {
+  is0 = n == 0,
+  peval = lit n
+}
+add e1 e2 = Exp {
+  is0 = false,
+  peval = let (e1', e2') = (peval e1, peval e2)
+          in if (is0 e1') then e2' else add(e1', e2')
 }
 --END_EXPR4

### 2016-04-23 src/paper/Circuit.java

@@ -1,0 +1,151 @@
+package paper;
+
+import java.util.List;
+
+import lombok.Obj;
+
+@Obj
+//BEGIN_CIRCUIT
+interface Circuit { int width(); }
+//END_CIRCUIT
+@Obj
+//BEGIN_FAN
+interface Fan extends Circuit {
+  int _n();
+  default int width() { return _n(); }
+}
+//END_FAN
+
+//BEGIN_STRETCH
+interface Stretch extends Circuit {
+  List<Integer> _ns();
+  Circuit _c();
+  default int width() {
+    return _ns().stream().reduce(0, (a,b) -> a+b);
+  }
+}
+//END_STRETCH
+@Obj
+//BEGIN_BESIDE
+interface Beside extends Circuit {
+  Circuit _c1();
+  Circuit _c2();
+  default int width() {
+    return _c1().width() + _c2().width();
+  }
+}
+//END_BESIDE
+@Obj
+//BEGIN_ID
+interface Id extends Circuit {
+  int _n();
+  default int width() { return _n(); }
+}
+//END_ID
+@Obj
+//BEGIN_ABOVE
+interface Above extends Circuit {
+  CircuitWS _c1();
+  CircuitWS _c2();
+  default int width() { return _c2().width(); }
+}
+//END_ABOVE
+
+@Obj
+//BEGIN_CIRCUITWS
+interface CircuitWS extends Circuit {
+  boolean wellSized();
+}
+//END_CIRCUITWS
+@Obj
+//BEGIN_FANWS
+interface FanWS extends Fan, CircuitWS {
+  default boolean wellSize() { return true; }
+}
+//END_FANWS
+@Obj
+//BEGIN_STRETCHWS
+interface StretchWS extends Stretch, CircuitWS {
+  CircuitWS _c();
+  default boolean wellSized() {
+    return _ns().size() == _c().width();
+  }
+}
+//END_STRETCHWS
+@Obj
+//BEGIN_BESIDEWS
+interface BesideWS extends Beside, CircuitWS {
+  CircuitWS _c1();
+  CircuitWS _c2();
+  default boolean wellSized() {
+    return _c1().wellSized() && _c2().wellSized();
+  }
+}
+//END_BESIDEWS
+
+//BEGIN_FAMILY
+@Obj interface Family {
+  interface Circuit {
+    int width();
+  }
+  interface Fan extends Circuit {
+    int _n();
+    default int width() {
+      return _n();
+    }
+  }
[…]

### 2016-04-25 paper/sections/ShallowOO.tex
Text for S2

--- 2016-04-19 paper/sections/ShallowOO.tex
@@
 \section{Shallow Object-Oriented Programming}
+
+\begin{comment}
+Weixin writes this part.
 
@@
 
-\begin{verbatim}
-{-
-type Exp = Int
+Shallow embedding interpreter pattern
+Deep embedding visitor pattern
+\end{comment}
 
-lit x = x
+In this section, we argue that \emph{shallow embedding} and straightforward OO
+programming are essentially the same thing. We use a subset
+of the DSL presented in Gibbons and Wu's paper~\cite{} as the running example.
+We first show how to use shallow embedding to define it and then gradually
+rewrite the definition into an ``OO style''. And we also show how to do
+transformations in shallow embedding, which is typically thought can only be
+done in deep embedding.
 
-add e1 e2 = e1 + e2
--}
+\subsection{A DSL for Parallel Prefix Circuits}
+Consider an DSL named \dsl that models parallel circuits.
+Its BNF grammar is given below:
+$$
+\begin{array}{lcl}
+Circuit & ::= & Identity \\
+ & \mid & Fan \\
+ & \mid & Beside \\
+Fan & ::= & ``fan''\; PositiveNumber \\
+Identity & ::= & ``identity''\;  PositiveNumber \\
+Beside & ::= & Circuit ``beside'' Circuit \\
+PositiveNumber & ::= & [1-9] \\
+     & \mid & [1-9][0-9]^* \\
+\end{array}
+$$
+% bnf
+\weixin{the bnf grammar needs polishing}
 
-{-
-newtype Exp = E Int
+There are three constructs - two primitives
+\emph{fan} and \emph{identity} and one combinator \emph{beside} whose
+meanings are: \emph{identity n} contains \emph{n} isolated vertical wires;
+\emph{fan n} contains \emph{n} vertical wires and the first wire connects to
+all the remaining ones from top to bottom; $beside\ c_1\ c_2$ joins two circuits
+$c_1$ and $c_2$ horizontally.
+For example, the circuit \emph{fan 3 beside identity 3} can be visualized in Figure~\ref{}.
 
-lit x = E x
+\subsection{Shallow Embedding and Object-Oriented Programming}
+Different from deep embedding, which uses a data structure to represent the
+abstract syntax of a language, shallow embedding defines a language
+directly through encoding its semantics. In the case of \dsl,
+the shallow embedding implementation of language should conform to the types given below:
 
-add (E e1) (E e2) = E (e1 + e2)
--}
+\lstinputlisting[language=haskell,linerange=2-5]{./code/shallowCircuit.hs}%APPLY:CTYPE
+where \texttt{Circuit} models the semantic domain and should be later filled in with a
+concrete type according to the semantics of the language.
+For example, if the semantics of \dsl is to calculate the width of a
+circuit, the definition would be:
+\lstinputlisting[language=haskell,linerange=11-16]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
 
-data Exp = E {
-  eval :: Int,
-  transform :: Exp
-}
+As shallow embedding is semantics-oriented, it is already an integer after
+constructing a circuit. For example, we will immediately get $6$ for
+\texttt{fan 3 `beside` identity 3} and hence \texttt{width1} is just an identity function.
 
-lit x = E {eval = x, transform = lit x}
+The above implementation can be rewritten by wrapping the result into an
+datatype and get back the value by pattern matching:
+\lstinputlisting[language=haskell,linerange=20-26]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
+A question then arises - is this implementation shallow or deep?
 
-add e1 e2 = E {
-  eval = eval e1 + eval e2,
-  transform = add (lit 0) e2
-}
+We continue rewriting the definition. This time, we use a record to represent
+the result and give it a name:
+\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
+The implementation is apparently shallow since it is no different from the last version.
+At this stage, however, the definition looks so like an OO program that we can easily
+translate it into an OO language like Java:
+\lstinputlisting[linerange=5-24]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
+The record type maps to an interface with its field declared as an abstract method.
+Each case corresponds to an class and its parameters become fields of the class.
+And the class implements the \texttt{Circuit} interface by giving \texttt{width}
+a definition. This implementation, in fact, is an instance of the \interp pattern.
+In other words, shallow embedding and the \interp are essentially the same thing.
+It may also worth mentioning that deep embedding is closely related to the
+\textsc{Visitor} pattern.
 
-p = eval $ transform (add (lit 5) (lit 3))
-\end{verbatim}
+\subsection{Transformations in Shallow Embedding}
+One of the main criticism of shallow embedding is that how to define transformations
+becomes unclear since there is not a structure representing the Abstract Syntax Tree (AST).
+%In fact, transformations can still be defined in shallow embedding.
+For example, there usually are some language constructs that do not extend the
+expressive power of the language but ease programming in the language.
+Then, in implementation, we define a core language with less constructs and
+define a transformation that rewrites the AST to use core constructs.
+
+The \emph{identity n} construct in \dsl is a syntactic sugar, which can
+be rewritten using the following formula:
+$$identity\ n = (fan\ 1) \overbrace{`beside`\ (fan\ 1)}^{n-1}\ (n \geq 1)$$
+which states that \emph{identity n} equals to \emph{n} of \emph{fan 1} combined
+with \emph{beside}.
+%Using the above formula, we can define a transformation that desugars every
+%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.
+
+We can implement this desugaring operation in Java:
+\lstinputlisting[language=haskell,linerange=29-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
+The return type of \texttt{desugar} is \texttt{Circuit} reflecting that it is a
+transformation. The implementation is straightforward: \emph{fan n} just returns
+itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wrap back
+their results into an instance of \texttt{Beside}; \emph{identity n} encodes the above
+formula.
+
+\weixin{define a show function in the and show the result after desugaring?}
+Of course, defining \texttt{desugar} is also possible in Haskell:
+
+\lstinputlisting[language=haskell,linerange=40-55]{./code/shallowCircuit.hs}%APPLY:DESUGAR
+The definition is a little bit tricky since it makes use of Haskell's laziness.

### 2016-04-25 paper/code/shallowCircuit.hs

@@ -1,0 +1,68 @@
+{-BEGIN_CTYPE
+type Circuit = ...
+identity :: Int -> Circuit
+fan :: Int -> Circuit
+beside :: Circuit -> Circuit -> Circuit
+END_CTYPE-}
+
+-- stretch :: [Int] -> Circuit -> Circuit
+
+--BEGIN_CIRCUIT1
+type Circuit1 = Int
+identity1 n = n
+fan1 n = n
+beside1 c1 c2 = c1 + c2
+
+width1 = id
+--END_CIRCUIT1
+
+--BEGIN_CIRCUIT2
+newtype Circuit2 = Circuit2 Int
+identity2 n = Circuit2 n
+fan2 n = Circuit2 n
+beside2 (Circuit2 c1) (Circuit2 c2) =
+  Circuit2 $ c1 + c2
+
+width2 (Circuit2 n) = n
+--END_CIRCUIT2
+
+
+--BEGIN_CIRCUIT3
+newtype Circuit3 = Circuit3 { width3 :: Int }
+identity3 n = Circuit3 { width3 = n }
+fan3 n = Circuit3 { width3 = n }
+beside3 c1 c2 = Circuit3 {
+  width3 = width3 c1 + width3 c2
+}
+--END_CIRCUIT3
+
+--BEGIN_DESUGAR
+data Circuit = Circuit {
+  width :: Int,
+  desugar :: Circuit
+}
+identity n = Circuit {
+  width = width c,
+  desugar = c
+} where c = foldl1 beside $ replicate n (fan 1)
+fan n = Circuit {
+  width = n,
+  desugar = fan n
+}
+beside c1 c2 = Circuit {
+  width = width c1 + width c2,
+  desugar = beside (desugar c1) (desugar c2)
+}
+--END_DESUGAR
+
+--BEGIN_SYNTAX_HS
+--identity n = n
+above _ c2 = c2
+--END_SYNTAX_HS
+
+
+--BEGIN_SEMANTICS_HS
+--fan2 n = (n, n > 0)
+--beside2 c1 c2 = (fst c1 + fst c2, snd c1 && snd c2)
+--stretch2 ns c = (sum ns, snd c && length ns == fst c)
+--END_SEMANTICS_HS

### 2016-04-25 paper/computePositions.rb

@@ -49,6 +49,10 @@ Dir['../src/*/*.java'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end
 
+Dir['../src/*/*/*.java'].each do |file|
+  scan(file, beginpos, endpos, fileloc)
+end
+
 Dir['./code/*'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end

### 2016-04-25 src/paper/Circuit.java

@@ -4,6 +4,8 @@ import java.util.List;
 
 import lombok.Obj;
 
+
+interface namespace {
 @Obj
 //BEGIN_CIRCUIT
 interface Circuit { int width(); }
@@ -37,16 +39,29 @@ interface Beside extends Circuit {
 //END_BESIDE
 @Obj
 //BEGIN_ID
-interface Id extends Circuit {
-  int _n();
-  default int width() { return _n(); }
-}
+//interface Identity extends Circuit {
+//  int _n();
+//  default int width() { return _n(); }
+//}
 //END_ID
+interface Identity extends Circuit {
+  int _n();
+  default int width() {
+    return _n();
+  }
+  default Circuit desugar() {
+    Circuit res = Fan.of(1);
+    for (int i = 1; i < _n(); i++)
+      res = Beside.of(res, Fan.of(1));
+    return res;
+  }
+}
+
 @Obj
 //BEGIN_ABOVE
 interface Above extends Circuit {
-  CircuitWS _c1();
-  CircuitWS _c2();
+  Circuit _c1();
+  Circuit _c2();
   default int width() { return _c2().width(); }
 }
 //END_ABOVE
@@ -82,12 +97,17 @@ interface BesideWS extends Beside, CircuitWS {
   }
 }
 //END_BESIDEWS
-
+}
+/*
 //BEGIN_FAMILY
 @Obj interface Family {
-  interface Circuit {
-    int width();
-  }
+  // the same code in Figure .
+  ...
+}
+//END_FAMILY
+*/
+@Obj interface Family {
+  interface Circuit { int width(); }
   interface Fan extends Circuit {
     int _n();
     default int width() {
@@ -109,43 +129,37 @@ interface BesideWS extends Beside, CircuitWS {
     }
   }
 }
-//END_FAMILY
 
-//BEGIN_FAMILY_SEMANTICS
-@Obj interface NewSemantics extends Family {
-  interface Circuit {
-    boolean wellSize();
-  }
-  interface Fan extends Circuit {
-    default boolean wellSize() { return true; }
-  }
-  interface Stretch extends Circuit {
-    default boolean wellSize() {
-      return _ns().size() == _c().width();
-    }
-  }
-  interface Beside extends Circuit {
-    default boolean wellSize() {
-      return _c1().wellSize() && _c2().wellSize();
-    }
-  }
-}
-//END_FAMILY_SEMANTICS
 
 //BEGIN_FAMILY_SYNTAX
[…]

### 2016-04-25 src/paper/sec2/Circuit.java

@@ -1,0 +1,63 @@
+package paper.sec2;
+
+class A {
+//BEGIN_CIRCUIT_OO
+interface Circuit { int width(); }
+class Fan implements Circuit {
+  int _n;
+  Fan(int n) { _n = n; }
+  public int width() { return _n; }
+}
+class Beside implements Circuit {
+  Circuit _c1, _c2;
+  Beside(Circuit c1, Circuit c2) {
+    _c1 = c1; c2 = _c2;
+  }
+  public int width() {
+    return _c1.width() + _c2.width();
+  }
+}
+class Identity implements Circuit {
+  int _n;
+  Identity(int n) { _n = n; }
+  public int width() { return _n; }
+}
+//END_CIRCUIT_OO
+}
+class Sec2 {
+//BEGIN_DESUGAR_OO
+interface Circuit {
+  int width();
+  Circuit desugar();
+}
+class Fan implements Circuit {
+  int _n;
+  Fan(int n) { _n = n; }
+  public int width() { return _n; }
+  public Circuit desugar() { return this; }
+}
+class Beside implements Circuit {
+  Circuit _c1, _c2;
+  Beside(Circuit c1, Circuit c2) {
+    _c1 = c1; c2 = _c2;
+  }
+  public int width() {
+    return _c1.width() + _c2.width();
+  }
+  public Circuit desugar() {
+    return new Beside(_c1.desugar(), _c2.desugar());
+  }
+}
+class Identity implements Circuit {
+  int _n;
+  Identity(int n) { _n = n; }
+  public int width() { return _n; }
+  public Circuit desugar() {
+    Circuit res = new Fan(1);
+    for (int i = 1; i < _n; i++)
+      res = new Beside(res, new Fan(1));
+    return res;
+  }
+}
+//END_DESUGAR_OO
+}

### 2016-04-26 paper/sections/Overview.tex
Text on S3, revised

--- 2016-04-19 paper/sections/Overview.tex
@@
 \section{Overview}
+\weixin{title: Evolution on DSLs?}
+
+\begin{comment}
+Weixin writes this one.
 
@@
 or introducing before with some other examples?
+
+\end{comment}
+
+%As language evolves, the need of new syntax often arises.
+As DSLs evolve along the time, the demand for new syntax and and semantics may arise.
+It would be good if we can introduce these new features to the DSL \emph{modularly}.
+%This is actually a hard problem, known as the Expression Problem (EP)~\cite{wadler}, which
+This, however, requires the host language equipped with two dimensions of
+extensibility.
+In this section, we argue that from extensibility perspective, OO languages are
+better candidates as host languages than FP languages.
+To demonstrate this, we try to extend \dsl with new syntax and new semantics in
+the setting of shallow embedding. Adding new syntax is both easy for both FP
+and OOP. Adding new semantics, however, is hard in FP. Although it is possible
+in OOP, the solution requires some boilerplate code. We hence developed \name
+for defining modular extensions easily.
+
+\subsection{Initial System}
+Before introducing any extensions, we need to rewrite the initial implementation for \dsl shown in Figure~\ref{}:
+
+\lstinputlisting[linerange=11-26]{../src/paper/Circuit.java}%APPLY:INIT
+We change the implementation language constructs from classes to interfaces and
+their fields to unimplemented getter methods. The purpose of this modification
+is to allow types of the fields to be refined in future, which is vital for
+retaining extensibility.
+
+\subsection{Adding New Syntax}
+Shallow embedding makes it easy to add new syntax, both in FP and OOP.
+Suppose that we would like to add two new constructs to the language:
+the first is \emph{stretch ns c}, which inserts additional wires into the circuit \emph{c} by summing up \emph{ns};
+the second is $above c_1 c_2$, which combines two circuits of the same width vertically.
+With these new constructs, more complex circuit can be constructed.
+For example, Figure~\ref{} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.
+
+To accomplish this goal in Haskell, we can simply define two more cases,
+\texttt{stretch} and \texttt{above}, for the semantic function:
+\lstinputlisting[linerange=59-60]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
+Similarly, defining two new interfaces
+\texttt{Stretch} and \texttt{Above} that both extend \texttt{Circuit} and
+implement the \texttt{width} method is all we need to do in Java:
+
+\lstinputlisting[linerange=29-40]{../src/paper/Circuit.java}%APPLY:SYNTAX
+
+\subsection{Adding New Semantics}
+The new combinators, however, can not apply to arbitrary circuits - they have
+some invariants in their definitions.
+To make sure that a circuit is constructed correctly, we need to expand the
+semantics of the language for doing such checks.
+
+Adding new semantics, however, becomes hard for shallow embedding.
+Gibbons and Wu worked around this problem in the following way:
+\lstinputlisting[linerange=47-51]{../src/paper/Circuit.java}%APPLY:HS_SEMANTICS
+
+which is similar to how we add \texttt{desugar} to the implementation in Section~\ref{}.
+The difference is that they use a tuple instead of a record to merge the
+semantic domains, then define the two semantic functions simultaneously for each
+case, and split the definitions through projections on the tuple in the end.
+This solution, however, modifies existing code, breaking the requirement of EP.
+
+Conversely, the support of covariant type-refinements and inheritance for OO
+languages allows us to add new semantics in a \emph{modular} way:
+%Different from records or tuples, interfaces are extensible.
+\lstinputlisting[linerange=47-75]{../src/paper/Circuit.java}%APPLY:SEMANTICS
+
+Interface \texttt{CircuitWS} extends the original interface and declare a new
+semantic function \texttt{wellSized} inside.
+Then all existing cases should extend both their corresponding original
+implementation and \texttt{CircuitWS} and implement the new method
+\texttt{wellSized}. Also, all the occurrences of \texttt{Circuit} are
+refined as \texttt{CircuitWS} so that we can call \texttt{wellSized} on inner circuits returned by getters.
+As Java does not support type-refinements on fields, we hence implement these
+constructs as interfaces rather than classes.
+
+\subsection{\name's Support}
+There exists some boilerplate for the Java solution presented above:
+\begin{itemize}
+  \item Interfaces that represent language constructs should be instantiated for
+    creating objects;
+  \item Type-refinements should be done manually
+  \item The inheritance relation needs to be repeatedly stated for all interfaces.
+\end{itemize}
+Also, programmers would not get warned if they forget to extend all the
+constructs with new semantics.
+
+\name addresses all these problems through embracing family polymorphism and code instrument.
+By using \name, we can refactor the extensions in the way shown in Figure\ref{}.
+\lstinputlisting[linerange=80-83]{../src/paper/Circuit.java}%APPLY:FAMILY
+\lstinputlisting[linerange=112-122]{../src/paper/Circuit.java}%APPLY:FAMILY_SYNTAX
+\lstinputlisting[linerange=126-141]{../src/paper/Circuit.java}%APPLY:FAMILY_SEMANTICS
+
+Note that \name generates \texttt{of} method for class-like interfaces for
+constructing objects. The static method \texttt{of} returns an instance of an anonymous class
+that implements the interface with all
+getters implemented and its instance is returned.
+Moreover, \name has a good support family polymorphism.
+Client users need only explicitly declared the dependencies on other families
+through (...). \name will re-declare all members from the base families in the extended family.
+And the dependencies between members of families are automatically deducted and
+implicitly expressed in member re-declaration.
+
+% Multiple interpretation
+% Dependent interpretation
+% Context-sensitive interpretation

### 2016-04-26 src/paper/Circuit.java

@@ -5,30 +5,18 @@ import java.util.List;
 import lombok.Obj;
 
 
-interface namespace {
 @Obj
-//BEGIN_CIRCUIT
+interface Init {
+//BEGIN_INIT
 interface Circuit { int width(); }
-//END_CIRCUIT
-@Obj
-//BEGIN_FAN
 interface Fan extends Circuit {
   int _n();
   default int width() { return _n(); }
 }
-//END_FAN
-
-//BEGIN_STRETCH
-interface Stretch extends Circuit {
-  List<Integer> _ns();
-  Circuit _c();
-  default int width() {
-    return _ns().stream().reduce(0, (a,b) -> a+b);
-  }
+interface Identity extends Circuit {
+  int _n();
+  default int width() { return _n(); }
 }
-//END_STRETCH
-@Obj
-//BEGIN_BESIDE
 interface Beside extends Circuit {
   Circuit _c1();
   Circuit _c2();
@@ -36,59 +24,35 @@ interface Beside extends Circuit {
     return _c1().width() + _c2().width();
   }
 }
-//END_BESIDE
-@Obj
-//BEGIN_ID
-//interface Identity extends Circuit {
-//  int _n();
-//  default int width() { return _n(); }
-//}
-//END_ID
-interface Identity extends Circuit {
-  int _n();
+//END_INIT
+//BEGIN_SYNTAX
+interface Stretch extends Circuit {
+  List<Integer> _ns();
+  Circuit _c();
   default int width() {
-    return _n();
-  }
-  default Circuit desugar() {
-    Circuit res = Fan.of(1);
-    for (int i = 1; i < _n(); i++)
-      res = Beside.of(res, Fan.of(1));
-    return res;
+    return _ns().stream().reduce(0, (a,b) -> a+b);
   }
 }
-
-@Obj
-//BEGIN_ABOVE
 interface Above extends Circuit {
   Circuit _c1();
   Circuit _c2();
   default int width() { return _c2().width(); }
 }
-//END_ABOVE
+//END_SYNTAX
+}
 
 @Obj
-//BEGIN_CIRCUITWS
+interface Semantics extends Init {
+//BEGIN_SEMANTICS
 interface CircuitWS extends Circuit {
   boolean wellSized();
 }
-//END_CIRCUITWS
-@Obj
-//BEGIN_FANWS
+interface IdentityWS extends Identity, CircuitWS {
+  default boolean wellSize() { return true; }
+}
[…]

### 2016-04-26 paper/sections/Tool.tex
Delegating tasks
Text. Code generation.

@@
 
-How to generate code for family polymorphism!
+%Haoyuan should write this part.
+%How to generate code for family polymorphism!
+
+In this section, we present an overview of the code that \name generates. The syntax and type system are
+consistent with the Java language. We use translation functions to illustrate our code generation.
+To make it clearer, we split the process of code generation into two parts, in which case we introduce a
+new annotation \lstinline{@FamilyOf}, which is not defined in the source code but only helps to explain
+the translation. In our implementation, the annotation processing is a combination of \name
+and \lstinline{@FamilyOf}. Below are the translation functions:
+
+\begin{lstlisting}
+  (*$\llbracket$*)@Obj interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*)
+= @ObjOf interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth'}$*) (*$\overline{I'}$*)}
+\end{lstlisting}
+
+\subsection{@Obj}
+\lstinline{@Obj} builds the structure of family polymorphism for the annotated type. More specifically, \lstinline{@Obj} tackles two
+tasks: (1) building the dependencies (subtyping relations) between new family members and old ones; (2) refining field types. To this
+aim, \lstinline{@Obj} re-declares all member types and field methods from the base families. The detection runs recursively throughout
+the nested interfaces in base families. For member types with same names, the re-declaration in the annotated type is generated or modified
+with supertypes filled in for dependencies.
+
+\subsubsection{sameMembers}
+The auxiliary function \lstinline{sameMembers} generates the declaration of common member types from base families in the annotated type. It
+firstly checks if these members are homogeneous in their own hierarchies, and then fuses the old declarations into a new re-declaration, where
+all field types are automatically refined. Meanwhile, the code generation fills in the ``extends'' for the re-declaration, to build the inheritance
+relationship to the sources.
+\begin{itemize}
+\item $I_0\in$ \lstinline{sameMembers(}$\overline{I}$\lstinline{)}
+    \begin{itemize}
+    \item \lstinline{n = name(}$I_0$\lstinline{)}
+    \item $\overline{sup}$\lstinline{ = consistent(n, }$\overline{I}$\lstinline{)}
+    \item $I_0$\lstinline{ = fuse(n, }$\overline{sup}$\lstinline{, }$\overline{I}$\lstinline{)}
+    \end{itemize}
+\item $\overline{sup}$\lstinline{ = consistent(n, }$\overline{I}$\lstinline{)}
+    \begin{itemize}
+    \item \lstinline{I.n = interface n extends }$\overline{sup}$ \{...\} %style
+    \item $\forall$\lstinline{I.n}
+    \end{itemize}
+\item $I_0$\lstinline{ = fuse(n, }$\overline{sup}$\lstinline{, }$\overline{I}$\lstinline{)}
+    \begin{itemize}
+    \item $I_0$\lstinline{ = interface n extends }$\overline{sup}$ $\{\overline{meth}\ \ \overline{I_1}\}$ %style
+    \item $\overline{meth}$\lstinline{ = fuse(}$\overline{I.\overline{meth}}$\lstinline{)}
+    \item $\overline{I_1}$\lstinline{ = fuse(}$\overline{I.\overline{I_1}}$\lstinline{)}
+    \end{itemize}
+\end{itemize}
+
+\subsubsection{$\overline{I'}$}
+\begin{itemize}
+\item $\forall I_0\in$ \lstinline{sameMembers(}$\overline{I_s}$\lstinline{)}
+    \begin{itemize}
+    \item If $\exists I\in\overline{I}$, \lstinline{name(}$I_0$\lstinline{) = name(}$I$\lstinline{)}, then \lstinline{fuse2(}$I_0,I$\lstinline{)}$\in\overline{I'}$
+    \item Otherwise $I_0\in\overline{I'}$
+    \item More: generating \lstinline{extends}
+    \end{itemize}
+\end{itemize}
+
+\subsubsection{$\overline{meth'}$}
+\begin{itemize}
+\item $\overline{meth'}$ = $\overline{meth}$\lstinline{ + copyFields}
+\end{itemize}
+
+\subsection{@ObjOf}
+The second part of annotation processing is abstracted here using the newly introduced annotation
+\lstinline{@ObjOf}. It recursively generates constructer methods \lstinline{of} for the annotated interface
+and all nested interfaces. Furthermore, \lstinline{@ObjOf} supports getters, void and fluent setters, withers and
+the general \lstinline{with} methods. That is to say, \lstinline{@ObjOf} generates a static \lstinline{of} method
+for a type, returning an instance of that type. Such an instance is implemented using an anonymous class with auto-generated
+implementations for all the methods it supports (stated above). On the other hand, users are expected to put underscores as the
+prefix of field methods, and consequently \lstinline{of} identifies these field methods and takes them as its arguments.
+
+\begin{itemize}
+\item Recursively
+\item Use ``\_'' to distinguish fields and non-field methods
+\item Getters, setters, withers, etc
+\item \lstinline{of()} method
+\end{itemize}

### 2016-04-26 paper/sections/Overview.tex
More text on S3

@@
 
-\lstinputlisting[linerange=11-26]{../src/paper/Circuit.java}%APPLY:INIT
+\lstinputlisting[linerange=11-26]{../src/paper/sec3/Circuit.java}%APPLY:INIT
 We change the implementation language constructs from classes to interfaces and
@@
 Suppose that we would like to add two new constructs to the language:
-the first is \emph{stretch ns c}, which inserts additional wires into the circuit \emph{c} by summing up \emph{ns};
-the second is $above c_1 c_2$, which combines two circuits of the same width vertically.
-With these new constructs, more complex circuit can be constructed.
+
+\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
+summing up \emph{ns} and $above c_1 c_2$ combines two circuits of the same width vertically.
+With these new constructs, more complex circuits can be constructed.
 For example, Figure~\ref{} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.
 
-To accomplish this goal in Haskell, we can simply define two more cases,
+To accomplish this task in Haskell, we can simply define two more cases,
 \texttt{stretch} and \texttt{above}, for the semantic function:
-\lstinputlisting[linerange=59-60]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
+\lstinputlisting[language=haskell,linerange=45-46]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
 Similarly, defining two new interfaces
@@
 
-\lstinputlisting[linerange=29-40]{../src/paper/Circuit.java}%APPLY:SYNTAX
+\lstinputlisting[linerange=29-40]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
 
@@
 Gibbons and Wu worked around this problem in the following way:
-\lstinputlisting[linerange=47-51]{../src/paper/Circuit.java}%APPLY:HS_SEMANTICS
-
+\lstinputlisting[language=haskell,linerange=2-6]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
 which is similar to how we add \texttt{desugar} to the implementation in Section~\ref{}.
-The difference is that they use a tuple instead of a record to merge the
-semantic domains, then define the two semantic functions simultaneously for each
-case, and split the definitions through projections on the tuple in the end.
+The difference is that they use a tuple instead of a record to merge semantic
+domains, then define the two semantic functions simultaneously for each
+case, and split the semantic functions through projections on the tuple in the end.
 This solution, however, modifies existing code, breaking the requirement of EP.
@@
 %Different from records or tuples, interfaces are extensible.
-\lstinputlisting[linerange=47-75]{../src/paper/Circuit.java}%APPLY:SEMANTICS
+\lstinputlisting[linerange=47-75]{../src/paper/sec3/Circuit.java}%APPLY:SEMANTICS
 
@@
     creating objects;
-  \item Type-refinements should be done manually
-  \item The inheritance relation needs to be repeatedly stated for all interfaces.
+  \item Type-refinements should be done manually;
+  \item Similar inheritance relationships have to be repeatedly stated for all interfaces.
 \end{itemize}
-Also, programmers would not get warned if they forget to extend all the
+Worse still, programmers would not get warned if they forget to extend all the
 constructs with new semantics.
@@
 By using \name, we can refactor the extensions in the way shown in Figure\ref{}.
-\lstinputlisting[linerange=80-83]{../src/paper/Circuit.java}%APPLY:FAMILY
-\lstinputlisting[linerange=112-122]{../src/paper/Circuit.java}%APPLY:FAMILY_SYNTAX
-\lstinputlisting[linerange=126-141]{../src/paper/Circuit.java}%APPLY:FAMILY_SEMANTICS
+\lstinputlisting[linerange=80-82]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY
+\lstinputlisting[linerange=110-112]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SYNTAX
+\lstinputlisting[linerange=131-156]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SEMANTICS
 
-Note that \name generates \texttt{of} method for class-like interfaces for
-constructing objects. The static method \texttt{of} returns an instance of an anonymous class
-that implements the interface with all
-getters implemented and its instance is returned.
-Moreover, \name has a good support family polymorphism.
-Client users need only explicitly declared the dependencies on other families
-through (...). \name will re-declare all members from the base families in the extended family.
-And the dependencies between members of families are automatically deducted and
-implicitly expressed in member re-declaration.
+Note that initially all the definitions are put inside a \texttt{Family} interface.
+Extensions are defined inside another interface that extends base families.
+For syntax extension, we just move the definitions of new constructs into
+\texttt{SyntaxExt} interface.
+Semantic extension is more interesting, let us look at it in detail.
+Since all the definitions are nested interfaces, their names are local and can
+be reused in other families.
+The inheritance relationships is stated only once at family level, then \name can infer the inheritance relationships for members inside the family according to their names.
+Also, there is no need to
+manually refine the return type, as it is done by \name. Moreover, a static \texttt{of}
+method would be generated for class-like interfaces, serving as the constructor
+for the interface. For example, the \texttt{Beside} inside \texttt{Semantics} after code instrumentation will look like this:
+
+\lstinputlisting[linerange=164-176]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
+% TODO: not implemented yet
+Family polymorphism on only gives more safety on the client code but also helps
+language implementers catch bug early.
+For example, if one forgets to implement, say cases from \texttt{Family}, in
+\texttt{Semantics}, she would get warned they have not implemented the
+\texttt{wellSized} method because class-like interfaces should not contain any
+unimplemented methods except for getters.
+These errors are captured by automatically re-declaring members with inheritance relationships.
+%The \texttt{of} method is  returns an instance of an anonymous class
+%that implements the interface with all getters implemented and its instance is returned.
 

### 2016-04-26 paper/sections/ShallowOO.tex
More text on S3

@@
 
-\lstinputlisting[language=haskell,linerange=2-5]{./code/shallowCircuit.hs}%APPLY:CTYPE
+\lstinputlisting[language=haskell,linerange=2-2]{./code/shallowCircuit.hs}%APPLY:CIRCUIT_TYPE
+\lstinputlisting[language=haskell,linerange=5-7]{./code/shallowCircuit.hs}%APPLY:TYPES
 where \texttt{Circuit} models the semantic domain and should be later filled in with a
@@
 And the class implements the \texttt{Circuit} interface by giving \texttt{width}
-a definition. This implementation, in fact, is an instance of the \interp pattern.
+a definition. This whole implementation is, in fact, an instance of the \interp
+pattern and the combinator class \texttt{Beside} employs the \textsc{Composite} pattern.
 In other words, shallow embedding and the \interp are essentially the same thing.
@@
 We can implement this desugaring operation in Java:
-\lstinputlisting[language=haskell,linerange=29-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
+\lstinputlisting[linerange=29-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
 The return type of \texttt{desugar} is \texttt{Circuit} reflecting that it is a
@@
 
-\lstinputlisting[language=haskell,linerange=40-55]{./code/shallowCircuit.hs}%APPLY:DESUGAR
+\lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
 The definition is a little bit tricky since it makes use of Haskell's laziness.

### 2016-04-26 paper/code/Desugar.hs

@@ -1,0 +1,23 @@
+identity :: Int -> Circuit
+fan :: Int -> Circuit
+beside :: Circuit -> Circuit -> Circuit
+
+--BEGIN_DESUGAR
+data Circuit = Circuit {
+  width :: Int,
+  desugar :: Circuit
+}
+identity n = Circuit {
+  width = width c,
+  desugar = c
+} where c = foldl1 beside $ replicate n (fan 1)
+fan n = Circuit {
+  width = n,
+  desugar = fan n
+}
+beside c1 c2 = Circuit {
+  width = width c1 + width c2,
+  desugar = beside (desugar c1) (desugar c2)
+}
+--END_DESUGAR
+

### 2016-04-26 paper/code/NewSemantics.hs

@@ -1,0 +1,7 @@
+--BEGIN_SEMANTICS_HS
+identity n = (True,n)
+fan n = (True,n)
+beside c1 c2 = (fst c1 && fst c2,fst c1 + fst c2)
+stretch ns c = (fst c && length ns == snd c)
+above c1 c2 = (fst c1 && fst c2 && snd c1 == snd c2,snd c1)
+--END_SEMANTICS_HS

### 2016-04-26 paper/code/shallowCircuit.hs

@@ -1,17 +1,17 @@
-{-BEGIN_CTYPE
+{-BEGIN_CIRCUIT_TYPE
 type Circuit = ...
+END_CIRCUIT_TYPE-}
+--BEGIN_TYPES
 identity :: Int -> Circuit
 fan :: Int -> Circuit
 beside :: Circuit -> Circuit -> Circuit
-END_CTYPE-}
-
--- stretch :: [Int] -> Circuit -> Circuit
+--END_TYPES
 
 --BEGIN_CIRCUIT1
-type Circuit1 = Int
-identity1 n = n
-fan1 n = n
-beside1 c1 c2 = c1 + c2
+type Circuit = Int
+identity n = n
+fan n = n
+beside c1 c2 = c1 + c2
 
 width1 = id
 --END_CIRCUIT1
@@ -36,33 +36,12 @@ beside3 c1 c2 = Circuit3 {
 }
 --END_CIRCUIT3
 
---BEGIN_DESUGAR
-data Circuit = Circuit {
-  width :: Int,
-  desugar :: Circuit
-}
-identity n = Circuit {
-  width = width c,
-  desugar = c
-} where c = foldl1 beside $ replicate n (fan 1)
-fan n = Circuit {
-  width = n,
-  desugar = fan n
-}
-beside c1 c2 = Circuit {
-  width = width c1 + width c2,
-  desugar = beside (desugar c1) (desugar c2)
-}
---END_DESUGAR
+--BEGIN_SYNTAX_TYPES
+stretch :: [Int] -> Circuit -> Circuit
+above :: Circuit -> Circuit -> Circuit
+--END_SYNTAX_TYPES
 
 --BEGIN_SYNTAX_HS
---identity n = n
-above _ c2 = c2
+stretch ns c = sum ns
+above c1 c2 = c2
 --END_SYNTAX_HS
-
-
---BEGIN_SEMANTICS_HS
---fan2 n = (n, n > 0)
---beside2 c1 c2 = (fst c1 + fst c2, snd c1 && snd c2)
---stretch2 ns c = (sum ns, snd c && length ns == fst c)
---END_SEMANTICS_HS

### 2016-04-26 src/paper/sec3/Circuit.java

@@ -1,4 +1,4 @@
-package paper;
+package paper.sec3;
 
 import java.util.List;
 
@@ -36,13 +36,13 @@ interface Stretch extends Circuit {
 interface Above extends Circuit {
   Circuit _c1();
   Circuit _c2();
-  default int width() { return _c2().width(); }
+  default int width() { return _c1().width(); }
 }
 //END_SYNTAX
 }
 
 @Obj
-interface Semantics extends Init {
+interface SemanticsExt extends Init {
 //BEGIN_SEMANTICS
 interface CircuitWS extends Circuit {
   boolean wellSized();
@@ -78,24 +78,22 @@ interface AboveWS extends Beside, CircuitWS {
 /*
 //BEGIN_FAMILY
 @Obj interface Family {
-  // the same code in Figure .
-  ...
+  // same code of Figure ?
 }
 //END_FAMILY
 */
 @Obj interface Family {
   interface Circuit { int width(); }
-  interface Fan extends Circuit {
+  interface Identity extends Circuit {
     int _n();
     default int width() {
       return _n();
     }
   }
-  interface Stretch extends Circuit {
-    List<Integer> _ns();
-    Circuit _c();
+  interface Fan extends Circuit {
+    int _n();
     default int width() {
-      return _ns().stream().reduce(0, (a,b) -> a+b);
+      return _n();
     }
   }
   interface Beside extends Circuit {
@@ -107,12 +105,20 @@ interface AboveWS extends Beside, CircuitWS {
   }
 }
 
-
+/*
 //BEGIN_FAMILY_SYNTAX
-@Obj interface SyntaxExt extends Family {
-  interface Identity extends Circuit {
-    int _n();
-    default int width() { return _n(); }
+@Obj interface Syntax extends Family {
+  // same code of Figure ?
+}
+//END_FAMILY_SYNTAX
+*/
+@Obj interface Syntax extends Family {
+  interface Stretch extends Circuit {
+    List<Integer> _ns();
+    Circuit _c();
+    default int width() {
+      return _ns().stream().reduce(0, (a,b) -> a+b);
+    }
   }
   interface Above extends Circuit {
     Circuit _c1();
@@ -120,23 +126,53 @@ interface AboveWS extends Beside, CircuitWS {
     default int width() { return _c2().width(); }
   }
 }
-//END_FAMILY_SYNTAX
 
 //BEGIN_FAMILY_SEMANTICS
-@Obj interface SemanticsExt extends Family {
[…]

### 2016-05-10 paper/sections/Tool.tex
Translation rules.

@@
 
-\subsubsection{sameMembers}
-The auxiliary function \lstinline{sameMembers} generates the declaration of common member types from base families in the annotated type. It
-firstly checks if these members are homogeneous in their own hierarchies, and then fuses the old declarations into a new re-declaration, where
-all field types are automatically refined. Meanwhile, the code generation fills in the ``extends'' for the re-declaration, to build the inheritance
-relationship to the sources.
-\begin{itemize}
-\item $I_0\in$ \lstinline{sameMembers(}$\overline{I}$\lstinline{)}
-    \begin{itemize}
-    \item \lstinline{n = name(}$I_0$\lstinline{)}
-    \item $\overline{sup}$\lstinline{ = consistent(n, }$\overline{I}$\lstinline{)}
-    \item $I_0$\lstinline{ = fuse(n, }$\overline{sup}$\lstinline{, }$\overline{I}$\lstinline{)}
-    \end{itemize}
-\item $\overline{sup}$\lstinline{ = consistent(n, }$\overline{I}$\lstinline{)}
-    \begin{itemize}
-    \item \lstinline{I.n = interface n extends }$\overline{sup}$ \{...\} %style
-    \item $\forall$\lstinline{I.n}
-    \end{itemize}
-\item $I_0$\lstinline{ = fuse(n, }$\overline{sup}$\lstinline{, }$\overline{I}$\lstinline{)}
-    \begin{itemize}
-    \item $I_0$\lstinline{ = interface n extends }$\overline{sup}$ $\{\overline{meth}\ \ \overline{I_1}\}$ %style
-    \item $\overline{meth}$\lstinline{ = fuse(}$\overline{I.\overline{meth}}$\lstinline{)}
-    \item $\overline{I_1}$\lstinline{ = fuse(}$\overline{I.\overline{I_1}}$\lstinline{)}
-    \end{itemize}
-\end{itemize}
-
-\subsubsection{$\overline{I'}$}
-\begin{itemize}
-\item $\forall I_0\in$ \lstinline{sameMembers(}$\overline{I_s}$\lstinline{)}
-    \begin{itemize}
-    \item If $\exists I\in\overline{I}$, \lstinline{name(}$I_0$\lstinline{) = name(}$I$\lstinline{)}, then \lstinline{fuse2(}$I_0,I$\lstinline{)}$\in\overline{I'}$
-    \item Otherwise $I_0\in\overline{I'}$
-    \item More: generating \lstinline{extends}
-    \end{itemize}
-\end{itemize}
-
-\subsubsection{$\overline{meth'}$}
-\begin{itemize}
-\item $\overline{meth'}$ = $\overline{meth}$\lstinline{ + copyFields}
-\end{itemize}
-
 \subsection{@ObjOf}
@@
 
+\begin{figure*}
+\begin{lstlisting}
+  (*$\llbracket$*)@Family interface (*$I_m$*) extends (*$I_1,\cdots,I_n$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*) = (*$\llbracket$*)@Obj interface (*$I_m$*) extends (*$I_1,\cdots,I_n$*) {(*$\overline{meth'}$*) (*$\overline{I'}$*)}(*$\rrbracket$*)
+\end{lstlisting}
+\hspace{.3in}where $\overline{meth'}=\overline{meth}\ \cup\ $\textsf{fieldMethods(}$I_1$\textsf{)}$\ \cup\cdots\cup\ $\textsf{fieldMethods(}$I_n$\textsf{)}, $\overline{I'}$ = \textsf{newChilds(}$I_1,\cdots,I_n,\overline{I}$\textsf{)}
+~\\~\\
+(1) $\overline{I'}$ = \textsf{newChilds(}$I_1,\cdots,I_n,\overline{I}$\textsf{)} \textcolor{red}{(definition of newChilds)}
     \begin{itemize}
-\item Recursively
-\item Use ``\_'' to distinguish fields and non-field methods
-\item Getters, setters, withers, etc
-\item \lstinline{of()} method
+    \item $\forall I_0\in\overline{I}$, if $\exists I'_0\in\ $\textsf{collectMembers(}$I_1,\cdots,I_n$\textsf{)} and \textsf{name(}$I_0$\textsf{)} = \textsf{name(}$I'_0$\textsf{)}, then \textsf{combine(}$I_0,I'_0$\textsf{)}$\ \in\overline{I'}$
+    \item Otherwise, $I_0\in\overline{I'}$
     \end{itemize}
+(2) $\llbracket$\lstinline{@Family}\textsf{ interface }$I$\textsf{ extends }$\overline{I_s},\ \overline{I_t}$\textsf{ \{\}}$\rrbracket\ \in$ \textsf{collectMembers(}$I_1,\cdots,I_n$\textsf{)} \textcolor{red}{(definition of collectMembers)}
+    \begin{itemize}
+    \item $\forall i,$ s.t. $I\in$\textsf{ childs(}$I_i$\textsf{)},
+        \begin{itemize}
+        \item $\overline{I_s}$ = \textsf{suptypes(}$I_i.I$\textsf{)}
+        \item $I_i.I\in\overline{I_t}$
+        \end{itemize}
+    \end{itemize}
+(3) $\overline{I_0}$ = \textsf{childs(}$I$\textsf{)}, $\overline{I}$ = \textsf{suptypes(}$I$\textsf{)} \textcolor{red}{(definition of childs and suptypes)}
+    \begin{itemize}
+    \item \textsf{ibody(}$I$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I}$\textsf{ \{}$\overline{meth}\ \overline{I_0}$\textsf{\}}
+    \item \textcolor{red}{need to use ibody(I) to find the declaration. formalize?}
+    \end{itemize}
+(4) $meth\in\ $\textsf{fieldMethods(}$I$\textsf{)} \textcolor{red}{(definition of fieldMethods)}
+    \begin{itemize}
+    \item $meth\in\ $\textsf{childs(}$I$\textsf{)}
+    \item $meth$ = $I_0$\textsf{ m();}
+    \end{itemize}
+(5) \textsf{interface }$I$\textsf{ extends }$\overline{I_s}$\textsf{ \{}$\overline{meth}\ \overline{I}$\textsf{\}} = \textsf{combine(}$I_m,I_n$\textsf{)} \textcolor{red}{(definition of combine)}
+    \begin{itemize}
+    \item \textsf{ibody(}$I_m$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I_{s1}}$\textsf{ \{}$\overline{meth_1}\ \overline{I_1}$\textsf{\}}
+    \item \textsf{ibody(}$I_n$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I_{s2}}$\textsf{ \{}$\overline{meth_2}\ \overline{I_2}$\textsf{\}}
+    \item $\overline{I_s}$ = $\overline{I_{s1}}\ \cup\ \overline{I_{s2}}$
+    \item $\overline{meth}$ = $\overline{meth_1}\ \cup\ \overline{meth_2}$
+    \item If $\exists I_1\in\overline{I_1}, I_2\in\overline{I_2}$, \textsf{name(}$I_1$\textsf{)} = \textsf{name(}$I_2$\textsf{)}, then \textsf{combine(}$I_1,I_2$\textsf{)}$\ \in\overline{I}$
+    \item Otherwise $(I\in I_1\ \Delta\ I_2)$, $I\in\overline{I}$.
+    \end{itemize}
+\caption{Translation of \lstinline{@Family}.}
+\end{figure*}
+
+\begin{figure*}
+\begin{lstlisting}
+  (*$\llbracket$*)@Obj interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*) = (*$\llbracket$*)@ObjOf interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) (*$\overline{meth'}$*) (*$\overline{\llbracket\textsf{@Obj}\ I\rrbracket}$*)}(*$\rrbracket$*)
+\end{lstlisting}
+\hspace{.3in}where $\overline{meth'}$ = \textsf{refine(}$I_0,\overline{meth}$\textsf{)}
+\begin{lstlisting}
+  (*$\llbracket$*)@ObjOf interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*) = interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) ofMethod((*$I_0$*)) (*$\overline{I}$*)}
+\end{lstlisting}
+\hspace{.3in}with \textsf{valid(}$I_0$\textsf{)}, \textsf{of} $\notin$ \textsf{dom(}$I_0$\textsf{)}
+~\\~\\
+(1) $I_0$ \textsf{with}$\#m$\textsf{(}$I\ \_$\textsf{val);} $\in$ \textsf{refine(}$I_0,\overline{meth}$\textsf{)} \textcolor{red}{(part I definition of newChilds) fields with underscore, field and isField}
+    \begin{itemize}
+    \item \textsf{isWith(mbody(with}$\#m,I_0$\textsf{)}$,I_0$\textsf{)}
+    \item \textsf{with}$\#m$ $\notin$ \textsf{dom(}$\overline{meth}$\textsf{)}
+    \end{itemize}
+(2) $I_0\ \_m$\textsf{(}$I\ \_$\textsf{val);} $\in$ \textsf{refine(}$I_0,\overline{meth}$\textsf{)} \textcolor{red}{(part II definition of newChilds)}
+    \begin{itemize}
+    \item \textsf{isSetter(mbody(}$\_m,I_0$\textsf{)}$,I_0$\textsf{)}
+    \item $\_m$ $\notin$ \textsf{dom(}$\overline{meth}$\textsf{)}
+    \end{itemize}
+(3) \textsf{valid(}$I_0$\textsf{)} if $\forall m\ \in\ $\textsf{dom(}$I_0$\textsf{)}, let $meth$ = \textsf{mbody(}$m,I_0$\textsf{)}, one of the following cases is satisfied:  \textcolor{red}{(definition of valid)}
+    \begin{itemize}
+    \item \textsf{isField(}$meth$\textsf{)}, where \textsf{isField(}$I\ m$\textsf{();)} = not \textsf{special(}$m$\textsf{)}
+    \item \textsf{isWith(}$meth,I_0$\textsf{)}, where \textsf{isWith(}$I'$ \textsf{with}$\#m$\textsf{(}$I\ x$\textsf{);}$,I_0$\textsf{)} = $I_0 :< I'$, \textsf{mbody(}$m,I_0$\textsf{)} = $I\ m$\textsf{();} and not \textsf{special(}$m$\textsf{)}
+    \item \textsf{isSetter(}$meth,I_0$\textsf{)}, where \textsf{isSetter(}$I'$ $\_m$\textsf{(}$I\ x$\textsf{);}$,I_0$\textsf{)} = $I_0 :< I'$, \textsf{mbody(}$m,I_0$\textsf{)} = $I\ m$\textsf{();} and not \textsf{special(}$m$\textsf{)}
+    \end{itemize}
+(4) \textsf{ofMethod(}$I_0$\textsf{)} = \textsf{static }$I_0$\textsf{ of(}$I_1\ \_m_1,\cdots,I_n\ \_m_n$\textsf{) \{} \textcolor{red}{(definition of ofMethod)}
+    \\ \textsf{return new }$I_0$\textsf{() \{}
+    \\ $I_1\ m_1$ = $\_m_1$\textsf{;}$\cdots I_n\ m_n$ = $\_m_n$\textsf{;}
+    \\ $I_1\ m_1$\textsf{() \{return }$m_1$\textsf{;\}}$\cdots I_n\ m_n$\textsf{() \{return }$m_n$\textsf{;\}}
+    \\ \textsf{withMethod(}$I_1,m_1,I_0,\overline{e}_1$\textsf{)}$\cdots$\textsf{withMethod(}$I_n,m_n,I_0,\overline{e}_n$\textsf{)}
+    \\ \textsf{setterMethod(}$I_1,m_1,I_0$\textsf{)}$\cdots$\textsf{setterMethod(}$I_n,m_n,I_0$\textsf{)}
+    \\ \textsf{\};\}}
+    \begin{itemize}
+    \item $I_1\ m_1$\textsf{();}$\cdots I_n\ m_n$\textsf{();} = \textsf{fields(}$I_0$\textsf{)}
+    \item $\overline{e}_i$ = $m_1,\cdots,m_{i-1},\_$\textsf{val}$,m_{i+1},\cdots,m_n$
+    \end{itemize}
+(5) $meth$ $\in$ \textsf{fields(}$I_0$\textsf{)} \textcolor{red}{(definition of fields)}
+    \begin{itemize}
+    \item \textsf{isField(}$meth$\textsf{)}
+    \item $meth$ = \textsf{mbody(}$m^{meth},I_0$\textsf{)}
+    \end{itemize}
+(6) $I_0$ \textsf{with}$\#m$\textsf{(}$I\ \_$\textsf{val) \{return }$I_0$\textsf{.of(}$\overline{e}$\textsf{);\}} = \textsf{withMethod(}$I,m,I_0,\overline{e}$\textsf{)} \textcolor{red}{(definition of withMethod)}
+    \begin{itemize}
+    \item \textsf{mbody(with}$\#m,I_0$\textsf{)} has the form \textsf{mh;}
+    \end{itemize}
+(7) $I_0$ $\_m$\textsf{(}$I\ \_$\textsf{val) \{}$m$ = $\_$\textsf{val;return this;\}} = \textsf{setterMethod(}$I,m,I_0$\textsf{)} \textcolor{red}{(definition of setterMethod)}
+    \begin{itemize}
+    \item \textsf{mbody(}$\_m,I_0$\textsf{)} has the form \textsf{mh;}
+    \end{itemize}
+\caption{Translation of \lstinline{@Obj} and \lstinline{@ObjOf}.}
+\end{figure*}

### 2016-05-14 paper/sections/Discussion.tex
Delegating tasks
More text.

@@
 \section{Discussion and Limitations}
+
+Haoyuan writes this one.
 
@@
   - Other languages?
+
+\paragraph{Transformations and binary methods} Some other patterns, however, are not captured by our approach.
+
+\paragraph{Limitations} Our tool has certain limitations at a few aspects. As Lombok only provides experimental support
+for separate compilation, our \textsf{@Family} does not support separate compilation at this stage, so all the related
+interfaces should be included in a single Java file. As for generics, we provide comparatively nice support for generic interfaces
+and methods, however, only method typing of non-generic methods is captured, and that of generic ones is delegated to the Java compiler.
+
+On the other hand, it would be nice to have operator overloading and infix operators in Java syntax, so that the code could be
+written in a more concise and elegant way. Nevertheless, this is limited by the programming language, and we expect our approach to be oriented to
+other OO languages.

### 2016-05-14 paper/sections/Tool.tex
More text.

@@
 
-In this section, we present an overview of the code that \name generates. The syntax and type system are
+In this section, we present an overview of the code that \lstinline{@Family} generates. The syntax and type system are
 consistent with the Java language. We use translation functions to illustrate our code generation.
-To make it clearer, we split the process of code generation into two parts, in which case we introduce a
-new annotation \lstinline{@FamilyOf}, which is not defined in the source code but only helps to explain
-the translation. In our implementation, the annotation processing is a combination of \name
-and \lstinline{@FamilyOf}. Below are the translation functions:
+To make it clearer, we split the process of code generation into three parts, in which case we introduce two
+new annotations \lstinline{@Obj} and \lstinline{@ObjOf}, which are not implemented in the source code but just help to explain
+the process of translation. In our implementation, the annotation processing is a combination of the three parts, corresponding to
+\lstinline{@Family}, \lstinline{@Obj} and \lstinline{@ObjOf} respectively. The translation rules are presented in Figure \ref{fig:trans1}
+and Figure \ref{fig:trans2}, together with some auxiliary functions.
 
-\begin{lstlisting}
-  (*$\llbracket$*)@Obj interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*)
-= @ObjOf interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth'}$*) (*$\overline{I'}$*)}
-\end{lstlisting}
+\subsection{\textsf{@Family}}
+\lstinline{@Family} builds the structure of family polymorphism for the annotated type. More specifically, \lstinline{@Family} tackles two
+tasks: (1) building the dependencies (subtyping relations) between new family members and old ones; (2) refining field types. To this
+aim, \lstinline{@Family} re-declares all member types and field methods from the base families (see \textsf{collectMembers} and \textsf{fieldMethods} in Figure \ref{fig:trans1}).
+\lstinline{@Family} recognizes field methods by checking if they are non-void, no-argument methods, and if the name starts with an underscore ``\_''.
+The function \textsf{collectMembers} finds all the direct member types from base families $I_1,\cdots,I_n$, creates new types with the same names for re-declaration, and builds the subtyping relations among them.
+The newly created types are also annotated with \lstinline{@Family}, leading to such a recursive process throughout the nested interfaces in base families.
+If users have already declared these member types in order to introduce new operations, behaviors or whatever, the function \textsf{combine} will integrate them to the re-declaration, by combining methods and supertypes following ``extends''.
 
-\subsection{@Obj}
-\lstinline{@Obj} builds the structure of family polymorphism for the annotated type. More specifically, \lstinline{@Obj} tackles two
-tasks: (1) building the dependencies (subtyping relations) between new family members and old ones; (2) refining field types. To this
-aim, \lstinline{@Obj} re-declares all member types and field methods from the base families. The detection runs recursively throughout
-the nested interfaces in base families. For member types with same names, the re-declaration in the annotated type is generated or modified
-with supertypes filled in for dependencies.
+\subsection{\textsf{@Obj}}
+The second part of annotation processing is abstracted here using the newly introduced annotation
+\lstinline{@Obj}. The translation of \lstinline{@Obj} is like a preprocessing of the third part \lstinline{@ObjOf},
+which generates constructor method \textsf{of} for the annotated type. Besides the generation of constructors, \lstinline{@ObjOf} also supports
+a series of field and non-field methods, including getters, void and fluent setters, \textsf{with}- methods and the general \textsf{with} method.
+\textcolor{red}{Haoyuan: Details on these operations?
+Can draw a table.} Among them, the fluent setters, \textsf{with}- methods and the general \textsf{with()} method all take the annotated type as their return types, and these types can be refined in some subtypes of the annotated type.
+What \lstinline{@Obj} precisely does is that it collects all these refinable methods from the super interfaces, and refines them to the annotated type, so that \lstinline{@ObjOf} will later provide a default implementation for them during the generation of constructor methods.
 
-\subsection{@ObjOf}
-The second part of annotation processing is abstracted here using the newly introduced annotation
-\lstinline{@ObjOf}. It recursively generates constructer methods \lstinline{of} for the annotated interface
-and all nested interfaces. Furthermore, \lstinline{@ObjOf} supports getters, void and fluent setters, withers and
-the general \lstinline{with} methods. That is to say, \lstinline{@ObjOf} generates a static \lstinline{of} method
-for a type, returning an instance of that type. Such an instance is implemented using an anonymous class with auto-generated
+In Figure \ref{fig:trans2}, the first translation function presents the processing of \lstinline{@Obj}.
+Besides delegating \lstinline{@Obj} to the member types in a recursive way, it invokes the function \textsf{refine} to handle the type refinements.
+The definition of \textsf{refine} is shown in (1) and (2), where return types of \textsf{with-} methods and fluent setters are refined respectively.
+They are both state operations, and for simplicity we omit the case for general \textsf{with} method in our translation rules.
+
+\subsection{\textsf{@ObjOf}}
+The third part of translation relies on \textsf{@ObjOf}, which directly generates the static \textsf{of} method, serving as a constructor method.
+So it returns an instance of the annotated type.
+This method takes one argument for each field method from the domain of the interface.
+Similarly, it recursively generates \lstinline{of} methods for the annotated interface and all nested interfaces.
+
+The translation function of \textsf{@ObjOf} is shown in the second rule of Figure \ref{fig:trans2}.
+It firstly uses \textsf{valid} to check whether all the abstract methods in the interface are valid; that is to say, any one of them is either a field (getter) method, a \textsf{with-} method or a setter method.
+We capture these patterns so that \textsf{@ObjOf} can provide default implementations for them.
+Then, if \textsf{valid} is satisfied, a static \textsf{of} method is generated in the interface to return an instance of it.
+Such an instance is implemented in the return statement using an anonymous class with auto-generated
 implementations for all the methods it supports (stated above). On the other hand, users are expected to put underscores as the
@@
     \item \textsf{ibody(}$I$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I}$\textsf{ \{}$\overline{meth}\ \overline{I_0}$\textsf{\}}
-    \item \textcolor{red}{need to use ibody(I) to find the declaration. formalize?}
+    \item \textcolor{red}{need to use ibody and mbody. formalize?}
     \end{itemize}
@@
     \item $meth\in\ $\textsf{childs(}$I$\textsf{)}
-    \item $meth$ = $I_0$\textsf{ m();}
+    \item $meth$ = $I_0$\textsf{ \_m();}
     \end{itemize}
@@
 \caption{Translation of \lstinline{@Family}.}
+\label{fig:trans1}
 \end{figure*}
@@
 \caption{Translation of \lstinline{@Obj} and \lstinline{@ObjOf}.}
+\label{fig:trans2}
 \end{figure*}

### 2016-05-22 paper/sections/Discussion.tex
More text Discussion.

--- 2016-04-19 paper/sections/Discussion.tex
@@
 
-- Transformations and Binary methods (but argue that 
-these types of operations are not used in shallow DSLs 
-any way!).
+%Haoyuan writes this one.
 
-- Java syntax not great for DSLs. 
-operator overloding; infix operators (but this is a Java limitation; 
-not a limitation of our approach).
+%- Transformations and Binary methods (but argue that
+%these types of operations are not used in shallow DSLs
+%any way!).
 
-- Limitations of the tool: separate compiltions and ... 
-generics not fully implemented.
+%- Java syntax not great for DSLs.
+%operator overloding; infix operators (but this is a Java limitation;
+%not a limitation of our approach).
 
-- How to apply our approach to other OO language 
-  - C\# has annotations, but can you do AST rewriting?
-  - Scala has macros;
-  - Other languages?
+%- Limitations of the tool: separate compiltions and ...
+%generics not fully implemented.
+
+%- How to apply our approach to other OO language
+%  - C\# has annotations, but can you do AST rewriting?
+%  - Scala has macros;
+%  - Other languages?
+
+\paragraph{Transformations} Our approach tends to give a taste that shallow embeddings are convenient to
+use in object-oriented languages, and extensible with family polymorphism. Nevertheless, shallow embeddings are also under
+criticism in terms of its not supporting ASTs explicitly, leading to the fact that some operations like transformations
+cannot be easily used with shallow embeddings. In Section ?(2.3) we give an example of desugaring in Java using shallow
+embeddings, however when we try to apply family polymorphism for inheritance and extensibility, code duplication can be introduced.
+Suppose we firstly integrate the four members \textsf{Circuit}, \textsf{Fan}, \textsf{Beside} and \textsf{Identity} in a base family,
+and then define a new family that extends the base one, then inside the new member \textsf{Identity}, how can we implement the transformation \textsf{desugar()}?
+Certainly we cannot use \textsf{super} to call the old implementation, otherwise we are invoking old constructors of \textsf{Beside},
+leading to type errors afterwards. For sure we can just copy and paste the code for the old implementation, but too much
+code duplication would depress the users. Another solution is to use F-bounded polymorphism (\textcolor{red}{Haoyuan: refs. Need example?
+We do have an example in the Pretty Printer case study}),
+in that case code duplication is avoided, but accordingly, it arouses explosively heavy uses of types and parameterisations.
+
+\paragraph{Binary methods} Binary methods are indeed another sort of operations that prevent users from using shallow embeddings in Java
+comfortably. \textcolor{red}{Haoyuan: refs?} Java supports covariant return types, hence we can refine the types of field methods and 
+simple interpretations, but with binary methods like \textsf{equals}, method subtyping in inheritance is a big problem since the argument type
+is changed. In Scala, however, we can deal with binary methods with the help of ? \textcolor{red}{Haoyuan: Could you help Weixin?}.
+
+\paragraph{Shallow or deep embeddings?} We know the fact that our approach integrates shallow embeddings and simple family polymorphism, and that in some object-oriented languages like Java,
+some trouble is introduced when dealing with extensible transformations or binary methods. But on the other hand, these types of operations are not used in
+shallow DSLs anyway; with a single interpretation we use shallow embeddings, but since shallow embeddings do not represent abstract syntax trees,
+it is really hard to realize transformations, and similarly, with binary methods, users tend to use deep embeddings instead to build data hierarchies.
+
+\paragraph{Limitations} Our tool has its certain limitations at a few aspects. As Lombok only provides experimental support
+for separate compilation, our \textsf{@Family} does not support separate compilation at this stage, so all the related
+interfaces should be included in a single Java file. On generics, our implementation of \textsf{@Family} provides support
+for generic interfaces and methods without bounds, on the other hand, generic method typing is not explicitly captured
+by the annotation but delegated to the Java compiler. \textcolor{red}{Haoyuan: We still need to talk about Lombok in detail?
+Another limitation is that only Eclipse handler is supported.}
+
+Regarding syntax, it would be nice to have operator overloading and infix operators in Java, so that the code could be
+written in a more concise and elegant way. Nevertheless, this is limited by the programming language, and we expect our approach to be oriented to
+other OO languages. \textcolor{red}{Haoyuan: How to apply our approach to other OO language? C\# has annotations, but can you do AST rewriting? Scala has macros;  Other languages?}

### 2016-05-22 paper/sections/Tool.tex
More text Discussion.

@@
 \lstinline{@Family}, \lstinline{@Obj} and \lstinline{@ObjOf} respectively. The translation rules are presented in Figure \ref{fig:trans1}
-and Figure \ref{fig:trans2}, together with some auxiliary functions.
+and Figure \ref{fig:trans2}, together with some auxiliary functions. Note that in practice, our \textsf{@Family} supports simple
+generics without bounds, but for simplicity reason it is not included in the formalization.
+
+\begin{figure*}
+\begin{tabular}{|l|c|c|}
+\hline
+\begin{lstlisting}
+interface Base {
+	interface Exp {}
+	interface Lit extends Exp {
+		int i();
+		Lit withI(int val);
+	}
+	interface Add extends Exp {
+		Exp e1(); Exp e2();
+	}
+}
+\end{lstlisting} & & \\
+\hline
+\begin{lstlisting}
+@Family interface New extends Base {
+	interface Exp { int _eval(); }
+	interface Lit {
+		default int _eval() { return i(); }
+	}
+	interface Add {
+		default int _eval() {
+			return e1()._eval() + e2()._eval();
+		}
+	}
+}
+\end{lstlisting} & \begin{tabular}{c}$\Rightarrow$\\Family\\Desugaring\end{tabular} & \begin{lstlisting}
+@Obj interface New extends Base {
+	@Obj interface Exp extends Base.Exp {
+		int _eval();
+	}
+	@Obj interface Lit extends Exp, Base.Lit {
+		int i();
+		default int _eval() { return i(); }
+	}
+	@Obj interface Add extends Exp, Base.Add {
+		Exp e1(); Exp e2();
+		default int _eval() {
+			return e1()._eval() + e2()._eval();
+		}
+	}
+}
+\end{lstlisting} \\
+\hline
+& & $\Downarrow$ Obj Desugaring \\
+\hline
+\begin{lstlisting}
+interface New extends Base {
+	interface Exp extends Base.Exp {
+		int _eval();
+	}
+	interface Lit extends Exp, Base.Lit {
+		int i(); Lit withI(int val);
+		default int _eval() { return i(); }
+		static Lit of(int _i) {
+			return new Lit() {
+				int i = _i;
+				public int i() { return i; }
+				public Lit withI(int val) {
+					return Lit.of(val);
+				}
+			};
+		}
+	}
+	interface Add extends Exp, Base.Add {
+		Exp e1(); Exp e2();
+		default int _eval() {
+			return e1()._eval() + e2()._eval();
+		}
+		static Add of(Exp _e1, Exp _e2) {
+			return new Add() {
+				Exp e1 = _e1; Exp e2 = _e2;
+				public Exp e1() { return e1; }
+				public Exp e2() { return e2; }
+			};
+		}
+	}
+	static New of() { return new New(){}; }
+}
+\end{lstlisting} & \begin{tabular}{c}$\Leftarrow$\\ObjOf\\Desugaring\end{tabular} & \begin{tabular}{l}\begin{lstlisting}
+@ObjOf interface New extends Base {
+	@ObjOf interface Exp extends Base.Exp {
+		int _eval();
+	}
+	@ObjOf interface Lit extends Exp, Base.Lit {
+		int i();
+\end{lstlisting}\vspace{-.05in}\\ \begin{lstlisting}[basicstyle=\ttfamily\small\color{red}]
+		Lit withI(int val);
+\end{lstlisting}\\ \begin{lstlisting}
+		default int _eval() { return i(); }
+	}
+	@ObjOf interface Add extends Exp, Base.Add {
+		Exp e1(); Exp e2();
+		default int _eval() {
+			return e1()._eval() + e2()._eval();
+		}
+	}
+}
+\end{lstlisting}\end{tabular} \\
+\hline
+\end{tabular}
+\caption{An example showing the flow of translation. \textcolor{red}{Haoyuan: fields with underscore}}\label{fig:flow}
+\end{figure*}
+
+\subsection{Flow of Translation}
+Figure~\ref{fig:flow} illustrates the flow of translation with a simple example, showing intermediate results of desugaring.
+The base family \textsf{Base} contains three members, \textsf{Exp}, and its subtypes \textsf{Lit} and \textsf{Add}.
+\textsf{Lit} has a field method \textsf{i()} together with a wither method, where \textsf{Add} has two field methods \textsf{e1()} and \textsf{e2()}.
+At this time, another interface \textsf{New} is defined as the new family that extends the old one, where we add an operation \textsf{\_eval()} for the member types.
+In the client code, the user simply re-declares the three members with the addition of \textsf{\_eval()}, without building subtyping relations with the old members, and a simple annotation \textsf{@Family} will deal with all the stuff.
+
+In the first step of translation, namely family desugaring, \textsf{@Family} collects all the members from old families, and fills in the ``extends'' to build the inheritance relations.
+At the same time, it re-declares all the field methods, for example, although the type name of \textsf{e1()} is still \textsf{Exp}, it refers to the new member type, where we address field type refinements, and this new \textsf{e1()} has the operation \textsf{\_eval()}.
+Then it delegates the work to \textsf{@Obj}.
+In the second phase, \textsf{@Obj} detects fluent setters and withers from the supertypes, then refines their return types to have consistency with the annotated type.
+In our example, the wither method \textsf{withI} is refined to have the new \textsf{Lit} as its return type.
+In the last step of translation, \textsf{@ObjOf} generates constructor methods for all the interfaces, except \textsf{Exp}, where its \textsf{\_eval()} is an abstract method and we cannot provide a default implementation for it.
+In the following subsections we will respectively explain the three steps of translation in detail.
 
@@
 If users have already declared these member types in order to introduce new operations, behaviors or whatever, the function \textsf{combine} will integrate them to the re-declaration, by combining methods and supertypes following ``extends''.
+Note that we use $\textsf{ibody(}I_m$\textsf{)} to find the exact declaration body of interface $I_m$.
+\textcolor{red}{Haoyuan: Acutally in the implementation, \textsf{ibody} takes two arguments, one is the name of type reference (interface name $I_m$), and the other is the environment, or where the type reference $I_m$ locates.}
 
@@
 In Figure \ref{fig:trans2}, the first translation function presents the processing of \lstinline{@Obj}.
+Note that we use $\textsf{mbody(}m,I\textsf{)}$ to get the full declaration of method $m$ seen by $I$, such a method can either be directly defined in $I$ or inherited from its supertypes.
 Besides delegating \lstinline{@Obj} to the member types in a recursive way, it invokes the function \textsf{refine} to handle the type refinements.
@@
     \item \textsf{ibody(}$I$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I}$\textsf{ \{}$\overline{meth}\ \overline{I_0}$\textsf{\}}
-    \item \textcolor{red}{need to use ibody and mbody. formalize?}
     \end{itemize}

### 2016-05-23 paper/sections/CaseStudyPPrinter.tex
Case study: PPrinter.

\subsection{A Prettier Printer}

This case study refactors the Haskell code from \textcolor{red}{(Haoyuan: ref, a prettier printer)}
which implements a functional pretty printer library with high efficiency. In the original code, two
data structures are defined for documents. Here they are defined using shallow DSLs, and packaged into
two base families with \textsf{@Family} annotation applied. These two families \textsf{Family\_Doc} and
\textsf{Family\_Document} correspond to data types \textsf{Doc} and \textsf{DOC} respectively in original code.

\begin{lstlisting}
@Family interface Family_Doc {
	interface Doc {}
	interface Nil extends Doc {}
	interface Text extends Doc {
		String s(); Doc d();
	}
	interface Line extends Doc { int i(); Doc d(); }
}

@Family interface Family_Document {
	interface Document {}
	interface DNil extends Document {}
	interface DConcat extends Document {
		Document d1(); Document d2();
	}
	interface DNest extends Document {
		int i(); Document d();
	}
	interface DText extends Document { String s(); }
	interface DLine extends Document {}
	interface DUnion extends Document {
		Document d1(); Document d2();
	}
}
\end{lstlisting}

With the help of family polymorphism and \textsf{@Family}, we can easily add new operations and integrate them
in child families. Below is an example of encoding two original functions \textsf{layout} and \textsf{fits} in
the shallow embeddings. Note that with the annotation processing, verbose code for building inheritance relations is
automatically generated.
\begin{lstlisting}
@Family interface Family_Doc_LayoutFits extends Family_Doc {
	interface Doc {
		String _layout(); boolean _fits(int w);
	}
	interface Nil {
		default String _layout() { return ""; }
		default boolean _fits(int w) {
			return w >= 0;
		}
	}
	...
\end{lstlisting}
Furthermore, some functions like \textsf{be} and \textsf{beaux} need refactoring before they are encoded in shallow embeddings.
Finally we refactored the pretty printer library using a set of extensible families with shallow DSLs and operations, and our
\textsf{@Family} annotation. However there is a special operation called \textsf{group}, which works like a transformation:
\begin{lstlisting}
@Family interface Family_Document_Flatten extends Family_Document {
	interface Document {
		Document _flatten();
		default Document _group() {
			return DUnion.of(this._flatten(), this);
		}
	}
	...
}
\end{lstlisting}
The method \textsf{\_group} calls the constructor method from \textsf{DUnion}, whereas such a member type will be updated in the
child families, together with its constructors. To ensure type safety we are unable to reuse the code but just copy it and paste
to the child families, and hence code duplication is introduced. The matter of extensible transformations will be further discussed
in the next section.

### 2016-05-24 paper/sections/Overview.tex
Revision on Sec 2 & 3

@@
-\section{Overview}
-\weixin{title: Evolution on DSLs?}
+\section{Evolution on DSLs with \name}
 
@@
 %As language evolves, the need of new syntax often arises.
-As DSLs evolve along the time, the demand for new syntax and and semantics may arise.
+As DSLs evolve along the time, the demand for new syntax and new semantics may arise.
 It would be good if we can introduce these new features to the DSL \emph{modularly}.
 %This is actually a hard problem, known as the Expression Problem (EP)~\cite{wadler}, which
-This, however, requires the host language equipped with two dimensions of
-extensibility.
-In this section, we argue that from extensibility perspective, OO languages are
-better candidates as host languages than FP languages.
-To demonstrate this, we try to extend \dsl with new syntax and new semantics in
-the setting of shallow embedding. Adding new syntax is both easy for both FP
-and OOP. Adding new semantics, however, is hard in FP. Although it is possible
-in OOP, the solution requires some boilerplate code. We hence developed \name
-for defining modular extensions easily.
+%This, however, requires the host language equipped with two dimensions of extensibility.
+For shallow embedding, adding new syntax is trivial - just adding new cases in
+the semantic function. On the other hand, adding new semantics becomes hard as it allows
+only one definition for each case. Gibbons and Wu summarized how to define
+various type of semantic functions (interpretations) in shallow embedding by
+applying \emph{fold theory}. However, the encoding is based on tuples, which is
+not modular extensible. Though they also presented modular solution based on
+datatype a lar carte\cite{}. But such solution changes the encoding style
+dramatically and may be too complex to use practically.
+In contrast, the OO approach allows all kinds of
+interpretations to be defined in an modular intuitive way.
+In this section, we will show how various types of interpretations can be
+defined using the OO way with the help of \name. Along the way,
+some features of \name will be introduced, which simplify our implementation significantly.
+
+%argue that from extensibility perspective, OO languages are
+%better candidates as host languages than FP languages.
+%To demonstrate this, we try to extend \dsl with new syntax and new semantics in
+%the setting of shallow embedding. Adding new syntax is both easy for both FP
+%and OOP. Adding new semantics, however, is hard in FP. Although it is possible
+%in OOP, the solution requires some boilerplate code. We hence developed \name
+%for defining modular extensions easily.
 
 \subsection{Initial System}
-Before introducing any extensions, we need to rewrite the initial implementation for \dsl shown in Figure~\ref{}:
+Recall the \dsl language discussed in Section~\ref{}.
+We redefine it with \name:
 
-\lstinputlisting[linerange=11-26]{../src/paper/sec3/Circuit.java}%APPLY:INIT
-We change the implementation language constructs from classes to interfaces and
-their fields to unimplemented getter methods. The purpose of this modification
-is to allow types of the fields to be refined in future, which is vital for
-retaining extensibility.
+\lstinputlisting[linerange=19-35]{../src/paper/sec3/Circuit.java}%APPLY:INIT
+
+There are three main changes:
+1) All the original implementations are moved inside an interface \texttt{Init}; and
+2) The constructor classes are now interfaces for the purpose of employing
+multiple interface inheritance of Java;
+3) Fields of the classes become getter methods (starting with an underscore) so that their types can be refined for retaining extensibility~\cite{}.
+
+% introduce of method
+As classes are now interfaces, how can we create objects on them?
+\name automatically generates a \emph{static} \texttt{of} method for each
+class-like interface, i.e. an interface that contains only getters, for creating objects.
+For instance, the \texttt{of} method for \texttt{Beside} is:
+\begin{lstlisting}
+static Beside of(Circuit c1, Circuit c2) {
+  return new Beside() {
+    Circuit _c1 = c1; Circuit _c2 = c2;
+    public Circuit _c1() { return c1; }
+    public Circuit _c2() { return c2; }
+  }
+}
+\end{lstlisting}
+Inside the \texttt{of} method, an anonymous class that implements
+\texttt{Beside} is created and its instance is returned.
+
+\subsection{Multiple Interpretations}
+Defining additional interpretations is not so trivial for shallow embedding,
+especially for functional languages.
+Suppose that we want to have a function that checks whether a circuit is constructed correctly.
+
+Here comes the solution from Gibbons and Wu.
+\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
+where tuples are used to aggregate multiple interpretations in one definition.
+The way to introduce \texttt{desugar} in previous section is similar to this
+approach, where we use records instead.
+However, neither tuples nor records are extensible - whenever we need a new
+interpretation, we have to revise the original code, either expanding the arith of the
+tuple or inserting a new field in the record.
+
+Conversely, the OO approach allows us to define multiple interpretations in a
+\emph{modular} way through extending existing code:
+\lstinputlisting[linerange=39-52]{../src/paper/sec3/Circuit.java}%APPLY:MULTIPLE
+We expand the semantics in an extended family \texttt{Multiple}.
+Inside \texttt{Multiple}, the \texttt{Circuit} is re-declared
+with a new method \texttt{wellSized} and the \texttt{Circuit} hierarchy
+implements the \texttt{wellSized} method.
+Through extending the \texttt{Init} family, the \texttt{width} method is
+inherited without duplicating its definition. This way we modularly extend the
+semantics.
+
+Note that we do not explicitly write down \texttt{extends} clauses in the
+extended family.
+Instead, they are inferred by \name according to their
+names and the inheritance relationship between families.
+After instrumentation done by \name, the definitions of \texttt{Circuit} and \texttt{Beside}
+would be:
+
+\begin{lstlisting}
+interface Circuit extends Init.Circuit {
+  ... // wellSized declaration
+}
+interface Beside extends Circuit, Init.Beside {
+  Circuit _c1(); Circuit _c2(); // type refinements
+  ... // wellSized implementation
+}
+\end{lstlisting}
+Interfaces from base families of the same name will be put after \texttt{extends} clause and the
+interface hierarchy of \texttt{Circuit} is preserved in the extended family.
+Moreover, \name re-declare getters from the base \texttt{Beside} in the extended
+\texttt{Beside} so that their return types are refined to be the new
+\texttt{Circuit}.
+
+\subsection{Dependent Interpretations}
+Interpretations which depend on other interpretations are called \emph{dependent
+  interpretation}. %They are hard to express modularly in traditional shallow embedding.
+Before we introduce an example of such interpretations, we first add two new
+constructs to our language:
+\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
+summing up \emph{ns} and $above\ c_1\ c_2$ combines two circuits of the same width vertically.
+Their types are given below:
+\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+
+%\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
+%With these new constructs, more complex circuits can be constructed.
+%For example, Figure~\ref{} shows the circuit constructed by \emph{stretch
+%  [2,2,2] (fan 3) `beside` fan 1}.
+
+To add these two new constructs to \dsl, we can declare two class-like interfaces \texttt{Above} and
+\texttt{Stretch} that extends \texttt{Circuit} with \texttt{width} and
+\texttt{wellSized} implemented
+
+\lstinputlisting[linerange=56-76]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT
+Definitions of \texttt{wellSized} for the new cases make it a dependent interpretation, as they
+rely on the \texttt{width} method for calculating the width of a circuit.
+
+%Since \texttt{Circuit} in \texttt{Dependent} family extends the one from
+%\texttt{Init}, we can simply call \texttt{width} method on a
+%\texttt{Circuit} object when the width of a circuit is needed in the definition
+%of \texttt{wellSized}.
+
+\subsection{Context-sensitive Interpretations}
+% add tlayout
+Connections between vertical wires of a circuit can be modeled by dividing the
+circuit into different layers, with local connections only go rightwards from
+one layer to the next. Each layer is a collection of pairs $(i,j)$ denoting a
+connection from wire $i$ on this layer to wire $j$ on the next. The circuit shown in
+Figure\ref{} has the following connections:
+
+$$[[(0,1),(2,3)],[(1,3)],[(1,2)]]$$
+
+The above shows that the circuit has 3 layers: the first layer has connections
+from the first wire to the second, and from the second to the third; the second
+layer has only one connection from the first wire to the third wire; while the
+third layer has a connection from the first wire to the second.
+
+The method \texttt{tlayout} deduce the connections of a circuit. It additionally
+takes a parameter \texttt{}.
+
+\lstinputlisting[linerange=98-130]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE
+
+\texttt{tlayout} takes a transformation over \texttt{IntUnaryOperator} is considered as the context which might be changed.
+The accumulating argument \texttt{f} is a transformation function over the indices of wires in a circuit.
+
+Some definitions of helper methods are omitted for space reason.
+Their functionalities are: \texttt{lzw} stands for 'long zip with', which takes two lists and a binary operator for
+combining elements on the same position; if one list is shorter then the
+remaining elements of the other are appended to the result list; \texttt{concat}
+concatenates two lists with the first tail connected the second head;
+The indices of wires are calculated by accumulating the sum of the list \texttt{\_ns} using the \texttt{scanl1}.
+
+% in \texttt{CtxSensitive}
+
+\subsection{Interpretation Combination}
+One appealing feature that \name approach provides is that we can modularly
+merge interpretations defined for the same data structure together.
+This is done through from different families
+together using multiple interface inheritance.
+
+For example, we can merge the interpretations we have defined so far in this way:
+
+\lstinputlisting[linerange=161-163]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
+Simply defining a new family \texttt{Merge} that extends other families we want
+to combine. \name will implicitly re-declare all the interfaces and inject the
+inheritance relationships. For example, the implicit redeclaration of
+\texttt{Circuit} looks like this:
+
+\begin{lstlisting}
+interface Circuit extends Multiple.Circuit, Dependent.Circuit, CtxSensitive.Circuit {}
+\end{lstlisting}
+
+So do other interfaces.
+
+\begin{comment}
+\subsection{Parametrized interpretations}
+\subsection{Implicitly Parametrized interpretations}
+\subsection{Modular Interpretations}
 
@@
 
-\lstinputlisting[linerange=29-40]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
+\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
 
@@
 Gibbons and Wu worked around this problem in the following way:
-\lstinputlisting[language=haskell,linerange=2-6]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
+\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
 which is similar to how we add \texttt{desugar} to the implementation in Section~\ref{}.
@@
 By using \name, we can refactor the extensions in the way shown in Figure\ref{}.
-\lstinputlisting[linerange=80-82]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY
-\lstinputlisting[linerange=110-112]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SYNTAX
-\lstinputlisting[linerange=131-156]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SEMANTICS
+\lstinputlisting[linerange=46-48]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY
+\lstinputlisting[linerange=76-78]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SYNTAX
+\lstinputlisting[linerange=97-122]{../src/paper/sec3/Circuit.java}%APPLY:FAMILY_SEMANTICS
 
@@
 
-\lstinputlisting[linerange=164-176]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
+\lstinputlisting[linerange=168-180]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
 % TODO: not implemented yet
@@
 %that implements the interface with all getters implemented and its instance is returned.
-
-% Multiple interpretation
-% Dependent interpretation
-% Context-sensitive interpretation
+\end{comment}

### 2016-05-24 paper/sections/ShallowOO.tex
Revision on Sec 2 & 3

--- 2016-04-19 paper/sections/ShallowOO.tex
@@
 
+\begin{comment}
 Weixin writes this part.
@@
 
-\lstinputlisting[language=haskell,linerange=2-5]{./code/shallow.hs}%APPLY:EXPR1
-\lstinputlisting[language=haskell,linerange=9-12]{./code/shallow.hs}%APPLY:EXPR2
-\lstinputlisting[language=haskell,linerange=16-18]{./code/shallow.hs}%APPLY:EXPR3
-\lstinputlisting[language=haskell,linerange=22-33]{./code/shallow.hs}%APPLY:EXPR4
+Shallow embedding interpreter pattern
+Deep embedding visitor pattern
+\end{comment}
+
+In this section, we argue that \emph{shallow embedding} and straightforward OO
+programming are closely related. To demonstrate it, we use a subset
+of the DSL presented in Gibbons and Wu's paper~\cite{} as the running example.
+We first give the original implementation of the language in shallow embedding
+and then gradually rewrite the definition into an ``OO style''. Also, we show how to do
+transformations in shallow embedding, which is typically thought to be the
+strength of deep embedding.
+
+\subsection{A DSL for Parallel Prefix Circuits}
+Consider an DSL named \dsl that models parallel circuits.
+Its BNF grammar is given below:
+$$
+\begin{array}{lcl}
+Circuit & ::= & Identity \\
+ & \mid & Fan \\
+ & \mid & Beside \\
+Fan & ::= & ``fan''\; PositiveNumber \\
+Identity & ::= & ``identity''\;  PositiveNumber \\
+Beside & ::= & Circuit ``beside'' Circuit \\
+PositiveNumber & ::= & [1-9] \\
+     & \mid & [1-9][0-9]^* \\
+\end{array}
+$$
+% bnf
+\weixin{the bnf grammar needs polishing}
+
+% BNF grammar should not contain left recursion and left factor, rewrite it
+% using descent recursion
+
+There are three constructs - two primitives
+\emph{identity} and \emph{fan} and one combinator \emph{beside}, whose
+meanings are: \emph{identity n} contains \emph{n} isolated vertical wires;
+\emph{fan n} contains \emph{n} vertical wires and the first wire connects to
+all the remaining wires from top to bottom; $beside\ c_1\ c_2$ joins two circuits
+$c_1$ and $c_2$ horizontally.
+For example, the circuit \emph{fan 3 beside identity 3} can be visualized in Figure~\ref{}.
+
+\subsection{Shallow Embedding and Object-Oriented Programming}
+Different from deep embedding, which uses a data structure to represent the
+abstract syntax of a language, shallow embedding defines a language
+directly through encoding its semantics. In the case of \dsl,
+the shallow embedding implementation should conform to the following types:
+
+\lstinputlisting[language=haskell,linerange=2-2]{./code/shallowCircuit.hs}%APPLY:CIRCUIT_TYPE
+\lstinputlisting[language=haskell,linerange=5-7]{./code/shallowCircuit.hs}%APPLY:TYPES
+The type \texttt{Circuit}, representing the semantic domain, is not defined yet
+and should be filled in later with a concrete type according to the semantics.
+For example, if the semantics of \dsl is to calculate the width of a
+circuit, the semantic function would be:
+\lstinputlisting[language=haskell,linerange=11-16]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
+As shallow embedding is semantics-oriented, it is already an integer after
+constructing a circuit and hence \texttt{width1} is an identity function.
+For example, we will immediately get $6$ for
+\texttt{fan 3 `beside` identity 3} and hence \texttt{width1} is just an identity function.
+
+The above implementation can be rewritten by wrapping the result into an
+datatype and get back the value through pattern matching:
+\lstinputlisting[language=haskell,linerange=20-26]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
+Then circuit construction gives back a \texttt{Circuit2} which contains an
+integer value representing the width and \texttt{width2} extracts out the value.
+%One may be curious whether this variant implementation is still shallow embedding or not. The answer is yes because ... % why it is shallow
+
+The rewrite goes on by using a record with a sole field to represent the result:
+\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
+We no longer need a separate \texttt{width3} function as the field name is the extractor.
+The implementation is still shallow since essentially it is nothing different
+from the last one.
+However, the definition already approaches an OO program such that we can easily
+translate it into an OO language like Java:
+\lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
+where the datatype maps to the abstract class \texttt{Circuit} with its field declared as an abstract method.
+Each case in the semantic function corresponds to an class and its parameters become fields of the class.
+And all the classes extends \texttt{Circuit} by giving \texttt{width} a definition.
+
+This implementation is exactly how we model \dsl in an OO language for the first
+place, following the \interp pattern.
+In other words, shallow embedding and straightforward OO programming are closely
+related (deep embedding is closely related to the \textsc{Visitor} pattern).
+
+\subsection{Transformations in Shallow Embedding}
+One of the main criticism of shallow embedding is that how to define transformations
+becomes unclear since there is not a data structure representing the Abstract Syntax Tree (AST).
+%In fact, transformations can still be defined in shallow embedding.
+
+An instance of this kind of operations could be desugaring, which eliminates some
+language constructs that could be expressed by other language constructs without
+sacrificing the expressiveness of the language.
+In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
+be rewritten via the following formula:
+$$identity\ n = (fan\ 1) \overbrace{\ beside\ (fan\ 1)}^{n-1}\ (n > 0)$$
+which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
+%Using the above formula, we can define a transformation that desugars every
+%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.
+
+Implementing the desugaring method in Java could be:
+\lstinputlisting[linerange=31-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
+The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is
+straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
+their results into an instance of \texttt{Beside}; \emph{identity n} mimics the
+formula.
+
+\weixin{define a show function in the and show the result after desugaring?}
+Inspired by the Java implementation, we can also define \texttt{desugar} in Haskell:
+
+\lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
+The definition relies on the laziness of Haskell.

### 2016-05-24 paper/code/Desugar.hs

@@ -8,9 +8,9 @@ data Circuit = Circuit {
   desugar :: Circuit
 }
 identity n = Circuit {
-  width = width c,
-  desugar = c
-} where c = foldl1 beside $ replicate n (fan 1)
+  width = n,
+  desugar = foldr1 beside $ replicate n (fan 1)
+}
 fan n = Circuit {
   width = n,
   desugar = fan n

### 2016-05-24 paper/code/NewSemantics.hs

@@ -1,7 +1,9 @@
 --BEGIN_SEMANTICS_HS
-identity n = (True,n)
-fan n = (True,n)
-beside c1 c2 = (fst c1 && fst c2,fst c1 + fst c2)
-stretch ns c = (fst c && length ns == snd c)
-above c1 c2 = (fst c1 && fst c2 && snd c1 == snd c2,snd c1)
+identity n = (n,True)
+fan n = (n,True)
+beside c1 c2 = 
+  (width c1 + width c2,wellSized c1 && wellSized c2)
+
+width = fst
+wellSized = snd
 --END_SEMANTICS_HS

### 2016-05-24 src/paper/sec2/Circuit.java

@@ -1,62 +1,62 @@
 package paper.sec2;
 
+import static java.util.stream.IntStream.range;
+
 class A {
 //BEGIN_CIRCUIT_OO
-interface Circuit { int width(); }
-class Fan implements Circuit {
+abstract class Circuit {
+  abstract int width();
+}
+class Fan extends Circuit {
   int _n;
   Fan(int n) { _n = n; }
-  public int width() { return _n; }
+  int width() { return _n; }
 }
-class Beside implements Circuit {
+class Beside extends Circuit {
   Circuit _c1, _c2;
   Beside(Circuit c1, Circuit c2) {
     _c1 = c1; c2 = _c2;
   }
-  public int width() {
-    return _c1.width() + _c2.width();
-  }
+  int width() { return _c1.width() + _c2.width(); }
 }
-class Identity implements Circuit {
+class Identity extends Circuit {
   int _n;
   Identity(int n) { _n = n; }
-  public int width() { return _n; }
+  int width() { return _n; }
 }
 //END_CIRCUIT_OO
 }
 class Sec2 {
 //BEGIN_DESUGAR_OO
-interface Circuit {
-  int width();
-  Circuit desugar();
+abstract class Circuit {
+  abstract int width();
+  abstract Circuit desugar();
 }
-class Fan implements Circuit {
+class Fan extends Circuit {
   int _n;
   Fan(int n) { _n = n; }
-  public int width() { return _n; }
-  public Circuit desugar() { return this; }
+  int width() { return _n; }
+  Circuit desugar() { return this; }
 }
-class Beside implements Circuit {
+class Beside extends Circuit {
   Circuit _c1, _c2;
   Beside(Circuit c1, Circuit c2) {
     _c1 = c1; c2 = _c2;
   }
-  public int width() {
-    return _c1.width() + _c2.width();
-  }
-  public Circuit desugar() {
+  int width() { return _c1.width() + _c2.width(); }
+  Circuit desugar() {
     return new Beside(_c1.desugar(), _c2.desugar());
   }
 }
-class Identity implements Circuit {
+class Identity extends Circuit {
   int _n;
   Identity(int n) { _n = n; }
-  public int width() { return _n; }
-  public Circuit desugar() {
-    Circuit res = new Fan(1);
-    for (int i = 1; i < _n; i++)
-      res = new Beside(res, new Fan(1));
-    return res;
+  int width() { return _n; }
[…]

### 2016-05-24 src/paper/sec3/Circuit.java

@@ -1,163 +1,167 @@
 package paper.sec3;
 
+import static java.lang.Integer.min;
+import static java.util.Collections.emptyList;
+import static java.util.Collections.singletonList;
+import static java.util.stream.Collectors.toList;
+import static java.util.stream.IntStream.range;
+
+import java.util.ArrayList;
 import java.util.List;
+import java.util.function.BinaryOperator;
+import java.util.function.IntUnaryOperator;
+import java.util.stream.Stream;
 
 import lombok.Obj;
 
 
-@Obj
-interface Init {
 //BEGIN_INIT
-interface Circuit { int width(); }
-interface Fan extends Circuit {
-  int _n();
-  default int width() { return _n(); }
-}
-interface Identity extends Circuit {
-  int _n();
-  default int width() { return _n(); }
-}
-interface Beside extends Circuit {
-  Circuit _c1();
-  Circuit _c2();
-  default int width() {
-    return _c1().width() + _c2().width();
-  }
-}
-//END_INIT
-//BEGIN_SYNTAX
-interface Stretch extends Circuit {
-  List<Integer> _ns();
-  Circuit _c();
-  default int width() {
-    return _ns().stream().reduce(0, (a,b) -> a+b);
-  }
-}
-interface Above extends Circuit {
-  Circuit _c1();
-  Circuit _c2();
-  default int width() { return _c1().width(); }
-}
-//END_SYNTAX
-}
-
-@Obj
-interface SemanticsExt extends Init {
-//BEGIN_SEMANTICS
-interface CircuitWS extends Circuit {
-  boolean wellSized();
-}
-interface IdentityWS extends Identity, CircuitWS {
-  default boolean wellSize() { return true; }
-}
-interface FanWS extends Fan, CircuitWS {
-  default boolean wellSize() { return true; }
-}
-interface BesideWS extends Beside, CircuitWS {
-  CircuitWS _c1();
-  CircuitWS _c2();
-  default boolean wellSized() {
-    return _c1().wellSized() && _c2().wellSized();
-  }
-}
-interface StretchWS extends Stretch, CircuitWS {
-  CircuitWS _c();
-  default boolean wellSized() {
-    return _ns().size() == _c().width();
-  }
-}
-interface AboveWS extends Beside, CircuitWS {
-  CircuitWS _c1();
-  CircuitWS _c2();
-  default boolean wellSized() {
[…]

### 2016-05-30 paper/sections/CaseStudyPPrinter.tex
Pretty printer.

@@
 This case study refactors the Haskell code from \textcolor{red}{(Haoyuan: ref, a prettier printer)}
-which implements a functional pretty printer library with high efficiency. In the original code, two
-data structures are defined for documents. Here they are defined using shallow DSLs, and packaged into
+which uses deep embeddings to implement a functional pretty printer library with high efficiency. In the original code, two
+data structures are defined for documents:
+\begin{lstlisting}[keywords={data,Int,String}]
+data DOC = NIL
+            | DOC :<> DOC
+            | NEST Int DOC
+            | TEXT String
+            | LINE
+            | DOC :<> DOC
+data Doc = Nil
+            | String `Text` Doc
+            | Int `Line` Doc
+\end{lstlisting}
+Here they are encoded with shallow DSLs, and packaged into
 two base families with \textsf{@Family} annotation applied. These two families \textsf{Family\_Doc} and
-\textsf{Family\_Document} correspond to data types \textsf{Doc} and \textsf{DOC} respectively in original code.
+\textsf{Family\_Document} correspond to data types \textsf{Doc} and \textsf{DOC} respectively.
 
@@
 	interface Text extends Doc {
-		String s(); Doc d();
+		String _s(); Doc _d();
 	}
-	interface Line extends Doc { int i(); Doc d(); }
+	interface Line extends Doc { int _i(); Doc _d(); }
 }
@@
 	interface DConcat extends Document {
-		Document d1(); Document d2();
+		Document _d1(); Document _d2();
 	}
 	interface DNest extends Document {
-		int i(); Document d();
+		int _i(); Document _d();
 	}
-	interface DText extends Document { String s(); }
+	interface DText extends Document { String _s(); }
 	interface DLine extends Document {}
 	interface DUnion extends Document {
-		Document d1(); Document d2();
+		Document _d1(); Document _d2();
 	}
@@
 	interface Doc {
-		String _layout(); boolean _fits(int w);
+		String layout(); boolean fits(int w);
 	}
 	interface Nil {
-		default String _layout() { return ""; }
-		default boolean _fits(int w) {
+		default String layout() { return ""; }
+		default boolean fits(int w) {
 			return w >= 0;
@@
 \end{lstlisting}
-Furthermore, some functions like \textsf{be} and \textsf{beaux} need refactoring before they are encoded in shallow embeddings.
-Finally we refactored the pretty printer library using a set of extensible families with shallow DSLs and operations, and our
+Furthermore, the function \textsf{be} needs refactoring before it is encoded in shallow embeddings.
+The original \textsf{be} is defined as follows:
+\begin{lstlisting}
+be w k [] = Nil
+be w k ((i,NIL):z) = be w k z
+be w k ((i,x :<> y):z) = be w k ((i,x):(i,y):z)
+be w k ((i,NEST j x):z) = be w k ((i+j,x):z)
+be w k ((i,TEXT s):z) = s `Text` be w (k+length s) z
+be w k ((i,LINE):z) = i `Line` be w i z
+be w k ((i,x :<|> y):z) = better w k (be w k ((i,x):z)) (be w k ((i,y):z))
+\end{lstlisting}
+We refactor it by introducing a helper function \textsf{beaux}:
+\begin{lstlisting}
+beaux w k i NIL z = be w k z
+beaux w k i (x :<> y) z = be w k ((i,x):(i,y):z)
+beaux w k i (NEST j x) z = be w k ((i+j,x):z)
+beaux w k i (TEXT s) z = s `Text` be w (k+length s) z
+beaux w k i LINE z = i `Line` be w i z
+beaux w k i (x :<|> y) z = better w k (be w k ((i,x):z)) (be w k ((i,y):z))
+
+be w k [] = Nil
+be w k ((i,x):z) = beaux w k i x z
+\end{lstlisting}
+In this case, \textsf{beaux} can be defined as an operation method \lstinline{beaux(int, int, int, List<Pair<int, Document>>)} inside the type \textsf{Document}, and
+its pattern matching corresponds to the default implementations in different member types (like \textsf{DNil}, \textsf{DConcat}, etc). On the other hand, \textsf{be} is implemented as a static method 
+\lstinline{be(int, int, List<Pair<int, Document>>)} outside the member types.
+
+Finally we finished the refactoring of the pretty printer library using a set of extensible families with shallow DSLs and operations, and our
 \textsf{@Family} annotation. However there is a special operation called \textsf{group}, which works like a transformation:
@@
 	interface Document {
-		Document _flatten();
-		default Document _group() {
-			return DUnion.of(this._flatten(), this);
+		Document flatten();
+		default Document group() {
+			return DUnion.of(this.flatten(), this);
 		}
@@
 \end{lstlisting}
-The method \textsf{\_group} calls the constructor method from \textsf{DUnion}, whereas such a member type will be updated in the
+The method \textsf{group} calls the constructor method from \textsf{DUnion}, whereas such a member type will be updated in the
 child families, together with its constructors. To ensure type safety we are unable to reuse the code but just copy it and paste

### 2016-05-30 paper/sections/Tool.tex
Pretty printer.

--- 2016-05-14 paper/sections/Tool.tex
@@
 \lstinline{@Family}, \lstinline{@Obj} and \lstinline{@ObjOf} respectively. The translation rules are presented in Figure \ref{fig:trans1}
-and Figure \ref{fig:trans2}, together with some auxiliary functions.
+and Figure \ref{fig:trans2}, together with some auxiliary functions. Note that in practice, our \textsf{@Family} supports simple
+generics without bounds, but for simplicity reason it is not included in the formalization.
+
+\begin{figure}
+\begin{tabular}{|l|}
+\hline
+\begin{lstlisting}
+@Family interface New extends Base {
+	interface Circuit { int width(); }
+	interface Beside {
+		default int width() {
+			return _c1().width() + _c2().width();
+		}
+	}
+}
+\end{lstlisting} \\
+\hline
+$\Downarrow$ Family desugaring \\
+\hline
+\begin{lstlisting}
+@Obj interface New extends Base {
+	@Obj interface Circuit extends Base.Circuit {
+		int width();
+	}
+	@Obj interface Beside extends
+			Circuit, Base.Beside {
+		Circuit _c1(); Circuit _c2();
+		default int width() {
+			return _c1().width() + _c2().width();
+		}
+	}
+}
+\end{lstlisting} \\
+\hline
+$\Downarrow$ Obj desugaring \\
+\hline
+\begin{lstlisting}
+@ObjOf interface New extends Base {
+	@ObjOf interface Circuit extends Base.Circuit {
+		int width();
+	}
+	@ObjOf interface Beside extends
+			Circuit, Base.Beside {
+		Circuit _c1(); Circuit _c2();
+\end{lstlisting}\vspace{-.05in}\\ \begin{lstlisting}[basicstyle=\ttfamily\small\color{red}]
+		Beside withC1(Circuit val);
+\end{lstlisting}\\ \begin{lstlisting}
+		default int width() {
+			return _c1().width() + _c2().width();
+		}
+	}
+}
+\end{lstlisting} \\
+\hline
+$\Downarrow$ ObjOf desugaring \\
+\hline
+\begin{lstlisting}
+interface New extends Base {
+	interface Circuit extends Base.Circuit {
+		int width();
+	}
+	interface Beside extends Circuit, Base.Beside {
+		Circuit _c1(); Circuit _c2();
+		Beside withC1(Circuit val);
+		default int width() {
+			return _c1().width() + _c2().width();
+		}
+		static Beside of(Circuit _c1, Circuit _c2) {
+			return new Beside() {
+				Circuit c1 = _c1; Circuit c2 = _c2;
+				public Circuit _c1() { return c1; }
+				public Circuit _c2() { return c2; }
+				public Beside withC1(Circuit val) {
+					return Beside.of(val, this._c2());
+				}
+			};
+		}
+	}
+	static New of() { return new New(){}; }
+}
+\end{lstlisting} \\
+\hline
+\end{tabular}
+\caption{An example showing the flow of translation.}\label{fig:flow}
+\end{figure}
+
+\subsection{Flow of Translation}
+Figure~\ref{fig:flow} illustrates the flow of translation using a simplified version of the Scans example, with intermediate results of desugaring shown.
+A base family \textsf{Base} is predefined as follows:
+\begin{lstlisting}
+interface Base {
+	interface Circuit {}
+	interface Beside extends Circuit {
+		Circuit _c1(); Circuit _c2();
+		Beside withC1(Circuit val);
+	}
+}
+\end{lstlisting}
+\textsf{Base} contains two members, \textsf{Circuit}, and its subtype \textsf{Beside}.
+\textsf{Beside} has two field methods \textsf{\_c1()} and \textsf{\_c2()} together with a wither method \textsf{withC1}.
+At this time, another interface \textsf{New} is defined as the new family that extends the old one, where we add an operation \textsf{width()} to the member types.
+This is precisely the interface that we are going to translate.
+In the client code, the user simply re-declares the two members with the addition of \textsf{width()}, without building subtyping relations among new and old members, and a simple annotation \textsf{@Family} will deal with all the stuff.
+
+In the first step of translation, namely family desugaring, \textsf{@Family} collects all the members from old families, and fills in the ``extends'' to build the inheritance relations.
+At the same time, it re-declares all the field methods, for example, although the type name of \textsf{\_c1()} is still \textsf{Circuit}, it refers to the new member type, where we address field type refinements, and hence this new \textsf{\_c1()} has the operation \textsf{width()}.
+Then it delegates the work to \textsf{@Obj}.
+In the second phase, \textsf{@Obj} detects fluent setters and withers from the supertypes, then refines their return types to have consistency with the annotated type.
+In our example, the wither method \textsf{withC1} is refined to have the new \textsf{Beside} as its return type.
+In the last step of translation, \textsf{@ObjOf} generates constructor methods for all the interfaces, except \textsf{Circuit}, where its \textsf{width()} is an abstract method and we cannot provide a default implementation for it.
+In the following subsections we will respectively explain the three steps of translation in detail.
 

### 2016-05-31 paper/sections/Overview.tex
Split

@@
 interface hierarchy of \texttt{Circuit} is preserved in the extended family.
-Moreover, \name re-declare getters from the base \texttt{Beside} in the extended
-\texttt{Beside} so that their return types are refined to be the new
-\texttt{Circuit}.
+Moreover, \name re-declare getters in the extended \texttt{Beside} so that their return types are refined to be the new \texttt{Circuit}.
 
@@
 
-To add these two new constructs to \dsl, we can declare two class-like interfaces \texttt{Above} and
-\texttt{Stretch} that extends \texttt{Circuit} with \texttt{width} and
-\texttt{wellSized} implemented
+Adding new constructs in \name is easy: just we declare new class-like
+interfaces in an extended family:
+\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
+Only definitions of the new constructs \texttt{Above} and \texttt{Stretch} are given.
+Their peer constructs and \texttt{Circuit} are implicitly re-declared by \name,
+e.g. \texttt{Fan}:
 
-\lstinputlisting[linerange=56-76]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT
-Definitions of \texttt{wellSized} for the new cases make it a dependent interpretation, as they
-rely on the \texttt{width} method for calculating the width of a circuit.
+\begin{lstlisting}
+interface Fan extends Init.Fan {}
+\end{lstlisting}
 
+Not arbitrary circuits can be combined using \texttt{Above} and \texttt{Stretch}.
+We hence define the \texttt{wellSized} method for these two new constructs to
+verify the constraints they imply:
+\lstinputlisting[linerange=73-85]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT
+
+Definitions of \texttt{wellSized} for the extended cases make it a dependent interpretation, as they
+rely on another method defined on the \texttt{Circuit}, \texttt{width}, for calculating the width of a circuit.
+
+Here, we show yet another appealing feature of \name - interpretation combination.
+The \texttt{Dependent} family combines two base families
+
+merge interpretations defined for the same data structure together.
 %Since \texttt{Circuit} in \texttt{Dependent} family extends the one from
@@
 
-\lstinputlisting[linerange=98-130]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE
+\lstinputlisting[linerange=107-139]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE
 
@@
 Their functionalities are: \texttt{lzw} stands for 'long zip with', which takes two lists and a binary operator for
-combining elements on the same position; if one list is shorter then the
+combining elements of the same position and if one list is shorter then the
 remaining elements of the other are appended to the result list; \texttt{concat}
@@
 
+\begin{comment}
 \subsection{Interpretation Combination}
@@
 
-\lstinputlisting[linerange=161-163]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
+\lstinputlisting[linerange=170-172]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
 Simply defining a new family \texttt{Merge} that extends other families we want
@@
 
-\begin{comment}
 \subsection{Parametrized interpretations}
@@
 
-\lstinputlisting[linerange=168-180]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
+\lstinputlisting[linerange=177-189]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
 % TODO: not implemented yet

### 2016-05-31 paper/sections/CaseStudy.tex
Delegating tasks
Case study: PPrinter.
Some text on case study

@@
 \section{Case Study}
+
+Weixin and Haoyuan.
 
@@
 extensible transformations.
+
+\subsection{Scans}
+% statistics, pictures, implementation
+Section~\ref{} is based on our first case study, hence we discuss it briefly here.
+In the case study, we strictly follows the flow of the paper and implement all kinds of
+interpretations modularly, including \texttt{depth} and \texttt{layout} that are
+not discussed in Section~\ref{}.
+
+Starting from a data structure that describes the five constructs without any interpretations, we gradually add
+\texttt{depth}, \texttt{width}, \texttt{wellSized}, \texttt{layout} and
+\texttt{tlayout} through extending the family.
+
+To make our DSL more user-friendly, we defined some wrappers for constructing
+circuits conveniently. Afterwards, we can construct a circuit this way:
+\begin{lstlisting}
+Circuit circuit =
+  fan(2).beside(fan(2))
+  .above(fan(2).stretch(2,2))
+  .above(id(1).beside(fan(2)).beside(id(1)));
+\end{lstlisting}
+For primitives like \texttt{fan}, we define a static method and for binary
+combinators like \texttt{above}, we define a default method on the \texttt{Circuit}.
+
+Additionally, we extend the language with a \texttt{draw} method which renders a
+circuit using Java Swing.
+Calling \texttt{circuit.draw()} on the \texttt{circuit} we just defined will
+display what Figure~\ref{} shows.
+
+% show Java swing output
+\begin{figure}
+\end{figure}
+
+\input{sections/CaseStudyPPrinter.tex}
+
+\subsection{Diagrams}
+\emph{Diagrams}~\cite{} is a powerful, flexible, declarative Haskell DSL for creating vector graphics.
+This case study implements a simplified version of \emph{Diagrams} from
+Gibbons's lecture notes ~\cite{}.
+The DSL consists of three simpler sub-languages for describing shapes, colors, styles.
+% Haskell code
+\begin{lstlisting}[language=haskell]
+data Shape
+  = Rectangle Double Double
+  | Ellipse Double Double
+  | Triangle Double
+
+type StyleSheet = [Styling]
+data Styling
+  = FillColour Col
+  | StrokeColour Col
+  | StrokeWidth Double
+
+data Col = Red | Blue | Bisque | Black | Green | Yellow | Brown
+
+data Picture
+  = Place StyleSheet Shape
+  | Above Picture Picture
+  | Beside Picture Picture
+\end{lstlisting}
+
+\paragraph{Shapes}
+There are three primitive shapes - rectangle, ellipse and triangle.
+
+\begin{lstlisting}
+interface Shape {}
+interface Rectangle extends Shape {
+    double _x(); double _y();
+}
+interface Ellipse extends Shape {
+    double _rx(); double _ry();
+}
+interface Triangle extends Shape {
+    double _l();
+}
+\end{lstlisting}
+
+\paragraph{Styles}
+A shape can be decorated with drawing styles. A \texttt{StyleSheet} contains a
+list of \texttt{Stylings} where fill color, stroke color or stroke width can be specified.
+\begin{lstlisting}
+interface StyleSheet {
+    List<? extends Styling> _stylings();
+}
+interface Styling {}
+interface FillColor extends Styling {
+    Col _color();
+}
+interface StrokeColor extends Styling {
+    Col _color();
+}
+interface StrokeWidth extends Styling {
+    double _width();
+}
+\end{lstlisting}
+\texttt{Col} is also a data structure that describes the supported colors.
+\begin{lstlisting}
+interface Col {}
+interface Red extends Col {}
+interface Blue extends Col {}
+interface Green extends Col {}
+interface Yellow extends Col {}
+interface Bisque extends Col {}
+interface Black extends Col {}
+\end{lstlisting}
+
+\paragraph{Pictures}
+\begin{lstlisting}
+interface Picture {}
+interface Place extends Picture {
+    Shape _s(); StyleSheet _ss();
+}
+interface Above extends Picture {
+    Picture _p1(); Picture _p2();
+}
+interface Beside extends Picture {
+    Picture _p1(); Picture _p2();
+}
+\end{lstlisting}
+
+Provided with a \texttt{StyleSheet}, a shape can be lifted to a picture using
+\texttt{Place}. And pictures can be merged using \texttt{Above} or
+\texttt{Beside}: the former combines two pictures vertically whereas the latter
+combines two pictures horizontally.
+
+With these three main data structure, we can draw a picture.
+
+\subsubsection{Rendering}
+We render our fancy picture using scalable vector graphics (SVG) as our backend.
+To do so, we need to flatten the picture which is recursively constructed.
+
+SVG is based on XML and hence we define to model a XML file.
+
+\paragraph{Transformations}
+
+\subsubsection{Extensions}
+\cite{} contains exercises about extensions on the language, e.g.
+adding new shapes. The deep embedding approach in Haskell requires modifications on the ADTs
+whereas our approach can modularly introduce new language constructs.
+.
+Each sub-language retains independent extensibility.
+For example, we can add a \texttt{show} method for the purpose of debugging.

### 2016-06-08 paper/sections/CaseStudyPPrinter.tex
related work on family polymorphism

@@
 child families, together with its constructors. To ensure type safety we are unable to reuse the code but just copy it and paste
-to the child families, and hence code duplication is introduced. The matter of extensible transformations will be further discussed
-in the next section.
+to the child families, and hence code duplication is introduced by the use of factory methods. The matter of extensible
+transformations will be further discussed in the next section.

### 2016-06-08 paper/sections/RelatedWork.tex
Delegating tasks
related work on family polymorphism

@@
 \section{Related Work}
+
+Weixin and Haoyuan.
 
@@
 
+There has been a lot of related work on family polymorphism[1], including various lightweight encodings[2,3,4,5]. In the original
+paper[1], nested classes are represented as attributes of an object, which involves a dependent type system. In [4], Saito and Igarashi
+proposed a lightweight variant of family polymorphism, which uses classes rather than objects to represent families. Later work in [4]
+proposes a simple extension to Featherweight Generic Java[6], introducing self-type variables. There are also some existing languages,
+like Scala, which supports symmetric mixin compositions and self-type annotations, hence can provide better support for encoding family
+polymorphism. Our lightweight encoding relies entirely on the existing Java language without language extensions, and certainly some features from
+the original paper are sacrificed, including the mismatching problem of recursive class definitions (also known as binary methods).
+
+On the other hand, our approach uses nested interfaces to build the relationship among families and members, which is different from
+path-dependent types in [1]. The work in [8] inspires us with the concept of higher-order hierarchies, and hence our \textsf{@Family} annotation
+provides support for nested families. Furthermore, our lightweight encoding of family polymorphism is polished with the help of annotation processing
+to generate constructor methods automatically.
+
+[1] family polymorphism
+[2] Lightweight scalable components
+[3] The essence of lightweight family polymorphism
+[4] Lightweight family polymorphism
+[5] Lightweight dependent classes
+[6] Featherweight Java
+[7] On binary method
+[8] Higher-order hierarchies
+
 \paragraph{ThisType}

### 2016-06-09 paper/sections/CaseStudy.tex
More case study

--- 2016-04-19 paper/sections/CaseStudy.tex
@@
 extensible transformations.
+
+\subsection{Scans}
+% statistics, pictures, implementation
+The first case study re-implements \dsl, which is discussed quite thoroughly in
+Section~\ref{}.
+We strictly follows the flow of the paper and introduce all kinds of
+interpretations modularly.
+Starting from a plain AST without any interpretations in the base family, we gradually add
+\texttt{depth}, \texttt{width}, \texttt{wellSized}, \texttt{layout} and
+\texttt{tlayout} methods through extending the family.
+
+The following shows the client code:
+\begin{lstlisting}
+Circuit circuit =
+  above(beside(fan(2), fan(2)),
+  above(stretch(fan(2),2,2),
+  beside(id(1), beside(fan(2), id(1)))));
+\end{lstlisting}
+
+To simplify client code, we defined wrappers around \texttt{of} methods
+for conveniently constructing circuits. For instance, the wrapper for
+\texttt{Fan} is defined as follows:
+\begin{lstlisting}
+static Fan fan(int n) {
+  return Fan.of(n);
+}
+\end{lstlisting}
+The \texttt{circuit} object supports all the methods stated above.
+Additionally, we introduce a \texttt{draw} method which renders a
+circuit using Java Swing.
+Calling \texttt{circuit.draw()} on the \texttt{circuit} will
+display what Figure~\ref{sec:} shows.
+
+% show Java Swing figure
+\begin{figure}
+\end{figure}
+
+\input{sections/CaseStudyPPrinter.tex}
+
+\subsection{Diagrams}
+\emph{diagrams}~\cite{yates2015diagrams} is a declarative Haskell EDSL for creating vector graphics.
+This case study implements a simplified version of \emph{diagrams} from
+Gibbons's lecture notes~\cite{gibbonsexercises}.
+
+\subsubsection{The Language}
+The DSL consists of three simpler sub-languages for describing shapes, styles
+and pictures. Note that all these definitions are put inside a base family.
+
+% Haskell code
+\begin{comment}
+\begin{lstlisting}[language=haskell]
+data Shape
+  = Rectangle Double Double
+  | Ellipse Double Double
+  | Triangle Double
+
+type StyleSheet = [Styling]
+data Styling
+  = FillColour Col
+  | StrokeColour Col
+  | StrokeWidth Double
+
+data Col = Red | Blue | Bisque | Black | Green | Yellow | Brown
+
+data Picture
+  = Place StyleSheet Shape
+  | Above Picture Picture
+  | Beside Picture Picture
+\end{lstlisting}
+\end{comment}
+
+\paragraph{Shapes}
+Shapes are fundamental components of a picture. The sub-language provides three
+primitives for constructing rectangles, ellipses and triangles:
+
+\lstinputlisting[linerange=111-120]{../src/diagrams/Diagrams.java}%APPLY:SHAPE
+
+\paragraph{Styles}
+A shape can be decorated with drawing styles. A \texttt{StyleSheet} contains a
+list of \texttt{Stylings} for specifying fill color, stroke color or stroke width:
+
+\lstinputlisting[linerange=124-136]{../src/diagrams/Diagrams.java}%APPLY:STYLE
+\lstinputlisting[linerange=140-145]{../src/diagrams/Diagrams.java}%APPLY:COLOR
+
+The definition of \texttt{StyleSheet} reveals how to deal with lists or more generally
+collection types in \name. We capture the element type using a wildcard with an upper
+bound (\texttt{? extends Styling}) to model covariant lists. This allows the
+upper bound to be refined with its subtype in extensions. As a result, we are
+able to invoke new interpretations on elements of the list.
+
+\paragraph{Pictures}
+Provided with a \texttt{StyleSheet}, a \texttt{Shape} can be lifted to a
+\texttt{Picture} using \texttt{Place}. Meanwhile, two pictures can be
+combined together vertically or horizontally via binary combinator
+\texttt{Above} and \texttt{Beside} respectively:
+
+\lstinputlisting[linerange=151-160]{../src/diagrams/Diagrams.java}%APPLY:PICTURE
+
+We can already draw simple pictures based on existing definitions.
+For example, Figure~\ref{woman} (with auxiliary wrappers defined) shows
+a woman in a red dress and blue stockings.
+To actually visualize the picture as shown in the right hand side of
+Figure~\ref{woman}, we give semantics to the language for accomplishing this goal.
+
+\begin{figure*}
+  \begin{tabular}{lr}
+\begin{minipage}[t]{.9\textwidth}
+\lstinputlisting[linerange=530-537]{../src/diagrams/Diagrams.java}%APPLY:WOMAN
+\end{minipage}
+&
+\begin{minipage}[t]{.1\textwidth}
+  \vspace{0.2em}
+\includegraphics[width=1.3cm]{woman.png}
+\end{minipage} \\
+\end{tabular}
+\caption{A woman drawn in \emph{diagrams}}
+\label{woman}
+\end{figure*}
+
+\subsubsection{Animation}
+We use scalable vector graphics (SVG) as the backend for rendering
+pictures.
+SVG aligns styled shapes by center, which requires us to determine a center
+point of a picture as the origin of coordinate, and calculate
+positions of all the shapes inside the picture accordingly.
+To represent a position in the coordinates we define \texttt{Pos}:
+\begin{lstlisting}
+@Obj interface Pos {
+    double _x(); double _y();
+}
+\end{lstlisting}
+The position of a shape can then be determined by a pair of \texttt{Pos}, called
+an extent,
+locating the lower left and upper right conners:
+\begin{lstlisting}
+@Obj interface Extent {
+    Pos _p1(); Pos _p2();
+}
+\end{lstlisting}
+Their definitions are stable so we put them out of the family and annotated
+with \texttt{@Obj}.
+
+Given an individual shape, its extent can be calculated according to its type.
+We therefore extend \texttt{Shape} with a \texttt{toExtent} method:
+
+\begin{lstlisting}
+    interface Shape {
+        Extent toExtent();
+    }
+\end{lstlisting}
+
+When pictures are combined together, the center point is changed and hence the
+positions of shapes in both pictures should all be recalculated.
+The transformation sub-language records how to adjust those positions,
+which includes identity transformation, translations, and compositions of the two:
+
+\lstinputlisting[linerange=300-319]{../src/diagrams/Diagrams.java}%APPLY:TRANSFORM
+
+\paragraph{Drawing}
+To easily translate a \texttt{Picture} into SVG format, we flatten the recursively structured pictures into a non-empty
+sequence of transformed styled shapes.
+\texttt{Drawing} contains three lists of the same length carrying all the
+information needed for generating a SVG file:
+
+\begin{lstlisting}
+interface Drawing {
+    List<? extends Shape> _shapes();
+    List<? extends StyleSheet> _styles();
+    List<? extends Transform> _transforms();
+}
+\end{lstlisting}
+
+% Step 1: Picture to Drawing
+To generate a \texttt{Drawing} from a \texttt{Picture}, we extend the
+\texttt{Picture} hierarchy with a \texttt{draw} method:
+\lstinputlisting[linerange=268-270]{../src/diagrams/Diagrams.java}%APPLY:DRAW
+For Place,
+
+% The implementation for each case is omitted here for space reasons.
+
+% Step 2: Drawing to XML
+SVG is based on XML and the structure of an XML file can be captured by following definitions:
+\begin{lstlisting}
+@Obj interface XML {
+    String _tag();
+    List<Attr> _attrs();
+    List<XML> _xmls();
+}
+@Obj interface Attr {
+    String _name();
+    String _value();
+}
+\end{lstlisting}
+We then define \texttt{toXML} method both in \texttt{Drawing} and \texttt{Shape}.
+\texttt{Drawing} builds up a complete XML by calling the \texttt{toXML} on each
+\texttt{Shape}.
+
+The final step is to pretty print the content of an \texttt{XML} object and
+write it to an file with extension
+\emph{.svg}.
+Then open the file using a modern browser, you will see the right side of
+Figure~\ref{woman} appear on the screen.
+
+\subsubsection{Discussion}
+\cite{} contains exercises about extensions on the language, e.g.
+adding new shapes. The deep embedding approach in Haskell requires modifications on the ADTs
+whereas our approach can modularly introduce new language constructs.
+Each sub-language retains independent extensibility.
+For example, we can add a \texttt{show} method for the purpose of debugging.

### 2016-06-09 paper/sections/Overview.tex
More case study

@@
     Circuit _c1 = c1; Circuit _c2 = c2;
-    public Circuit _c1() { return c1; }
-    public Circuit _c2() { return c2; }
+    public Circuit _c1() { return _c1; }
+    public Circuit _c2() { return _c2; }
   }
@@
 
-Here comes the solution from Gibbons and Wu.
+Here comes the solution from Gibbons and Wu:
 \lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
 where tuples are used to aggregate multiple interpretations in one definition.
-The way to introduce \texttt{desugar} in previous section is similar to this
+The way to introduce \texttt{desugar} in Section~\ref{sec:oo} is similar to this
 approach, where we use records instead.
@@
 
-Note that we do not explicitly write down \texttt{extends} clauses in the
+Note that we do not explicitly write down \texttt{extends} clauses for known members in the
 extended family.
-Instead, they are inferred by \name according to their
-names and the inheritance relationship between families.
+Instead, \texttt{extends} clauses are inferred by \name according to their
+names and the inheritance relationship stated between families.
 After instrumentation done by \name, the definitions of \texttt{Circuit} and \texttt{Beside}
@@
 \end{lstlisting}
-Interfaces from base families of the same name will be put after \texttt{extends} clause and the
-interface hierarchy of \texttt{Circuit} is preserved in the extended family.
-Moreover, \name re-declare getters in the extended \texttt{Beside} so that their return types are refined to be the new \texttt{Circuit}.
+The corresponding definitions from \texttt{Init} are inherited. For class-like
+interfaces like \texttt{Beside}, the new \texttt{Circuit} is also extended.
+Moreover, \name re-declare its getters so that their return types are refined to be the new \texttt{Circuit}.
 
@@
 Here, we show yet another appealing feature of \name - interpretation combination.
+That is, we can modularly merge interpretations defined for the same data structure together.
 The \texttt{Dependent} family combines two base families
 
-merge interpretations defined for the same data structure together.
 %Since \texttt{Circuit} in \texttt{Dependent} family extends the one from
@@
 With these new constructs, more complex circuits can be constructed.
-For example, Figure~\ref{} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.
+For example, Figure~\ref{sec:oo} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.
 

### 2016-06-09 paper/sections/ShallowOO.tex
More case study

@@
-\section{Shallow Object-Oriented Programming}
+\section{Shallow Object-Oriented Programming}\label{sec:oo}
 
@@
 with transformation that we can port back to Haskell.
-
-Shallow embedding interpreter pattern
-Deep embedding visitor pattern
 \end{comment}
@@
 transformations in shallow embedding, which is typically thought to be the
-strength of deep embedding.
+privilege of deep embedding.
 
@@
 \lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
-where the datatype maps to the abstract class \texttt{Circuit} with its field declared as an abstract method.
+where the record type maps to the abstract class \texttt{Circuit} and field
+declaration becomes method declaration.
 Each case in the semantic function corresponds to an class and its parameters become fields of the class.
@@
 be rewritten via the following formula:
-$$identity\ n = (fan\ 1) \overbrace{\ beside\ (fan\ 1)}^{n-1}\ (n > 0)$$
+$$identity\ n = (fan\ 1) \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (n > 0)$$
 which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.

### 2016-06-09 src/diagrams/Diagrams.java

@@ -57,14 +57,14 @@ interface Attr extends Show {
 
 @Obj
 interface Extent extends Show {
-    Pos _p1();
-    Pos _p2();
+    Pos _l();
+    Pos _r();
     default Extent union(Extent e) {
-        return Extent.of(Pos.of(Math.min(_p1()._1(), e._p1()._1()), Math.min(_p1()._2(), e._p1()._2())),
-                Pos.of(Math.max(_p2()._1(), e._p2()._1()), Math.max(_p2()._2(), e._p2()._2())));
+        return Extent.of(Pos.of(Math.min(_l()._x(), e._l()._x()), Math.min(_l()._y(), e._l()._y())),
+                Pos.of(Math.max(_r()._x(), e._r()._x()), Math.max(_r()._y(), e._r()._y())));
     }
     default String show() {
-        return "(" + _p1().show() + "," + _p2().show() + ")";
+        return "(" + _l().show() + "," + _r().show() + ")";
     }
 }
 
@@ -86,37 +86,89 @@ interface XML extends Show {
 
 @Obj
 interface Pos extends Show {
-    double _1();
-    double _2();
+    double _x();
+    double _y();
     default Pos add(Pos other) {
-        return Pos.of(_1() + other._1(), _2() + other._2());
+        return Pos.of(_x() + other._x(), _y() + other._y());
     }
     default Pos resize(double scale) {
-        return Pos.of(scale*_1(), scale*_2());
+        return Pos.of(scale*_x(), scale*_y());
     }
     default String show() {
-        return _1() + "," + _2();
+        return _x() + "," + _y();
     }
     default List<Attr> toAttrs(String x, String y) {
-        return Arrays.asList(Attr.of(x, ""+_1()), Attr.of(y, ""+_2()));
+        return Arrays.asList(Attr.of(x, ""+_x()), Attr.of(y, ""+_y()));
     }
     default Pos conjugate() {
-        return Pos.of(_1(), -_2());
+        return Pos.of(_x(), -_y());
     }
 }
+
 @Obj
-interface Family {
-    /**
-     * Shape
-     */
+interface AST {
+    //BEGIN_SHAPE
+    interface Shape { }
+    interface Rectangle extends Shape {
+        double _x(); double _y();
+    }
+    interface Ellipse extends Shape {
+        double _rx(); double _ry();
+    }
+    interface Triangle extends Shape {
+        double _l();
+    }
+    //END_SHAPE
+
+    //BEGIN_STYLE
[…]

### 2016-06-09 src/paper/sec3/Circuit.java

@@ -52,22 +52,31 @@ import lombok.Obj;
 }
 //END_MULTIPLE
 
-//BEGIN_DEPENDENT
-@Obj interface Dependent extends Multiple {
-  interface Stretch {
+//BEGIN_SYNTAX
+@Obj interface Syntax extends Init {
+  interface Stretch extends Circuit {
     List<Integer> _ns();
     Circuit _c();
     default int width() {
       return _ns().stream()
         .mapToInt(Integer::intValue).sum();
     }
+  }
+  interface Above extends Circuit {
+    Circuit _c1(); Circuit _c2();
+    default int width() { return _c1().width(); }
+  }
+}
+//END_SYNTAX
+
+//BEGIN_DEPENDENT
+@Obj interface Dependent extends Syntax, Multiple {
+  interface Stretch {
     default boolean wellSized() {
       return _ns().size() == _c().width();
     }
   }
   interface Above {
-    Circuit _c1(); Circuit _c2();
-    default int width() { return _c1().width(); }
     default boolean wellSized() {
       return _c1().wellSized() && _c2().wellSized()
         && _c1().width() == _c2().width();

### 2016-06-09 paper/sections/RelatedWork.tex
Some text on related work

--- 2016-04-19 paper/sections/RelatedWork.tex
@@
 \paragraph{Deep and Shallow Embeddings}
+When designing embedded DSLs, there always exists a choice between deep
+embedding and shallow embedding.
+However, either approach has its own strengths and weakness.
+Deep embedding uses an AST to represent programs, allowing
+analyses, optimizations or further translation on the programs.
+Nonetheless, the AST definition is typically not extensible, as a result
+adding new language constructs becomes difficult.
 
-\paragraph{Modularity of DSLs?}
-Data types a la Carte; Finally Tagless; 
-Object Algebras
+Shallow embedding directly encodes the source language using the semantics of
+the host language bypassing an AST, which yields flexible and concise implementations.
+The dual problem in shallow embedding is that the semantic domain is fixed,
+making it hard to introduce new interpretations.
+Although Gibbons and Wu ~\cite{gibbons} showed that it is possible to define multiple
+interpretations via tuples, the encoding requires modifications on existing code.
+
+To enjoy the benefits of both techniques, \cite{svenningsson2012combining}
+combines deep and shallow embedding together, which uses shallow embedding to
+encode the surface language and translates it to a deeply embedded core language.
+Similarly, Yin-Yang~\cite{Jovanovic:2014:YCD:2658761.2658771} allows user
+program written in shallow embedding and automatically generates the deeply
+embedded version through Scala macros. Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at
+the same time so that clients can choose for a
+particular interpretation.
+Our approach, however, can be viewed as a hybrid of shallow and deep embedding -
+the encoding looks like shallow embedding while preserving the AST.
+
+%Hudak 1. building DSL 2. modular DSL
+
+% Our approach. The simplicity . The AST is preserved, allowing transformations
+
+\paragraph{Modularity of DSLs}
+How to add variants (language constructs) as well as operations
+(interpretations) modularly, a.k.a the Expression Problem (EP), is a hot research
+topic. There are many solutions to EP. We discuss some of them in
+OOP and in FP respectively.
+
+In OOP, the Vistor pattern is widely used for modeling languages, which is
+well-known for adding new operations easily. Many works
+have been done for making it possible to extend variants~\cite{oliveira07genericity,oliveira09modular}.
+%In fact, the Visitor pattern has a strong connection with deep and shallow embedding.
+% Specifically, the internal variant corresponds to shallow embedding whereas the external variant is equivalent to deep embedding.
+Object Algebras~\cite{bruno12oa} simplifies the solution, which is an variant of
+internal visitor pattern.
+ % However, this approach requires heavy encoding and advanced type system features in Scala.
+Another line of work is on OO decomposition. ~\cite{zenger} capture recursive
+arguments using virtual types with bounds. Wang and Oliveira~\cite{eptrivially16}
+uses covariant types, a common feature available in most OO languages, to achieve the same goal which is available in most OO languages.
+Compared to the Visitor pattern, the OO decomposition approach is simpler and our encoding is based on ~\cite{eptrivially16}.
+
+In FP, the two main solutions are \emph{finally
+  tagless}~\cite{carette2009finally} and \emph{data types a la
+  carte}~\cite{swierstra2008data} (DTC).
+Finally tagless approach uses a type class to abstract over all interpretations
+of the language. Concrete interpretations are given through creating a data type and
+making it an instance of that type.
+DTC represents language constructs separately and composes them together using
+extensible sums. However, not like OO languages which come with subtyping, one
+has to manually implement the subtyping machinery for variants.
+Moreover, neither DTC nor finally tagless approach can add new interpretations
+that depend on existing interpretations modularly like our approach.
 

### 2016-06-09 paper/code/NewSemantics.hs

@@ -1,6 +1,6 @@
 --BEGIN_SEMANTICS_HS
-identity n = (n,True)
-fan n = (n,True)
+identity n = (n,n > 0)
+fan n = (n,n > 0)
 beside c1 c2 = 
   (width c1 + width c2,wellSized c1 && wellSized c2)
 

### 2016-06-09 paper/sections/RelatedWork.tex
Multiple inheritance.

@@
 proposes a simple extension to Featherweight Generic Java[6], introducing self-type variables. There are also some existing languages,
-like Scala, which supports symmetric mixin compositions and self-type annotations, hence can provide better support for encoding family
+like Scala, which supports symmetric mixin compositions and self-type annotations[?], hence can provide better support for encoding family
 polymorphism. Our lightweight encoding relies entirely on the existing Java language without language extensions, and certainly some features from
-the original paper are sacrificed, including the mismatching problem of recursive class definitions (also known as binary methods).
+the original paper are sacrificed, including the mismatching problem of recursive class definitions (also known as binary methods[7]). Relying
+on the Java semantics, covariant return types are supported and used for our automatic type refinements. Nevertheless, in the case of binary methods,
+we cannot address those with recursive member types as parameters. This problem is however handled in some other work like MyType[?], ThisType[?] and self-type
+annotations in Scala as we said.
 
@@
 
-\paragraph{ThisType}
+%\paragraph{ThisType}
 
 \paragraph{Multiple Inheritance}
+
+Multiple inheritance is an expressive and useful feature in programming languages, yet difficult to model in relation to the famous diamond problem[?].
+Many models have been developed to integrate multiple inheritance in some languages, including C++ virtual inheritance[?], mixins[?], traits[?] and
+hybrid models such as CZ[?]. In our approach, introducing new variants or operations requires subtyping relations among member types to be built, and
+hence we rely on the Java multiple interface inheritance. A difference from the other models is that our \textsf{@Family} annotation can check method
+typing and detect conflicts during the annotation processing, and on the other hand, subtyping relations are automatically generated for the users, which
+makes client code more concise. Furthermore, we have mentioned that interfaces are updated with auto-generated constructor methods, whereas object initialization is a big problem in some other models.

### 2016-06-10 paper/sections/CaseStudy.tex
Some text

@@
 circuit using Java Swing.
-Calling \texttt{circuit.draw()} on the \texttt{circuit} will
+Calling \texttt{draw} on the \texttt{circuit} will
 display what Figure~\ref{sec:} shows.
@@
 To actually visualize the picture as shown in the right hand side of
-Figure~\ref{woman}, we give semantics to the language for accomplishing this goal.
+Figure~\ref{woman}, we need to give semantics to the language for animation.
 
@@
 \begin{minipage}[t]{.9\textwidth}
-\lstinputlisting[linerange=530-537]{../src/diagrams/Diagrams.java}%APPLY:WOMAN
+\lstinputlisting[linerange=537-544]{../src/diagrams/Diagrams.java}%APPLY:WOMAN
 \end{minipage}
@@
 \end{lstlisting}
-Their definitions are stable so we put them out of the family and annotated
+These definitions are stable so we put them out of the family and annotated
 with \texttt{@Obj}.
@@
 
-\lstinputlisting[linerange=300-319]{../src/diagrams/Diagrams.java}%APPLY:TRANSFORM
+\lstinputlisting[linerange=164-171]{../src/diagrams/Diagrams.java}%APPLY:TRANSFORM
 
@@
 To easily translate a \texttt{Picture} into SVG format, we flatten the recursively structured pictures into a non-empty
-sequence of transformed styled shapes.
-\texttt{Drawing} contains three lists of the same length carrying all the
-information needed for generating a SVG file:
+sequence of transformed styled shapes, represented by \texttt{Drawing}:
 
@@
 \texttt{Picture} hierarchy with a \texttt{draw} method:
-\lstinputlisting[linerange=268-270]{../src/diagrams/Diagrams.java}%APPLY:DRAW
-For Place,
+\lstinputlisting[linerange=279-281]{../src/diagrams/Diagrams.java}%APPLY:DRAW
 
@@
 % Step 2: Drawing to XML
-SVG is based on XML and the structure of an XML file can be captured by following definitions:
+With all the necessary information contained in drawing, we can now generate the
+SVG file. SVG is based on XML and the structure of an XML file can be captured by following definitions:
 \begin{lstlisting}
@@
 @Obj interface Attr {
-    String _name();
-    String _value();
+    String _name(); String _value();
 }
 \end{lstlisting}
-We then define \texttt{toXML} method both in \texttt{Drawing} and \texttt{Shape}.
-\texttt{Drawing} builds up a complete XML by calling the \texttt{toXML} on each
-\texttt{Shape}.
+We then extend a \texttt{toXML} method both in \texttt{Drawing} and \texttt{Shape}.
+\begin{lstlisting}
+interface Shape {
+  XML toXML(List<Attr> styleAttrs, Transform trans);
+}
+interface Drawing {
+  XML toXML();
+}
+\end{lstlisting}
+Calling \texttt{toXML} on a \texttt{Drawing} will builds up a complete XML by calling the \texttt{toXML} on each \texttt{Shape}.
 
 The final step is to pretty print the content of an \texttt{XML} object and
-write it to an file with extension
-\emph{.svg}.
-Then open the file using a modern browser, you will see the right side of
+save it to an file with the extension \emph{svg}.
+Then open the file with a modern browser, you will see the right side of
 Figure~\ref{woman} appear on the screen.
@@
 \subsubsection{Discussion}
+Each sub-language retains independent extensibility.
 \cite{} contains exercises about extensions on the language, e.g.
-adding new shapes. The deep embedding approach in Haskell requires modifications on the ADTs
+adding new shapes.
+The deep embedding approach in Haskell requires modifications on the ADTs
 whereas our approach can modularly introduce new language constructs.
-Each sub-language retains independent extensibility.
+At the same time, new interpretations can be added modularly.
 For example, we can add a \texttt{show} method for the purpose of debugging.

### 2016-06-10 paper/sections/Overview.tex
Some text

@@
 not modular extensible. Though they also presented modular solution based on
-datatype a lar carte\cite{}. But such solution changes the encoding style
+\emph{data types a la carte}\cite{swierstra2008data}. But such solution changes the encoding style
 dramatically and may be too complex to use practically.
@@
 multiple interface inheritance of Java;
-3) Fields of the classes become getter methods (starting with an underscore) so that their types can be refined for retaining extensibility~\cite{}.
+3) Fields of the classes become getter methods (starting with an underscore) so that their types can be refined for retaining extensibility~\cite{eptrivially16}.
 
@@
 especially for functional languages.
-Suppose that we want to have a function that checks whether a circuit is constructed correctly.
-
-Here comes the solution from Gibbons and Wu:
+Suppose that we want to have a function that checks whether a circuit is constructed correctly. Here comes the solution from Gibbons and Wu:
 \lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
-where tuples are used to aggregate multiple interpretations in one definition.
-The way to introduce \texttt{desugar} in Section~\ref{sec:oo} is similar to this
-approach, where we use records instead.
+Tuples are used to aggregate multiple interpretations in one definition.
+The way to introduce \texttt{desugar} demonstrated in Section~\ref{sec:oo} is similar to this
+approach, where records are used instead.
 However, neither tuples nor records are extensible - whenever we need a new
@@
   interpretation}. %They are hard to express modularly in traditional shallow embedding.
-Before we introduce an example of such interpretations, we first add two new
-constructs to our language:
-\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
-summing up \emph{ns} and $above\ c_1\ c_2$ combines two circuits of the same width vertically.
+Before we introduce an example of such interpretations, we first show how to add new
+constructs in \name.
+
+The new constructs are: \emph{stretch ns c} which inserts additional wires into the circuit \emph{c} by
+summing up \emph{ns} and $above\ c_1\ c_2$ which combines two circuits of the same width vertically.
 Their types are given below:
@@
 
-%\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
+%\lstinputlisting[linerange=56-68]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
 %With these new constructs, more complex circuits can be constructed.
@@
 
-Adding new constructs in \name is easy: just we declare new class-like
+Adding new constructs in \name is easy - just we define new class-like
 interfaces in an extended family:
-\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
+\lstinputlisting[linerange=56-68]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
 Only definitions of the new constructs \texttt{Above} and \texttt{Stretch} are given.
@@
 
-Not arbitrary circuits can be combined using \texttt{Above} and \texttt{Stretch}.
+Not arbitrary circuits can be combined using \texttt{Above} and \texttt{Stretch}
+as stated in their specifications.
 We hence define the \texttt{wellSized} method for these two new constructs to
 verify the constraints they imply:
-\lstinputlisting[linerange=73-85]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT
+\lstinputlisting[linerange=72-84]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT
 
@@
 rely on another method defined on the \texttt{Circuit}, \texttt{width}, for calculating the width of a circuit.
+Most importantly, our approach do not force dependent interpretation to be
+defined together with what they depend on. In the case of \texttt{width} and
+\texttt{wellSized}, they can be defined in separate families as long as the
+signature of \texttt{width} is exposed to the definition of \texttt{wellSized}.
 
-Here, we show yet another appealing feature of \name - interpretation combination.
+The definition of \texttt{Dependent} shows yet another appealing feature of \name - interpretation combination.
 That is, we can modularly merge interpretations defined for the same data structure together.
@@
 
-\lstinputlisting[linerange=107-139]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE
+\lstinputlisting[linerange=106-138]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE
 
@@
 
-\lstinputlisting[linerange=170-172]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
+\lstinputlisting[linerange=169-171]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
 Simply defining a new family \texttt{Merge} that extends other families we want
@@
 
-\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
+\lstinputlisting[linerange=56-68]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
 
@@
 
-\lstinputlisting[linerange=177-189]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
+\lstinputlisting[linerange=176-188]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
 % TODO: not implemented yet

### 2016-06-10 paper/sections/RelatedWork.tex
Some text

@@
 Another line of work is on OO decomposition. ~\cite{zenger} capture recursive
-arguments using virtual types with bounds. Wang and Oliveira~\cite{eptrivially16}
-uses covariant types, a common feature available in most OO languages, to achieve the same goal which is available in most OO languages.
-Compared to the Visitor pattern, the OO decomposition approach is simpler and our encoding is based on ~\cite{eptrivially16}.
+arguments using virtual types with bounds using Scala. Wang and Oliveira~\cite{eptrivially16}
+uses covariant types, a common feature available in most OO languages, to
+achieve same goal, which is the basis of our approach.
+Compared to the Visitor pattern, the OO decomposition approach is simpler and
+our encoding is based on ~\cite{eptrivially16}. The
+naming problem for extensions is solved by family polymorphism.
 
@@
 has to manually implement the subtyping machinery for variants.
-Moreover, neither DTC nor finally tagless approach can add new interpretations
-that depend on existing interpretations modularly like our approach.
+Moreover, both DTC and finally tagless approach force dependent interpretations
+to be defined along with what they depend on.
+This prevents new interpretation that depend on existing interpretations from
+being defined modularly while our approach does not have such requirement.
 

### 2016-06-10 src/circuit/CircuitDSL.java

@@ -41,7 +41,7 @@ interface AST {
 }
 
 @Obj
-interface DepthFeature extends AST {
+interface Depth extends AST {
     interface Circuit {
         int depth();
     }
@@ -73,7 +73,7 @@ interface DepthFeature extends AST {
 }
 
 @Obj
-interface WidthFeature extends AST {
+interface Width extends AST {
     interface Circuit {
         int width();
     }
@@ -99,46 +99,46 @@ interface WidthFeature extends AST {
     }
     interface Stretch {
         default int width() {
-            return _ns().stream().reduce(0, (a, b) -> a + b);
+            return _ns().stream().reduce(0, Integer::sum);
         }
     }
 }
 
 
 @Obj
-interface WellSizeFeature extends WidthFeature, DepthFeature {
+interface WellSized extends Width, Depth {
     interface Circuit {
-        boolean wellSize();
+        boolean wellSized();
     }
     interface Identity {
-        default boolean wellSize() {
+        default boolean wellSized() {
             return true;
         }
     }
     interface Fan {
-        default boolean wellSize() {
+        default boolean wellSized() {
             return true;
         }
     }
     interface Above {
-        default boolean wellSize() {
-            return _c1().wellSize() && _c2().wellSize() && _c1().width() == _c2().width();
+        default boolean wellSized() {
+            return _c1().wellSized() && _c2().wellSized() && _c1().width() == _c2().width();
         }
     }
     interface Beside {
-        default boolean wellSize() {
-            return _c1().wellSize() && _c2().wellSize();
+        default boolean wellSized() {
+            return _c1().wellSized() && _c2().wellSized();
         }
     }
     interface Stretch {
-        default boolean wellSize() {
-            return _c().wellSize() && _ns().size() == _c().width();
+        default boolean wellSized() {
+            return _c().wellSized() && _ns().size() == _c().width();
         }
     }
 }
 
 @Obj
-interface ExtendedAST extends WellSizeFeature {
+interface ExtendedAST extends WellSized {
     interface RStretch extends Stretch {}
 }
 
@@ -169,13 +169,7 @@ interface LayoutFeature extends ExtendedAST {
     }
     interface Stretch {
         default Layout layout() {
-            int acc = 0;
-            List<Integer> _ns = new ArrayList<>();
-            for (int n : _ns()) {
[…]

### 2016-06-10 src/circuit/Test.java

@@ -1,41 +1,20 @@
 package circuit;
 
 import static circuit.CircuitDSL.fan;
-import static circuit.CircuitDSL.identity;
+import static circuit.CircuitDSL.id;
 
-import java.util.Arrays;
-
-import circuit.CircuitDSL.Above;
-import circuit.CircuitDSL.Beside;
 import circuit.CircuitDSL.Circuit;
-import circuit.CircuitDSL.Fan;
-import circuit.CircuitDSL.Identity;
-import circuit.CircuitDSL.Stretch;
 
 public class Test {
     public static void main(String[] args) {
-        // Figure 3
-        Circuit c3 = fan(2).beside(identity(1))
-                .above(identity(1).beside(fan(2)));
-        // Figure 4
-        Circuit c4 = fan(3).stretch(3,2,3);
-        Circuit c4r = fan(3).rStretch(2,3,1);
-        // Figure 1
-        Circuit c1 =
-                Above.of(
-                        Above.of(
-                                Beside.of(fan(2), Fan.of(2)),
-                                Stretch.of(Fan.of(2), IntList.of(Arrays.asList(2,2)))),
-                        Beside.of(
-                                Beside.of(
-                                        Identity.of(1),
-                                        Fan.of(2)),
-                                Identity.of(1)));
         Circuit circuit =
             fan(2).beside(fan(2))
             .above(fan(2).stretch(2,2))
-            .above(identity(1).beside(fan(2)).beside(identity(1)));
+            .above(id(1).beside(fan(2)).beside(id(1)));
+//        circuit.width();
+//        circuit.depth();
+//        circuit.layout();
+//        circuit.tlayout(f)
         circuit.draw();
-        System.out.println(circuit.layout().show());
     }
 }

### 2016-06-10 src/diagrams/Diagrams.java

@@ -159,6 +159,17 @@ interface AST {
         Picture _p1(); Picture _p2();
     }
     //END_PICTURE
+
+    //BEGIN_TRANSFORM
+    interface Transform {}
+    interface Identity extends Transform {}
+    interface Translate extends Transform {
+        Pos _p();
+    }
+    interface Compose extends Transform {
+        Transform _t1(); Transform _t2();
+    }
+    //END_TRANSFORM
 }
 
 @Obj
@@ -296,28 +307,24 @@ interface SVG extends AST {
         }
     }
 
-    //BEGIN_TRANSFORM
     interface Transform {
         Pos transform(Pos p);
     }
-    interface Identity extends Transform {
+    interface Identity {
         @Override default Pos transform(Pos p) {
             return p;
         }
     }
-    interface Translate extends Transform {
-        Pos _p();
+    interface Translate {
         @Override default Pos transform(Pos p) {
             return _p().add(p);
         }
     }
-    interface Compose extends Transform {
-        Transform _t1(); Transform _t2();
+    interface Compose {
         @Override default Pos transform(Pos p) {
             return _t1().transform(_t2().transform(p));
         }
     }
-    //END_TRANSFORM
 
     interface Drawing {
         List<? extends Transform> _transforms();

### 2016-06-10 src/paper/sec3/Circuit.java

@@ -55,8 +55,7 @@ import lombok.Obj;
 //BEGIN_SYNTAX
 @Obj interface Syntax extends Init {
   interface Stretch extends Circuit {
-    List<Integer> _ns();
-    Circuit _c();
+    List<Integer> _ns(); Circuit _c();
     default int width() {
       return _ns().stream()
         .mapToInt(Integer::intValue).sum();

### 2016-06-11 paper/paper.tex
Add case studies
paper structure
Some text for S3
Text for S2
Text. Code generation.
abstract

@@
-\documentclass[10pt,numbers,preprint,nocopyrightspace]{sigplanconf}
+\documentclass[10pt,preprint,nocopyrightspace]{sigplanconf}
 
@@
 \usepackage{enumitem}
+\usepackage{stmaryrd}
 
-\setlist[itemize]{noitemsep,nolistsep}
-\setlist[enumerate]{noitemsep,nolistsep}
+\newcommand{\name}{\textbf{@Family}\xspace}
+\newcommand{\dsl}{\emph{Scans}\xspace}
+\newcommand{\interp}{\textsc{Interpreter}\xspace}
+
+%\newcommand{\authornote}[3]{}
+\newcommand{\authornote}[3]{{\color{#2} {\sc #1}:#3}}
+\newcommand{\weixin}[1]{\authornote{weixin}{cyan}{#1}}
+% \setlist[itemize]{noitemsep,nolistsep}
+% \setlist[enumerate]{noitemsep,nolistsep}
 
@@
 
-\title{Title}
+\title{Shallow EDSLs and Object-Oriented Programming}
 
@@
 \begin{abstract}
-TODO
+
+Shallow Embedded Domain Specific Languages (EDSLs) use
+\emph{procedural abstraction} to directly encode a Domain Specific
+Language into an existing host language. Procedural abstraction has
+been argued to be the essence of Object-Oriented Programming (OOP). Given
+that Object-Oriented Languages have evolved over more than 50 years 
+to optimize the use of procedural abstraction, they ought to have some 
+advantages to encode shallow EDSLs.
+
+This paper argues that OOP abstractions, including \emph{inheritance}
+and \emph{subtyping}, increase the modularity and reuse of shallow
+DSLs when compared to classical procedural abstraction. We make this
+argument by taking a recent paper by Gibbons, where procedural
+abstraction is used in Haskell to model a simple EDSL, and we recode
+that EDSL in Java. Although from the \emph{syntactical} point of view
+there are obvious disadvantages in the Java version, from the semantic
+and modularity point of view the Java version has clear advantages.
+To alleviate some of the syntactical disadvantages of Java, we create
+an annotation that uses transparent code generation techniques to
+automatically eliminate large portions of boilerplate code in Java.
+To further illustrate the applicability of our techniques, we conduct 
+several case studies using larger DSLs from the literature.
+
 \end{abstract}
@@
 \input{sections/Introduction.tex}
+\input{sections/ShallowOO.tex}
 \input{sections/Overview.tex}
+\input{sections/Tool.tex}
 \input{sections/CaseStudy.tex}
 \input{sections/RelatedWork.tex}
+\input{sections/Discussion.tex}
 %\acks

### 2016-06-11 paper/paper.tex
abstract

@@
 that Object-Oriented Languages have evolved over more than 50 years 
-to optimize the use of procedural abstraction, they ought to have some 
+to improve the use of procedural abstraction, they ought to have some 
 advantages to encode shallow EDSLs.
@@
 DSLs when compared to classical procedural abstraction. We make this
-argument by taking a recent paper by Gibbons, where procedural
+argument by taking a recent paper by Gibbons and Wu, where procedural
 abstraction is used in Haskell to model a simple EDSL, and we recode
@@
 an annotation that uses transparent code generation techniques to
-automatically eliminate large portions of boilerplate code in Java.
+automatically eliminate large portions of boilerplate code.
 To further illustrate the applicability of our techniques, we conduct 

### 2016-06-13 paper/paper.tex
introduction

@@
 To alleviate some of the syntactical disadvantages of Java, we create
-an annotation that uses transparent code generation techniques to
+an annotation inspired by \emph{family polymorphism} and a recent
+solution to the Expression Problem. 
+The annotation uses transparent code generation techniques to
 automatically eliminate large portions of boilerplate code.
-To further illustrate the applicability of our techniques, we conduct 
+To further illustrate the applicability of our tool and techniques, we conduct 
 several case studies using larger DSLs from the literature. 
-
 \end{abstract}

### 2016-06-13 paper/sections/Introduction.tex
introduction

\section{Introduction}

Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{}, existing
languages (such as Haskell) have been used to directly encode DSLs. Two common approaches to EDSLs are the
so-called \emph{shallow} and \emph{deep} embeddings. The difference 
between these two styles of embeddings is commonly described as 
follows:

\begin{quote}
\emph{With a deep embedding, terms in the DSL are implemented simply to
construct an abstract syntax tree (AST), which is subsequently
transformed for optimization and traversed for evaluation. With a
shallow embedding, terms in the DSL are implemented directly by
their semantics, bypassing the intermediate AST and its traversal.}\cite{gibbons15folding}
\end{quote}

This definition seems to be widely accepted and similar definitions
appear in many works~\cite{}. We believe that this definition is a
vague and not very accurate, often leading to some contradicting
claims. For example, in their work on EDSLs~\cite{}, Gibbons and Wu
claim that deep embeddings (which encode ASTs using algebraic
datatypes in Haskell) allow adding new DSL interpretations easily, but
they make adding new language constructs difficult. In contrast
Gibbons and Wu claim that shallow embeddings have dual modularity
properties: new cases are easy to add, but new interpretations are
hard.  A contradiction happens if instead of using Haskell and
algebraic datatypes, one uses an OO language to encode an AST, for
example with the {\sc Composite} pattern. In that case, according to
definition of deep embedding above, the OO approach classifies as a
deep embedding. Nevertheless, as the Expression Problem~\cite{} tell
us, in the OO approach adding new cases is easy, but adding
interpretations is hard. Thus we have a contradiction, since we have
an AST representation (i.e. a deep embedding) with the modularity
properties of shallow embeddings.

We believe that the core of problem is that ASTs can be represented 
in multiple ways. In particular, it is well know that functions alone 
are enough to encode datastructures such as ASTs (via Church
encodings~\cite{}). Distinguishing deep and shallow
embeddings solely on whether a datastructure is being used 
or whether functions are used to encode interpretations directly 
falls short when the functions themselfves represent an AST.

In this paper we propose a different definition for deep and shallow
embeddings, by distinguishing on what kind of data abstraction is used
by each encoding. Deep embeddings use user-defined types
(i.e. disjoint unions or algebraic datatypes), whereas shallow
embeddings use procedural abstraction. This distinction is inspired by
Reynolds 1975 paper~\cite{} on user-defined types and procedural
abstraction being complementary approaches to data abstraction, which
many view~\cite{} as the first documented discussion on the Expression
Problem.

Shallow Embedded Domain Specific Languages (EDSLs) use
\emph{procedural abstraction} to directly encode a Domain Specific
Language into an existing host language. Procedural abstraction has
been argued to be the essence of Object-Oriented Programming (OOP). Given
that Object-Oriented Languages have evolved over more than 50 years 
to improve the use of procedural abstraction, they ought to have some 
advantages to encode shallow EDSLs.

This paper argues that OOP abstractions, including \emph{inheritance}
and \emph{subtyping}, increase the modularity and reuse of shallow
DSLs when compared to classical procedural abstraction. We make this
argument by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple EDSL, and we recode
that EDSL in Java. Although from the \emph{syntactical} point of view
there are obvious disadvantages in the Java version, from the semantic
and modularity point of view the Java version has clear advantages.
To alleviate some of the syntactical disadvantages of Java, we create
an annotation inspired by \emph{family polymorphism} and a recent
solution to the Expression Problem. 
The annotation uses transparent code generation techniques to
automatically eliminate large portions of boilerplate code.
To further illustrate the applicability of our tool and techniques, we conduct 
several case studies using larger DSLs from the literature.

### 2016-06-13 paper/sections/CaseStudyPPrinter.tex
refs.

@@
 
-This case study refactors the Haskell code from \textcolor{red}{(Haoyuan: ref, a prettier printer)}
+This case study refactors the Haskell code from \cite{Wadler98aprettier}
 which uses deep embeddings to implement a functional pretty printer library with high efficiency. In the original code, two

### 2016-06-13 paper/sections/RelatedWork.tex
refs.

@@
 
-There has been a lot of related work on family polymorphism[1], including various lightweight encodings[2,3,4,5]. In the original
-paper[1], nested classes are represented as attributes of an object, which involves a dependent type system. In [4], Saito and Igarashi
-proposed a lightweight variant of family polymorphism, which uses classes rather than objects to represent families. Later work in [4]
-proposes a simple extension to Featherweight Generic Java[6], introducing self-type variables. There are also some existing languages,
+There has been a lot of related work on family polymorphism~\cite{ernst2001family}, including various lightweight encodings~\cite{Kamina:2007:LSC:1289971.1289996,saito2007essence,igarashi2005lightweight,Kamina:2008:LDC:1449913.1449932}.
+In the original
+paper~\cite{ernst2001family}, nested classes are represented as attributes of an object, which involves a dependent type system. In~\cite{igarashi2005lightweight}, Saito and Igarashi
+proposed a lightweight variant of family polymorphism, which uses classes rather than objects to represent families. Later work in~\cite{igarashi2005lightweight}
+proposes a simple extension to Featherweight Generic Java~\cite{Igarashi:2001:FJM:503502.503505}, introducing self-type variables. There are also some existing languages,
 like Scala, which supports symmetric mixin compositions and self-type annotations[?], hence can provide better support for encoding family
 polymorphism. Our lightweight encoding relies entirely on the existing Java language without language extensions, and certainly some features from
-the original paper are sacrificed, including the mismatching problem of recursive class definitions (also known as binary methods[7]). Relying
+the original paper are sacrificed, including the mismatching problem of recursive class definitions (also known as binary methods~\cite{bruce1995binary}). Relying
 on the Java semantics, covariant return types are supported and used for our automatic type refinements. Nevertheless, in the case of binary methods,
@@
 On the other hand, our approach uses nested interfaces to build the relationship among families and members, which is different from
-path-dependent types in [1]. The work in [8] inspires us with the concept of higher-order hierarchies, and hence our \textsf{@Family} annotation
+path-dependent types in~\cite{ernst2001family}. The work in~\cite{ernst2003higher} inspires us with the concept of higher-order hierarchies, and hence our \textsf{@Family} annotation
 provides support for nested families. Furthermore, our lightweight encoding of family polymorphism is polished with the help of annotation processing
@@
 
-[1] family polymorphism
-[2] Lightweight scalable components
-[3] The essence of lightweight family polymorphism
-[4] Lightweight family polymorphism
-[5] Lightweight dependent classes
-[6] Featherweight Java
-[7] On binary method
-[8] Higher-order hierarchies
+%[1] family polymorphism
+%[2] Lightweight scalable components
+%[3] The essence of lightweight family polymorphism
+%[4] Lightweight family polymorphism
+%[5] Lightweight dependent classes
+%[6] Featherweight Java
+%[7] On binary method
+%[8] Higher-order hierarchies
 

### 2016-06-14 paper/paper.tex
refs.
typo.
Make space the same width as characters

@@
 language=Java,                % choose the language of the code
-columns=flexible,
+columns=fullflexible,
+keepspaces=true,
 lineskip=-1pt,
@@
 numbers=none,
-stepnumber=1,
 numbersep=5pt,                  % how far the line-numbers are from the code

### 2016-06-14 paper/code/Desugar.hs

@@ -4,19 +4,19 @@ beside :: Circuit -> Circuit -> Circuit
 
 --BEGIN_DESUGAR
 data Circuit = Circuit {
-  width :: Int,
+  width   :: Int,
   desugar :: Circuit
 }
 identity n = Circuit {
-  width = n,
+  width   = n,
   desugar = foldr1 beside $ replicate n (fan 1)
 }
 fan n = Circuit {
-  width = n,
+  width   = n,
   desugar = fan n
 }
 beside c1 c2 = Circuit {
-  width = width c1 + width c2,
+  width   = width c1 + width c2,
   desugar = beside (desugar c1) (desugar c2)
 }
 --END_DESUGAR

### 2016-06-14 paper/code/shallowCircuit.hs

@@ -3,34 +3,36 @@ type Circuit = ...
 END_CIRCUIT_TYPE-}
 --BEGIN_TYPES
 identity :: Int -> Circuit
-fan :: Int -> Circuit
-beside :: Circuit -> Circuit -> Circuit
+fan      :: Int -> Circuit
+beside   :: Circuit -> Circuit -> Circuit
 --END_TYPES
 
 --BEGIN_CIRCUIT1
 type Circuit = Int
-identity n = n
-fan n = n
+identity n   = n
+fan n        = n
 beside c1 c2 = c1 + c2
 
+width1 :: Int
 width1 = id
 --END_CIRCUIT1
 
 --BEGIN_CIRCUIT2
 newtype Circuit2 = Circuit2 Int
-identity2 n = Circuit2 n
-fan2 n = Circuit2 n
+identity2 n                         = Circuit2 n
+fan2 n                              = Circuit2 n
 beside2 (Circuit2 c1) (Circuit2 c2) =
   Circuit2 $ c1 + c2
 
+width2              :: Circuit2 -> Int
 width2 (Circuit2 n) = n
 --END_CIRCUIT2
 
 
 --BEGIN_CIRCUIT3
 newtype Circuit3 = Circuit3 { width3 :: Int }
-identity3 n = Circuit3 { width3 = n }
-fan3 n = Circuit3 { width3 = n }
+identity3 n   = Circuit3 { width3 = n }
+fan3 n        = Circuit3 { width3 = n }
 beside3 c1 c2 = Circuit3 {
   width3 = width3 c1 + width3 c2
 }
@@ -38,10 +40,10 @@ beside3 c1 c2 = Circuit3 {
 
 --BEGIN_SYNTAX_TYPES
 stretch :: [Int] -> Circuit -> Circuit
-above :: Circuit -> Circuit -> Circuit
+above   :: Circuit -> Circuit -> Circuit
 --END_SYNTAX_TYPES
 
 --BEGIN_SYNTAX_HS
 stretch ns c = sum ns
-above c1 c2 = c2
+above c1 c2  = c2
 --END_SYNTAX_HS

### 2016-06-14 paper/sections/Overview.tex
Add implicit staging in related work

@@
 Their types are given below:
-\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\lstinputlisting[language=haskell,linerange=42-43]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 
@@
 
-\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\lstinputlisting[language=haskell,linerange=42-43]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 \emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
@@
 \texttt{stretch} and \texttt{above}, for the semantic function:
-\lstinputlisting[language=haskell,linerange=45-46]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
+\lstinputlisting[language=haskell,linerange=47-48]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
 Similarly, defining two new interfaces

### 2016-06-14 paper/sections/RelatedWork.tex
Add implicit staging in related work

@@
 
-To enjoy the benefits of both techniques, \cite{svenningsson2012combining}
-combines deep and shallow embedding together, which uses shallow embedding to
+To enjoy the benefits of both techniques, Svenningsson and Axelsson~\cite{svenningsson2012combining}
+combine deep and shallow embedding together, which uses shallow embedding to
 encode the surface language and translates it to a deeply embedded core language.
-Similarly, Yin-Yang~\cite{Jovanovic:2014:YCD:2658761.2658771} allows user
+Similarly, \emph{Yin-Yang}~\cite{Jovanovic:2014:YCD:2658761.2658771} allows user
 program written in shallow embedding and automatically generates the deeply
 embedded version through Scala macros. Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at
-the same time so that clients can choose for a
-particular interpretation.
-Our approach, however, can be viewed as a hybrid of shallow and deep embedding -
-the encoding looks like shallow embedding while preserving the AST.
+the same time so that clients can choose for a particular interpretation.
+Scherr and Chiba~\cite{scherr2014implicit} uses \emph{implicit staging} to
+extract shallow EDSL part from a normal program and represent it in a form
+similar to AST, allowing optimizations to be performed.
+Our approach, however, supports transformations in the first place while retaining the
+simplicity of the shallow embedding.
 
@@
  % However, this approach requires heavy encoding and advanced type system features in Scala.
-Another line of work is on OO decomposition. ~\cite{zenger} capture recursive
+Another line of work is on OO decomposition. ~\cite{zenger05independentlyextensible} capture recursive
 arguments using virtual types with bounds using Scala. Wang and Oliveira~\cite{eptrivially16}
-uses covariant types, a common feature available in most OO languages, to
-achieve same goal, which is the basis of our approach.
+use covariant type refinements, a common feature available in conventional OO
+languages, to solve the EP, which is the basis of our approach.
 Compared to the Visitor pattern, the OO decomposition approach is simpler and
@@
 This prevents new interpretation that depend on existing interpretations from
-being defined modularly while our approach does not have such requirement.
+being defined modularly.
 

### 2016-06-14 paper/sections/ShallowOO.tex
Add implicit staging in related work

@@
 programming are closely related. To demonstrate it, we use a subset
-of the DSL presented in Gibbons and Wu's paper~\cite{} as the running example.
-We first give the original implementation of the language in shallow embedding
-and then gradually rewrite the definition into an ``OO style''. Also, we show how to do
-transformations in shallow embedding, which is typically thought to be the
-privilege of deep embedding.
+of the DSL presented in Gibbons and Wu's paper~\cite{gibbons2014folding} as the running example.
+We first give the original shallow embedded implementation of the language
+written in Haskell
+and gradually rewrite it approaching to an ``OO style''.
+Then translating the program into an OO version like Java becomes straightforward.
+We also illustrate how to do transformations in shallow embedding, which is typically regarded as the privilege of deep embedding.
 
@@
 $c_1$ and $c_2$ horizontally.
-For example, the circuit \emph{fan 3 beside identity 3} can be visualized in Figure~\ref{}.
+For example, the circuit \emph{beside (fan 3) (identity 3)} can be visualized in Figure~\ref{}.
 
@@
 circuit, the semantic function would be:
-\lstinputlisting[language=haskell,linerange=11-16]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
+\lstinputlisting[language=haskell,linerange=11-17]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
 As shallow embedding is semantics-oriented, it is already an integer after
@@
 For example, we will immediately get $6$ for
-\texttt{fan 3 `beside` identity 3} and hence \texttt{width1} is just an identity function.
+\texttt{beside (fan 3) (identity 3)} and hence \texttt{width1} is just an identity function.
 
@@
 datatype and get back the value through pattern matching:
-\lstinputlisting[language=haskell,linerange=20-26]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
-Then circuit construction gives back a \texttt{Circuit2} which contains an
-integer value representing the width and \texttt{width2} extracts out the value.
+\lstinputlisting[language=haskell,linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
+% Then circuit construction gives back a \texttt{Circuit2} which contains an
+% integer value representing the width and \texttt{width2} extracts out the value.
+% Is this definition still shallow embedding?
 %One may be curious whether this variant implementation is still shallow embedding or not. The answer is yes because ... % why it is shallow
@@
 The rewrite goes on by using a record with a sole field to represent the result:
-\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
+\lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
 We no longer need a separate \texttt{width3} function as the field name is the extractor.
@@
 
-This implementation is exactly how we model \dsl in an OO language for the first
+This implementation is exactly how we would model \dsl in an OO language in the first
 place, following the \interp pattern.
 In other words, shallow embedding and straightforward OO programming are closely
-related (deep embedding is closely related to the \textsc{Visitor} pattern).
+related.
+It may worth mentioning that deep embedding is closely related to the \textsc{Visitor} pattern.
 
@@
 be rewritten via the following formula:
-$$identity\ n = (fan\ 1) \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (n > 0)$$
+$$identity\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)\ (n > 0)$$
 which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.

### 2016-06-14 src/paper/sec3/Test.java
Scala code

@@ -1,0 +1,25 @@
+package paper.sec3;
+
+import java.util.List;
+
+interface A {
+    void f();
+}
+interface B extends A {
+    void g();
+}
+
+interface As {
+  List<? extends A> as();
+  default void f() {
+      as().get(0).f();
+  }
+}
+
+interface Bs extends As {
+  @Override List<? extends B> as();
+
+  default void g() {
+      as().get(0).g();
+  }
+}

### 2016-06-14 src/scala/base/Circuit.scala

@@ -1,0 +1,14 @@
+//BEGIN_CIRCUIT_SCALA
+package base
+trait Circuit { def width: Int }
+class Identity(n: Int) extends Circuit {
+  def width = n
+}
+class Fan(n: Int) extends Circuit {
+  def width = n
+}
+class Beside(c1: Circuit, c2: Circuit) 
+    extends Circuit {
+  def width = c1.width + c2.width
+}
+//END_CIRCUIT_SCALA

### 2016-06-14 src/scala/base/Set.scala

@@ -1,0 +1,22 @@
+package base
+
+trait Set {
+  def exists(m: Int, n: Int): Boolean
+}
+class None extends Set {
+  override def exists(m: Int, n: Int) = false
+}
+class All extends Set {
+  override def exists(m: Int, n: Int) = true
+}
+class Limit(m1: Int, n1: Int, s: Set) extends Set {
+  override def exists(m: Int, n: Int) = {
+    val maxM = Math.max(m, m1)
+    val minN = Math.min(n, n1)
+    maxM <= minN && s.exists(maxM, minN)
+  }
+}
+class Union(s1: Set, s2: Set) extends Set {
+  override def exists(m: Int, n: Int) =
+    s1.exists(m, n) || s2.exists(m, n)
+}

### 2016-06-14 src/scala/desugar/Circuit.scala

@@ -1,0 +1,21 @@
+package desugar
+//BEGIN_DESUGAR_SCALA
+trait Circuit extends base.Circuit {
+  def desugar: Circuit
+}
+class Fan(n: Int) 
+    extends base.Fan(n) with Circuit {
+  def desugar = this
+}
+class Beside(c1: Circuit, c2: Circuit) 
+    extends base.Beside(c1, c2) with Circuit {
+  def desugar = new Beside(c1.desugar, c2.desugar)
+}
+class Identity(n: Int) 
+    extends base.Identity(n) with Circuit {
+  def desugar = {
+    val fan1: Circuit = new Fan(1)
+    (1 until n) map{_=>fan1} reduce{new Beside(_,_)}
+  }
+}
+//END_DESUGAR_SCALA

### 2016-06-14 src/scala/desugar/Set.scala

@@ -1,0 +1,24 @@
+package desugar
+
+trait Set extends base.Set {
+  def opt(m: Int, n: Int): Set = this
+}
+class None extends base.None with Set {}
+class All extends base.All with Set {
+  override  def opt(m: Int, n: Int) = new Limit(m, n, this)
+}
+class Limit(m1:Int, n1:Int, s:Set) extends base.Limit(m1,n1,s) with Set {
+  override def opt(m: Int, n: Int): Set = {
+    val maxM = Math.max(m, m1)
+    val minN = Math.min(n, n1)
+    if (maxM < minN) new Limit(maxM, minN, s) else new None
+  }
+}
+class Union(s1: Set, s2: Set) extends base.Union(s1,s2) with Set {
+  override def opt(m: Int, n: Int) = new Limit(m, n, this)
+}
+
+object Test extends App {
+  val c = new Beside(new Fan(3), new Identity(3))
+  println(c.width)
+}

### 2016-06-14 paper/paper.tex
Format Scala code

@@
 
+\lstdefinelanguage{JavaScala}{
+  morekeywords={public,int,interface,implements,default,
+    abstract,case,void,catch,class,def,static,%
+    do,else,extends,false,final,finally,%
+    for,if,implicit,import,match,mixin,%
+    new,null,object,override,package,%
+    private,protected,requires,return,sealed,%
+    super,this,throw,trait,true,try,%
+    type,var,val,while,yield,with},
+  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
+  sensitive=true,
+  morecomment=[l]{//},
+  morecomment=[n]{/*}{*/},
+  morestring=[b]",
+  morestring=[b]',
+  morestring=[b]"""
+}
+
 \lstset{ %
-language=Java,                % choose the language of the code
+language=JavaScala,                % choose the language of the code
 columns=fullflexible,

### 2016-06-14 paper/sections/ShallowOO.tex
Format Scala code

@@
 translate it into an OO language like Java:
-\lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
+\lstinputlisting[linerange=2-13]{../src/scala/base/Circuit.scala}%APPLY:CIRCUIT_SCALA
 where the record type maps to the abstract class \texttt{Circuit} and field
@@
 Implementing the desugaring method in Java could be:
-\lstinputlisting[linerange=31-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
+\lstinputlisting[linerange=3-20]{../src/scala/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
 The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is

### 2016-06-14 paper/computePositions.rb

@@ -53,6 +53,10 @@ Dir['../src/*/*/*.java'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end
 
+Dir['../src/*/*/*.scala'].each do |file|
+  scan(file, beginpos, endpos, fileloc)
+end
+
 Dir['./code/*'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end

### 2016-06-16 paper/sections/Introduction.tex
more text

@@
 Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{}, existing
-languages (such as Haskell) have been used to directly encode DSLs. Two common approaches to EDSLs are the
-so-called \emph{shallow} and \emph{deep} embeddings. The difference 
-between these two styles of embeddings is commonly described as 
-follows:
+languages (such as Haskell) have been used to directly encode
+DSLs. Two common approaches to EDSLs are the so-called \emph{shallow}
+and \emph{deep} embeddings. The difference between these two styles of
+embeddings is commonly described as follows:
 
@@
 This definition seems to be widely accepted and similar definitions
-appear in many works~\cite{}. We believe that this definition is a
-vague and not very accurate, often leading to some contradicting
+appear in many works~\cite{}. We believe that this definition is
+vague, and often leads to some contradicting
 claims. For example, in their work on EDSLs~\cite{}, Gibbons and Wu
@@
 
-We believe that the core of problem is that ASTs can be represented 
-in multiple ways. In particular, it is well know that functions alone 
-are enough to encode datastructures such as ASTs (via Church
-encodings~\cite{}). Distinguishing deep and shallow
-embeddings solely on whether a datastructure is being used 
-or whether functions are used to encode interpretations directly 
-falls short when the functions themselfves represent an AST.
+We believe that the core of problem is that ASTs can be represented in
+multiple ways. In particular, it is well know that functions alone are
+enough to encode datastructures such as ASTs (via Church
+encodings~\cite{}). Distinguishing deep and shallow embeddings based
+solely on whether a datastructure is being used or whether
+interpretation functions are used is misleading. Moreover, it gives
+the false impression that shallow embeddings are somewhat less
+expressive than deep embeddings, because they do not have access to
+the AST.
 
-In this paper we propose a different definition for deep and shallow
+In this paper we propose to distinguish EDSLs in terms of the data
+abstraction used to model the language constructs instead.  We follow
+Reynold's classification~\cite{} of data abstractions:
+\emph{procedural abstraction} and \emph{user-defined types}. It is
+clear that shallow embeddings use \emph{procedural
+  abstraction}~\cite{}: the DSLs are modelled by interpretation
+functions. Therefore, the other implementation option for EDSLs is to use
+\emph{user-defined types}. In Reynolds terminology user-defined types
+mean disjoint union types, which are nowadays commonly available in
+modern languages as \emph{algebraic datatypes}. Disjoint union types 
+can also be emulated in OOP using the {\sc Visitor} pattern.
+
+As Cook~\cite{} argued
+procedural abstraction is the essence of Object-Oriented
+Programming. If we accept Cook's view, the implementation of a shallow
+DSL in OOP languages should simply correspond to a standard
+object-oriented program. By identifying shallow EDSLs with procedural
+abstraction it is clear that an EDSL implemented in a functional
+language using an interpretation function is closely related to an
+Object-Oriented Program. The new interpretation of shallow EDSL
+removes the contradiction in Gibbons and Wu's work. Both the Haskell
+implementation of a shallow EDSL and the OOP program using the {\sc
+  Composite} pattern are now under the definition of shallow
+EDSL. Following Reynold's work on data abstraction~\cite{}, the other
+implementation option for EDSLs is to use \emph{user-defined
+  types}. In his original work user-defined types mean disjoint union
+types, which are nowadays commonly available in modern languages as
+\emph{algebraic datatypes}.
+
+By using the kind of data abstraction as a way to distinguish 
+between different implementations strategies for DSLs it is clear 
+that the OOP program and a correponding
+
+Instead of shallow
+EDSLs, we prefer to talk about \emph{procedural} EDSLs. The new
+terminology is important for two reasons: 1) it is more precise; 
+2) procedural EDSLs emcompass a wider range
+
+ not all procedural EDSLs
+
+%However it is this, apparently harmless, translation step is
+%where the distinction between shallow and deep embeddings becomes
+%fuzzy, since it can be argued that the OO program represents an AST.
+
+Instead of deep embeddings, we prefer
+
+propose a different definition for deep and shallow
 embeddings, by distinguishing on what kind of data abstraction is used

### 2016-06-17 paper/paper.tex
more text

@@
 and modularity point of view the Java version has clear advantages.
-To alleviate some of the syntactical disadvantages of Java, we create
-an annotation inspired by \emph{family polymorphism} and a recent
-solution to the Expression Problem.
-The annotation uses transparent code generation techniques to
-automatically eliminate large portions of boilerplate code.
-To further illustrate the applicability of our tool and techniques, we conduct
-several case studies using larger DSLs from the literature.
+%To alleviate some of the syntactical disadvantages of Java, we create
+%an annotation inspired by \emph{family polymorphism} and a recent
+%solution to the Expression Problem.
+%The annotation uses transparent code generation techniques to
+%automatically eliminate large portions of boilerplate code.
+%To further illustrate the applicability of our tool and techniques, we conduct
+%several case studies using larger DSLs from the literature.
 \end{abstract}

### 2016-06-17 paper/sections/Introduction.tex
more text

--- 2016-06-13 paper/sections/Introduction.tex
@@
 Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{}, existing
-languages (such as Haskell) have been used to directly encode DSLs. Two common approaches to EDSLs are the
-so-called \emph{shallow} and \emph{deep} embeddings. The difference 
-between these two styles of embeddings is commonly described as 
-follows:
+languages (such as Haskell) have been used to directly encode
+DSLs. Two common approaches to EDSLs are the so-called \emph{shallow}
+and \emph{deep} embeddings. The origin of that terminology can be
+attributed to Boulton's work~\cite{}. The difference between these
+two styles of embeddings is commonly described as follows:
 
@@
 
-This definition seems to be widely accepted and similar definitions
-appear in many works~\cite{}. We believe that this definition is a
-vague and not very accurate, often leading to some contradicting
-claims. For example, in their work on EDSLs~\cite{}, Gibbons and Wu
-claim that deep embeddings (which encode ASTs using algebraic
-datatypes in Haskell) allow adding new DSL interpretations easily, but
-they make adding new language constructs difficult. In contrast
-Gibbons and Wu claim that shallow embeddings have dual modularity
-properties: new cases are easy to add, but new interpretations are
-hard.  A contradiction happens if instead of using Haskell and
-algebraic datatypes, one uses an OO language to encode an AST, for
-example with the {\sc Composite} pattern. In that case, according to
-definition of deep embedding above, the OO approach classifies as a
-deep embedding. Nevertheless, as the Expression Problem~\cite{} tell
-us, in the OO approach adding new cases is easy, but adding
-interpretations is hard. Thus we have a contradiction, since we have
-an AST representation (i.e. a deep embedding) with the modularity
-properties of shallow embeddings.
+%This definition is widely accepted and similar definitions appear in
+%many other works~\cite{}. 
+We argue that this definition is vague,
+and often leads to some contradicting claims. For example, in their
+work on EDSLs~\cite{}, Gibbons and Wu claim that deep embeddings
+(which encode ASTs using algebraic datatypes in Haskell) allow adding
+new DSL interpretations easily, but they make adding new language
+constructs difficult. In contrast Gibbons and Wu claim that shallow
+embeddings have dual modularity properties: new cases are easy to add,
+but new interpretations are hard.  A contradiction happens if, instead
+of using Haskell and algebraic datatypes, one uses an OO language to
+encode an AST, for example with the {\sc Composite} pattern. In that
+case, \emph{according to definition of deep embedding above, the OO
+  approach classifies as a deep embedding}. Nevertheless, as the
+Expression Problem~\cite{} tell us, in the OO approach adding new
+language constructs is easy, but adding interpretations is hard. Thus
+we have a contradiction, since we have an AST representation (i.e. a
+deep embedding) with the modularity properties of shallow embeddings.
 
-We believe that the core of problem is that ASTs can be represented 
-in multiple ways. In particular, it is well know that functions alone 
-are enough to encode datastructures such as ASTs (via Church
-encodings~\cite{}). Distinguishing deep and shallow
-embeddings solely on whether a datastructure is being used 
-or whether functions are used to encode interpretations directly 
-falls short when the functions themselfves represent an AST.
+We believe that the core of problem is that ASTs can be represented in
+multiple ways. In particular, it is well know that functions alone are
+enough to encode datastructures such as ASTs (via Church
+encodings~\cite{}).  Distinguishing deep and shallow embeddings based
+solely on whether a ``real'' datastructure is being used or not is
+misleading.  Moreover, it gives the impression that shallow embeddings
+are significantly less expressive than deep embeddings, because they
+do not have access to the datastructure.
+Gibbons and Wu themselves feel uneasy with the definition of shallow 
+embeddings when they say:
+``\emph{So it turns out that the syntax of the DSL is not really as ephemeral
+in a shallow embedding as Boulton's choice of terms suggests.}''
 
-In this paper we propose a different definition for deep and shallow
-embeddings, by distinguishing on what kind of data abstraction is used
-by each encoding. Deep embeddings use user-defined types
-(i.e. disjoint unions or algebraic datatypes), whereas shallow
-embeddings use procedural abstraction. This distinction is inspired by
-Reynolds 1975 paper~\cite{} on user-defined types and procedural
-abstraction being complementary approaches to data abstraction, which
-many view~\cite{} as the first documented discussion on the Expression
-Problem.
+A first goal in this paper is to propose distinguishing EDSLs in terms of the data
+abstraction used to model the language constructs instead.  We follow
+Reynold's classification~\cite{} of data abstractions:
+\emph{procedural abstraction} and \emph{user-defined types}. It is
+clear that shallow embeddings use \emph{procedural
+  abstraction}~\cite{}: the DSLs are modelled by interpretation
+functions. Therefore, the other implementation option for EDSLs is to use
+\emph{user-defined types}. In Reynolds terminology user-defined types
+mean disjoint union types, which are nowadays commonly available in
+modern languages as \emph{algebraic datatypes}. Disjoint union types 
+can also be emulated in OOP using the {\sc Visitor} pattern. A
+distinction based on data abstraction provides a remedy for Gibbons 
+and Wu's contradiction. An EDSL implemented with algebraic
+datatypes falls into the category of user-defined types, 
+while a Composite-based OO implementation falls under procedural
+abstraction. For the rest of the paper we identify shallow EDSLs 
+with EDSLs implemented using procedural abstraction.
 
-Shallow Embedded Domain Specific Languages (EDSLs) use
-\emph{procedural abstraction} to directly encode a Domain Specific
-Language into an existing host language. Procedural abstraction has
-been argued to be the essence of Object-Oriented Programming (OOP). Given
+The second and main goal of this paper is to show that OOP languages have
+advantages for the implementation of shallow embeddings.
+As Cook~\cite{} argued procedural abstraction is the essence of
+Object-Oriented Programming. If we accept Cook's view, the
+implementation of a shallow DSL in OOP languages should simply
+correspond to a standard object-oriented program. Given
 that Object-Oriented Languages have evolved over more than 50 years 
@@
 
-This paper argues that OOP abstractions, including \emph{inheritance}
+We show that OOP abstractions, including \emph{inheritance}
 and \emph{subtyping}, increase the modularity and reuse of shallow
@@
 and modularity point of view the Java version has clear advantages.
-To alleviate some of the syntactical disadvantages of Java, we create
-an annotation inspired by \emph{family polymorphism} and a recent
-solution to the Expression Problem. 
-The annotation uses transparent code generation techniques to
-automatically eliminate large portions of boilerplate code.
-To further illustrate the applicability of our tool and techniques, we conduct 
-several case studies using larger DSLs from the literature.
+
+This paper has two goals:
+
+\begin{itemize}
+
+\item To argue that
+
+\end{itemize}

### 2016-06-17 paper/paper.tex
revised abstract

@@
 abstraction is used in Haskell to model a simple EDSL, and we recode
-that EDSL in Java. Although from the \emph{syntactical} point of view
-there are obvious disadvantages in the Java version, from the semantic
-and modularity point of view the Java version has clear advantages.
+that EDSL in Scala. Although from the \emph{syntactical} point of view
+there are minor disadvantages in the Scala version, from the \emph{semantic}
+and modularity point of view the Scala version has clear advantages.
 %To alleviate some of the syntactical disadvantages of Java, we create

### 2016-06-17 paper/sections/Introduction.tex
revised abstract

@@
 
+%\begin{comment}
 %This definition is widely accepted and similar definitions appear in
 %many other works~\cite{}. 
-We argue that this definition is vague,
-and often leads to some contradicting claims. For example, in their
-work on EDSLs~\cite{}, Gibbons and Wu claim that deep embeddings
-(which encode ASTs using algebraic datatypes in Haskell) allow adding
-new DSL interpretations easily, but they make adding new language
-constructs difficult. In contrast Gibbons and Wu claim that shallow
-embeddings have dual modularity properties: new cases are easy to add,
-but new interpretations are hard.  A contradiction happens if, instead
-of using Haskell and algebraic datatypes, one uses an OO language to
-encode an AST, for example with the {\sc Composite} pattern. In that
-case, \emph{according to definition of deep embedding above, the OO
-  approach classifies as a deep embedding}. Nevertheless, as the
-Expression Problem~\cite{} tell us, in the OO approach adding new
-language constructs is easy, but adding interpretations is hard. Thus
-we have a contradiction, since we have an AST representation (i.e. a
-deep embedding) with the modularity properties of shallow embeddings.
+%We argue that this definition is vague,
+%and often leads to some contradicting claims.
+
+Although the above definition is quite reasonable and widely accepted,
+it leaves some space to (mis)interpretation. For example it is unclear 
+how to classify an EDSL using the {\sc Composite} or {\sc Interpreter} 
+patterns in OOP. Would this OO approach be
+classified as a shallow or deep embedding? We believe arguments can be
+made both ways. Since the {\sc Composite}
+pattern is normally accepted to be a way to encode ASTs, it would be
+reasonable to say that \emph{according to definition of deep embedding
+  above, the OO approach classifies as a deep
+  embedding}. However, as we shall argure in the remainder of the
+paper, another possible interpretation is that the OO approach is
+really a shallow embedding. At least some authors~\cite{} seem to agree with 
+the latter interpretation, but we believe that the current definition
+leaves some space open to interpretation.
+
+\begin{comment}
+For example, in their work
+on EDSLs~\cite{}, Gibbons and Wu claim that deep embeddings (which
+encode ASTs using algebraic datatypes in Haskell) allow adding new DSL
+interpretations easily, but they make adding new language constructs
+difficult. In contrast Gibbons and Wu claim that shallow embeddings
+have dual modularity properties: new cases are easy to add, but new
+interpretations are hard.  However what if, instead of using Haskell
+and algebraic datatypes, one uses an OO language to encode an AST, for
+example with the {\sc Composite} pattern.  Would this OO approach be
+classified as a shallow or deep embedding? We believe arguments can be
+made both ways. Since the {\sc Composite}
+pattern is normally accepted to be a way to encode ASTs, it would be
+reasonable to say that \emph{according to definition of deep embedding
+  above, the OO approach classifies as a deep
+  embedding}. Unfortunatelly this interpretation could be problematic.
+As the Expression Problem~\cite{} tell us,
+in the OO approach adding new language constructs is easy, but adding
+interpretations is hard. Thus this would contradict Gibbons and Wu's
+claims, since we have an AST representation (i.e. a deep embedding)
+with the modularity properties of shallow embeddings.
 
@@
 in a shallow embedding as Boulton's choice of terms suggests.}''
+\end{comment}
 
-A first goal in this paper is to propose distinguishing EDSLs in terms of the data
+To avoid confusion and make the terminology more precise, we first 
+propose distinguishing EDSLs in terms of the data
 abstraction used to model the language constructs instead.  We follow
@@
 can also be emulated in OOP using the {\sc Visitor} pattern. A
-distinction based on data abstraction provides a remedy for Gibbons 
-and Wu's contradiction. An EDSL implemented with algebraic
-datatypes falls into the category of user-defined types, 
+distinction based on data abstraction provides a remedy for possible
+misinterpretation. An EDSL implemented with algebraic
+datatypes falls into the category of user-defined types (deep embedding), 
 while a Composite-based OO implementation falls under procedural
-abstraction. For the rest of the paper we identify shallow EDSLs 
+abstraction (shallow embedding). For the rest of the paper we identify shallow EDSLs 
 with EDSLs implemented using procedural abstraction.
 
-The second and main goal of this paper is to show that OOP languages have
+This paper shows that OOP languages have
 advantages for the implementation of shallow embeddings.

### 2016-06-17 paper/paper.tex
abstract

@@
 Shallow Embedded Domain Specific Languages (EDSLs) use
-\emph{procedural abstraction} to directly encode a Domain Specific
-Language into an existing host language. Procedural abstraction has
+\emph{procedural abstraction} to directly encode a DSL into an existing host language. Procedural abstraction has
 been argued to be the essence of Object-Oriented Programming (OOP). Given
-that Object-Oriented Languages have evolved over more than 50 years
+that OO languages have evolved over more than 50 years
 to improve the use of procedural abstraction, they ought to have some
@@
 and \emph{subtyping}, increase the modularity and reuse of shallow
-DSLs when compared to classical procedural abstraction. We make this
+EDSLs when compared to classical procedural abstraction. We make this
 argument by taking a recent paper by Gibbons and Wu, where procedural
 abstraction is used in Haskell to model a simple EDSL, and we recode
-that EDSL in Scala. Although from the \emph{syntactical} point of view
-there are minor disadvantages in the Scala version, from the \emph{semantic}
-and modularity point of view the Scala version has clear advantages.
+that EDSL in Scala. From the \emph{semantic}
+and \emph{modularity} point of view the Scala version has clear advantages 
+over the Haskell version.
+
 %To alleviate some of the syntactical disadvantages of Java, we create

### 2016-06-18 paper/sections/ShallowOO.tex
Separate transformations from S2; Discuss sophisticated optimizations

--- 2016-06-14 paper/sections/ShallowOO.tex
@@
 
-In this section, we argue that \emph{shallow embedding} and straightforward OO
-programming are closely related. To demonstrate it, we use a subset
+In this section, we show that OO approach and procedural abstraction are closely related
+, and hence falls into the category of shallow embedding.
+To demonstrate it, we use a subset
 of the DSL presented in Gibbons and Wu's paper~\cite{gibbons2014folding} as the running example.
@@
 written in Haskell
-and gradually rewrite it approaching to an ``OO style''.
-Then translating the program into an OO version like Java becomes straightforward.
-We also illustrate how to do transformations in shallow embedding, which is typically regarded as the privilege of deep embedding.
+and gradually rewrite it towards an ``OO style''.
+Then translating the program into an OO language becomes straightforward.
+We also illustrate how to do transformations in shallow embedding,
 
@@
 Its BNF grammar is given below:
-$$
-\begin{array}{lcl}
-Circuit & ::= & Identity \\
- & \mid & Fan \\
- & \mid & Beside \\
-Fan & ::= & ``fan''\; PositiveNumber \\
-Identity & ::= & ``identity''\;  PositiveNumber \\
-Beside & ::= & Circuit ``beside'' Circuit \\
-PositiveNumber & ::= & [1-9] \\
-     & \mid & [1-9][0-9]^* \\
-\end{array}
-$$
-% bnf
-\weixin{the bnf grammar needs polishing}
-
 % BNF grammar should not contain left recursion and left factor, rewrite it
 % using descent recursion
+% \setlength{\grammarparsep}{20pt plus 1pt minus 1pt} % increase separation between rules
+\setlength{\grammarindent}{5em} % increase separation between LHS/RHS
 
-There are three constructs - two primitives
-\emph{identity} and \emph{fan} and one combinator \emph{beside}, whose
-meanings are: \emph{identity n} contains \emph{n} isolated vertical wires;
-\emph{fan n} contains \emph{n} vertical wires and the first wire connects to
+\begin{grammar}
+<circuit> ::= `fan' <positive-number>
+\alt `identity' <positive-number>
+\alt `beside' <circuit> <circuit>
+\end{grammar}
+
+The DSL consists of three constructs - two primitives
+\emph{identity} and \emph{fan} and one combinator \emph{beside}.
+Their meanings are: \emph{identity n} contains \emph{n} isolated vertical wires;
+\emph{fan n} has \emph{n} vertical wires with its first wire connected to
 all the remaining wires from top to bottom; $beside\ c_1\ c_2$ joins two circuits
 $c_1$ and $c_2$ horizontally.
-For example, the circuit \emph{beside (fan 3) (identity 3)} can be visualized in Figure~\ref{}.
+With these three constructs, simple circuits can be described.
+For example, Figure~\ref{} visualize the circuit \emph{beside (fan 3) (identity 3)}.
 
 \subsection{Shallow Embedding and Object-Oriented Programming}
-Different from deep embedding, which uses a data structure to represent the
-abstract syntax of a language, shallow embedding defines a language
-directly through encoding its semantics. In the case of \dsl,
-the shallow embedding implementation should conform to the following types:
+Shallow embedding defines a language directly through encoding its semantics
+using procedural abstraction. In the case of \dsl,
+an shallow embedded implementation should conform to the following types:
 
@@
 and should be filled in later with a concrete type according to the semantics.
-For example, if the semantics of \dsl is to calculate the width of a
-circuit, the semantic function would be:
+Suppose that the semantics of \dsl is to calculate the width of a
+circuit. The definitions would be:
 \lstinputlisting[language=haskell,linerange=11-17]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
-As shallow embedding is semantics-oriented, it is already an integer after
-constructing a circuit and hence \texttt{width1} is an identity function.
-For example, we will immediately get $6$ for
-\texttt{beside (fan 3) (identity 3)} and hence \texttt{width1} is just an identity function.
+%As shallow embedding is semantics-oriented,
+\texttt{width1} is an identity function because \texttt{Circuit} is a type
+synonym of \lstinline{Int} and it is already an integer after circuit construction.
+For example, we will immediately get $6$ for \texttt{beside (fan 3) (identity 3)}.
 
-The above implementation can be rewritten by wrapping the result into an
-datatype and get back the value through pattern matching:
+We start the rewrite by wrapping the result into an
+datatype and getting back the value through pattern matching:
 \lstinputlisting[language=haskell,linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
@@
 
-The rewrite goes on by using a record with a sole field to represent the result:
+The rewrite goes on by using a record with a sole field representing the result:
 \lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
@@
 However, the definition already approaches an OO program such that we can easily
-translate it into an OO language like Java:
-\lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
-where the record type maps to the abstract class \texttt{Circuit} and field
+translate it into an OO language like Scala, as shown in Figure~\ref{code:base}.
+
+\begin{figure}
+\lstinputlisting[linerange=2-13]{../src/scala/base/Circuit.scala}%APPLY:CIRCUIT_SCALA
+\caption{OO implementation of \dsl}
+\label{code:base}
+\end{figure}
+where the record type maps to the trait \texttt{Circuit} and field
 declaration becomes method declaration.
 Each case in the semantic function corresponds to an class and its parameters become fields of the class.
-And all the classes extends \texttt{Circuit} by giving \texttt{width} a definition.
+And all the classes extends \texttt{Circuit} by implementing \texttt{width}.
 
 This implementation is exactly how we would model \dsl in an OO language in the first
-place, following the \interp pattern.
+place, following the \interp pattern (which uses \textsc{Composite} pattern to
+organize classes).
 In other words, shallow embedding and straightforward OO programming are closely
@@
 It may worth mentioning that deep embedding is closely related to the \textsc{Visitor} pattern.
-
-\subsection{Transformations in Shallow Embedding}
-One of the main criticism of shallow embedding is that how to define transformations
-becomes unclear since there is not a data structure representing the Abstract Syntax Tree (AST).
-%In fact, transformations can still be defined in shallow embedding.
-
-An instance of this kind of operations could be desugaring, which eliminates some
-language constructs that could be expressed by other language constructs without
-sacrificing the expressiveness of the language.
-In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
-be rewritten via the following formula:
-$$identity\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)\ (n > 0)$$
-which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
-%Using the above formula, we can define a transformation that desugars every
-%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.
-
-Implementing the desugaring method in Java could be:
-\lstinputlisting[linerange=31-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
-The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is
-straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
-their results into an instance of \texttt{Beside}; \emph{identity n} mimics the
-formula.
-
-\weixin{define a show function in the and show the result after desugaring?}
-Inspired by the Java implementation, we can also define \texttt{desugar} in Haskell:
-
-\lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
-The definition relies on the laziness of Haskell.

### 2016-06-18 paper/sections/Transformations.tex
Separate transformations from S2; Discuss sophisticated optimizations

\section{Transformations in Shallow Embedding}
Transformations are typically regarded as the privilege of deep embedding.
Since there is not a data structure persisting the AST, how to define transformations
becomes unclear, which becomes one of the main criticism of shallow embedding.
It is not completely true especially for the OO setting.
This section illustrates how to define various transformations in shallow embedding.
%In fact, transformations can still be defined in shallow embedding.

\subsection{Desugaring}
Desugaring is one kind of transformations, which eliminates some
language constructs that could be expressed by other constructs without
sacrificing the expressiveness of the whole language.
In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
be rewritten via the following formula:
$$identity\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
%Using the above formula, we can define a transformation that desugars every
%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.

Figure~\ref{code:desugar} shows how to implement desugaring in Scala.
\begin{figure}
\lstinputlisting[linerange=3-18]{../src/scala/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
\caption{OO implementation of desugaring}
\label{code:desugar}
\end{figure}
The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is
straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
their results into an instance of \texttt{Beside}; \emph{identity n} mimics the
formula.

%\weixin{define a show function in the and show the result after desugaring?}
Inspired by the Scala implementation, we can also define \texttt{desugar} in Haskell:

\lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
The definition relies on the laziness of Haskell.
However, different from the Scala version, introducing \texttt{desugar} is done
through modifying the original code by inserting a new field in the record.
We will discuss modularity issues in detail in Section~\ref{}.

\subsection{Sophisticated Optimizations}
Sophisticated optimizations may need to inspect multiple internal
representations of values.
For example, we may want to merge consecutive vertical wires together:
$$
beside\ (identity\ m)\ (identity\ n) = identity\ (m + n)
$$
This rewrite rule applies only when both $c_1$ and $c_2$ of the $beside\ c_1\ c_2$
are $identity$, which has to inspect the internal representation of $c_1$ and $c_2$ simultaneously.
However, pattern matching can only be simulated only at the the top level, i.e. $beside$.
There is no way to tell what the class types of $c_1$ and $c_2$ are.
Fortunately, Scala has built-in support for pattern matching. Nevertheless, we
can not simply decorate classes with \lstinline{case} modifier as
it is not possible to have a case class that extends another case class.
Alternatively, we define the \lstinline{unapply} method, a.k.a \emph{extractor}, in the companion object
to tell the compiler how to destruct the object. Here comes the implementation
of the above optimization:
\lstinputlisting[linerange=2-20]{../src/scala/mergeIds/Circuit.scala}%APPLY:MERGEIDS_SCALA

For other OO languages, it is still possible to simulate
pattern matching through test methods or type test and type casts,
although they may not be as elegant as extractors~\cite{emir2007matching}.
Interested reader can refer to \ref{} for implementations using these approaches.

\begin{comment}
\paragraph{Test methods.}
We can introduce some test methods to the hierarchy for determining the class
type of object and extract information from it:
\begin{lstlisting}
  def fromId: Option[Int] = ...
\end{lstlisting}
Then
\begin{lstlisting}
(for {
  i1 <- c1.fromId; i2 <- c2.fromId
} yield new Id(i1+i2))
.orElse(Some(new Beside(t1,t2))).get
\end{lstlisting}

\paragraph{Type test and type cast.}
Each constructs is of different type. We can test its type and convert.
\begin{lstlisting}
(t1,t2) match {
  case (i1: Id,i2: Id) => new Id(i1.n + i2.n)
  case _ => new Beside(t1,t2)
}
\end{lstlisting}

\paragraph{Extractors.}
Scala supports pattern matching.
one can manually implement an extractor, a.k.a \texttt{unapply}
method, in the companion object to tell how to destruct an object.
\begin{lstlisting}
object Id {
  def unapply(c: Identity) = Some(c.n)
}

(t1,t2) match {
  case (Id(n1),Id(n2)) => Id(n1+n2)
  case _ => Beside(t1,t2)
}
\end{lstlisting}

Note that we can not simply decorate classes with \lstinline{case} modifier to
automatically generate the \lstinline{unapply} method -
it is not possible to have a case class that extends another case class.
\end{comment}

### 2016-06-18 paper/code/shallowCircuit.hs

@@ -24,7 +24,7 @@ fan2 n                              = Circuit2 n
 beside2 (Circuit2 c1) (Circuit2 c2) =
   Circuit2 $ c1 + c2
 
-width2              :: Circuit2 -> Int
+width2 :: Circuit2 -> Int
 width2 (Circuit2 n) = n
 --END_CIRCUIT2
 

### 2016-06-18 src/scala/base/Circuit.scala

@@ -1,13 +1,13 @@
 //BEGIN_CIRCUIT_SCALA
 package base
 trait Circuit { def width: Int }
-class Identity(n: Int) extends Circuit {
+class Identity(val n: Int) extends Circuit {
   def width = n
 }
-class Fan(n: Int) extends Circuit {
+class Fan(val n: Int) extends Circuit {
   def width = n
 }
-class Beside(c1: Circuit, c2: Circuit) 
+class Beside(val c1: Circuit, val c2: Circuit) 
     extends Circuit {
   def width = c1.width + c2.width
 }

### 2016-06-18 src/scala/desugar/Circuit.scala

@@ -1,19 +1,17 @@
 package desugar
 //BEGIN_DESUGAR_SCALA
 trait Circuit extends base.Circuit {
-  def desugar: Circuit
+  def desugar: Circuit = this
 }
-class Fan(n: Int) 
-    extends base.Fan(n) with Circuit {
-  def desugar = this
-}
-class Beside(c1: Circuit, c2: Circuit) 
+class Fan(val n: Int) 
+    extends base.Fan(n) with Circuit
+class Beside(val c1: Circuit, val c2: Circuit) 
     extends base.Beside(c1, c2) with Circuit {
-  def desugar = new Beside(c1.desugar, c2.desugar)
+  override def desugar = new Beside(c1.desugar, c2.desugar)
 }
-class Identity(n: Int) 
-    extends base.Identity(n) with Circuit {
-  def desugar = {
+class Id(val n: Int) 
+    extends base.Id(n) with Circuit {
+  override def desugar: Circuit = {
     val fan1: Circuit = new Fan(1)
     (1 until n) map{_=>fan1} reduce{new Beside(_,_)}
   }

### 2016-06-19 paper/sections/Modularity.tex
Draft Section 4

\section{Modular EDSLs in OOP}
This section shows that procedural abstraction mechanisms provided by OO
languages improves the modularity and reusability of EDSLs, in particular
subtyping and inheritance although most OO languages blend them together in one
syntactic form.

\subsection{Subtyping}
Subtyping allows us separate interfaces and implementations, which
encapsulates details of implementations and yields flexible design of EDSLs.
Also, it forms the foundation of the \textsc{Composite} pattern,
which enables us to composite instances of subtypes where a supertype is needed,
following the Liskov substitution principle.

Figure~\ref{code:base} illustrates the use of subtyping,
where all the classes are subtype of the trait \lstinline{Circuit}.

We can easily introduce new language constructs to \dsl by adding a new class to
the hierarchy.
For example, an \lstinline{above} combinator that combines two circuits vertically:
\begin{lstlisting}
class Above(val c1: Circuit, val c2: Circuit) extends Circuit {
  def width = c1.width
}
\end{lstlisting}
The new class works seamlessly with existing classes, as it is also a subtype of
\lstinline{Circuit}. We can construct a circuit by a combination of old
and new constructs:\\
\lstinline{new Beside(new Above(new Fan(2),new Id(2)),new Id(3))}

% type refinements
% record subtyping.

\subsection{Inheritance}
Inheritance is critical for code reuse, which allows
a class to obtain functionalities from other existing classes
without duplicating the code.
Moreover, the inherited functionality can be refined through method overriding
and new functionalities can be introduced via defining new methods and
fields.

We have already shown how to reuse an EDSL through inheritance in Figure~\ref{code:desugar}.
Instead of copying and pasting \lstinline{width} definition case by case
from the old implementation to the new one, we define the new language through inheriting existing
classes, e.g. \lstinline{class Fan extends base.Fan}.
Additionally, we introduce a new method \lstinline{desugar} to the hierarchy.
and gives a default implementation in \lstinline{Circuit} and override it when needed.
Note that we do covariant type-refinements on fields of type \lstinline{Circuit} so that
we can call \lstinline{desugar} method on them.

% Overriding in FP
One may argue that similar code reuse can be simulated in FP through reusing a record instance.
In fact, it is more like composition than inheritance from the OO perspective.
Inheritance plus dynamic dispatch maximize the code reuse for composition.
Consider a simple example:

\begin{lstlisting}[language=haskell]
data T = T { f :: Int, g :: Int }
a = T { f = g a, g = 0 }
b = a { g = 1 }
\end{lstlisting}
Here, we want to reuse \lstinline{a} but ``override'' its \lstinline{g} field when defining \lstinline{b}.
However, the expected result of calling \lstinline{f b} is 1 but it is indeed 0.
The reason is that the \lstinline{f} field ``inherited'' from \lstinline{a} calls
\lstinline{g} on the old instance rather than the new one.

To make it clear, Figure~\ref{code:compare} give the corresponding OO implementation using composition
and inheritance respectively.
\begin{figure}
\begin{tabular}{ll}
\begin{minipage}{.25\textwidth}
\begin{comment}
\begin{lstlisting}
trait T {
  def f: Int
  def g: Int
}
class A extends T {
  def f = g
  def g = 0
}
\end{lstlisting}
\end{comment}
\begin{lstlisting}
object Composition {
  class B(a: A) extends T {
    def f = a.f
    def g = 1
  }
  new B(new A).f // 0
}
\end{lstlisting}
\end{minipage}
&
\begin{minipage}{.25\textwidth}
\begin{lstlisting}
object Inheritance {
  class B extends A {
    override def g = 1
  }
  (new B).f // 1
}
\end{lstlisting}
\end{minipage}
\end{tabular}
\caption{Composition vs inheritance}
\label{code:compare}
\end{figure}
In the case of inheritance, calling \lstinline{f} will execute the overridden \lstinline{g}
as \lstinline{this} points to \lstinline{B} rather than \lstinline{A}.
For compositions, a method can not be reused if when the implementation of what
they depend on changes like \lstinline{f} in \lstinline{A}, even if the logic holds.
OO languages provide these two options for reusing code.

\subsection{Extensibility of EDSLs}
As EDSLs evolve along the time, the need for new syntax and new semantics arises.
This requires host languages equipped with extensibility mechanisms.
OO languages give simple solution to the Expression Problem~\cite{eptrivially16}
and hence are suitable host languages in terms of extensibility.

We have already shown how to extend \dsl on these two dimensions modularly in an
OO language: a new language construct \lstinline{Above} and a new interpretation function \lstinline{desugar}.

Gibbons and Wu introduce new interpretations in a way similar to
Figure~\ref{code:haskell} by appending the definition to a tuple,
which requires revision on the original code.
Though they also present a modular solution based on~\ref{swierstra2008data},
the complexity of encoding increases significantly. Moreover, how to dependent
interpretation modularly is not clear.

\subsection{Discussion}
There are certain kind of methods that are not reusable using the OO approach we
presented, including transformations (or producer methods) and binary
methods as discussed in~\ref{eptrivially16}.
For transformations, we have to refine their return type in extensions and hence
the original code can not be reused, resulting code duplications.
For binary methods like equality, there is no way to refine the argument type.
It is still possible to make it reusable by using virtual types in Scala to capture the type of \lstinline{this}~\ref{zenger}.
However, this may significantly complicate the encoding.

### 2016-06-19 paper/sections/ShallowOO.tex
Draft Section 4

@@
 
-In this section, we show that OO approach and procedural abstraction are closely related
-, and hence falls into the category of shallow embedding.
+In this section, we show that OO approach and procedural abstraction are closely
+related, and hence falls into the category of shallow embedding.
 To demonstrate it, we use a subset
 of the DSL presented in Gibbons and Wu's paper~\cite{gibbons2014folding} as the running example.
-We first give the original shallow embedded implementation of the language
-written in Haskell
+We first give the original shallow embedded implementation in Haskell
 and gradually rewrite it towards an ``OO style''.
 Then translating the program into an OO language becomes straightforward.
-We also illustrate how to do transformations in shallow embedding,
 
@@
 
-\subsection{Shallow Embedding and Object-Oriented Programming}
+\subsection{Shallow Embedding and OOP}
 Shallow embedding defines a language directly through encoding its semantics
@@
 \lstinputlisting[language=haskell,linerange=5-7]{./code/shallowCircuit.hs}%APPLY:TYPES
-The type \texttt{Circuit}, representing the semantic domain, is not defined yet
-and should be filled in later with a concrete type according to the semantics.
+The type \lstinline{Circuit}, representing the semantic domain, is to be filled in with a concrete type according to the semantics.
 Suppose that the semantics of \dsl is to calculate the width of a
@@
 %As shallow embedding is semantics-oriented,
-\texttt{width1} is an identity function because \texttt{Circuit} is a type
+\lstinline{width1} is an identity function because \lstinline{Circuit} is a type
 synonym of \lstinline{Int} and it is already an integer after circuit construction.
-For example, we will immediately get $6$ for \texttt{beside (fan 3) (identity 3)}.
+For example, we will immediately get $6$ for \lstinline{beside (fan 3) (identity 3)}.
 
@@
 \lstinputlisting[language=haskell,linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
-% Then circuit construction gives back a \texttt{Circuit2} which contains an
-% integer value representing the width and \texttt{width2} extracts out the value.
+% Then circuit construction gives back a \lstinline{Circuit2} which contains an
+% integer value representing the width and \lstinline{width2} extracts out the value.
 % Is this definition still shallow embedding?
@@
 \lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
-We no longer need a separate \texttt{width3} function as the field name is the extractor.
+We no longer need a separate \lstinline{width3} function as the field name is the extractor.
 The implementation is still shallow since essentially it is nothing different
@@
 \end{figure}
-where the record type maps to the trait \texttt{Circuit} and field
+where the record type maps to the trait \lstinline{Circuit} and field
 declaration becomes method declaration.
 Each case in the semantic function corresponds to an class and its parameters become fields of the class.
-And all the classes extends \texttt{Circuit} by implementing \texttt{width}.
+And all the classes extends \lstinline{Circuit} by implementing \lstinline{width}.
 
@@
 organize classes).
-In other words, shallow embedding and straightforward OO programming are closely
+In summary, shallow embedding and straightforward OO programming are closely
 related.

### 2016-06-19 paper/sections/Transformations.tex
Draft Section 4

@@
 \end{figure}
-The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is
+The newly introduced method \lstinline{desugar} in \lstinline{Circuit} returns a \lstinline{Circuit} reflecting that it is a transformation. The implementation of \lstinline{desugar} for each case is
 straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
-their results into an instance of \texttt{Beside}; \emph{identity n} mimics the
+their results into an instance of \lstinline{Beside}; \emph{identity n} mimics the
 formula.
@@
 %\weixin{define a show function in the and show the result after desugaring?}
-Inspired by the Scala implementation, we can also define \texttt{desugar} in Haskell:
+Inspired by the Scala implementation, we can also define \lstinline{desugar} in Haskell:
 
+\begin{figure}
 \lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
+\caption{Haskell implementation of desugaring}
+\label{code:haskell}
+\end{figure}
 The definition relies on the laziness of Haskell.
-However, different from the Scala version, introducing \texttt{desugar} is done
+However, different from the Scala version, introducing \lstinline{desugar} is done
 through modifying the original code by inserting a new field in the record.
@@
 Scala supports pattern matching.
-one can manually implement an extractor, a.k.a \texttt{unapply}
+one can manually implement an extractor, a.k.a \lstinline{unapply}
 method, in the companion object to tell how to destruct an object.

### 2016-06-19 src/scala/base/Circuit.scala

@@ -1,7 +1,7 @@
 //BEGIN_CIRCUIT_SCALA
 package base
 trait Circuit { def width: Int }
-class Identity(val n: Int) extends Circuit {
+class Id(val n: Int) extends Circuit {
   def width = n
 }
 class Fan(val n: Int) extends Circuit {

### 2016-06-19 src/scala/mergeIds/Circuit.scala

@@ -1,0 +1,21 @@
+//BEGIN_MERGEIDS_SCALA
+trait Circuit extends base.Circuit {
+  def mergeIds: Circuit = this
+}
+class Id(val n: Int) extends base.Id(n) with Circuit
+object Id {
+  def unapply(c: Id) = Some(c.n)
+}
+class Fan(val n: Int) 
+    extends base.Fan(n) with Circuit
+class Beside(val c1: Circuit, val c2: Circuit)
+    extends base.Beside(c1, c2) with Circuit {
+  override def mergeIds: Circuit = {
+    val (d1, d2) = (c1.mergeIds, c2.mergeIds)
+    (d1, d2) match {
+      case (Id(n1), Id(n2)) => new Id(n1 + n2)
+      case _ => new Beside(d1, d2)
+    }
+  }
+}
+//END_MERGEIDS_SCALA

### 2016-06-19 paper/sections/Modularity.tex
Short related work

@@
 \begin{figure}
-\begin{tabular}{ll}
-\begin{minipage}{.25\textwidth}
-\begin{comment}
+\begin{tabular}{lll}
+\begin{minipage}{.17\textwidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{comment}
+\end{minipage}
+&
+\begin{minipage}{.17\textwidth}
 \begin{lstlisting}
-object Composition {
 class B(a: A) extends T {
@@
 }
-  new B(new A).f // 0
-}
+new B(new A).f// 0
 \end{lstlisting}
@@
 &
-\begin{minipage}{.25\textwidth}
+\begin{minipage}{.17\textwidth}
 \begin{lstlisting}
-object Inheritance {
 class B extends A {
@@
 }
-  (new B).f // 1
-}
+(new B).f// 1
 \end{lstlisting}
@@
 \end{tabular}
-\caption{Composition vs inheritance}
+\caption{Composition (left) vs Inheritance (right)}
 \label{code:compare}

### 2016-06-19 paper/sections/RelatedWorkShort.tex
Short related work

\section{Related Work}

\paragraph{Deep and Shallow Embeddings}
Shallow embedding yields flexible and concise EDSLs while deep embedding makes
it easy to define optimizations.
There are a lot of work~\cite{svenningsson2012combining,
  Jovanovic:2014:YCD:2658761.2658771, scherr2014implicit} trying to blend these two
approaches to enjoy benefits from both by using the shallow embedding as the
surface language and generate deep embedding version to allow optimizations to
be defined.
%Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at the same time so that clients can choose for a particular interpretation;
The OO approach we present retains the simplicity of shallow embedding while
makes it possible to directly implement optimizations.

\paragraph{Modularity of EDSLs}
Languages that solves the EP are capable of being host languages for
implementing modular EDSLs. But many of these solutions require sophisticated
type parameterization, advanced features or heavy encoding which impose
complexity to the DSL implementers or even clients and may be too complicated to use in practice.
For example, solutions based on the \textsc{Visitor}
pattern~\cite{oliveira09modular,hofer2010modular} %more references
Relatively simple solution like Object Algebras~\cite{bruno12oa} may not support
optimizations well.
Our approach, however, is simple, intuitive and supports transformations.

### 2016-06-20 paper/paper.tex
Separate transformations from S2; Discuss sophisticated optimizations
Draft Section 4
Short related work
More text

@@
-\documentclass[10pt,preprint,nocopyrightspace]{sigplanconf}
+\documentclass[10pt,preprint,numbers,nocopyrightspace]{sigplanconf}
 
@@
 \usepackage{stmaryrd}
+\usepackage{syntax}
 
 \newcommand{\name}{\textbf{@Family}\xspace}
-\newcommand{\dsl}{\emph{Scans}\xspace}
+\newcommand{\dsl}{\emph{scans}\xspace}
 \newcommand{\interp}{\textsc{Interpreter}\xspace}
@@
 \emph{procedural abstraction} to directly encode a DSL into an existing host language. Procedural abstraction has
-been argued to be the essence of Object-Oriented Programming (OOP). Given
-that OO languages have evolved over more than 50 years
-to improve the use of procedural abstraction, they ought to have some
-advantages to encode shallow EDSLs.
-
-This paper argues that OOP abstractions, including \emph{inheritance}
-and \emph{subtyping}, increase the modularity and reuse of shallow
+been argued to be the essence of Object-Oriented Programming (OOP). 
+%Given
+%that OO languages have evolved over more than 50 years
+%to improve the use of procedural abstraction, they ought to have some
+%advantages to encode shallow EDSLs.
+This paper argues that OOP abstractions 
+(including \emph{inheritance}, \emph{subtyping} and
+\emph{type-refinement}) 
+increase the modularity and reuse of shallow
 EDSLs when compared to classical procedural abstraction. We make this
@@
 \input{sections/ShallowOO.tex}
-\input{sections/Overview.tex}
-\input{sections/Tool.tex}
-\input{sections/CaseStudy.tex}
-\input{sections/RelatedWork.tex}
-\input{sections/Discussion.tex}
+\input{sections/Transformations.tex}
+\input{sections/Modularity.tex}
+\input{sections/RelatedWorkShort.tex}
+%\input{sections/Overview.tex}
+%\input{sections/Tool.tex}
+%\input{sections/CaseStudy.tex}
+%\input{sections/RelatedWork.tex}
+%\input{sections/Discussion.tex}
 %\acks

### 2016-06-20 paper/sections/Introduction.tex
More text

@@
 it leaves some space to (mis)interpretation. For example it is unclear 
-how to classify an EDSL using the {\sc Composite} or {\sc Interpreter} 
+how to classify an EDSL implemented using the {\sc Composite} or {\sc Interpreter} 
 patterns in OOP. Would this OO approach be
 classified as a shallow or deep embedding? We believe arguments can be
-made both ways. Since the {\sc Composite}
-pattern is normally accepted to be a way to encode ASTs, it would be
+made both ways. Since the {\sc Composite} or {\sc Interpreter}
+patterns are normally accepted to provide a way to encode ASTs, it would be
 reasonable to say that \emph{according to definition of deep embedding
@@
 paper, another possible interpretation is that the OO approach is
-really a shallow embedding. At least some authors~\cite{} seem to agree with 
-the latter interpretation, but we believe that the current definition
-leaves some space open to interpretation.
+really a shallow embedding. At least some authors~\cite{} seem to
+implicitly agree with 
+the latter interpretation.
 
@@
 
-To avoid confusion and make the terminology more precise, we first 
-propose distinguishing EDSLs in terms of the data
+To avoid confusion and make the terminology more precise,
+this paper distinguishes EDSLs in terms of the data
 abstraction used to model the language constructs instead.  We follow
@@
 
-This paper shows that OOP languages have
-advantages for the implementation of shallow embeddings.
-As Cook~\cite{} argued procedural abstraction is the essence of
-Object-Oriented Programming. If we accept Cook's view, the
-implementation of a shallow DSL in OOP languages should simply
-correspond to a standard object-oriented program. Given
-that Object-Oriented Languages have evolved over more than 50 years 
-to improve the use of procedural abstraction, they ought to have some 
-advantages to encode shallow EDSLs.
+As Cook~\cite{} argued
+procedural abstraction is the essence of Object-Oriented
+Programming. If we accept Cook's view, the implementation of a shallow
+DSL in OOP languages should simply correspond to a standard
+object-oriented program. Given that OOP languages have
+evolved over more than 50 years to improve the use of procedural
+abstraction, they ought to have some advantages to encode shallow
+EDSLs.
 
-We show that OOP abstractions, including \emph{inheritance}
-and \emph{subtyping}, increase the modularity and reuse of shallow
-DSLs when compared to classical procedural abstraction. We make this
-argument by taking a recent paper by Gibbons and Wu, where procedural
-abstraction is used in Haskell to model a simple EDSL, and we recode
-that EDSL in Java. Although from the \emph{syntactical} point of view
-there are obvious disadvantages in the Java version, from the semantic
-and modularity point of view the Java version has clear advantages.
+The main goal of this paper is to show that OOP languages have advantages for the
+implementation of shallow embeddings.  
+In order to understand how OOP can help with shallow embeddings, lets
+first review some of the limitations of shallow embeddings commonly found in
+the literature:
 
-This paper has two goals:
+\begin{enumerate}
+
+\item {\bf Single Interpretation} An often stated limitation of
+  shallow embeddings is that they only support a single
+  interpretation.
+
+\item {\bf No Transformations} Another commonly stated limitation 
+of shallow embeddings is that they do not support transformations,
+preventing optimizations and other useful transformations.
+
+\end{enumerate}
+
+\noindent Both limitations are often used as motivators to switch to deep embeddings.
+
+We show that OOP abstractions, including \emph{(multiple)
+  inheritance}, \emph{subtyping} and \emph{type-refinement}, are
+helpful to address those problems. For the first problem, we can
+employ a recently proposed design pattern~\cite{}, which provides a simple
+solution to the \emph{Expression Problem}\cite{} in OOP languages. Thus
+using just standard OOP mechanims enables \emph{multiple modular
+  interpretations} to co-exist and be combined in shallow embeddings.
+For the second problem, we show that transformations can be encoded 
+with recursive objects. Interestingly enough, for transformations it
+is possible to port back the OO solution to Haskell.
+
+We make our arguments by taking a recent paper by Gibbons and Wu,
+where procedural abstraction is used in Haskell to model a simple
+EDSL, and we recode that EDSL in Scala. Although from the
+\emph{syntactical} point of view there are minor inconveniences in
+the Scala version, from the \emph{semantic} and \emph{modularity} point of view the
+Scala version has clear advantages.
+
+In summary, our contributions are:
 
@@
 
-\item To argue that
+\item {\bf Multiple Modular Interpretations for Shallow Embeddings:} 
+  We show that with standard OOP mechanisms it is easy to support multiple modular
+  interpretations for shallow embeddings.
+
+\item {\bf Transformations for Shallow Embeddings:} We show that
+  transformations are encodable with recursive objects. Moreover, this technique
+  can be ported back into functional programming as well.
 

### 2016-06-20 paper/sections/ShallowOO.tex
Revised Section 2

@@
 
-In this section, we show that OO approach and procedural abstraction are closely
-related, and hence falls into the category of shallow embedding.
-To demonstrate it, we use a subset
-of the DSL presented in Gibbons and Wu's paper~\cite{gibbons2014folding} as the running example.
-We first give the original shallow embedded implementation in Haskell
-and gradually rewrite it towards an ``OO style''.
-Then translating the program into an OO language becomes straightforward.
+This section shows that an OO approach and shallow embeddings using
+procedural abstraction are closely related.  We use a subset of the
+DSL presented in Gibbons and Wu's paper~\cite{gibbons2014folding} as
+the running example.  We first give the original shallow embedded
+implementation in Haskell and gradually rewrite it towards an ``OO
+style''.  Then translating the program into an OO language becomes
+straightforward.
 
@@
 With these three constructs, simple circuits can be described.
-For example, Figure~\ref{} visualize the circuit \emph{beside (fan 3) (identity 3)}.
+%For example, Figure~\ref{} visualize the circuit \emph{beside (fan 3) (identity 3)}.
 
-\subsection{Shallow Embedding and OOP}
-Shallow embedding defines a language directly through encoding its semantics
+\subsection{Shallow Embeddings and OOP}
+Shallow embeddings define a language directly through encoding its semantics
 using procedural abstraction. In the case of \dsl,
@@
 %As shallow embedding is semantics-oriented,
+\bruno{why is ``id'' in bold?}
 \lstinline{width1} is an identity function because \lstinline{Circuit} is a type
@@
 
-We start the rewrite by wrapping the result into an
-datatype and getting back the value through pattern matching:
+\begin{comment}
+
+A simple rewriting of the previous program is to wrap the result into an
+datatype, getting back the value through pattern matching:
 \lstinputlisting[language=haskell,linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
@@
 
-The rewrite goes on by using a record with a sole field representing the result:
+\end{comment}
+
+\paragraph{Towards OOP}
+A simple, \emph{semantics preserving}, rewriting of the above program 
+uses a record implemented as a newtype. The record has a sole field representing the result:
 \lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
-We no longer need a separate \lstinline{width3} function as the field name is the extractor.
-The implementation is still shallow since essentially it is nothing different
-from the last one.
-However, the definition already approaches an OO program such that we can easily
-translate it into an OO language like Scala, as shown in Figure~\ref{code:base}.
+\bruno{width3 should now be width2} 
+%We no longer need a separate \lstinline{width3} function as the field name is the extractor.
+The implementation is still shallow since essentially the same program
+as the previous one. In fact Haskell \lstinline{newtype} do not add any operational
+behaviour to the program, and the two programs are effectivelly the
+same.  However, having fields makes the program look more like an 
+OOP program.
+
+\paragraph{Porting to Scala}
+Indeed, we can easily translate the Haskell program into an OO
+language like Scala, as shown in Figure~\ref{code:base}.
+The record type maps to the trait \lstinline{Circuit} and field
+declaration becomes a method declaration.
+Each case in the semantic function corresponds to an class and its parameters become fields of the class.
+And all the classes extends \lstinline{Circuit} by implementing \lstinline{width}.
 
@@
 \end{figure}
-where the record type maps to the trait \lstinline{Circuit} and field
-declaration becomes method declaration.
-Each case in the semantic function corresponds to an class and its parameters become fields of the class.
-And all the classes extends \lstinline{Circuit} by implementing \lstinline{width}.
 
@@
 related.
-It may worth mentioning that deep embedding is closely related to the \textsc{Visitor} pattern.
+%It may worth mentioning that deep embedding is closely related to the \textsc{Visitor} pattern.

### 2016-06-20 paper/sections/ShallowOO.tex
Modifications according to the comment

@@
 
-\lstinputlisting[language=haskell,linerange=2-2]{./code/shallowCircuit.hs}%APPLY:CIRCUIT_TYPE
-\lstinputlisting[language=haskell,linerange=5-7]{./code/shallowCircuit.hs}%APPLY:TYPES
+\lstinputlisting[linerange=2-2]{./code/shallowCircuit.hs}%APPLY:CIRCUIT_TYPE
+\lstinputlisting[linerange=5-7]{./code/shallowCircuit.hs}%APPLY:TYPES
 The type \lstinline{Circuit}, representing the semantic domain, is to be filled in with a concrete type according to the semantics.
@@
 circuit. The definitions would be:
-\lstinputlisting[language=haskell,linerange=11-17]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
+\lstinputlisting[linerange=11-17]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
 %As shallow embedding is semantics-oriented,
@@
 datatype, getting back the value through pattern matching:
-\lstinputlisting[language=haskell,linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
+\lstinputlisting[linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
 % Then circuit construction gives back a \lstinline{Circuit2} which contains an
@@
 uses a record implemented as a newtype. The record has a sole field representing the result:
-\lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
+\lstinputlisting[linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
 \bruno{width3 should now be width2} 

### 2016-06-20 paper/sections/Introduction.tex
Revised Section 2
Latest

@@
   above, the OO approach classifies as a deep
-  embedding}. However, as we shall argure in the remainder of the
+  embedding}. However, as we shall argue in the remainder of the
 paper, another possible interpretation is that the OO approach is
@@
 
+\item {Identifying Advantages of OOP languages for Shallow Embeddings:}
+
 \item {\bf Multiple Modular Interpretations for Shallow Embeddings:} 

### 2016-06-20 paper/sections/Modularity.tex

@@
 \begin{figure}
+\setlength{\tabcolsep}{1pt}
 \begin{tabular}{lll}
-\begin{minipage}{.17\textwidth}
+\begin{minipage}[b]{.15\textwidth}
 \begin{lstlisting}
@@
 }
-class A extends T {
+class A
+    extends T {
   def f = g
@@
 &
-\begin{minipage}{.17\textwidth}
+\begin{minipage}[b]{.15\textwidth}
 \begin{lstlisting}
+// Composition
 class B(a: A) extends T {
@@
 }
-new B(new A).f// 0
+B(A).f // 0
 \end{lstlisting}
@@
 &
-\begin{minipage}{.17\textwidth}
+\begin{minipage}[b]{.2\textwidth}
 \begin{lstlisting}
+// Inheritance
 class B extends A {
@@
 }
-(new B).f// 1
+B.f // 1
 \end{lstlisting}
@@
 \end{tabular}
-\caption{Composition (left) vs Inheritance (right)}
+\caption{Composition vs Inheritance}
 \label{code:compare}

### 2016-06-20 paper/paper.tex
Revised Section 2
Add Haskell keywords

@@
 \newcommand{\weixin}[1]{\authornote{weixin}{cyan}{#1}}
+\newcommand{\bruno}[1]{\authornote{bruno}{red}{#1}}
 % \setlist[itemize]{noitemsep,nolistsep}
@@
 
-\lstdefinelanguage{JavaScala}{
+\lstdefinelanguage{JavaScalaHaskell}{
   morekeywords={public,int,interface,implements,default,
@@
     super,this,throw,trait,true,try,%
-    type,var,val,while,yield,with},
+    type,var,val,while,yield,with,data,type,newtype},
   otherkeywords={=>,<-,<\%,<:,>:,\#,@},
@@
 \lstset{ %
-language=JavaScala,                % choose the language of the code
+language=JavaScalaHaskell,                % choose the language of the code
 columns=fullflexible,

### 2016-06-20 paper/sections/Interpretations.tex
Section 3 in progress

\section{Interpretations in Shallow Embeddings}

\begin{comment}
Weixin writes this one.

Go over jeremy's examples, maybe having only 3 diagram 
constructs instead of 5 for space reasons.

Use five constructs to show extensibility.

Think of how to introduce our tool? Using the Jeremy's examples? 
or introducing before with some other examples?

\end{comment}

A well-known limitation of shallow embeddings is that they allow only a single
interpretation. Gibbons and Wu worked around this problem by accommodating
multiple interpretations using tuples and summarized how to define various types
of interpretations with a shallow embedding.
This section illustrates the OOP way of defining interpretations. Different from
their workaround, our approach is modular extensible.

\subsection{Multiple Interpretations}
%Defining additional interpretations is not trivial for shallow embedding,
%especially for functional languages.
Suppose that we want to have an additional function that checks whether a circuit is
constructed correctly. Here comes the Haskell solution:
\lstinputlisting[linerange=2-9]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
which is not modular because we add the definition of \lstinline{wellSized} by
modifying the original code. And whenever a new interpretation comes,
the arity of the tuple must be incremented and the new interpretation has to be
appended to each case.
Though Gibbons and Wu presented an modular way of defining multiple
interpretations in Haskell based on
\cite{swierstra2008data}. But such solution changes the encoding style
dramatically and may be too complex to use in practice.

Conversely, we can define multiple interpretations in a
modular and intuitive way with an OO language like Scala:
\lstinputlisting[linerange=2-15]{../src/wellsized/Circuit.scala}%APPLY:MULTIPLE_SCALA
Instead of modifying the original code, we define \lstinline{wellSized} modularly.
The encoding makes use of three OOP abstraction mechanisms:
\emph{inheritance}, \emph{subtyping} and \emph{type-refinement}.
Specifically, the new \lstinline{Circuit} is a subtype of
\lstinline{base.Circuit} and declares a new method \lstinline{wellSized}.
The hierarchy implements the new \lstinline{Circuit} by inheriting the corresponding class
from \lstinline{base} and
complementing the body of \lstinline{wellSized}.
Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit}
type to avoid type mismatch in creating objects.

\subsection{Dependent Interpretations}
 \emph{Dependent interpretations} can not be defined alone which use other
 interpretations in their definition.
%Such interpretations are non-compositional.
In Haskell a dependent interpretation must be defined together with what it
dependents on and makes no exceptions on modular approaches like~\cite{}.
This prevents a new interpretation that depends on existing
interpretations from being defined modularly.
Fortunately, OO approach can still modularize such interpretations.

Before giving an example of dependent interpretations, we first show how to add new
constructs. The new constructs are: \emph{stretch ns c} which inserts additional wires into the circuit \emph{c} by
summing up \emph{ns} and $above\ c_1\ c_2$ which combines two circuits of the same width vertically.
Their types are given below:
\lstinputlisting[linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES

Adding new constructs is easy - just defining new classes that implement \lstinline{Circuit}:
\lstinputlisting[linerange=19-30]{../src/wellsized/Circuit.scala}%APPLY:VARIANT
Not arbitrary circuits can be combined using $above$ and $stretch$
as stated in their specifications.
We hence define the \texttt{wellSized} method for these two new constructs to
verify the constraints they imply:

Definitions of \texttt{wellSized} for the extended cases make it a dependent interpretation, as they
rely on another method defined on the \texttt{Circuit}, \texttt{width}, for calculating the width of a circuit.
Most importantly, our approach do not force dependent interpretation to be
defined together with what they depend on. In the case of \texttt{width} and
\texttt{wellSized}, they can be defined in separate families as long as the
signature of \texttt{width} is exposed to the definition of \texttt{wellSized}.

\paragraph{Solving Expression Problem} The code above shows how to
define new variant, combining with

\subsection{Context-sensitive Interpretations}
by declaring contexts as arguments of the method.
For space reasons, we omit the.

### 2016-06-20 paper/sections/ShallowOO.tex
Section 3 in progress

@@
 <circuit> ::= `fan' <positive-number>
-\alt `identity' <positive-number>
+\alt `id' <positive-number>
 \alt `beside' <circuit> <circuit>
@@
 The DSL consists of three constructs - two primitives
-\emph{identity} and \emph{fan} and one combinator \emph{beside}.
-Their meanings are: \emph{identity n} contains \emph{n} isolated vertical wires;
+\emph{id} and \emph{fan} and one combinator \emph{beside}.
+Their meanings are: \emph{id n} contains \emph{n} isolated vertical wires;
 \emph{fan n} has \emph{n} vertical wires with its first wire connected to
@@
 With these three constructs, simple circuits can be described.
-%For example, Figure~\ref{} visualize the circuit \emph{beside (fan 3) (identity 3)}.
+%For example, Figure~\ref{} visualize the circuit \emph{beside (fan 3) (id 3)}.
 
@@
 using procedural abstraction. In the case of \dsl,
-an shallow embedded implementation should conform to the following types:
+an shallow embedded implementation should conform to the following
+types:
 
@@
 circuit. The definitions would be:
-\lstinputlisting[linerange=11-17]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
+\lstinputlisting[linerange=11-14]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
 %As shallow embedding is semantics-oriented,
-\bruno{why is ``id'' in bold?}
-\lstinline{width1} is an identity function because \lstinline{Circuit} is a type
+%\bruno{why is ``id'' in bold?}
+\lstinline{Circuit} is just a type
 synonym of \lstinline{Int} and it is already an integer after circuit construction.
-For example, we will immediately get $6$ for \lstinline{beside (fan 3) (identity 3)}.
+For example, we will immediately get $6$ for \lstinline{beside (fan 3) (id 3)}.
 
@@
 datatype, getting back the value through pattern matching:
-\lstinputlisting[linerange=21-28]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
+\lstinputlisting[linerange=18-25]{./code/shallowCircuit.hs}%APPLY:CIRCUIT2
 % Then circuit construction gives back a \lstinline{Circuit2} which contains an
@@
 uses a record implemented as a newtype. The record has a sole field representing the result:
-\lstinputlisting[linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
-\bruno{width3 should now be width2} 
+\lstinputlisting[linerange=30-35]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
+%\bruno{width3 should now be width2}
 %We no longer need a separate \lstinline{width3} function as the field name is the extractor.
@@
 \begin{figure}
-\lstinputlisting[linerange=2-13]{../src/scala/base/Circuit.scala}%APPLY:CIRCUIT_SCALA
+\lstinputlisting[linerange=2-15]{../src/width/Circuit.scala}%APPLY:CIRCUIT_SCALA
 \caption{OO implementation of \dsl}

### 2016-06-20 paper/computePositions.rb

@@ -45,15 +45,11 @@ fileloc = {}
 beginpos = {}
 endpos = {}
 
-Dir['../src/*/*.java'].each do |file|
+Dir['../src/*/*.{java,scala}'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end
 
-Dir['../src/*/*/*.java'].each do |file|
-  scan(file, beginpos, endpos, fileloc)
-end
-
-Dir['../src/*/*/*.scala'].each do |file|
+Dir['../src/*/*/*.{java,scala}'].each do |file|
   scan(file, beginpos, endpos, fileloc)
 end
 

### 2016-06-20 src/desugar/Circuit.scala

@@ -1,0 +1,24 @@
+package desugar
+
+//BEGIN_DESUGAR_SCALA
+trait Circuit extends width.Circuit {
+  def desugar: Circuit = this
+}
+trait Fan extends width.Fan with Circuit
+object Fan {
+  def apply(x: Int): Circuit = new Fan{val n=x}
+}
+trait Beside extends width.Beside with Circuit {
+  val c1, c2: Circuit
+  override def desugar =
+    Beside(c1.desugar, c2.desugar)
+}
+object Beside {
+  def apply(x: Circuit, y: Circuit): Circuit = 
+    new Beside{val c1=x; val c2=y}
+}
+trait Id extends width.Id with Circuit {
+  override def desugar =
+    List.fill(n)(Fan(1)).reduce(Beside(_,_))
+}
+//END_DESUGAR_SCALA

### 2016-06-20 src/optimizations/Circuit.scala

@@ -1,0 +1,23 @@
+package optimizations
+
+//BEGIN_MERGEIDS_SCALA
+trait Circuit {
+  def mergeIds: Circuit = this
+}
+trait Id extends Circuit { val n: Int }
+object Id {
+  def apply(i: Int) = new Id{val n=i}
+  def unapply(c: Id) = Some(c.n)
+}
+trait Fan extends Circuit { val n: Int }
+trait Beside extends Circuit {
+  val c1, c2: Circuit
+  override def mergeIds: Circuit = {
+    val (d1,d2) = (c1.mergeIds,c2.mergeIds)
+    (d1,d2) match {
+      case (Id(n1),Id(n2)) => Id(n1 + n2)
+      case _ => new Beside{val c1=d1; val c2=d2}
+    }
+  }
+}
+//END_MERGEIDS_SCALA

### 2016-06-20 src/wellsized/Circuit.scala

@@ -1,0 +1,31 @@
+//BEGIN_MULTIPLE_SCALA
+package wellsized
+trait Circuit extends width.Circuit { 
+  def wellSized: Boolean 
+}
+trait Id extends width.Id with Circuit {
+  def wellSized = n > 0
+}
+trait Fan extends width.Fan with Circuit {
+  def wellSized = n > 0
+}
+trait Beside extends width.Beside with Circuit {
+  val c1, c2: Circuit // type refined!
+  def wellSized = c1.wellSized && c2.wellSized
+}
+//END_MULTIPLE_SCALA
+
+//BEGIN_VARIANT
+trait Above extends Circuit {
+  val c1, c2: Circuit
+  def width = c1.width
+  def wellSized = 
+    c1.wellSized && c2.wellSized && c1.width == c2.width
+}
+trait Stretch extends Circuit {
+  val ns: List[Int]
+  val c: Circuit
+  def width = ns.sum
+  def wellSized = c.wellSized && ns.length == c.width
+}
+//END_VARIANT

### 2016-06-20 src/width/Circuit.scala

@@ -1,0 +1,16 @@
+//BEGIN_CIRCUIT_SCALA
+package width
+trait Circuit { def width: Int }
+trait Fan extends Circuit {
+  val n: Int 
+  def width = n
+}
+trait Id extends Circuit {
+  val n: Int
+  def width = n
+}
+trait Beside extends Circuit {
+  val c1, c2: Circuit
+  def width = c1.width + c2.width
+}
+//END_CIRCUIT_SCALA

### 2016-06-21 paper/sections/Interpretations.tex
More S3

@@
 
-Conversely, we can define multiple interpretations in a
+Conversely, we can define introduce new interpretation in a
 modular and intuitive way with an OO language like Scala:
+\begin{figure}
 \lstinputlisting[linerange=2-15]{../src/wellsized/Circuit.scala}%APPLY:MULTIPLE_SCALA
+\caption{Adding new interpretations}
+\label{code:operation}
+\end{figure}
 Instead of modifying the original code, we define \lstinline{wellSized} modularly.
@@
 summing up \emph{ns} and $above\ c_1\ c_2$ which combines two circuits of the same width vertically.
-Their types are given below:
-\lstinputlisting[linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\setlength{\grammarindent}{5em} % increase separation between LHS/RHS
+The extended grammar is given below:
+%\lstinputlisting[linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\begin{grammar}
+<circuit> ::= \ldots
+\alt `above' <circuit> <circuit>
+\alt `stretch' <positive-numbers> <circuit>
+\end{grammar}
 
-Adding new constructs is easy - just defining new classes that implement \lstinline{Circuit}:
-\lstinputlisting[linerange=19-30]{../src/wellsized/Circuit.scala}%APPLY:VARIANT
+Figure~\ref{code:variant} shows how to add new constructs - just defining new classes that implement \lstinline{Circuit}.
+\begin{figure}
+\lstinputlisting[linerange=2-11]{../src/width/Variants.scala}%APPLY:VARIANT_SCALA
+\caption{Adding new language constructs}
+\label{code:variant}
+\end{figure}
+\paragraph{Solving the Expression Problem (EP)} As EDSLs evolve along the time, the
+demand of new syntax and new semantics may arise. It would be good if we could
+implement these features modularly. This, however, requires that the host
+language should be capable of solving the EP. Figure~\ref{code:operation} and
+Figure~\ref{code:variant} together illustrate how elegant the OO solution is.
+In this sense, OO languages are good candidates for defining modular EDSLs.
+
 Not arbitrary circuits can be combined using $above$ and $stretch$
 as stated in their specifications.
-We hence define the \texttt{wellSized} method for these two new constructs to
+We hence define the \lstinline{wellSized} method for these two new constructs to
 verify the constraints they imply:
-
-Definitions of \texttt{wellSized} for the extended cases make it a dependent interpretation, as they
-rely on another method defined on the \texttt{Circuit}, \texttt{width}, for calculating the width of a circuit.
-Most importantly, our approach do not force dependent interpretation to be
-defined together with what they depend on. In the case of \texttt{width} and
-\texttt{wellSized}, they can be defined in separate families as long as the
-signature of \texttt{width} is exposed to the definition of \texttt{wellSized}.
-
-\paragraph{Solving Expression Problem} The code above shows how to
-define new variant, combining with
+\lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
+Definitions of \lstinline{wellSized} for the extended cases make it a dependent interpretation, as they
+rely \lstinline{width} for calculating the width of a circuit.
+What distinguishes the OO approach is that \lstinline{width} and \lstinline{wellSized}
+can be defined separately. Essentially, with only the signature of
+\lstinline{width} exposed to the \lstinline{Circuit} we can
+define \lstinline{wellSized}.
 
 \subsection{Context-sensitive Interpretations}
-by declaring contexts as arguments of the method.
-For space reasons, we omit the.
+Unlike \lstinline{width} and \lstinline{wellSized} which can be defined with
+only the given circuit, interpretations may need some mutable contexts for definition.
+These contexts can be captured by arguments of the method. For example, a
+function that collects all the connections between wires inside a circuit would have
+the following signature:
+\begin{lstlisting}
+type Layout = List[List[Tuple2[Int,Int]]]
+def tlayout(f: Int => Int): Layout
+\end{lstlisting}
+where the context \lstinline{f} may vary in recursive calls.
+For space reasons, we omit the implementation details.

### 2016-06-21 paper/sections/Transformations.tex
Modifications according to the comment
More S3

@@
-\section{Transformations in Shallow Embedding}
+\section{Transformations in Shallow Embeddings}
 Transformations are typically regarded as the privilege of deep embedding.
@@
 \begin{figure}
-\lstinputlisting[linerange=3-18]{../src/scala/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
+\lstinputlisting[linerange=4-25]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
 \caption{OO implementation of desugaring}
@@
 %\weixin{define a show function in the and show the result after desugaring?}
-Inspired by the Scala implementation, we can also define \lstinline{desugar} in Haskell:
-
-\begin{figure}
-\lstinputlisting[language=haskell,linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
-\caption{Haskell implementation of desugaring}
-\label{code:haskell}
-\end{figure}
-The definition relies on the laziness of Haskell.
-However, different from the Scala version, introducing \lstinline{desugar} is done
-through modifying the original code by inserting a new field in the record.
-We will discuss modularity issues in detail in Section~\ref{}.
+Inspired by the Scala implementation, we can port it back to Haskell:
+\lstinputlisting[linerange=6-21]{./code/desugar.hs}%APPLY:DESUGAR
+which relies on the laziness of Haskell.
 
@@
 representations of values.
-For example, we may want to merge consecutive vertical wires together:
+Suppose we want to merge consecutive vertical wires together:
 $$
@@
 of the above optimization:
-\lstinputlisting[linerange=2-20]{../src/scala/mergeIds/Circuit.scala}%APPLY:MERGEIDS_SCALA
+\lstinputlisting[linerange=4-22]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
 

### 2016-06-21 src/desugar/Circuit.scala

@@ -3,6 +3,8 @@ package desugar
 //BEGIN_DESUGAR_SCALA
 trait Circuit extends width.Circuit {
   def desugar: Circuit = this
+  type Layout = List[List[Tuple2[Int,Int]]]
+  def tlayout(f: Int => Int): Layout
 }
 trait Fan extends width.Fan with Circuit
 object Fan {

### 2016-06-21 src/desugar/Circuit.scala
Clean up code

@@ -3,8 +3,6 @@ package desugar
 //BEGIN_DESUGAR_SCALA
 trait Circuit extends width.Circuit {
   def desugar: Circuit = this
-  type Layout = List[List[Tuple2[Int,Int]]]
-  def tlayout(f: Int => Int): Layout
 }
 trait Fan extends width.Fan with Circuit
 object Fan {

### 2016-06-21 paper/sections/Interpretations.tex
Shorten S4

@@
 \section{Interpretations in Shallow Embeddings}
-
-\begin{comment}
-Weixin writes this one.
-
-Go over jeremy's examples, maybe having only 3 diagram 
-constructs instead of 5 for space reasons.
-
-Use five constructs to show extensibility.
-
-Think of how to introduce our tool? Using the Jeremy's examples? 
-or introducing before with some other examples?
-
-\end{comment}
 
@@
 
-Conversely, we can define introduce new interpretation in a
+Conversely, we can introduce new interpretation in a
 modular and intuitive way with an OO language like Scala:

### 2016-06-21 paper/sections/Transformations.tex
Clean up code
Shorten S4

@@
 \section{Transformations in Shallow Embeddings}
-Transformations are typically regarded as the privilege of deep embedding.
+Transformations are typically regarded as the privilege of deep embeddings.
 Since there is not a data structure persisting the AST, how to define transformations
-becomes unclear, which becomes one of the main criticism of shallow embedding.
-It is not completely true especially for the OO setting.
-This section illustrates how to define various transformations in shallow embedding.
+becomes unclear.
+It is not completely true, especially for the OO approach.
+This section illustrates how to define various transformations, the missing
+kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
 %In fact, transformations can still be defined in shallow embedding.
@@
 Desugaring is one kind of transformations, which eliminates some
-language constructs that could be expressed by other constructs without
-sacrificing the expressiveness of the whole language.
-In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
+language constructs without sacrificing the expressiveness of the whole language.
+In \dsl, the \emph{id} construct is such a syntactic sugar that can
 be rewritten via the following formula:
-$$identity\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
-which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
+$$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
+which states that \emph{id n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
 %Using the above formula, we can define a transformation that desugars every
@@
 
-Figure~\ref{code:desugar} shows how to implement desugaring in Scala.
+Figure~\ref{code:desugar} shows how to eliminate \lstinline{Id} construct in Scala.
 \begin{figure}
-\lstinputlisting[linerange=4-25]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
+\lstinputlisting[linerange=4-23]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
 \caption{OO implementation of desugaring}
@@
 \end{figure}
-The newly introduced method \lstinline{desugar} in \lstinline{Circuit} returns a \lstinline{Circuit} reflecting that it is a transformation. The implementation of \lstinline{desugar} for each case is
-straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
-their results into an instance of \lstinline{Beside}; \emph{identity n} mimics the
+The newly introduced method \lstinline{desugar} in \lstinline{Circuit} returns a
+\lstinline{Circuit} reflecting the nature of a transformation. The implementation of \lstinline{desugar} for each case is
+straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
+their results into an instance of \lstinline{Beside}; \emph{id n} just mimics the
 formula.
@@
 representations of values.
-Suppose we want to merge consecutive vertical wires together:
+Suppose we want to merge consecutive vertical wires:
 $$
-beside\ (identity\ m)\ (identity\ n) = identity\ (m + n)
+beside\ (id\ m)\ (id\ n) = id\ (m + n)
 $$
 This rewrite rule applies only when both $c_1$ and $c_2$ of the $beside\ c_1\ c_2$
-are $identity$, which has to inspect the internal representation of $c_1$ and $c_2$ simultaneously.
+are $id$, which has to inspect the internal representation of $c_1$ and $c_2$ simultaneously.
 However, pattern matching can only be simulated only at the the top level, i.e. $beside$.
-There is no way to tell what the class types of $c_1$ and $c_2$ are.
-Fortunately, Scala has built-in support for pattern matching. Nevertheless, we
-can not simply decorate classes with \lstinline{case} modifier as
-it is not possible to have a case class that extends another case class.
-Alternatively, we define the \lstinline{unapply} method, a.k.a \emph{extractor}, in the companion object
-to tell the compiler how to destruct the object. Here comes the implementation
+There is no easy way to destruct $c_1$ and $c_2$.
+Fortunately, Scala has built-in support for pattern matching. For extensibility reasons, we
+can not simply decorate classes with \lstinline{case} modifier.
+% as it is not possible to have a case class that extends another case class.
+Alternatively, we manually define the \lstinline{unapply} method, a.k.a \emph{extractor}, in the companion object. Here comes the implementation
 of the above optimization:

### 2016-06-21 src/optimizations/Circuit.scala

@@ -1,16 +1,16 @@
 package optimizations
 
 //BEGIN_MERGEIDS_SCALA
-trait Circuit {
+trait Circuit extends width.Circuit {
   def mergeIds: Circuit = this
 }
-trait Id extends Circuit { val n: Int }
+trait Id extends width.Id with Circuit
 object Id {
   def apply(i: Int) = new Id{val n=i}
   def unapply(c: Id) = Some(c.n)
 }
-trait Fan extends Circuit { val n: Int }
-trait Beside extends Circuit {
+trait Fan extends width.Fan with Circuit
+trait Beside extends width.Beside with Circuit {
   val c1, c2: Circuit
   override def mergeIds: Circuit = {
     val (d1,d2) = (c1.mergeIds,c2.mergeIds)

### 2016-06-21 paper/sections/Interpretations.tex
Minor changes

@@
 Before giving an example of dependent interpretations, we first show how to add new
-constructs. The new constructs are: \emph{stretch ns c} which inserts additional wires into the circuit \emph{c} by
-summing up \emph{ns} and $above\ c_1\ c_2$ which combines two circuits of the same width vertically.
+constructs. The extended grammar of \dsl is given below,
+which contains two extra constructs \emph{above} and \emph{stretch}:
 \setlength{\grammarindent}{5em} % increase separation between LHS/RHS
-The extended grammar is given below:
 %\lstinputlisting[linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
@@
 \end{grammar}
-
-Figure~\ref{code:variant} shows how to add new constructs - just defining new classes that implement \lstinline{Circuit}.
+$above\ c_1\ c_2$ combines two circuits of the same width vertically;
+\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
+summing up \emph{ns}.
+Figure~\ref{code:variant} shows how to add new constructs - simply defining new classes that implement \lstinline{Circuit}.
 \begin{figure}
@@
 \end{figure}
-\paragraph{Solving the Expression Problem (EP)} As EDSLs evolve along the time, the
-demand of new syntax and new semantics may arise. It would be good if we could
-implement these features modularly. This, however, requires that the host
-language should be capable of solving the EP. Figure~\ref{code:operation} and
-Figure~\ref{code:variant} together illustrate how elegant the OO solution is.
-In this sense, OO languages are good candidates for defining modular EDSLs.
 
@@
 
+\paragraph{Solving the Expression Problem (EP)} As EDSLs evolve along the time, the
+demand for new syntax and new semantics may arise. It would be good if
+these features can be introduced modularly. Then the capability of solving the EP
+becomes a requirement of the host language.
+Figure~\ref{code:operation} and Figure~\ref{code:variant} together illustrate
+how elegant the OO solution,
+indicating that OO languages are suitable host languages for defining modular EDSLs.
+
 \subsection{Context-sensitive Interpretations}

### 2016-06-22 paper/sections/Interpretations.tex
Mention multiple inheritance in S3

@@
 interpretation. Gibbons and Wu worked around this problem by accommodating
-multiple interpretations using tuples and summarized how to define various types
-of interpretations with a shallow embedding.
-This section illustrates the OOP way of defining interpretations. Different from
-their workaround, our approach is modular extensible.
+multiple interpretations using tuples. However, their encoding needs to modify
+the original code. Although a modular solution based on \cite{swierstra2008data}
+is also presented, it complicates the encoding dramatically and may prevent pratical use.
+OO approach, on the contrary, provides modular yet simple solution of defining
+multiple interpretations. This section illustrates how various type of
+interpretations can be defined in an OOP way.
 
@@
 Suppose that we want to have an additional function that checks whether a circuit is
-constructed correctly. Here comes the Haskell solution:
+constructed correctly. Here comes Gibbons and Wu's solution:
 \lstinputlisting[linerange=2-9]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
-which is not modular because we add the definition of \lstinline{wellSized} by
-modifying the original code. And whenever a new interpretation comes,
+which is not modular because
+whenever a new interpretation is needed (e.g. \lstinline{wellSized}), the
+original code has to be revised -
 the arity of the tuple must be incremented and the new interpretation has to be
 appended to each case.
-Though Gibbons and Wu presented an modular way of defining multiple
-interpretations in Haskell based on
-\cite{swierstra2008data}. But such solution changes the encoding style
-dramatically and may be too complex to use in practice.
+%we add the definition of \lstinline{wellSized} by modifying the original code.
 
-Conversely, we can introduce new interpretation in a
-modular and intuitive way with an OO language like Scala:
+Conversely, we can introduce new interpretations in a
+modular and intuitive way with an OO language like Scala, as shown in Figure~\ref{code:operation}.
 \begin{figure}
@@
 \end{figure}
-Instead of modifying the original code, we define \lstinline{wellSized} modularly.
-The encoding makes use of three OOP abstraction mechanisms:
+%Instead of modifying the original code, we define \lstinline{wellSized} modularly.
+The encoding relies on three OOP abstraction mechanisms:
 \emph{inheritance}, \emph{subtyping} and \emph{type-refinement}.
@@
 complementing the body of \lstinline{wellSized}.
-Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit}
-type to avoid type mismatch in creating objects.
+Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit} type
+to avoid type mismatch in creating objects.
+
+% Multiple inheritance
+We can even define \lstinline{wellSized} separately:
+\begin{lstlisting}
+trait Circuit { def wellSized: Boolean }
+trait Id extends Circuit { ... }
+...
+\end{lstlisting}
+And merge the two hierarchy through \emph{multiple inheritance} for providing
+multiple interpretations:
+\begin{lstlisting}
+trait Circuit
+    extends width.Circuit with wellSized.Circuit
+trait Id extends Circuit
+    with width.Id with wellSized.Circuit
+...
+\end{lstlisting}
 
 \subsection{Dependent Interpretations}
- \emph{Dependent interpretations} can not be defined alone which use other
- interpretations in their definition.
+ \emph{Dependent interpretations} use other interpretations in their definition.
 %Such interpretations are non-compositional.
@@
 interpretations from being defined modularly.
-Fortunately, OO approach can still modularize such interpretations.
+Fortunately, OO approach does not have such restriction.
 
 Before giving an example of dependent interpretations, we first show how to add new
-constructs. The extended grammar of \dsl is given below,
+constructs in OO approach. The extended grammar of \dsl is given below,
 which contains two extra constructs \emph{above} and \emph{stretch}:
@@
 summing up \emph{ns}.
-Figure~\ref{code:variant} shows how to add new constructs - simply defining new classes that implement \lstinline{Circuit}.
+Figure~\ref{code:variant} shows how to add them to \dsl - simply defining new classes that implement \lstinline{Circuit}.
 \begin{figure}
@@
 
-Not arbitrary circuits can be combined using $above$ and $stretch$
-as stated in their specifications.
-We hence define the \lstinline{wellSized} method for these two new constructs to
+From the specifications we can see that not arbitrary circuits can be combined using $above$ and $stretch$.
+We hence define the \lstinline{wellSized} method for them to
 verify the constraints they imply:
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
-Definitions of \lstinline{wellSized} for the extended cases make it a dependent interpretation, as they
-rely \lstinline{width} for calculating the width of a circuit.
-What distinguishes the OO approach is that \lstinline{width} and \lstinline{wellSized}
-can be defined separately. Essentially, with only the signature of
-\lstinline{width} exposed to the \lstinline{Circuit} we can
-define \lstinline{wellSized}.
+Definitions of \lstinline{wellSized} for the extended cases make it a dependent
+interpretation, as they can not be defined without \lstinline{width}.
+Note that \lstinline{width} and \lstinline{wellSized} are defined separately.
+Essentially, it is sufficient to define \lstinline{wellSized} with only the signature of \lstinline{width} in \lstinline{Circuit}.
 

### 2016-06-22 paper/sections/Introduction.tex

@@
 
-Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{}, existing
+Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{hudak1998modular}, existing
 languages (such as Haskell) have been used to directly encode
@@
 and \emph{deep} embeddings. The origin of that terminology can be
-attributed to Boulton's work~\cite{}. The difference between these
+attributed to Boulton et al.'s work~\cite{Boulton92dsl}. The difference between these
 two styles of embeddings is commonly described as follows:
@@
 shallow embedding, terms in the DSL are implemented directly by
-their semantics, bypassing the intermediate AST and its traversal.}\cite{gibbons15folding}
+their semantics, bypassing the intermediate AST and its traversal.}\cite{gibbons2014folding}
 \end{quote}
@@
   above, the OO approach classifies as a deep
-  embedding}. However, as we shall argue in the remainder of the
-paper, another possible interpretation is that the OO approach is
-really a shallow embedding. At least some authors~\cite{} seem to
-implicitly agree with 
-the latter interpretation.
+  embedding}.
 
@@
 
-To avoid confusion and make the terminology more precise,
-this paper distinguishes EDSLs in terms of the data
-abstraction used to model the language constructs instead.  We follow
-Reynold's classification~\cite{} of data abstractions:
-\emph{procedural abstraction} and \emph{user-defined types}. It is
-clear that shallow embeddings use \emph{procedural
+In this paper, however, we prefer to interpret shallow embeddings
+differently. We argue that a straightforward OO approach is really a
+shallow embedding. The reason is that shallow embeddings are typically
+implemented using procedural abstraction and, as Cook~\cite{} argued,
+procedural abstraction is the essence of Object-Oriented
+Programming. Thus the implementation of a shallow
+DSL in OOP languages should simply correspond to a standard
+object-oriented program.
+
+If we accept this view, 
+a natural way to distinguish implementations of EDSLs is 
+in terms of the data abstraction used to model the language
+constructs instead. We can follow Reynold's~\cite{} classification of
+data abstractions: \emph{procedural abstraction} and \emph{user-defined
+  types}. It is clear that shallow embeddings use \emph{procedural
   abstraction}~\cite{}: the DSLs are modelled by interpretation
-functions. Therefore, the other implementation option for EDSLs is to use
-\emph{user-defined types}. In Reynolds terminology user-defined types
-mean disjoint union types, which are nowadays commonly available in
-modern languages as \emph{algebraic datatypes}. Disjoint union types 
-can also be emulated in OOP using the {\sc Visitor} pattern. A
-distinction based on data abstraction provides a remedy for possible
-misinterpretation. An EDSL implemented with algebraic
-datatypes falls into the category of user-defined types (deep embedding), 
-while a Composite-based OO implementation falls under procedural
-abstraction (shallow embedding). For the rest of the paper we identify shallow EDSLs 
-with EDSLs implemented using procedural abstraction.
+functions. Thus, the other implementation option for EDSLs is to
+use \emph{user-defined types}. In Reynolds terminology user-defined
+types mean disjoint union types, which are nowadays commonly available
+in modern languages as \emph{algebraic datatypes}. Disjoint union
+types can also be emulated in OOP using the {\sc Visitor} pattern.
+
+A distinction based on data abstraction is more precise and provides a
+remedy for possible misinterpretation. An EDSL implemented with
+algebraic datatypes falls into the category of user-defined types
+(deep embedding), while a Composite-based OO implementation falls
+under procedural abstraction (shallow embedding).
+
+%At least some authors~\cite{} seem to implicitly
+%agree with the latter interpretation.
 

### 2016-06-22 paper/sections/Overview.tex

@@
 Suppose that we want to have a function that checks whether a circuit is constructed correctly. Here comes the solution from Gibbons and Wu:
-\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
+\lstinputlisting[language=haskell,linerange=2-9]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
 Tuples are used to aggregate multiple interpretations in one definition.
@@
 Their types are given below:
-\lstinputlisting[language=haskell,linerange=42-43]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\lstinputlisting[language=haskell,linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 
@@
 
-\lstinputlisting[language=haskell,linerange=42-43]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\lstinputlisting[language=haskell,linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 \emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
@@
 \texttt{stretch} and \texttt{above}, for the semantic function:
-\lstinputlisting[language=haskell,linerange=47-48]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
+\lstinputlisting[language=haskell,linerange=44-45]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
 Similarly, defining two new interfaces
@@
 Gibbons and Wu worked around this problem in the following way:
-\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
+\lstinputlisting[language=haskell,linerange=2-9]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
 which is similar to how we add \texttt{desugar} to the implementation in Section~\ref{}.

### 2016-06-22 paper/sections/ShallowOO.tex

@@
 Indeed, we can easily translate the Haskell program into an OO
-language like Scala, as shown in Figure~\ref{code:base}.
+language like Scala:
+\lstinputlisting[linerange=2-15]{../src/width/Circuit.scala}%APPLY:CIRCUIT_SCALA
 The record type maps to the trait \lstinline{Circuit} and field
@@
 And all the classes extends \lstinline{Circuit} by implementing \lstinline{width}.
-
-\begin{figure}
-\lstinputlisting[linerange=2-15]{../src/width/Circuit.scala}%APPLY:CIRCUIT_SCALA
-\caption{OO implementation of \dsl}
-\label{code:base}
-\end{figure}
 

### 2016-06-22 paper/sections/Transformations.tex

@@
 
-Figure~\ref{code:desugar} shows how to eliminate \lstinline{Id} construct in Scala.
-\begin{figure}
+The following shows how to eliminate \lstinline{Id} construct in Scala:
 \lstinputlisting[linerange=4-23]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
-\caption{OO implementation of desugaring}
-\label{code:desugar}
-\end{figure}
-The newly introduced method \lstinline{desugar} in \lstinline{Circuit} returns a
-\lstinline{Circuit} reflecting the nature of a transformation. The implementation of \lstinline{desugar} for each case is
-straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
-their results into an instance of \lstinline{Beside}; \emph{id n} just mimics the
-formula.
+The newly introduced method \lstinline{desugar} returns a
+\lstinline{Circuit} reflecting the nature of transformations. The implementation of \lstinline{desugar} for each case is
+straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
+$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the formula.
 
 %\weixin{define a show function in the and show the result after desugaring?}
-Inspired by the Scala implementation, we can port it back to Haskell:
+Inspired by the Scala implementation, we can port it back to Haskell thanks to
+the laziness of Haskell:
 \lstinputlisting[linerange=6-21]{./code/desugar.hs}%APPLY:DESUGAR
-which relies on the laziness of Haskell.
 
@@
 $$
-This rewrite rule applies only when both $c_1$ and $c_2$ of the $beside\ c_1\ c_2$
-are $id$, which has to inspect the internal representation of $c_1$ and $c_2$ simultaneously.
-However, pattern matching can only be simulated only at the the top level, i.e. $beside$.
-There is no easy way to destruct $c_1$ and $c_2$.
-Fortunately, Scala has built-in support for pattern matching. For extensibility reasons, we
-can not simply decorate classes with \lstinline{case} modifier.
-% as it is not possible to have a case class that extends another case class.
-Alternatively, we manually define the \lstinline{unapply} method, a.k.a \emph{extractor}, in the companion object. Here comes the implementation
-of the above optimization:
+This rewrite rule applies only when both $c_1$ and $c_2$ of $beside\ c_1\ c_2$
+are $id$ and also accesses their inner value.
+This kind of optimizations are hard to define in shallow embeddings since it
+requires deep pattern matching on the AST.
+Fortunately, Scala has built-in support for pattern matching on objects.
+We can not simply decorate classes with \lstinline{case} modifier
+as case class to case class extension is not supported.
+Alternatively, we could manually define an \lstinline{unapply} method, a.k.a
+\emph{extractor}, in the companion object:
+%Here comes the implementation of the optimization:
 \lstinputlisting[linerange=4-22]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
 
-For other OO languages, it is still possible to simulate
-pattern matching through test methods or type test and type casts,
+For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type test and type casts},
 although they may not be as elegant as extractors~\cite{emir2007matching}.

### 2016-06-22 paper/sections/Introduction.tex
Intro

@@
 classified as a shallow or deep embedding? We believe arguments can be
-made both ways. Since the {\sc Composite} or {\sc Interpreter}
-patterns are normally accepted to provide a way to encode ASTs, it would be
-reasonable to say that \emph{according to definition of deep embedding
+made both ways. The {\sc Composite} or {\sc Interpreter}
+patterns are normally accepted to provide a way to encode ASTs. Thus, 
+one possible interpretation is that \emph{according to definition of deep embedding
   above, the OO approach classifies as a deep
@@
 
-In this paper, however, we prefer to interpret shallow embeddings
+In this paper, however, we prefer to interpret shallow and deep embeddings
 differently. We argue that a straightforward OO approach is really a
-shallow embedding. The reason is that shallow embeddings are typically
-implemented using procedural abstraction and, as Cook~\cite{} argued,
+shallow embedding. The reason is that shallow embeddings are 
+implemented using \emph{procedural abstraction} and, as Cook~\cite{cook09abstraction} argued,
 procedural abstraction is the essence of Object-Oriented
@@
 
-If we accept this view, 
+If we accept Cook's view on OOP, 
 a natural way to distinguish implementations of EDSLs is 
 in terms of the data abstraction used to model the language
-constructs instead. We can follow Reynold's~\cite{} classification of
-data abstractions: \emph{procedural abstraction} and \emph{user-defined
+constructs instead. As Reynold's~\cite{reynolds94proceduralabstraction} suggested there are two
+types of data abstraction: procedural abstraction and \emph{user-defined
   types}. It is clear that shallow embeddings use \emph{procedural
-  abstraction}~\cite{}: the DSLs are modelled by interpretation
+  abstraction}: the DSLs are modelled by interpretation
 functions. Thus, the other implementation option for EDSLs is to
@@
 algebraic datatypes falls into the category of user-defined types
-(deep embedding), while a Composite-based OO implementation falls
+(deep embedding), while a {\sc Composite}-based OO implementation falls
 under procedural abstraction (shallow embedding).
@@
 
-As Cook~\cite{} argued
-procedural abstraction is the essence of Object-Oriented
-Programming. If we accept Cook's view, the implementation of a shallow
-DSL in OOP languages should simply correspond to a standard
-object-oriented program. Given that OOP languages have
-evolved over more than 50 years to improve the use of procedural
-abstraction, they ought to have some advantages to encode shallow
-EDSLs.
+The main goal of this paper is to argue that OOP languages have
+advantages for the implementation of shallow embeddings. This should
+not came as a surprise, as OOP languages have evolved over more than 50
+years to improve the use of procedural abstraction.
 
-The main goal of this paper is to show that OOP languages have advantages for the
-implementation of shallow embeddings.  
 In order to understand how OOP can help with shallow embeddings, lets
-first review some of the limitations of shallow embeddings commonly found in
-the literature:
+first review some of the limitations of shallow embeddings commonly
+found in the literature:
 
@@
 helpful to address those problems. For the first problem, we can
-employ a recently proposed design pattern~\cite{}, which provides a simple
-solution to the \emph{Expression Problem}\cite{} in OOP languages. Thus
+employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
+solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
 using just standard OOP mechanims enables \emph{multiple modular
@@
 with recursive objects. Interestingly enough, for transformations it
-is possible to port back the OO solution to Haskell.
+is possible to port back the OO solution to
+Haskell.
+ \bruno{double-check if the paper really shows this!}
 
-We make our arguments by taking a recent paper by Gibbons and Wu,
+We make our arguments by taking a recent paper by Gibbons and Wu~\cite{gibbons2014folding},
 where procedural abstraction is used in Haskell to model a simple
@@
 
+\begin{comment}
 In summary, our contributions are:
@@
 
-\item {Identifying Advantages of OOP languages for Shallow Embeddings:}
+\item {}
 
@@
 \end{itemize}
+\end{comment}

### 2016-06-22 paper/sections/Interpretations.tex
revised part of Section 3

@@
 multiple interpretations using tuples. However, their encoding needs to modify
-the original code. Although a modular solution based on \cite{swierstra2008data}
+the original code. This section illustrates how various type of
+interpretations can be defined in OOP.
+\begin{comment}
+Although a modular solution based on \cite{swierstra2008data}
 is also presented, it complicates the encoding dramatically and may prevent pratical use.
 OO approach, on the contrary, provides modular yet simple solution of defining
-multiple interpretations. This section illustrates how various type of
-interpretations can be defined in an OOP way.
+multiple interpretations. 
+\end{comment}
 
@@
 %especially for functional languages.
+\paragraph{Multiple Interpretations in Haskell}
 Suppose that we want to have an additional function that checks whether a circuit is
-constructed correctly. Here comes Gibbons and Wu's solution:
+constructed correctly. Gibbons and Wu's solution is:
 \lstinputlisting[linerange=2-9]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
-which is not modular because
-whenever a new interpretation is needed (e.g. \lstinline{wellSized}), the
+\noindent This solution is not modular because it relies 
+on defining the two interpretations (\lstinline{width} and
+\lstinline{wellSized}) simultaneously, using a pair. It is not
+possible to define \lstinline{width} independently and then reuse it.
+Whenever a new interpretation is needed (e.g. \lstinline{wellSized}), the
 original code has to be revised -
@@
 
-Conversely, we can introduce new interpretations in a
+\paragraph{Multiple Interpretations in Scala}
+In contrast, new interpretations are introduced in a 
 modular and intuitive way with an OO language like Scala, as shown in Figure~\ref{code:operation}.
@@
 Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit} type
-to avoid type mismatch in creating objects.
+to avoid type mismatch in creating objects~\cite{eptrivially16}.
 
 % Multiple inheritance
-We can even define \lstinline{wellSized} separately:
+We can even define \lstinline{wellSized} independently:
 \begin{lstlisting}
@@
 \end{lstlisting}
-And merge the two hierarchy through \emph{multiple inheritance} for providing
+And merge the two hierarchies through \emph{multiple inheritance} for providing
 multiple interpretations:

### 2016-06-22 paper/sections/Interpretations.tex
Section 3

@@
 
-\subsection{Multiple Interpretations}
+\subsection{Multiple Interpretations}\label{subsec:multiple}
 %Defining additional interpretations is not trivial for shallow embedding,
@@
 Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit} type
-to avoid type mismatch in creating objects~\cite{eptrivially16}.
+to avoid type mismatches in methods~\cite{eptrivially16}.
 
@@
 \subsection{Dependent Interpretations}
- \emph{Dependent interpretations} use other interpretations in their definition.
+ \emph{Dependent interpretations} are a generalization of multiple
+interpretations, where one or more interpretations can use other
+interpretations in their definition. Typically such interpretations 
+would be said to be \emph{non-compositional} since they do not 
+dependent only on themselves. 
+In Haskell dependent interpretations are once again defined with
+tuples in a non-modular way. We omit the Haskell code here, since 
+it is quite similar to the example in Section~\ref{subsec:multiple}
+
 %Such interpretations are non-compositional.
-In Haskell a dependent interpretation must be defined together with what it
-dependents on and makes no exceptions on modular approaches like~\cite{}.
-This prevents a new interpretation that depends on existing
-interpretations from being defined modularly.
-Fortunately, OO approach does not have such restriction.
+%In Haskell a dependent interpretation must be defined together with what it
+%dependents on and makes no exceptions on modular approaches like~\cite{}.
+%This prevents a new interpretation that depends on existing
+%interpretations from being defined modularly.
+%Fortunately, OO approach does not have such restriction.
 
-Before giving an example of dependent interpretations, we first show how to add new
-constructs in OO approach. The extended grammar of \dsl is given below,
-which contains two extra constructs \emph{above} and \emph{stretch}:
+To illustrate dependent interpretations, we first add two new language constructs
+constructs \emph{above} and \emph{stretch}:
 \setlength{\grammarindent}{5em} % increase separation between LHS/RHS
@@
 summing up \emph{ns}.
-Figure~\ref{code:variant} shows how to add them to \dsl - simply defining new classes that implement \lstinline{Circuit}.
+Figure~\ref{code:variant} shows how to add them to \dsl, simply by
+\emph{modularly} defining new classes that implement \lstinline{Circuit}.
 \begin{figure}
@@
 
-From the specifications we can see that not arbitrary circuits can be combined using $above$ and $stretch$.
-We hence define the \lstinline{wellSized} method for them to
-verify the constraints they imply:
+\paragraph{Dependent Interpretations in Scala}
+The definitions of \lstinline{wellSized} for $above$ and $stretch$
+require dependent interpretations.
+Circuits written with $above$ and $stretch$ have to obey to certain
+size constrains, which are captured by the \lstinline{wellSized} method:
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
 Definitions of \lstinline{wellSized} for the extended cases make it a dependent
-interpretation, as they can not be defined without \lstinline{width}.
+interpretation, as they need \lstinline{width} in their definition.
 Note that \lstinline{width} and \lstinline{wellSized} are defined separately.
-Essentially, it is sufficient to define \lstinline{wellSized} with only the signature of \lstinline{width} in \lstinline{Circuit}.
+Essentially, it is sufficient to define \lstinline{wellSized} while
+knowing only the signature of \lstinline{width} in \lstinline{Circuit}.
 
-\paragraph{Solving the Expression Problem (EP)} As EDSLs evolve along the time, the
-demand for new syntax and new semantics may arise. It would be good if
-these features can be introduced modularly. Then the capability of solving the EP
-becomes a requirement of the host language.
-Figure~\ref{code:operation} and Figure~\ref{code:variant} together illustrate
-how elegant the OO solution,
-indicating that OO languages are suitable host languages for defining modular EDSLs.
+\subsection{Discussion} 
+Gibbons and Wu~\cite{gibbons2014folding} claim that in shallow
+embeddings new language constructs are easy to add, but new
+interpretations are hard. As our OOP approach shows, in OOP both
+language constructs and new interpretations are easy to add in shallow
+embeddings. In other words, the circuit DSL presented so far does not
+suffer from the Expression Problem. The key point is that procedural
+abstraction combined with OOP features (subtyping, inheritance and
+type-refinement) adds expressiveness over traditional procedural
+abstraction.
 
-\subsection{Context-sensitive Interpretations}
+Gibbons and Wu, also show some different variants of interpretations, 
+such as context-sensitive interpretations. 
 Unlike \lstinline{width} and \lstinline{wellSized} which can be defined with
-only the given circuit, interpretations may need some mutable contexts for definition.
+only the given circuit, context-dependent interpretations may need some context.
 These contexts can be captured by arguments of the method. For example, a
@@
 \end{lstlisting}
-where the context \lstinline{f} may vary in recursive calls.
-For space reasons, we omit the implementation details.
+where the context \lstinline{f} may vary in recursive
+calls. Context-sensitive transformations do not pose any particular
+challenge. For space reasons, we omit the implementation details here. Full details
+are available online.

### 2016-06-22 paper/sections/Introduction.tex
shorter intro

@@
 
-In this paper, however, we prefer to interpret shallow and deep embeddings
+In this paper, however, we prefer to interpret shallow embeddings
 differently. We argue that a straightforward OO approach is really a
@@
 
+\begin{comment}
 If we accept Cook's view on OOP, 
@@
 under procedural abstraction (shallow embedding). 
+\end{comment}
 
@@
 In order to understand how OOP can help with shallow embeddings, lets
-first review some of the limitations of shallow embeddings commonly
+first review some of the limitations commonly
 found in the literature:
@@
 Haskell.
- \bruno{double-check if the paper really shows this!}
 

### 2016-06-22 src/circuit/CircuitDSL.java
Clean up code

@@ -1,298 +1,0 @@
-package circuit;
-import static java.util.Collections.emptyList;
-import static java.util.Collections.singletonList;
-import static java.util.stream.Collectors.toList;
-
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-import java.util.function.BinaryOperator;
-import java.util.function.IntUnaryOperator;
-import java.util.stream.IntStream;
-import java.util.stream.Stream;
-
-import javax.swing.JFrame;
-import javax.swing.SwingUtilities;
-
-import lombok.Obj;
-
-@Obj
-interface AST {
-    interface Circuit {
-    }
-    interface Identity extends Circuit {
-        int _n();
-    }
-    interface Fan extends Circuit {
-        int _n();
-    }
-    interface Above extends Circuit {
-        Circuit _c1();
-        Circuit _c2();
-    }
-    interface Beside extends Circuit {
-        Circuit _c1();
-        Circuit _c2();
-    }
-    interface Stretch extends Circuit {
-        List<Integer> _ns();
-        Circuit _c();
-    }
-}
-
-@Obj
-interface Depth extends AST {
-    interface Circuit {
-        int depth();
-    }
-    interface Identity{
-        default int depth() {
-            return 0;
-        }
-    }
-    interface Fan {
-        default int depth() {
-            return 1;
-        }
-    }
-    interface Above {
-        default int depth() {
-            return _c1().depth() + _c2().depth();
-        }
-    }
-    interface Beside {
-        default int depth() {
-            return Math.max(_c1().depth(), _c2().depth());
-        }
-    }
-    interface Stretch {
-        default int depth() {
-            return _c().depth();
-        }
-    }
-}
-
-@Obj
-interface Width extends AST {
-    interface Circuit {
-        int width();
-    }
-    interface Identity {
-        default int width() {
-            return _n();
-        }
-    }
-    interface Fan {
-        default int width() {
-            return _n();
-        }
-    }
-    interface Above {
-        default int width() {
-            return _c1().width();
-        }
-    }
-    interface Beside {
-        default int width() {
-            return _c1().width() + _c2().width();
-        }
-    }
-    interface Stretch {
-        default int width() {
-            return _ns().stream().reduce(0, Integer::sum);
-        }
-    }
-}
-
-
-@Obj
-interface WellSized extends Width, Depth {
[…]

### 2016-06-22 src/circuit/IntPair.java

@@ -1,18 +1,0 @@
-package circuit;
-
-import java.util.function.IntUnaryOperator;
-
-import lombok.Obj;
-
-@Obj
-public interface IntPair {
-    int _x();
-    int _y();
-
-    default String show() {
-        return "(" + _x() + "," + _y() + ")";
-    }
-    default IntPair map(IntUnaryOperator f) {
-        return IntPair.of(f.applyAsInt(_x()), f.applyAsInt(_y()));
-    }
-}

### 2016-06-22 src/circuit/Test.java

@@ -1,20 +1,0 @@
-package circuit;
-
-import static circuit.CircuitDSL.fan;
-import static circuit.CircuitDSL.id;
-
-import circuit.CircuitDSL.Circuit;
-
-public class Test {
-    public static void main(String[] args) {
-        Circuit circuit =
-            fan(2).beside(fan(2))
-            .above(fan(2).stretch(2,2))
-            .above(id(1).beside(fan(2)).beside(id(1)));
-//        circuit.width();
-//        circuit.depth();
-//        circuit.layout();
-//        circuit.tlayout(f)
-        circuit.draw();
-    }
-}

### 2016-06-22 src/depth/Circuit.scala

@@ -1,0 +1,22 @@
+package depth
+trait Circuit extends width.Circuit { 
+  def depth: Int 
+}
+trait Id extends width.Id with Circuit {
+  def depth = 0
+}
+trait Fan extends width.Fan with Circuit {
+  def depth = 1
+}
+trait Above extends width.Above with Circuit {
+  val c1, c2: Circuit
+  def depth = c1.depth + c2.depth
+}
+trait Beside extends width.Beside with Circuit {
+  val c1, c2: Circuit
+  def depth = Math.max(c1.depth, c2.depth)
+}
+trait Stretch extends width.Stretch with Circuit {
+  val c: Circuit
+  def depth = c.depth
+}

### 2016-06-22 src/diagrams/Diagrams.java

@@ -1,559 +1,0 @@
-package diagrams;
-
-import static diagrams.ListUtilities.concat;
-import static java.util.Arrays.asList;
-import static java.util.Collections.emptyList;
-import static java.util.Collections.singletonList;
-import static java.util.stream.Collectors.joining;
-
-import java.io.IOException;
-import java.nio.file.Files;
-import java.nio.file.Path;
-import java.nio.file.Paths;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-import java.util.stream.Collectors;
-import java.util.stream.IntStream;
-import java.util.stream.Stream;
-
-import diagrams.Debug.Above;
-import diagrams.Debug.Beside;
-import diagrams.Debug.Bisque;
-import diagrams.Debug.Black;
-import diagrams.Debug.Blue;
-import diagrams.Debug.Col;
-import diagrams.Debug.Ellipse;
-import diagrams.Debug.FillColor;
-import diagrams.Debug.Green;
-import diagrams.Debug.Picture;
-import diagrams.Debug.Place;
-import diagrams.Debug.Rectangle;
-import diagrams.Debug.Red;
-import diagrams.Debug.Shape;
-import diagrams.Debug.StrokeColor;
-import diagrams.Debug.StrokeWidth;
-import diagrams.Debug.StyleSheet;
-import diagrams.Debug.Styling;
-import diagrams.Debug.Triangle;
-import lombok.Obj;
-
-
-interface Show {
-    String show();
-    static String showList(List<? extends Show> xs) {
-        return xs.stream().map(x -> x.show()).collect(Collectors.joining(",", "[", "]"));
-    }
-}
-
-@Obj
-interface Attr extends Show {
-    String _name();
-    String _value();
-    default String show() {
-        return _name() + "=\"" + _value() + "\"";
-    }
-}
-
-@Obj
-interface Extent extends Show {
-    Pos _l();
-    Pos _r();
-    default Extent union(Extent e) {
-        return Extent.of(Pos.of(Math.min(_l()._x(), e._l()._x()), Math.min(_l()._y(), e._l()._y())),
-                Pos.of(Math.max(_r()._x(), e._r()._x()), Math.max(_r()._y(), e._r()._y())));
-    }
-    default String show() {
-        return "(" + _l().show() + "," + _r().show() + ")";
-    }
-}
-
-@Obj
-interface XML extends Show {
-    String _tag();
-    List<Attr> _attrs();
[…]

### 2016-06-22 src/layout/Circuit.scala

@@ -1,0 +1,30 @@
+package layout
+
+trait Circuit extends wellsized.Circuit with depth.Circuit {
+  def layout: Layout
+  def tlayout(f: Int => Int): Layout
+}
+trait Id extends Circuit with wellsized.Id with depth.Id {
+  def layout = List()
+  def tlayout(f: Int => Int) = List()
+}
+trait Fan extends Circuit with wellsized.Fan with depth.Fan {
+  def layout = List(for (i <- List.range(1,n)) yield (0,i))
+  def tlayout(f: Int => Int) = List(for (i <- List.range(1,n)) yield (f(0),f(i)))
+}
+trait Above extends Circuit with wellsized.Above with depth.Above {
+  val c1, c2: Circuit
+  def layout = c1.layout ++ c2.layout
+  def tlayout(f: Int => Int) = c1.tlayout(f) ++ c2.tlayout(f)
+}
+trait Beside extends Circuit with wellsized.Beside with depth.Beside {
+  val c1, c2: Circuit
+  def layout = lzw (c1.layout, shift(c1.width)(c2.layout)) (_ ++ _) 
+  def tlayout(f: Int => Int) = 
+    lzw (c1.tlayout(f), c2.tlayout(f.andThen(c1.width + _))) (_ ++ _)
+}
+trait Stretch extends Circuit with wellsized.Stretch with depth.Stretch {
+  val c: Circuit
+  def layout = c.layout.map(_.map(connect(ns)))
+  def tlayout(f: Int => Int) = c.tlayout(f.andThen(partialSum(ns)(_) - 1))
+}

### 2016-06-22 src/layout/package.scala

@@ -1,0 +1,16 @@
+package object layout {
+  type IntPair = Tuple2[Int,Int]
+  type Layout = List[List[IntPair]] 
+  def lzw[A](xs: List[A], ys: List[A])(f: (A, A) => A): List[A] = (xs, ys) match {
+    case (Nil,_) => ys
+    case (_,Nil) => xs
+    case (x::xs,y::ys) => f(x,y)::lzw(xs,ys)(f)
+  }
+  def pmap[A,B](f: A => B)(p: Tuple2[A,A]): Tuple2[B,B] = p match {
+    case (x, y) => (f(x), f (y))
+  }
+  def shift(w: Int)(layout: Layout) = layout.map(_.map(pmap(w+_)))
+  def connect(ns: List[Int])(p: IntPair) =
+    pmap ((i: Int) => partialSum(ns)(i) - 1) (p)
+  def partialSum(ns: List[Int]): List[Int] = ns.scanLeft(0)(_ + _).tail
+}

### 2016-06-22 src/optimizations/Circuit.scala

@@ -12,12 +12,13 @@ object Id {
 trait Fan extends width.Fan with Circuit
 trait Beside extends width.Beside with Circuit {
   val c1, c2: Circuit
-  override def mergeIds: Circuit = {
-    val (d1,d2) = (c1.mergeIds,c2.mergeIds)
-    (d1,d2) match {
+  override def mergeIds = 
+    (c1.mergeIds,c2.mergeIds) match {
       case (Id(n1),Id(n2)) => Id(n1 + n2)
-      case _ => new Beside{val c1=d1; val c2=d2}
-    }
+      case _ => Beside(c1.mergeIds, c2.mergeIds)
   }
 }
 //END_MERGEIDS_SCALA
+object Beside {
+  def apply(x: Circuit, y: Circuit): Circuit = new Beside{val c1=x; val c2=y}
+}

### 2016-06-22 src/paper/sec2/Circuit.java

@@ -1,63 +1,0 @@
-package paper.sec2;
-
-import static java.util.stream.IntStream.range;
-
-class A {
-//BEGIN_CIRCUIT_OO
-abstract class Circuit {
-  abstract int width();
-}
-class Fan extends Circuit {
-  int _n;
-  Fan(int n) { _n = n; }
-  int width() { return _n; }
-}
-class Beside extends Circuit {
-  Circuit _c1, _c2;
-  Beside(Circuit c1, Circuit c2) {
-    _c1 = c1; c2 = _c2;
-  }
-  int width() { return _c1.width() + _c2.width(); }
-}
-class Identity extends Circuit {
-  int _n;
-  Identity(int n) { _n = n; }
-  int width() { return _n; }
-}
-//END_CIRCUIT_OO
-}
-class Sec2 {
-//BEGIN_DESUGAR_OO
-abstract class Circuit {
-  abstract int width();
-  abstract Circuit desugar();
-}
-class Fan extends Circuit {
-  int _n;
-  Fan(int n) { _n = n; }
-  int width() { return _n; }
-  Circuit desugar() { return this; }
-}
-class Beside extends Circuit {
-  Circuit _c1, _c2;
-  Beside(Circuit c1, Circuit c2) {
-    _c1 = c1; c2 = _c2;
-  }
-  int width() { return _c1.width() + _c2.width(); }
-  Circuit desugar() {
-    return new Beside(_c1.desugar(), _c2.desugar());
-  }
-}
-class Identity extends Circuit {
-  int _n;
-  Identity(int n) { _n = n; }
-  int width() { return _n; }
-  Circuit desugar() {
-    Circuit fan1 = new Fan(1);
-    return range(1, _n)
-      .mapToObj(i -> fan1)
-      .reduce(fan1, (x,y) -> new Beside(x,y));
-  }
-}
-//END_DESUGAR_OO
-}

### 2016-06-22 src/paper/sec3/Test.java

@@ -1,25 +1,0 @@
-package paper.sec3;
-
-import java.util.List;
-
-interface A {
-    void f();
-}
-interface B extends A {
-    void g();
-}
-
-interface As {
-  List<? extends A> as();
-  default void f() {
-      as().get(0).f();
-  }
-}
-
-interface Bs extends As {
-  @Override List<? extends B> as();
-
-  default void g() {
-      as().get(0).g();
-  }
-}

### 2016-06-22 src/scans/DrawCircuit.scala

@@ -1,0 +1,42 @@
+package scans
+
+import javax.swing.JPanel
+import java.awt.Graphics2D
+import java.awt.Graphics
+import java.awt.Dimension
+import layout._
+
+class DrawCircuit(layout: Layout) extends JPanel {
+  val L = 30;
+  val W = 30;
+  val R = 6;
+  def toCoord(c: Int, r: Int) = ((c+1)*W, (r+1)*L)
+
+  def drawDot(g: Graphics2D, p: IntPair) = 
+    g.fillOval(p._1 - R/2, p._2 - R/2, R, R)
+
+  def drawLine(g: Graphics2D, from: IntPair, to: IntPair) = 
+    g.drawLine(from._1, from._2, to._1, to._2);
+
+  override def paintComponent(g: Graphics) {
+    super.paintComponent(g);
+    val g2d = g.asInstanceOf[Graphics2D];
+    val col = layout.flatMap { x => x.map { pr => pr._2 } }.max
+
+    // draw vertical lines
+    for (c <- 0 to col) 
+      drawLine(g2d, toCoord(c, 0), toCoord(c, layout.size))
+    for {
+        r <- 0 to (layout.size-1)
+        (x,y) <- layout(r)
+    } {
+      val from = toCoord(x, r)
+      val to = toCoord(y, r+1)
+      drawDot(g2d, from)
+      drawDot(g2d, to);
+      drawLine(g2d, from, to);
+    }
+  }
+
+  override def getPreferredSize = new Dimension(300, 300)
+}

### 2016-06-22 src/scans/Test.scala

@@ -1,0 +1,12 @@
+package scans
+
+object Test extends App {
+  val c = above(beside(fan(2), fan(2)),
+          above(stretch(List(2,2), fan(2)),
+          beside(id(1), beside(fan(2), id(1)))))
+  println(c.width)
+  println(c.wellSized)
+  println(c.depth)
+  println(c.layout)
+  println(draw(c.tlayout(i=>i)))
+}

### 2016-06-22 src/scans/package.scala

@@ -1,0 +1,23 @@
+import javax.swing.SwingUtilities
+import javax.swing.JFrame
+
+package object scans {
+  import layout._  
+  def id(x: Int) = new Id{val n=x}
+  def fan(x: Int) = new Fan{val n=x}
+  def above(x: Circuit, y: Circuit) = new Above{val c1=x; val c2=y}
+  def beside(x: Circuit, y: Circuit) = new Beside{val c1=x; val c2=y}
+  def stretch(xs: List[Int], x: Circuit) = new Stretch{val ns=xs; val c=x}
+
+  def draw(layout: Layout) = SwingUtilities.invokeLater(new Runnable() {
+    override def run {
+      val frame = new JFrame
+      frame.setTitle("Draw Circuit")
+      frame.setResizable(false)
+      frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
+      frame.getContentPane().add(new DrawCircuit(layout));
+      frame.pack();
+      frame.setVisible(true);
+    }
+  })
+}

### 2016-06-24 paper/sections/Transformations.tex
Clean up code
Revise S4

@@
 \section{Transformations in Shallow Embeddings}
-Transformations are typically regarded as the privilege of deep embeddings.
-Since there is not a data structure persisting the AST, how to define transformations
-becomes unclear.
-It is not completely true, especially for the OO approach.
-This section illustrates how to define various transformations, the missing
-kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
-%In fact, transformations can still be defined in shallow embedding.
+Transformations are typically regarded as the privilege of deep embeddings as no
+persistent AST obscures how to define transformations in shallow embeddings.
+The \textsc{Composite} pattern applied in our OO approach encodes the AST, making
+it possible to implement transformations. More interestingly, we can port back
+some simple transformations to Haskell.
+This sections illustrates how to define transformations, the missing kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
 
 \subsection{Desugaring}
-Desugaring is one kind of transformations, which eliminates some
-language constructs without sacrificing the expressiveness of the whole language.
-In \dsl, the \emph{id} construct is such a syntactic sugar that can
-be rewritten via the following formula:
+Desugaring is an transformation, which eliminates some
+language constructs without sacrificing the expressiveness of the language.
+\emph{id} construct is such a syntactic sugar:
 $$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
-which states that \emph{id n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
-%Using the above formula, we can define a transformation that desugars every
-%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.
+The formula states that \emph{id n} can be rewritten as \emph{n} of \emph{fan 1}
+combined with \emph{beside}.
 
-The following shows how to eliminate \lstinline{Id} construct in Scala:
-\lstinputlisting[linerange=4-23]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
-The newly introduced method \lstinline{desugar} returns a
-\lstinline{Circuit} reflecting the nature of transformations. The implementation of \lstinline{desugar} for each case is
+A corresponding Scala implementation is given below:
+\lstinputlisting[linerange=4-17]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
+The method \lstinline{desugar} returns a
+\lstinline{Circuit} reflecting the nature of transformations.
+Implementing it for each case is
 straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
 $beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the formula.
+For constructing circuits easily, we define some companion objects with an \lstinline{apply} method implemented:
+\lstinputlisting[linerange=20-26]{../src/desugar/Circuit.scala}%APPLY:COMPANION
 
 %\weixin{define a show function in the and show the result after desugaring?}
-Inspired by the Scala implementation, we can port it back to Haskell thanks to
-the laziness of Haskell:
+We can port the implementation back to Haskell thanks to its laziness:
 \lstinputlisting[linerange=6-21]{./code/desugar.hs}%APPLY:DESUGAR
@@
 Sophisticated optimizations may need to inspect multiple internal
-representations of values.
-Suppose we want to merge consecutive vertical wires:
+representations of values, which imposes difficulty to shallow embeddings.
+For instance, consecutive \emph{id}s put side by side can be merged together:
 $$
@@
 $$
-This rewrite rule applies only when both $c_1$ and $c_2$ of $beside\ c_1\ c_2$
-are $id$ and also accesses their inner value.
-This kind of optimizations are hard to define in shallow embeddings since it
-requires deep pattern matching on the AST.
-Fortunately, Scala has built-in support for pattern matching on objects.
-We can not simply decorate classes with \lstinline{case} modifier
-as case class to case class extension is not supported.
-Alternatively, we could manually define an \lstinline{unapply} method, a.k.a
-\emph{extractor}, in the companion object:
-%Here comes the implementation of the optimization:
-\lstinputlisting[linerange=4-22]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
+This rule applies to a $beside\ c_1\ c_2$ only when both $c_1$ and $c_2$ are
+$id$, also their inner values are extracted for constructing the term
+on the right hand side of the equation.
+Fortunately, Scala has built-in support for pattern matching by decorating
+classes with \lstinline{case} modifiers or defining extractor objects.
+For extensibility reasons, case classes can not be used
+because case class to case class extension is not supported.
+Instead, we manually define \lstinline{unapply} methods in companion
+objects:
+\lstinputlisting[linerange=19-22]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
+Then we can define this optimization:
+\lstinputlisting[linerange=4-16]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
 
 For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type test and type casts},
-although they may not be as elegant as extractors~\cite{emir2007matching}.
-Interested reader can refer to \ref{} for implementations using these approaches.
+though they may not be as elegant as extractors~\cite{emir2007matching}.
+%Interested reader can refer to \ref{} for implementations using these approaches.
 

### 2016-06-24 src/desugar/Circuit.scala

@@ -2,23 +2,26 @@ package desugar
 
 //BEGIN_DESUGAR_SCALA
 trait Circuit extends width.Circuit {
-  def desugar: Circuit = this
+  def desugar: Circuit
 }
-trait Fan extends width.Fan with Circuit
-object Fan {
-  def apply(x: Int): Circuit = new Fan{val n=x}
+trait Fan extends width.Fan with Circuit {
+  def desugar = this
 }
 trait Beside extends width.Beside with Circuit {
   val c1, c2: Circuit
-  override def desugar =
-    Beside(c1.desugar, c2.desugar)
-}
-object Beside {
-  def apply(x: Circuit, y: Circuit): Circuit = 
-    new Beside{val c1=x; val c2=y}
+  def desugar = Beside(c1.desugar, c2.desugar)
 }
 trait Id extends width.Id with Circuit {
-  override def desugar =
+  def desugar = 
     List.fill(n)(Fan(1)).reduce(Beside(_,_))
 }
 //END_DESUGAR_SCALA
+//BEGIN_COMPANION
+object Fan {
+  def apply(x: Int): Circuit = new Fan{val n=x}
+}
+object Beside {
+  def apply(x: Circuit, y: Circuit): Circuit = 
+    new Beside{ val c1 = x; val c2 = y }
+}
+//END_COMPANION

### 2016-06-24 src/optimizations/Circuit.scala

@@ -5,10 +5,6 @@ trait Circuit extends width.Circuit {
   def mergeIds: Circuit = this
 }
 trait Id extends width.Id with Circuit
-object Id {
-  def apply(i: Int) = new Id{val n=i}
-  def unapply(c: Id) = Some(c.n)
-}
 trait Fan extends width.Fan with Circuit
 trait Beside extends width.Beside with Circuit {
   val c1, c2: Circuit
@@ -19,6 +15,12 @@ trait Beside extends width.Beside with Circuit {
   }
 }
 //END_MERGEIDS_SCALA
+//BEGIN_EXTRACTOR
+object Id {
+  def apply(i: Int) = new Id { val n = i }
+  def unapply(c: Id): Option[Int] = Some(c.n)
+}
+//END_EXTRACTOR
 object Beside {
   def apply(x: Circuit, y: Circuit): Circuit = new Beside{val c1=x; val c2=y}
 }

### 2016-06-24 paper/sections/ShallowOO.tex
Clean up Haskell code

@@
 the running example.  We first give the original shallow embedded
-implementation in Haskell and gradually rewrite it towards an ``OO
-style''.  Then translating the program into an OO language becomes
-straightforward.
+implementation in Haskell and rewrite it towards an ``OO style''.
+Then translating the program into an OO language becomes straightforward.
 
@@
 uses a record implemented as a newtype. The record has a sole field representing the result:
-\lstinputlisting[linerange=30-35]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
+\lstinputlisting[linerange=2-5]{./code/Rewrite.hs}%APPLY:CIRCUIT3
 %\bruno{width3 should now be width2}
 %We no longer need a separate \lstinline{width3} function as the field name is the extractor.
-The implementation is still shallow since essentially the same program
-as the previous one. In fact Haskell \lstinline{newtype} do not add any operational
-behaviour to the program, and the two programs are effectivelly the
+The implementation is still shallow because \lstinline{newtype} does not add any operational
+behaviour to the program, and hence the two programs are effectivelly the
 same.  However, having fields makes the program look more like an 

### 2016-06-24 paper/sections/Transformations.tex
Clean up Haskell code

@@
 some simple transformations to Haskell.
-This sections illustrates how to define transformations, the missing kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
+This section illustrates how to define transformations, the missing kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
 
@@
 We can port the implementation back to Haskell thanks to its laziness:
-\lstinputlisting[linerange=6-21]{./code/desugar.hs}%APPLY:DESUGAR
+\lstinputlisting[linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
 

### 2016-06-24 paper/code/Desugar.hs

@@ -3,21 +3,21 @@ fan :: Int -> Circuit
 beside :: Circuit -> Circuit -> Circuit
 
 --BEGIN_DESUGAR
-data Circuit2 = Circuit2 {
-  width2   :: Int,
-  desugar2 :: Circuit2
+data Circuit = Circuit {
+  width   :: Int,
+  desugar :: Circuit
 }
-id2 n = Circuit2 {
-  width2   = n,
-  desugar2 = foldr1 beside $ replicate n (fan 1)
+id n = Circuit {
+  width   = n,
+  desugar = foldr1 beside $ replicate n (fan 1)
 }
-fan2 n = Circuit2 {
-  width2   = n,
-  desugar2 = fan n
+fan n = Circuit {
+  width   = n,
+  desugar = fan n
 }
-beside2 c1 c2 = Circuit {
-  width2   = width2 c1 + width2 c2,
-  desugar2 = beside2 (desugar2 c1) (desugar2 c2)
+beside c1 c2 = Circuit {
+  width   = width c1 + width c2,
+  desugar = beside (desugar c1) (desugar c2)
 }
 --END_DESUGAR
 

### 2016-06-24 paper/code/Rewrite.hs

@@ -1,0 +1,6 @@
+--BEGIN_CIRCUIT3
+newtype Circuit = Circuit {width :: Int}
+id n         = Circuit {width = n}
+fan n        = Circuit {width = n}
+beside c1 c2 = Circuit {width = width c1 + width c2}
+--END_CIRCUIT3

### 2016-06-27 paper/sections/Interpretations.tex
Clean up Haskell code
One pass

@@
 multiple interpretations using tuples. However, their encoding needs to modify
-the original code. This section illustrates how various type of
-interpretations can be defined in OOP.
+the original code, and thus is non-modular. This section illustrates how various type of
+interpretations can be \emph{modularly} defined in OOP.
 \begin{comment}
@@
 In contrast, new interpretations are introduced in a 
-modular and intuitive way with an OO language like Scala, as shown in Figure~\ref{code:operation}.
+modular way with an OO language like Scala, as shown in Figure~\ref{code:operation}.
 \begin{figure}
@@
 % Multiple inheritance
+\begin{comment}
 We can even define \lstinline{wellSized} independently:
@@
 \end{lstlisting}
+\end{comment}
 
@@
 
-To illustrate dependent interpretations, we first add two new language constructs
+To illustrate dependent interpretations, we first add two new language 
 constructs \emph{above} and \emph{stretch}:
 \setlength{\grammarindent}{5em} % increase separation between LHS/RHS
-%\lstinputlisting[linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+%\lstinputlisting[linerange=29-30]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 \begin{grammar}
@@
 type-refinement) adds expressiveness over traditional procedural
-abstraction.
-
+abstraction.\bruno{Need to briefly mention that some they discuss 
+some advanced solutions in Haskell to address some of the problems:
+example: DTC and finally tagless}
 Gibbons and Wu, also show some different variants of interpretations, 
 such as context-sensitive interpretations. 
+Context-sensitive transformations do not pose any particular
+challenge. For space reasons, we omit the implementation details here. Full details
+are available online.
+
+\begin{comment}
 Unlike \lstinline{width} and \lstinline{wellSized} which can be defined with
@@
 are available online.
+\end{comment}

### 2016-06-27 paper/sections/ShallowOO.tex
One pass

@@
 
-This implementation is exactly how we would model \dsl in an OO language in the first
+This implementation is essentially how we would model \dsl in an OO language in the first
 place, following the \interp pattern (which uses \textsc{Composite} pattern to
-organize classes).
+organize classes). A minor difference is the use of
+traits, instead of classes. As we shall see using traits instead of
+classes will enable some additional modularity via multiple (trait-)inheritance.
 In summary, shallow embedding and straightforward OO programming are closely

### 2016-06-27 paper/sections/Transformations.tex
One pass

@@
 Transformations are typically regarded as the privilege of deep embeddings as no
-persistent AST obscures how to define transformations in shallow embeddings.
+persistent AST makes it non-obvious how to define transformations in shallow embeddings.
 The \textsc{Composite} pattern applied in our OO approach encodes the AST, making
@@
 some simple transformations to Haskell.
-This section illustrates how to define transformations, the missing kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
+This section illustrates how to define transformations, a missing kind
+of interpretations in Gibbons and Wu's paper, in shallow embeddings.
 
 \subsection{Desugaring}
-Desugaring is an transformation, which eliminates some
+Desugaring is a transformation, which eliminates some
 language constructs without sacrificing the expressiveness of the language.
-\emph{id} construct is such a syntactic sugar:
+The \emph{id} construct is an example of syntactic sugar:
 $$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
@@
 
+\paragraph{Desugaring in Scala}
 A corresponding Scala implementation is given below:
@@
 straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
-$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the formula.
+$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the
+formula.
 For constructing circuits easily, we define some companion objects with an \lstinline{apply} method implemented:
 \lstinputlisting[linerange=20-26]{../src/desugar/Circuit.scala}%APPLY:COMPANION
+\bruno{I would try to avoid Scala-specific things like extractors as
+  much as possible here. You can just define, for example:
 
-%\weixin{define a show function in the and show the result after desugaring?}
-We can port the implementation back to Haskell thanks to its laziness:
+def Fan (x : Int) = new Fan\{val n=x\}
+
+This is shorter, and less language-specific.
+}
+
+%\weixin{define a show function in the and show the result after
+%desugaring?}
+\paragraph{Desugaring in Haskell}
+Interestingly we can port the implementation back to Haskell:
 \lstinputlisting[linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
+\bruno{Please try to put more code in single lines:
+
+data Circuit = Circuit \{width::Int, desugar::Circuit\}
+
+}
+
+\noindent This code follows well-known encodings of objects into 
+\emph{functional} calculi with recursive records~\cite{}. Thanks 
+to Haskell's laziness, the port of recursive definitions such as
+\lstinline{beside} is straightforward.
 
 \subsection{Sophisticated Optimizations}
-Sophisticated optimizations may need to inspect multiple internal
-representations of values, which imposes difficulty to shallow embeddings.
+A final example is a more sophisticated type of optimization, 
+which requires inspecting multiple internal
+representations of values. Such type of optimizations 
+ imposes extra difficulty to shallow embeddings, but both 
+Cook~\cite{cook09abstraction} and Reynolds~\cite{reynolds94proceduralabstraction} have shown how it 
+is possible to overcome such difficulties. 
+%which imposes difficulty to shallow embeddings.
 For instance, consecutive \emph{id}s put side by side can be merged together:
@@
 Fortunately, Scala has built-in support for pattern matching by decorating
-classes with \lstinline{case} modifiers or defining extractor objects.
-For extensibility reasons, case classes can not be used
-because case class to case class extension is not supported.
+classes with \lstinline{case} modifiers or defining extractor objects~\cite{emir2007matching}.
+%For extensibility reasons, case classes can not be used
+%because case class to case class extension is not supported.
 Instead, we manually define \lstinline{unapply} methods in companion
@@
 For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type test and type casts},
-though they may not be as elegant as extractors~\cite{emir2007matching}.
+though they may not be as elegant as extractors.
 %Interested reader can refer to \ref{} for implementations using these approaches.

### 2016-06-27 paper/sections/Interpretations.tex
Address comments

@@
 example: DTC and finally tagless}
-Gibbons and Wu, also show some different variants of interpretations, 
+Of course, there are solutions to EP in Haskell, e.g. \emph{finally
+  tagless}~\cite{carette2009finally} and \emph{data types a la
+  carte}~\cite{swierstra2008data} (DTC).
+Finally tagless approach uses a type class to abstract over all interpretations
+of the language. Concrete interpretations are given through creating a data type and
+making it an instance of that type. However, it forces dependent interpretations to be defined along with what they depend on.
+DTC represents language constructs separately and composes them together using
+extensible sums. However, not like OO languages which come with subtyping, one
+has to manually implement the subtyping machinery for variants.
+
+Gibbons and Wu also show some different variants of interpretations,
 such as context-sensitive interpretations. 

### 2016-06-27 paper/sections/Transformations.tex
Address comments

@@
 formula.
-For constructing circuits easily, we define some companion objects with an \lstinline{apply} method implemented:
-\lstinputlisting[linerange=20-26]{../src/desugar/Circuit.scala}%APPLY:COMPANION
+For constructing circuits easily, we define some helper methods:
+\lstinputlisting[linerange=3-5]{../src/desugar/package.scala}%APPLY:COMPANION
 \bruno{I would try to avoid Scala-specific things like extractors as
@@
 Interestingly we can port the implementation back to Haskell:
-\lstinputlisting[linerange=6-21]{./code/Desugar.hs}%APPLY:DESUGAR
+\lstinputlisting[linerange=6-15]{./code/Desugar.hs}%APPLY:DESUGAR
 \bruno{Please try to put more code in single lines:

### 2016-06-27 paper/code/Desugar.hs

@@ -3,18 +3,12 @@ fan :: Int -> Circuit
 beside :: Circuit -> Circuit -> Circuit
 
 --BEGIN_DESUGAR
-data Circuit = Circuit {
-  width   :: Int,
-  desugar :: Circuit
-}
+data Circuit = Circuit{width::Int, desugar::Circuit}
 id n = Circuit {
   width   = n,
   desugar = foldr1 beside $ replicate n (fan 1)
 }
-fan n = Circuit {
-  width   = n,
-  desugar = fan n
-}
+fan n = Circuit { width = n, desugar = fan n }
 beside c1 c2 = Circuit {
   width   = width c1 + width c2,
   desugar = beside (desugar c1) (desugar c2)

### 2016-06-27 src/desugar/Circuit.scala

@@ -16,12 +16,3 @@ trait Id extends width.Id with Circuit {
     List.fill(n)(Fan(1)).reduce(Beside(_,_))
 }
 //END_DESUGAR_SCALA
-//BEGIN_COMPANION
-object Fan {
-  def apply(x: Int): Circuit = new Fan{val n=x}
-}
-object Beside {
-  def apply(x: Circuit, y: Circuit): Circuit = 
-    new Beside{ val c1 = x; val c2 = y }
-}
-//END_COMPANION

### 2016-06-27 src/desugar/package.scala

@@ -1,0 +1,7 @@
+package object desugar {
+//BEGIN_COMPANION
+def Fan(x: Int): Circuit = new Fan{val n=x}
+def Beside(x: Circuit, y: Circuit): Circuit = 
+    new Beside{ val c1 = x; val c2 = y}
+//END_COMPANION
+}

### 2016-06-27 paper/sections/Conclusion.tex
Conclusion

\section{Conclusion}
This paper shows how OO programming improves the modularity of shallow EDSLs.
With the procedural abstraction mechanisms provided by OO languages, various types of
interpretations can be defined modularly. We also show that defining
transformations in shallow embeddings is possible.

There are still certain interpretations that can not be modeled modularly,
including transformations and binary methods.
For transformations, we have to refine their return type in extensions and
supply a new implementation by duplicating the original code.
For binary methods like equality, there is no way to refine the argument type.
How to define these methods modularly is a possible line of future work.
Also, it is still tedious to express the inheritance relationships in extensions,
especially when multiple inheritance is used. Another line of future work is to
use some meta-programming mechanisms to eliminate these boilerplate.

### 2016-06-27 paper/sections/Transformations.tex
Shorten method name

@@
 Interestingly we can port the implementation back to Haskell:
-\lstinputlisting[linerange=6-15]{./code/Desugar.hs}%APPLY:DESUGAR
+\lstinputlisting[linerange=6-13]{./code/Desugar.hs}%APPLY:DESUGAR
 \bruno{Please try to put more code in single lines:
@@
 which requires inspecting multiple internal
-representations of values. Such type of optimizations 
- imposes extra difficulty to shallow embeddings, but both 
+representations of abstract values at the same time.
+Such type of optimizations imposes extra difficulty to shallow embeddings, but both
 Cook~\cite{cook09abstraction} and Reynolds~\cite{reynolds94proceduralabstraction} have shown how it 
@@
 objects:
-\lstinputlisting[linerange=19-22]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
+\lstinputlisting[linerange=18-21]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
 Then we can define this optimization:
-\lstinputlisting[linerange=4-16]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
+\lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
 

### 2016-06-27 src/desugar/package.scala

@@ -1,7 +1,7 @@
 package object desugar {
 //BEGIN_COMPANION
-def Fan(x: Int): Circuit = new Fan{val n=x}
+def Fan(x: Int): Circuit = new Fan { val n = x }
 def Beside(x: Circuit, y: Circuit): Circuit = 
-    new Beside{ val c1 = x; val c2 = y}
+    new Beside { val c1 = x; val c2 = y}
 //END_COMPANION
 }

### 2016-06-27 src/optimizations/Circuit.scala

@@ -2,16 +2,15 @@ package optimizations
 
 //BEGIN_MERGEIDS_SCALA
 trait Circuit extends width.Circuit {
-  def mergeIds: Circuit = this
+  def merge: Circuit = this
 }
 trait Id extends width.Id with Circuit
 trait Fan extends width.Fan with Circuit
 trait Beside extends width.Beside with Circuit {
   val c1, c2: Circuit
-  override def mergeIds = 
-    (c1.mergeIds,c2.mergeIds) match {
-      case (Id(n1),Id(n2)) => Id(n1 + n2)
-      case _ => Beside(c1.mergeIds, c2.mergeIds)
+  override def merge = (c1.merge, c2.merge) match { 
+    case (Id(n1), Id(n2)) => Id(n1 + n2) 
+    case _ => Beside(c1.merge, c2.merge)
   }
 }
 //END_MERGEIDS_SCALA

### 2016-06-27 paper/sections/Interpretations.tex
paper

@@
 type-refinement) adds expressiveness over traditional procedural
-abstraction.\bruno{Need to briefly mention that some they discuss 
-some advanced solutions in Haskell to address some of the problems:
-example: DTC and finally tagless}
-Of course, there are solutions to EP in Haskell, e.g. \emph{finally
-  tagless}~\cite{carette2009finally} and \emph{data types a la
+abstraction. Gibbons and Wu do discuss a number of advanced techniques that 
+can solve some of the modularity problems. For example, using type
+classes, the \emph{finally
+  tagless}~\cite{carette2009finally} can deal with the example in
+Section~\ref{subsec:multiple}. However
+tuples are still needed 
+to deal with dependent interpretations. In contrast the approach
+proposed here is just straightforward OOP, and dependent
+interpretations are not a problem.
+\begin{comment}
+and \emph{data types a la
   carte}~\cite{swierstra2008data} (DTC).
@@
 has to manually implement the subtyping machinery for variants.
-
+\end{comment}
 Gibbons and Wu also show some different variants of interpretations,
 such as context-sensitive interpretations. 
-Context-sensitive transformations do not pose any particular
-challenge. For space reasons, we omit the implementation details here. Full details
-are available online.
+Context-sensitive transformations are unproblematic as well. 
+For space reasons, full details are only available online.
 

### 2016-06-27 paper/sections/Introduction.tex
paper

@@
 
-In this paper, however, we prefer to interpret shallow embeddings
-differently. We argue that a straightforward OO approach is really a
-shallow embedding. The reason is that shallow embeddings are 
-implemented using \emph{procedural abstraction} and, as Cook~\cite{cook09abstraction} argued,
-procedural abstraction is the essence of Object-Oriented
-Programming. Thus the implementation of a shallow
+In this paper, to avoid ambiguity, we prefer a more precise meaning
+for shallow embeddings as EDSLs implemented using \emph{procedural abstraction}~\cite{reynolds94proceduralabstraction}. Such
+interpretation arises from the domain of shallow DSLs being
+typically a function, and procedural abstraction being a way to encode
+data abstractions using functions. As Cook~\cite{cook09abstraction} argued,
+procedural abstraction is also the essence of Object-Oriented
+Programming. Thus, according to our definition, the implementation of a shallow
 DSL in OOP languages should simply correspond to a standard
@@
 The main goal of this paper is to argue that OOP languages have
-advantages for the implementation of shallow embeddings. This should
-not came as a surprise, as OOP languages have evolved over more than 50
-years to improve the use of procedural abstraction.
-
+advantages for the implementation of shallow embeddings. 
+%This should
+%not came as a surprise, as OOP languages have evolved over more than 50
+%years to improve the use of procedural abstraction. 
 In order to understand how OOP can help with shallow embeddings, lets

### 2016-06-27 paper/sections/Transformations.tex
paper

@@
 \lstinputlisting[linerange=3-5]{../src/desugar/package.scala}%APPLY:COMPANION
-\bruno{I would try to avoid Scala-specific things like extractors as
-  much as possible here. You can just define, for example:
-
-def Fan (x : Int) = new Fan\{val n=x\}
-
-This is shorter, and less language-specific.
-}
 
@@
 \lstinputlisting[linerange=6-13]{./code/Desugar.hs}%APPLY:DESUGAR
-\bruno{Please try to put more code in single lines:
-
-data Circuit = Circuit \{width::Int, desugar::Circuit\}
-
-}
 

### 2016-06-27 paper/sections/Interpretations.tex
Rewording

@@
 A well-known limitation of shallow embeddings is that they allow only a single
-interpretation. Gibbons and Wu worked around this problem by accommodating
+interpretation. Gibbons and Wu~\cite{gibbons2014folding} work around this problem by accommodating
 multiple interpretations using tuples. However, their encoding needs to modify
@@
 interpretations, where one or more interpretations can use other
-interpretations in their definition. Typically such interpretations 
+interpretations in their definition. \weixin{The definition is not very clear,
+  starting from where}
+Typically such interpretations 
 would be said to be \emph{non-compositional} since they do not 
@@
 tuples in a non-modular way. We omit the Haskell code here, since 
-it is quite similar to the example in Section~\ref{subsec:multiple}
+it is quite similar to the example in Section~\ref{subsec:multiple}.
 
@@
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
-Definitions of \lstinline{wellSized} for the extended cases make it a dependent
-interpretation, as they need \lstinline{width} in their definition.
+The definitions of \lstinline{wellSized} for extended cases depend on \lstinline{width}.
 Note that \lstinline{width} and \lstinline{wellSized} are defined separately.
@@
 \subsection{Discussion} 
-Gibbons and Wu~\cite{gibbons2014folding} claim that in shallow
+Gibbons and Wu claim that in shallow
 embeddings new language constructs are easy to add, but new

### 2016-06-27 paper/sections/Introduction.tex
Rewording

@@
 
-Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{hudak1998modular}, existing
+Since Hudak's seminal paper on \weixin{EDSLs}~\cite{hudak1998modular}, existing
 languages (such as Haskell) have been used to directly encode
@@
 Programming. Thus, according to our definition, the implementation of a shallow
-DSL in OOP languages should simply correspond to a standard
+EDSL in OOP languages should simply correspond to a standard
 object-oriented program.

### 2016-06-27 paper/sections/RelatedWorkShort.tex
Rewording

@@
   Jovanovic:2014:YCD:2658761.2658771, scherr2014implicit} trying to blend these two
-approaches to enjoy benefits from both by using the shallow embedding as the
-surface language and generate deep embedding version to allow optimizations to
-be defined.
+approaches to enjoy benefits from both.
+They typically encode the surface language with a shallow embedding and
+then generate or translate to a deep embedded version for allowing optimizations.
 %Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at the same time so that clients can choose for a particular interpretation;
 The OO approach we present retains the simplicity of shallow embedding while
-makes it possible to directly implement optimizations.
+makes it possible to implement optimizations.
 
@@
 type parameterization, advanced features or heavy encoding which impose
-complexity to the DSL implementers or even clients and may be too complicated to use in practice.
+complexity to the DSL implementers or even clients, preventing practical use.
 For example, solutions based on the \textsc{Visitor}
@@
 optimizations well.
-Our approach, however, is simple, intuitive and supports transformations.
+Our approach, however, is simple and supports transformations.

### 2016-06-27 paper/sections/ShallowOO.tex
Rewording

@@
 
-The DSL consists of three constructs - two primitives
+\noindent The DSL consists of three constructs - two primitives
 \emph{id} and \emph{fan} and one combinator \emph{beside}.
@@
 organize classes). A minor difference is the use of
-traits, instead of classes. As we shall see using traits instead of
-classes will enable some additional modularity via multiple (trait-)inheritance.
-In summary, shallow embedding and straightforward OO programming are closely
+traits, instead of classes. Using traits instead of
+classes enables some additional modularity via multiple (trait-)inheritance.
+In summary, shallow embeddings and straightforward OO programming are closely
 related.

### 2016-06-28 paper/sections/Interpretations.tex
Revision

--- 2016-06-27 paper/sections/Interpretations.tex
@@
 
-A well-known limitation of shallow embeddings is that they allow only a single
-interpretation. Gibbons and Wu worked around this problem by accommodating
-multiple interpretations using tuples. However, their encoding needs to modify
+A often stated limitation of shallow embeddings is that they allow only a single
+interpretation. Gibbons and Wu worked around this problem by using tuples. However, their encoding needs to modify
 the original code, and thus is non-modular. This section illustrates how various type of
@@
 on defining the two interpretations (\lstinline{width} and
-\lstinline{wellSized}) simultaneously, using a pair. It is not
-possible to define \lstinline{width} independently and then reuse it.
+\lstinline{wellSized}) simultaneously, using a tuple. It is not
+possible reuse the independently defined \lstinline{width} function in
+Section~\ref{subsec:shallow}.
 Whenever a new interpretation is needed (e.g. \lstinline{wellSized}), the
-original code has to be revised -
+original code has to be revised:
 the arity of the tuple must be incremented and the new interpretation has to be
@@
 \paragraph{Multiple Interpretations in Scala}
-In contrast, new interpretations are introduced in a 
-modular way with an OO language like Scala, as shown in Figure~\ref{code:operation}.
+In contrast, Scala allows new interpretations to be introduced in a 
+modular way, as shown in Figure~\ref{code:operation}.
 \begin{figure}
@@
 from \lstinline{base} and
-complementing the body of \lstinline{wellSized}.
+implementing \lstinline{wellSized}.
 Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit} type
@@
 would be said to be \emph{non-compositional} since they do not 
-dependent only on themselves. 
+depend only on themselves. 
 In Haskell dependent interpretations are once again defined with
@@
 can solve some of the modularity problems. For example, using type
-classes, the \emph{finally
+classes, \emph{finally
   tagless}~\cite{carette2009finally} can deal with the example in

### 2016-06-28 paper/sections/Introduction.tex
Revision

--- 2016-06-27 paper/sections/Introduction.tex
@@
 
-Although the above definition is quite reasonable and widely accepted,
+\noindent Although the above definition is quite reasonable and widely accepted,
 it leaves some space to (mis)interpretation. For example it is unclear 
@@
 classified as a shallow or deep embedding? We believe arguments can be
-made both ways. The {\sc Composite} or {\sc Interpreter}
-patterns are normally accepted to provide a way to encode ASTs. Thus, 
-one possible interpretation is that \emph{according to definition of deep embedding
-  above, the OO approach classifies as a deep
-  embedding}.
+made both ways. 
+%The {\sc Composite} or {\sc Interpreter}
+%patterns are normally accepted to provide a way to encode ASTs. Thus, 
+%one possible interpretation is that \emph{according to definition of deep embedding
+%  above, the OO approach classifies as a deep
+%  embedding}.
 
@@
 %years to improve the use of procedural abstraction. 
-In order to understand how OOP can help with shallow embeddings, lets
+In order to understand how OOP can help with shallow embeddings, it is
+helpful to
 first review some of the limitations commonly
 found in the literature:
-
+\vspace{-5pt}
 \begin{enumerate}
@@
 \end{enumerate}
+\vspace{-5pt}
 
@@
 
-We show that OOP abstractions, including \emph{(multiple)
+We show that OOP abstractions, including \emph{
   inheritance}, \emph{subtyping} and \emph{type-refinement}, are
@@
 where procedural abstraction is used in Haskell to model a simple
-EDSL, and we recode that EDSL in Scala. Although from the
-\emph{syntactical} point of view there are minor inconveniences in
-the Scala version, from the \emph{semantic} and \emph{modularity} point of view the
-Scala version has clear advantages.
+EDSL, and we recode that EDSL in Scala. From the \emph{semantic} and \emph{modularity} point of view the
+Scala version has clear advantages over the Haskell version.
 

### 2016-06-28 paper/sections/ShallowOO.tex
Revision

--- 2016-06-27 paper/sections/ShallowOO.tex
@@
 
-The DSL consists of three constructs - two primitives
+\noindent The DSL consists of three constructs: two primitives
 \emph{id} and \emph{fan} and one combinator \emph{beside}.
@@
 
-\subsection{Shallow Embeddings and OOP}
+\subsection{Shallow Embeddings and OOP}\label{subsec:shallow}
 Shallow embeddings define a language directly through encoding its semantics
@@
 For example, we will immediately get $6$ for \lstinline{beside (fan 3) (id 3)}.
+
+Note that, for this simple circuit DSL, the Haskell domain is simply
+\lstinline{Int}. This domain is a degenerate case of
+procedural abstraction, where \lstinline{Int} can be viewed 
+as a no argument function. In Haskell, due to laziness, \lstinline{Int}
+is a good representation. In a call-by-value language 
+a no-argument function \lstinline{() -> Int} would be more
+appropriate to deal correctly with potential control-flow 
+language constructs. More realistic shallow DSLs, such as parser 
+combinators~\cite{leijen01parsec}, tend to have more complex functional domains.
 

### 2016-06-28 paper/sections/Interpretations.tex

@@
 A often stated limitation of shallow embeddings is that they allow only a single
-interpretation. Gibbons and Wu worked around this problem by using tuples. However, their encoding needs to modify
+interpretation. Gibbons and Wu~\cite{gibbons2014folding}  work around this problem by using tuples. However, their encoding needs to modify
 the original code, and thus is non-modular. This section illustrates how various type of
@@
 interpretations, where one or more interpretations can use other
-interpretations in their definition. Typically such interpretations 
-would be said to be \emph{non-compositional} since they do not 
-depend only on themselves. 
+interpretations in their definition. 
 In Haskell dependent interpretations are once again defined with
 tuples in a non-modular way. We omit the Haskell code here, since 
-it is quite similar to the example in Section~\ref{subsec:multiple}
+it is quite similar to the example in Section~\ref{subsec:multiple}.
 

### 2016-06-28 paper/sections/Introduction.tex

@@
 
-Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{hudak1998modular}, existing
+Since Hudak's seminal paper on EDSLs~\cite{hudak1998modular}, existing
 languages (such as Haskell) have been used to directly encode

### 2016-06-28 paper/sections/Transformations.tex

@@
 \subsection{Desugaring}
-Desugaring is a transformation, which eliminates some
+Desugaring is a transformation which eliminates some
 language constructs without sacrificing the expressiveness of the language.
@@
 This rule applies to a $beside\ c_1\ c_2$ only when both $c_1$ and $c_2$ are
-$id$, also their inner values are extracted for constructing the term
-on the right hand side of the equation.
+$id$, also their inner values are extracted for constructing a new $id$.
 Fortunately, Scala has built-in support for pattern matching by decorating
@@
 %because case class to case class extension is not supported.
-Instead, we manually define \lstinline{unapply} methods in companion
-objects:
+Here, we define the extractor for \lstinline{Id}:
 \lstinputlisting[linerange=18-21]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
-Then we can define this optimization:
+Then we can implement this optimization:
 \lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
 
-For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type test and type casts},
+For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type-tests/type-casts},
 though they may not be as elegant as extractors.

### 2016-06-28 paper/sections/Conclusion.tex
Conclusion

@@
 
-There are still certain interpretations that can not be modeled modularly,
-including transformations and binary methods.
+There are still certain interpretations that can not be modeled modularly using
+our approach, including transformations and binary methods.
 For transformations, we have to refine their return type in extensions and
@@
 For binary methods like equality, there is no way to refine the argument type.
-How to define these methods modularly is a possible line of future work.
-Also, it is still tedious to express the inheritance relationships in extensions,
-especially when multiple inheritance is used. Another line of future work is to
-use some meta-programming mechanisms to eliminate these boilerplate.
+Though with advanced type system features of Scala, these can still be defined
+modularly~\ref{zenger05independentlyextensible}. However, the encoding would
+become complicated.
+Also, we can further eliminate some boilerplate through code generation.
+For example, similar inheritance relationships are repeated in each constructs
+for extensions.

### 2016-06-28 paper/sections/Conclusion.tex
More changes

--- 2016-06-27 paper/sections/Conclusion.tex
@@
-\section{Conclusion}
+\section{Discussion and Conclusion}
 This paper shows how OO programming improves the modularity of shallow EDSLs.
@@
 
-There are still certain interpretations that can not be modeled modularly,
-including transformations and binary methods.
-For transformations, we have to refine their return type in extensions and
-supply a new implementation by duplicating the original code.
-For binary methods like equality, there is no way to refine the argument type.
-How to define these methods modularly is a possible line of future work.
-Also, it is still tedious to express the inheritance relationships in extensions,
-especially when multiple inheritance is used. Another line of future work is to
-use some meta-programming mechanisms to eliminate these boilerplate.
+Existing work showed that shallow embeddings yield flexible and concise EDSLs, while deep embedding makes
+it easy to define optimizations.
+There is a lot of work~\cite{svenningsson2012combining,
+  Jovanovic:2014:YCD:2658761.2658771, scherr2014implicit} trying to blend these two
+approaches to enjoy benefits from both.
+They typically encode the surface language with a shallow embedding and
+then generate or translate to a deep embedded version for allowing optimizations.
+%Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at the same time so that clients can choose for a particular interpretation;
+The OO approach we present retains the simplicity of shallow embedding while
+making it possible to implement optimizations. It would be interesting
+to conduct larger case studies to assess whether the techniques
+presented here are enough to avoiding deep embeddings for various DSLs
+in the literature.
+
+One limitation of the techniques presented here is that
+transformations require code duplication in extensions.  For
+transformations, we have to refine their return type in extensions and
+supply a new implementation by duplicating the original code.  How to
+define these transformations modularly, and without code duplication
+is a possible line of future work.  Also, it is still tedious to
+express the inheritance relationships in extensions, especially when
+multiple inheritance is used. Another line of future work is to use
+some meta-programming mechanisms to eliminate such boilerplate.

### 2016-06-28 paper/sections/Interpretations.tex
More changes

@@
 $above\ c_1\ c_2$ combines two circuits of the same width vertically;
-\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
+\emph{stretch ns c} inserts more wires into the circuit \emph{c} by
 summing up \emph{ns}.
-Figure~\ref{code:variant} shows how to add them to \dsl, simply by
-\emph{modularly} defining new classes that implement \lstinline{Circuit}.
+Figure~\ref{code:variant} shows the Scala implementation, which
+\emph{modularly} defines new traits implementing \lstinline{Circuit}.
 \begin{figure}

### 2016-06-28 paper/sections/Transformations.tex
Conclusion
More changes

@@
 \subsection{Desugaring}
-Desugaring is a transformation which eliminates some
+Desugaring is a transformation which eliminates 
 language constructs without sacrificing the expressiveness of the language.
@@
 $beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the
-formula.
-For constructing circuits easily, we define some helper methods:
+formula. Circuits are constructed using the helper methods:
 \lstinputlisting[linerange=3-5]{../src/desugar/package.scala}%APPLY:COMPANION
@@
 \lstinputlisting[linerange=6-13]{./code/Desugar.hs}%APPLY:DESUGAR
-
+\bruno{Code is cut-off in my pdf! Are the lines correct; beside is not
+  fully defined?}
 \noindent This code follows well-known encodings of objects into 
-\emph{functional} calculi with recursive records~\cite{}. Thanks 
+\emph{functional} calculi with recursive records~\cite{bruce97comparing}. Thanks 
 to Haskell's laziness, the port of recursive definitions such as
@@
 representations of abstract values at the same time.
-Such type of optimizations imposes extra difficulty to shallow embeddings, but both
-Cook~\cite{cook09abstraction} and Reynolds~\cite{reynolds94proceduralabstraction} have shown how it 
-is possible to overcome such difficulties. 
+Such optimizations impose extra difficulty to shallow
+embeddings.
+%, but both
+%Cook~\cite{cook09abstraction} and Reynolds~\cite{reynolds94proceduralabstraction} have shown how it 
+%is possible to overcome such difficulties. 
 %which imposes difficulty to shallow embeddings.

### 2016-06-28 paper/sections/Conclusion.tex

@@
 transformations, we have to refine their return type in extensions and
-supply a new implementation by duplicating the original code.  How to
-define these transformations modularly, and without code duplication
+supply a new implementation by duplicating the original code. 
+Though with advanced type system features of Scala, these can still be defined
+modularly~\cite{zenger05independentlyextensible}, the encoding would
+become significantly more complicated.
+How to define these transformations modularly, without code
+duplication, and without sphiticated types
 is a possible line of future work.  Also, it is still tedious to

### 2016-06-28 paper/sections/ShallowOO.tex
Another pass on Section 2

@@
 %\bruno{why is ``id'' in bold?}
-\lstinline{Circuit} is just a type
-synonym of \lstinline{Int} and it is already an integer after circuit construction.
-For example, we will immediately get $6$ for \lstinline{beside (fan 3) (id 3)}.
-
-Note that, for this simple circuit DSL, the Haskell domain is simply
+%\lstinline{Circuit} is just a type
+%synonym of \lstinline{Int} and it is already an integer after circuit construction.
+%For example, we will immediately get $6$ for \lstinline{beside (fan 3) (id 3)}.
+Note that, for this tiny DSL, the Haskell domain is simply
 \lstinline{Int}. This domain is a degenerate case of
@@
 \paragraph{Towards OOP}
-A simple, \emph{semantics preserving}, rewriting of the above program 
-uses a record implemented as a newtype. The record has a sole field representing the result:
+A simple, \emph{semantics preserving}, rewriting of the above program is given
+below, where a record with a sole field captures the domain and is declared as a \lstinline{newtype}:
 \lstinputlisting[linerange=2-5]{./code/Rewrite.hs}%APPLY:CIRCUIT3
@@
 declaration becomes a method declaration.
-Each case in the semantic function corresponds to an class and its parameters become fields of the class.
-And all the classes extends \lstinline{Circuit} by implementing \lstinline{width}.
+Each case in the semantic function corresponds to a trait and its parameters become fields of that trait.
+And these traits extend \lstinline{Circuit} and implement \lstinline{width}.
 
-This implementation is essentially how we would model \dsl in an OO language in the first
+This implementation is essentially how we would model \dsl with an OO language in the first
 place, following the \interp pattern (which uses \textsc{Composite} pattern to

### 2016-06-28 paper/sections/Transformations.tex
Code

@@
 formula. Circuits are constructed using the helper methods:
-\lstinputlisting[linerange=3-5]{../src/desugar/package.scala}%APPLY:COMPANION
+\lstinputlisting[linerange=3-6]{../src/desugar/package.scala}%APPLY:COMPANION
 
@@
 Interestingly we can port the implementation back to Haskell:
-\lstinputlisting[linerange=6-13]{./code/Desugar.hs}%APPLY:DESUGAR
-\bruno{Code is cut-off in my pdf! Are the lines correct; beside is not
-  fully defined?}
+\lstinputlisting[linerange=6-15]{./code/Desugar.hs}%APPLY:DESUGAR
 \noindent This code follows well-known encodings of objects into 
@@
 Here, we define the extractor for \lstinline{Id}:
-\lstinputlisting[linerange=18-21]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
+\lstinputlisting[linerange=18-18]{../src/optimizations/Circuit.scala}%APPLY:EXTRACTOR
 Then we can implement this optimization:

### 2016-06-28 src/layout/package.scala

@@ -1,4 +1,12 @@
 package object layout {
+  import javax.swing._
+
+  def id(x: Int) = new Id{val n=x}
+  def fan(x: Int) = new Fan{val n=x}
+  def above(x: Circuit, y: Circuit) = new Above{val c1=x; val c2=y}
+  def beside(x: Circuit, y: Circuit) = new Beside{val c1=x; val c2=y}
+  def stretch(xs: List[Int], x: Circuit) = new Stretch{val ns=xs; val c=x}
+
   type IntPair = Tuple2[Int,Int]
   type Layout = List[List[IntPair]] 
   def lzw[A](xs: List[A], ys: List[A])(f: (A, A) => A): List[A] = (xs, ys) match {
@@ -13,4 +21,16 @@ package object layout {
   def connect(ns: List[Int])(p: IntPair) =
     pmap ((i: Int) => partialSum(ns)(i) - 1) (p)
   def partialSum(ns: List[Int]): List[Int] = ns.scanLeft(0)(_ + _).tail
+
+  def draw(layout: Layout) = SwingUtilities.invokeLater(new Runnable() {
+    override def run {
+      val frame = new JFrame
+      frame.setTitle("Draw Circuit")
+      frame.setResizable(false)
+      frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
+      frame.getContentPane().add(new DrawCircuit(layout));
+      frame.pack();
+      frame.setVisible(true);
+    }
+  })
 }

### 2016-06-28 src/scans/package.scala

@@ -1,23 +1,0 @@
-import javax.swing.SwingUtilities
-import javax.swing.JFrame
-
-package object scans {
-  import layout._  
-  def id(x: Int) = new Id{val n=x}
-  def fan(x: Int) = new Fan{val n=x}
-  def above(x: Circuit, y: Circuit) = new Above{val c1=x; val c2=y}
-  def beside(x: Circuit, y: Circuit) = new Beside{val c1=x; val c2=y}
-  def stretch(xs: List[Int], x: Circuit) = new Stretch{val ns=xs; val c=x}
-
-  def draw(layout: Layout) = SwingUtilities.invokeLater(new Runnable() {
-    override def run {
-      val frame = new JFrame
-      frame.setTitle("Draw Circuit")
-      frame.setResizable(false)
-      frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
-      frame.getContentPane().add(new DrawCircuit(layout));
-      frame.pack();
-      frame.setVisible(true);
-    }
-  })
-}

### 2016-06-28 src/test/Test.scala

@@ -1,9 +1,11 @@
-package scans
+package test
 
 object Test extends App {
+  import layout._
   val c = above(beside(fan(2), fan(2)),
           above(stretch(List(2,2), fan(2)),
           beside(id(1), beside(fan(2), id(1)))))
+
   println(c.width)
   println(c.wellSized)
   println(c.depth)

### 2016-06-28 paper/paper.tex
local
One pass
Address comments
Conclusion
More changes
Author

@@
 \usepackage{syntax}
+
+\setlist[itemize]{noitemsep,nolistsep}
+\setlist[enumerate]{noitemsep,nolistsep}
 
@@
 \titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{short description of paper}   % 'preprint' option specified.
+\preprintfooter{Shallow EDSLs and Object-Oriented Programming}   % 'preprint' option specified.
 
@@
 
-\authorinfo{}
-           {}
-           {}
+\authorinfo{Weixin Zhang \and Bruno C.d.S.Oliveira}
+           {The University of Hong Kong, Pokfulam, Hong Kong, China}
+           {\{wxzhang2,bruno\}@cs.hku.hk}
 %\authorinfo{Name}
@@
 \input{sections/ShallowOO.tex}
+\input{sections/Interpretations.tex}
 \input{sections/Transformations.tex}
-\input{sections/Modularity.tex}
-\input{sections/RelatedWorkShort.tex}
+%\input{sections/RelatedWorkShort.tex}
+\input{sections/Conclusion.tex}
+%\input{sections/Modularity.tex}
 %\input{sections/Overview.tex}

### 2016-06-28 paper/sections/Conclusion.tex
url + client code

--- 2016-06-28 paper/sections/Conclusion.tex
@@
 approaches to enjoy benefits from both.
-They typically encode the surface language with a shallow embedding and
+They typically encode the surface language in shallow embeddings and
 then generate or translate to a deep embedded version for allowing optimizations.
@@
 transformations, we have to refine their return type in extensions and
-supply a new implementation by duplicating the original code.  How to
-define these transformations modularly, and without code duplication
-is a possible line of future work.  Also, it is still tedious to
+supply a new implementation by duplicating the original code. 
+Though these methods can still be defined modularly with advanced type
+system features of Scala~\cite{zenger05independentlyextensible},
+the encoding would become significantly more complicated.
+How to define these transformations modularly, without code
+duplication, and without sophisticated types
+is a possible line of future work.  Also, it is tedious to
 express the inheritance relationships in extensions, especially when

### 2016-06-28 paper/sections/Interpretations.tex
url + client code

@@
 \lstinline{base.Circuit} and declares a new method \lstinline{wellSized}.
-The hierarchy implements the new \lstinline{Circuit} by inheriting the corresponding class
-from \lstinline{base} and
+The hierarchy implements the new \lstinline{Circuit} by inheriting the corresponding trait from \lstinline{base} and
 implementing \lstinline{wellSized}.
@@
 Context-sensitive transformations are unproblematic as well. 
-For space reasons, full details are only available online.
+For space reasons, full details are only available online~\url{https://github.com/wxzh/shallow-dsl}.
 

### 2016-06-28 src/test/Test.scala

@@ -1,5 +1,12 @@
 package test
 
+object SimpleTest {
+//BEGIN_CLIENT
+import width._
+beside(fan(3), id(3)).width // 6
+//END_CLIENT
+}
+
 object Test extends App {
   import layout._
   val c = above(beside(fan(2), fan(2)),

### 2016-06-28 src/width/package.scala

@@ -1,0 +1,6 @@
+package object width {
+def id(x: Int): Circuit = new Id { val n = x }
+def fan(x: Int): Circuit = new Fan { val n = x }
+def beside(x: Circuit, y: Circuit): Circuit = 
+    new Beside { val c1 = x; val c2 = y}
+}

### 2016-06-28 paper/sections/Transformations.tex

@@
 \lstinputlisting[linerange=3-6]{../src/desugar/package.scala}%APPLY:COMPANION
+These methods are interfaces exposed to DSL clients, e.g.
+\lstinputlisting[linerange=5-6]{../src/test/Test.scala}%APPLY:CLIENT
+By importing a different package, clients can change the semantics with the majority of the code untouched.
 

### 2016-06-28 paper/sections/Conclusion.tex
Conclusion
4 pages

@@
 One limitation of the techniques presented here is that
-transformations require code duplication in extensions.  For
-transformations, we have to refine their return type in extensions and
-supply a new implementation by duplicating the original code. 
-Though these methods can still be defined modularly with advanced type
+transformations require code duplication in extensions.
+We have to refine their return type and
+supply a new implementation by copying the original code.
+Though transformations can still be defined modularly with advanced type
 system features of Scala~\cite{zenger05independentlyextensible},

### 2016-06-28 paper/sections/Introduction.tex
4 pages

@@
 for shallow embeddings as EDSLs implemented using \emph{procedural abstraction}~\cite{reynolds94proceduralabstraction}. Such
-interpretation arises from the domain of shallow DSLs being
+interpretation arises from the domain of shallow EDSLs being
 typically a function, and procedural abstraction being a way to encode
@@
 found in the literature:
-\vspace{-5pt}
 \begin{enumerate}
@@
 \end{enumerate}
-\vspace{-5pt}
 
@@
 solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
-using just standard OOP mechanims enables \emph{multiple modular
+using just standard OOP mechanisms enables \emph{multiple modular
   interpretations} to co-exist and be combined in shallow embeddings.
@@
 where procedural abstraction is used in Haskell to model a simple
-EDSL, and we recode that EDSL in Scala. From the \emph{semantic} and \emph{modularity} point of view the
+EDSL, and we recode that EDSL in Scala\footnote{Available online: \url{https://github.com/wxzh/shallow-dsl}}.
+From the \emph{semantic} and \emph{modularity} point of view the
 Scala version has clear advantages over the Haskell version.

### 2016-06-28 paper/sections/Transformations.tex
4 pages

--- 2016-06-28 paper/sections/Transformations.tex
@@
 \lstinline{Circuit} reflecting the nature of transformations.
-Implementing it for each case is
-straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
-$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} just mimics the
-formula. Circuits are constructed using the helper methods:
+Implementing each case is straightforward:
+\emph{fan n} returns itself; $beside\ c_1\ c_2$ returns a new
+$beside$ with $c_1$ and $c_2$ desugared; \emph{id n} mimics the
+formula.
+
+Circuits are constructed via some helper methods:
 \lstinputlisting[linerange=3-6]{../src/desugar/package.scala}%APPLY:COMPANION
+These methods would be intensively used in client code, e.g.
+\lstinputlisting[linerange=5-6]{../src/test/Test.scala}%APPLY:CLIENT
+By importing a different package, clients can obtain new semantics with the code
+describing circuits untouched.
 

### 2016-06-28 paper/sections/Conclusion.tex
Explain merge

@@
 %Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at the same time so that clients can choose for a particular interpretation;
-The OO approach we present retains the simplicity of shallow embedding while
-making it possible to implement optimizations. It would be interesting
+Our OO approach retains the simplicity of shallow embeddings while
+allowing optimizations. It would be interesting
 to conduct larger case studies to assess whether the techniques
-presented here are enough to avoiding deep embeddings for various DSLs
+presented here are enough to avoid deep embeddings for various DSLs
 in the literature.
 
-One limitation of the techniques presented here is that
-transformations require code duplication in extensions.
-We have to refine their return type and
-supply a new implementation by copying the original code.
+One limitation of our approach is that
+transformations require code duplication in extensions,
+as we must give an implementation when refining their return type.
 Though transformations can still be defined modularly with advanced type

### 2016-06-29 paper/sections/Interpretations.tex

@@
 
-A often stated limitation of shallow embeddings is that they allow only a single
+An often stated limitation of shallow embeddings is that they allow only a single
 interpretation. Gibbons and Wu~\cite{gibbons2014folding} work around this problem by using tuples. However, their encoding needs to modify
@@
 
-\paragraph{Multiple Interpretations in Scala}
-In contrast, Scala allows new interpretations to be introduced in a 
-modular way, as shown in Figure~\ref{code:operation}.
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure}
+\lstinputlisting[linerange=2-11]{../src/width/Variants.scala}%APPLY:VARIANT_SCALA
+\caption{Adding new language constructs}
+\label{code:variant}
+\end{figure}
+
+\paragraph{Multiple Interpretations in Scala}
+In contrast, Scala allows new interpretations to be introduced in a 
+modular way, as shown in Figure~\ref{code:operation}.
 %Instead of modifying the original code, we define \lstinline{wellSized} modularly.
@@
 \emph{modularly} defines new traits implementing \lstinline{Circuit}.
-\begin{figure}
-\lstinputlisting[linerange=2-11]{../src/width/Variants.scala}%APPLY:VARIANT_SCALA
-\caption{Adding new language constructs}
-\label{code:variant}
-\end{figure}
 
@@
 Circuits written with $above$ and $stretch$ have to obey to certain
-size constrains, which are captured by the \lstinline{wellSized} method:
+size constrains%, which are captured by the \lstinline{wellSized} method:
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
@@
 such as context-sensitive interpretations. 
-Context-sensitive transformations are unproblematic as well. 
-For space reasons, full details are only available online~\url{https://github.com/wxzh/shallow-dsl}.
+These interpretations are unproblematic as well.
+Implementation details can be found online.
 

### 2016-06-29 paper/sections/Transformations.tex

@@
 \lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
+where we pattern matching on the two merged inner circuits of \lstinline{Beside}
+simultaneously and applies the rule when both of them are \lstinline{Id}.
 

### 2016-06-29 paper/sections/Interpretations.tex
Rephrase

@@
  \emph{Dependent interpretations} are a generalization of multiple
-interpretations, where one or more interpretations can use other
-interpretations in their definition. 
+interpretations, where other interpretations are used for definition.
 In Haskell dependent interpretations are once again defined with

### 2016-06-29 paper/sections/ShallowOO.tex
4 pages
Typo

@@
 \subsection{A DSL for Parallel Prefix Circuits}
-Consider an DSL named \dsl that models parallel circuits.
+Consider a DSL named \dsl that models parallel circuits.
 Its BNF grammar is given below:
@@
 The implementation is still shallow because \lstinline{newtype} does not add any operational
-behaviour to the program, and hence the two programs are effectivelly the
+behaviour to the program, and hence the two programs are effectively the
 same.  However, having fields makes the program look more like an 

### 2016-06-29 paper/sections/Transformations.tex
Typo

--- 2016-06-28 paper/sections/Transformations.tex
@@
 \lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
+where we pattern matching on the two merged inner circuits of \lstinline{Beside}
+simultaneously and apply the rule when both of them are \lstinline{Id}.
 
-For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type-tests/type-casts},
+For other OO languages, pattern match can still be simulated through \emph{test methods} or \emph{type-tests/type-casts},
 though they may not be as elegant as extractors.

### 2016-06-29 paper/sections/Transformations.tex
Typo

--- 2016-06-28 paper/sections/Transformations.tex
@@
 \lstinputlisting[linerange=4-15]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
+where we pattern match on the two merged inner circuits of \lstinline{Beside}
+simultaneously and apply the rule when both of them are \lstinline{Id}.
 

### 2016-06-30 paper/sections/Interpretations.tex
Dependent interpretation

@@
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
-The definitions of \lstinline{wellSized} for extended cases depend on \lstinline{width}.
+\lstinline{wellSized} becomes a dependent interpretation
+because it depends not only on itself but also on \lstinline{width}.
 Note that \lstinline{width} and \lstinline{wellSized} are defined separately.

### 2016-06-30 paper/sections/Interpretations.tex
Abbreviations & Dependent interpretation

--- 2016-06-29 paper/sections/Interpretations.tex
@@
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
-The definitions of \lstinline{wellSized} for extended cases depend on \lstinline{width}.
+The \lstinline{wellSized} method becomes a dependent interpretation
+because it depends not only on itself but also on \lstinline{width} in the
+definition of \lstinline{Beside} and \lstinline{Above}.
 Note that \lstinline{width} and \lstinline{wellSized} are defined separately.

### 2016-06-30 paper/sections/Introduction.tex
Abbreviations & Dependent interpretation

@@
 
-Since Hudak's seminal paper on EDSLs~\cite{hudak1998modular}, existing
-languages (such as Haskell) have been used to directly encode
+Since Hudak's seminal paper on Embedded Domain Specific Languages (EDSLs)~\cite{hudak1998modular}, existing
+languages (e.g. Haskell) have been used to directly encode
 DSLs. Two common approaches to EDSLs are the so-called \emph{shallow}
@@
 how to classify an EDSL implemented using the {\sc Composite} or {\sc Interpreter} 
-patterns in OOP. Would this OO approach be
+patterns in Object-Oriented Programming (OOP). Would this OO approach be
 classified as a shallow or deep embedding? We believe arguments can be
@@
 data abstractions using functions. As Cook~\cite{cook09abstraction} argued,
-procedural abstraction is also the essence of Object-Oriented
-Programming. Thus, according to our definition, the implementation of a shallow
+procedural abstraction is also the essence of OOP.
+Thus, according to our definition, the implementation of a shallow
 EDSL in OOP languages should simply correspond to a standard

### 2016-06-30 paper/sections/Introduction.tex
paper
intro

@@
 patterns in Object-Oriented Programming (OOP). Would this OO approach be
-classified as a shallow or deep embedding? We believe arguments can be
-made both ways. 
+classified as a shallow or deep embedding? Some authors working on
+OOP EDSLs ~\cite{} consider a {\sc Composite} to be a deep
+embedding. Some other authors~\cite{gibbons2014folding} 
+consider implementations using tuples and/or the  {\sc Composite}
+pattern to be a shallow embedding.
+
 %The {\sc Composite} or {\sc Interpreter}
@@
 
-In this paper, to avoid ambiguity, we prefer a more precise meaning
+In this paper, we avoid ambiguitity by using a more precise meaning
 for shallow embeddings as EDSLs implemented using \emph{procedural abstraction}~\cite{reynolds94proceduralabstraction}. Such
-interpretation arises from the domain of shallow EDSLs being
-typically a function, and procedural abstraction being a way to encode
+interpretation arises naturally from the domain of shallow EDSLs being
+typically functions, and procedural abstraction being a way to encode
 data abstractions using functions. As Cook~\cite{cook09abstraction} argued,
@@
 
-The main goal of this paper is to argue that OOP languages have
+The main goal of our paper is to show that OOP languages have
 advantages for the implementation of shallow embeddings. 
-%This should
-%not came as a surprise, as OOP languages have evolved over more than 50
-%years to improve the use of procedural abstraction. 
-In order to understand how OOP can help with shallow embeddings, it is
-helpful to
-first review some of the limitations commonly
-found in the literature:
+It is perhaps due to those advantages that some authors 
+have considered  {\sc Composite}-based implementations to be deep
+embeddings, rather than shallow. However, we argue that the OOP
+mechanisms do not change the essence of the implementation, which is
+still shallow (i.e. using procedural abstraction). To understand how
+OOP is helpful for shallow embeddings, we should first look at the
+limitations of shallow embeddings commonly found in the literature:
 \begin{enumerate}
@@
 \end{enumerate}
-
-\noindent Both limitations are often used as motivators to switch to deep embeddings.
 
@@
 EDSL, and we recode that EDSL in Scala\footnote{Available online: \url{https://github.com/wxzh/shallow-dsl}}.
-From the \emph{semantic} and \emph{modularity} point of view the
+From the \emph{modularity} point of view the
 Scala version has clear advantages over the Haskell version.

### 2016-06-30 paper/paper.tex
url + client code
4 pages
More rewritting for intro

@@
 \setlist[enumerate]{noitemsep,nolistsep}
+\setlength{\floatsep}{6.0pt plus 2.0pt minus 2.0pt}
+\setlength{\textfloatsep}{10.0pt plus 2.0pt minus 4.0pt}
+\setlength{\intextsep}{6.0pt plus 2.0pt minus 2.0pt}
 
@@
 argument by taking a recent paper by Gibbons and Wu, where procedural
-abstraction is used in Haskell to model a simple EDSL, and we recode
+abstraction is used in Haskell to model a simple shallow EDSL, and we recode
 that EDSL in Scala. From the \emph{semantic}

### 2016-06-30 paper/sections/Introduction.tex
More rewritting for intro

@@
 patterns in Object-Oriented Programming (OOP). Would this OO approach be
-classified as a shallow or deep embedding? Some authors working on
+classified as a shallow or deep embedding? We feel there is a rather
+fuzzy line here, and the literature allows for both interpretations. Some authors working on
 OOP EDSLs ~\cite{} consider a {\sc Composite} to be a deep
@@
 pattern to be a shallow embedding.
+
+To avoid ambiguitity we propose defining shallow embeddings as EDSLs implemented using \emph{procedural abstraction}~\cite{reynolds94proceduralabstraction}. Such
+interpretation arises naturally from the domain of shallow EDSLs being
+functions, and procedural abstraction being a way to encode
+data abstractions using functions. As Cook~\cite{cook09abstraction} argued,
+procedural abstraction is also the essence of OOP.
+Thus, according to our definition, the implementation of a shallow
+EDSL in OOP languages should simply correspond to a standard
+object-oriented program.
+
+The main goal of our paper is to show the close relationship between 
+shallow embeddings and OOP, and argue that OOP languages have
+advantages for the implementation of shallow embeddings. 
+It is perhaps partly due to those advantages that some authors 
+have considered  {\sc Composite}-based implementations to be deep
+embeddings. However, we argue that the OOP
+mechanisms do not change the essence of the implementation, which is
+still shallow (i.e. using procedural abstraction). 
+To understand how
+OOP is helpful for shallow embeddings, we should first look at the
+limitations of shallow embeddings commonly found in the literature:
+\begin{enumerate}
+
+\item {\bf Single Interpretation} An often stated limitation of
+  shallow EDSLs is that they only support one
+  interpretation.
+
+\item {\bf No Transformations} Another commonly stated limitation 
+of shallow embeddings is the lack of support for transformations,
+preventing optimizations.
+
+\end{enumerate}
+
+We show that OOP abstractions, including \emph{
+  inheritance}, \emph{subtyping} and \emph{type-refinement}, are
+helpful to address those problems. For the first problem, we can
+employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
+solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
+using just standard OOP mechanisms enables \emph{multiple modular
+  interpretations} to co-exist and be combined in shallow embeddings.
+For the second problem, we show that transformations can be encoded 
+with recursive objects. Interestingly enough, for transformations it
+is possible to port back the OO solution to
+Haskell.
+
+We make our arguments by taking a recent paper by Gibbons and Wu~\cite{gibbons2014folding},
+where procedural abstraction is used in Haskell to model a simple \emph{shallow}
+EDSL, and we recode that EDSL in Scala\footnote{Available online: \url{https://github.com/wxzh/shallow-dsl}}.
+From the \emph{modularity} point of view the
+Scala version has clear advantages over the Haskell version.
+
+\begin{comment}
+In summary, our contributions are:
+
+\begin{itemize}
+
+\item {}
+
+\item {\bf Multiple Modular Interpretations for Shallow Embeddings:} 
+  We show that with standard OOP mechanisms it is easy to support multiple modular
+  interpretations for shallow embeddings.
+
+\item {\bf Transformations for Shallow Embeddings:} We show that
+  transformations are encodable with recursive objects. Moreover, this technique
+  can be ported back into functional programming as well.
+
+\end{itemize}
+\end{comment}
+
+\begin{comment}
+If we accept Cook's view on OOP, 
+a natural way to distinguish implementations of EDSLs is 
+in terms of the data abstraction used to model the language
+constructs instead. As Reynold's~\cite{reynolds94proceduralabstraction} suggested there are two
+types of data abstraction: procedural abstraction and \emph{user-defined
+  types}. It is clear that shallow embeddings use \emph{procedural
+  abstraction}: the DSLs are modelled by interpretation
+functions. Thus, the other implementation option for EDSLs is to
+use \emph{user-defined types}. In Reynolds terminology user-defined
+types mean disjoint union types, which are nowadays commonly available
+in modern languages as \emph{algebraic datatypes}. Disjoint union
+types can also be emulated in OOP using the {\sc Visitor} pattern.
+
+A distinction based on data abstraction is more precise and provides a
+remedy for possible misinterpretation. An EDSL implemented with
+algebraic datatypes falls into the category of user-defined types
+(deep embedding), while a {\sc Composite}-based OO implementation falls
+under procedural abstraction (shallow embedding). 
+\end{comment}
+
+%At least some authors~\cite{} seem to implicitly
+%agree with the latter interpretation.
 
@@
 \end{comment}
-
-In this paper, we avoid ambiguitity by using a more precise meaning
-for shallow embeddings as EDSLs implemented using \emph{procedural abstraction}~\cite{reynolds94proceduralabstraction}. Such
-interpretation arises naturally from the domain of shallow EDSLs being
-typically functions, and procedural abstraction being a way to encode
-data abstractions using functions. As Cook~\cite{cook09abstraction} argued,
-procedural abstraction is also the essence of OOP.
-Thus, according to our definition, the implementation of a shallow
-EDSL in OOP languages should simply correspond to a standard
-object-oriented program.
-
-\begin{comment}
-If we accept Cook's view on OOP, 
-a natural way to distinguish implementations of EDSLs is 
-in terms of the data abstraction used to model the language
-constructs instead. As Reynold's~\cite{reynolds94proceduralabstraction} suggested there are two
-types of data abstraction: procedural abstraction and \emph{user-defined
-  types}. It is clear that shallow embeddings use \emph{procedural
-  abstraction}: the DSLs are modelled by interpretation
-functions. Thus, the other implementation option for EDSLs is to
-use \emph{user-defined types}. In Reynolds terminology user-defined
-types mean disjoint union types, which are nowadays commonly available
-in modern languages as \emph{algebraic datatypes}. Disjoint union
-types can also be emulated in OOP using the {\sc Visitor} pattern.
-
-A distinction based on data abstraction is more precise and provides a
-remedy for possible misinterpretation. An EDSL implemented with
-algebraic datatypes falls into the category of user-defined types
-(deep embedding), while a {\sc Composite}-based OO implementation falls
-under procedural abstraction (shallow embedding). 
-\end{comment}
-
-%At least some authors~\cite{} seem to implicitly
-%agree with the latter interpretation.
-
-The main goal of our paper is to show that OOP languages have
-advantages for the implementation of shallow embeddings. 
-It is perhaps due to those advantages that some authors 
-have considered  {\sc Composite}-based implementations to be deep
-embeddings, rather than shallow. However, we argue that the OOP
-mechanisms do not change the essence of the implementation, which is
-still shallow (i.e. using procedural abstraction). To understand how
-OOP is helpful for shallow embeddings, we should first look at the
-limitations of shallow embeddings commonly found in the literature:
-\begin{enumerate}
-
-\item {\bf Single Interpretation} An often stated limitation of
-  shallow embeddings is that they only support a single
-  interpretation.
-
-\item {\bf No Transformations} Another commonly stated limitation 
-of shallow embeddings is that they do not support transformations,
-preventing optimizations and other useful transformations.
-
-\end{enumerate}
-
-We show that OOP abstractions, including \emph{
-  inheritance}, \emph{subtyping} and \emph{type-refinement}, are
-helpful to address those problems. For the first problem, we can
-employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
-solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
-using just standard OOP mechanisms enables \emph{multiple modular
-  interpretations} to co-exist and be combined in shallow embeddings.
-For the second problem, we show that transformations can be encoded 
-with recursive objects. Interestingly enough, for transformations it
-is possible to port back the OO solution to
-Haskell.
-
-We make our arguments by taking a recent paper by Gibbons and Wu~\cite{gibbons2014folding},
-where procedural abstraction is used in Haskell to model a simple
-EDSL, and we recode that EDSL in Scala\footnote{Available online: \url{https://github.com/wxzh/shallow-dsl}}.
-From the \emph{modularity} point of view the
-Scala version has clear advantages over the Haskell version.
-
-\begin{comment}
-In summary, our contributions are:
-
-\begin{itemize}
-
-\item {}
-
-\item {\bf Multiple Modular Interpretations for Shallow Embeddings:} 
-  We show that with standard OOP mechanisms it is easy to support multiple modular
-  interpretations for shallow embeddings.
-
-\item {\bf Transformations for Shallow Embeddings:} We show that
-  transformations are encodable with recursive objects. Moreover, this technique
-  can be ported back into functional programming as well.
-
-\end{itemize}
-\end{comment}

### 2016-06-30 paper/sections/Interpretations.tex
Save more space

--- 2016-06-29 paper/sections/Interpretations.tex
@@
 interpretations, where other interpretations are used for definition.
-In Haskell dependent interpretations are once again defined with
-tuples in a non-modular way. We omit the Haskell code here, since 
-it is quite similar to the example in Section~\ref{subsec:multiple}.
+In Haskell dependent interpretations are again defined with
+tuples in a non-modular way. We omit the Haskell code here since
+it is similar to the example in Section~\ref{subsec:multiple}.
 
@@
 Circuits written with $above$ and $stretch$ have to obey to certain
-size constrains%, which are captured by the \lstinline{wellSized} method:
+size constrains:%, which are captured by the \lstinline{wellSized} method:
 \lstinputlisting[linerange=2-11]{../src/wellsized/Variants.scala}%APPLY:DEPENDENT_SCALA
-The definitions of \lstinline{wellSized} for extended cases depend on \lstinline{width}.
+The \lstinline{wellSized} method becomes a dependent interpretation
+because it depends not only on itself but also on \lstinline{width} in the
+definition of \lstinline{Above} and \lstinline{Stretch}.
 Note that \lstinline{width} and \lstinline{wellSized} are defined separately.

### 2016-06-30 paper/sections/Transformations.tex
saving some space
Save more space

@@
 The \emph{id} construct is an example of syntactic sugar:
+\vspace{-8pt}
 $$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
@@
 simultaneously and apply the rule when both of them are \lstinline{Id}.
-
 For other OO languages, pattern matching can still be simulated through \emph{test methods} or \emph{type-tests/type-casts},

### 2016-06-30 paper/sections/Conclusion.tex
saving some space
Conclusion rewording

@@
 \section{Discussion and Conclusion}
-This paper shows how OO programming improves the modularity of shallow EDSLs.
-With the procedural abstraction mechanisms provided by OO languages, various types of
-interpretations can be defined modularly. We also show that defining
-transformations in shallow embeddings is possible.
+This paper shows the close correspondence between OOP and shallow
+embeddings, and how OOP improves the modularity of shallow EDSLs.
+With OOP, various types of interpretations can be defined modularly. 
+Moreover defining transformations is also possible.
 
-Existing work showed that shallow embeddings yield flexible and concise EDSLs, while deep embedding makes
-it easy to define optimizations.
-There is a lot of work~\cite{svenningsson2012combining,
+Existing work showed that shallow embeddings yield flexible and
+concise EDSLs, while deep embedding makes it easy to define
+optimizations. There is a lot of work~\cite{svenningsson2012combining,
   Jovanovic:2014:YCD:2658761.2658771, scherr2014implicit} trying to blend these two
@@
 to conduct larger case studies to assess whether the techniques
-presented here are enough to avoid deep embeddings for various DSLs
+presented here are enough to avoid sophisticated deep embeddings for various DSLs
 in the literature.
 
-One limitation of our approach is that
+One limitation is that
 transformations require code duplication in extensions,
@@
 duplication, and without sophisticated types
-is a possible line of future work.  Also, it is tedious to
-express the inheritance relationships in extensions, especially when
-multiple inheritance is used. Another line of future work is to use
-some meta-programming mechanisms to eliminate such boilerplate.
+is a possible line of future work.  
+%Also, it is tedious to
+%express the inheritance relationships in extensions, especially when
+%multiple inheritance is used. Another line of future work is to use
+%some meta-programming mechanisms to eliminate such boilerplate.

### 2016-06-30 paper/sections/Introduction.tex
Add references

@@
 fuzzy line here, and the literature allows for both interpretations. Some authors working on
-OOP EDSLs ~\cite{} consider a {\sc Composite} to be a deep
-embedding. Some other authors~\cite{gibbons2014folding} 
+OOP EDSLs ~\cite{rompf2012scala,scherrc2015} consider a {\sc Composite} to be a deep
+embedding. Some other authors~\cite{gibbons2014folding,barringer2011tracecontract}
 consider implementations using tuples and/or the  {\sc Composite}

### 2017-01-17 paper/sections/Introduction.tex
JFP format

@@
 \begin{quote}
-\emph{With a deep embedding, terms in the DSL are implemented simply to
+With a deep embedding, terms in the DSL are implemented simply to
 construct an abstract syntax tree (AST), which is subsequently
@@
 shallow embedding, terms in the DSL are implemented directly by
-their semantics, bypassing the intermediate AST and its traversal.}\cite{gibbons2014folding}
+their semantics, bypassing the intermediate AST and its traversal.\cite{gibbons2014folding}
 \end{quote}
@@
 still shallow (i.e. using procedural abstraction).
-To understand how
-OOP is helpful for shallow embeddings, we should first look at the
-limitations of shallow embeddings commonly found in the literature:
-\begin{enumerate}
-
-\item {\bf Single Interpretation} An often stated limitation of
-  shallow EDSLs is that they only support one
-  interpretation.
-
-\item {\bf No Transformations} Another commonly stated limitation 
-of shallow embeddings is the lack of support for transformations,
-preventing optimizations.
-
-\end{enumerate}
-
-We show that OOP abstractions, including \emph{
-  inheritance}, \emph{subtyping} and \emph{type-refinement}, are
-helpful to address those problems. For the first problem, we can
-employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
+An often stated limitation of shallow EDSLs is that they only support \emph{single} interpretation.
+We show that OOP abstractions, including \emph{inheritance}, \emph{subtyping} and \emph{type-refinement}, are
+helpful to address this problem.
+We can employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
 solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
@@
   interpretations} to co-exist and be combined in shallow embeddings.
-For the second problem, we show that transformations can be encoded 
-with recursive objects. Interestingly enough, for transformations it
-is possible to port back the OO solution to
-Haskell.
 

### 2017-01-17 paper/code/ContextSensitive.hs

@@ -1,0 +1,19 @@
+--BEGIN_TLAYOUT_HS
+lzw::(a -> a -> a) -> [a] -> [a] -> [a]
+lzw f [ ] ys            = ys
+lzw f xs [ ]            = xs
+lzw f (x : xs) (y : ys) = f x y : lzw f xs ys
+
+type Layout = [[(Int, Int)]]
+type Circuit = (Int,(Int -> Int) -> Layout)
+id n         = (n,\f -> [])
+fan n        = (n,\f -> [[(f 0,f j) | j <- [1..n-1]]])
+above c1 c2  = (width c1,\f -> tlayout c1 f ++ tlayout c2 f)
+beside c1 c2 = (width c1 + width c2
+               ,\f -> lzw (++) (tlayout c1 f) (tlayout c2 ((width c1+) . f)))
+stretch ns c = (sum ns,\f -> tlayout c (pred . (vs!!) . f))
+  where vs = scanl1 (+) ns
+
+width  = fst
+tlayout = snd
+--END_TLAYOUT_HS

### 2017-02-11 ICFP2017/Main.tex
Switch to ICFP format

\documentclass[acmlarge, anonymous, review]{acmart}\settopmatter{printfolios=true}

\usepackage{xspace}
\usepackage{url}
\usepackage{syntax}
\newcommand{\dsl}{\textsc{Scans}\xspace}
\newcommand{\interp}{\textsc{Interpreter}\xspace}
\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
\newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}

%% for comments
\input{Preamble}

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

% Document starts
\begin{document}

% Title portion
\title{Shallow EDSLs and Object-Oriented Programming}
\subtitle{Functional Pearl}
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

\begin{abstract}
Shallow Embedded Domain Specific Languages (EDSLs) use
\emph{procedural abstraction} to directly encode a DSL into an existing host language. Procedural abstraction has
been argued to be the essence of Object-Oriented Programming (OOP).
%Given
%that OO languages have evolved over more than 50 years
%to improve the use of procedural abstraction, they ought to have some
%advantages to encode shallow EDSLs.
This paper argues that OOP abstractions
(including \emph{inheritance}, \emph{subtyping} and
\emph{type-refinement})
increase the modularity and reuse of shallow
EDSLs when compared to classical procedural abstraction. We make this
argument by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple shallow EDSL, and we recode
that EDSL in Scala. From the \emph{semantic}
and \emph{modularity} point of view the Scala version has clear advantages
over the Haskell version.
\end{abstract}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}

\keywords{keyword1, keyword2, keyword3}

\maketitle

\input{sections/Introduction}

\input{sections/ShallowOO}

\input{sections/Interpretations}

\input{sections/Casestudy}

% Bibliography
\bibliography{Main}

% History dates
%\received{February 2007}{March 2009}{June 2009}

\end{document}

### 2017-02-20 paper/sections/Introduction.tex
local

--- 2016-06-30 paper/sections/Introduction.tex
@@
 
-Since Hudak's seminal paper on Embedded Domain Specific Languages (EDSLs)~\cite{hudak1998modular}, existing
-languages (e.g. Haskell) have been used to directly encode
+Since Hudak's seminal paper on Embedded DSLs (EDSLs)~\cite{hudak1998modular}, existing
+languages (such as Haskell) have been used to directly encode
 DSLs. Two common approaches to EDSLs are the so-called \emph{shallow}
@@
 
-\noindent Although the above definition is quite reasonable and widely accepted,
+Although the above definition is quite reasonable and widely accepted,
 it leaves some space to (mis)interpretation. For example it is unclear 
 how to classify an EDSL implemented using the {\sc Composite} or {\sc Interpreter} 
-patterns in Object-Oriented Programming (OOP). Would this OO approach be
-classified as a shallow or deep embedding? We feel there is a rather
-fuzzy line here, and the literature allows for both interpretations. Some authors working on
-OOP EDSLs ~\cite{} consider a {\sc Composite} to be a deep
-embedding. Some other authors~\cite{gibbons2014folding} 
-consider implementations using tuples and/or the  {\sc Composite}
-pattern to be a shallow embedding.
-
-To avoid ambiguitity we propose defining shallow embeddings as EDSLs implemented using \emph{procedural abstraction}~\cite{reynolds94proceduralabstraction}. Such
-interpretation arises naturally from the domain of shallow EDSLs being
-functions, and procedural abstraction being a way to encode
-data abstractions using functions. As Cook~\cite{cook09abstraction} argued,
-procedural abstraction is also the essence of OOP.
-Thus, according to our definition, the implementation of a shallow
-EDSL in OOP languages should simply correspond to a standard
-object-oriented program.
-
-The main goal of our paper is to show the close relationship between 
-shallow embeddings and OOP, and argue that OOP languages have
-advantages for the implementation of shallow embeddings. 
-It is perhaps partly due to those advantages that some authors 
-have considered  {\sc Composite}-based implementations to be deep
-embeddings. However, we argue that the OOP
-mechanisms do not change the essence of the implementation, which is
-still shallow (i.e. using procedural abstraction). 
-To understand how
-OOP is helpful for shallow embeddings, we should first look at the
-limitations of shallow embeddings commonly found in the literature:
-\begin{enumerate}
-
-\item {\bf Single Interpretation} An often stated limitation of
-  shallow EDSLs is that they only support one
-  interpretation.
-
-\item {\bf No Transformations} Another commonly stated limitation 
-of shallow embeddings is the lack of support for transformations,
-preventing optimizations.
-
-\end{enumerate}
-
-We show that OOP abstractions, including \emph{
-  inheritance}, \emph{subtyping} and \emph{type-refinement}, are
-helpful to address those problems. For the first problem, we can
-employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
-solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
-using just standard OOP mechanisms enables \emph{multiple modular
-  interpretations} to co-exist and be combined in shallow embeddings.
-For the second problem, we show that transformations can be encoded 
-with recursive objects. Interestingly enough, for transformations it
-is possible to port back the OO solution to
-Haskell.
-
-We make our arguments by taking a recent paper by Gibbons and Wu~\cite{gibbons2014folding},
-where procedural abstraction is used in Haskell to model a simple \emph{shallow}
-EDSL, and we recode that EDSL in Scala\footnote{Available online: \url{https://github.com/wxzh/shallow-dsl}}.
-From the \emph{modularity} point of view the
-Scala version has clear advantages over the Haskell version.
-
-\begin{comment}
-In summary, our contributions are:
-
-\begin{itemize}
-
-\item {}
-
-\item {\bf Multiple Modular Interpretations for Shallow Embeddings:} 
-  We show that with standard OOP mechanisms it is easy to support multiple modular
-  interpretations for shallow embeddings.
-
-\item {\bf Transformations for Shallow Embeddings:} We show that
-  transformations are encodable with recursive objects. Moreover, this technique
-  can be ported back into functional programming as well.
-
-\end{itemize}
-\end{comment}
-
-\begin{comment}
-If we accept Cook's view on OOP, 
-a natural way to distinguish implementations of EDSLs is 
-in terms of the data abstraction used to model the language
-constructs instead. As Reynold's~\cite{reynolds94proceduralabstraction} suggested there are two
-types of data abstraction: procedural abstraction and \emph{user-defined
-  types}. It is clear that shallow embeddings use \emph{procedural
-  abstraction}: the DSLs are modelled by interpretation
-functions. Thus, the other implementation option for EDSLs is to
-use \emph{user-defined types}. In Reynolds terminology user-defined
-types mean disjoint union types, which are nowadays commonly available
-in modern languages as \emph{algebraic datatypes}. Disjoint union
-types can also be emulated in OOP using the {\sc Visitor} pattern.
-
-A distinction based on data abstraction is more precise and provides a
-remedy for possible misinterpretation. An EDSL implemented with
-algebraic datatypes falls into the category of user-defined types
-(deep embedding), while a {\sc Composite}-based OO implementation falls
-under procedural abstraction (shallow embedding). 
-\end{comment}
-
-%At least some authors~\cite{} seem to implicitly
-%agree with the latter interpretation.
-
-%The {\sc Composite} or {\sc Interpreter}
-%patterns are normally accepted to provide a way to encode ASTs. Thus, 
-%one possible interpretation is that \emph{according to definition of deep embedding
-%  above, the OO approach classifies as a deep
-%  embedding}.
+patterns in OOP. Would this OO approach be
+classified as a shallow or deep embedding? We believe arguments can be
+made both ways. The {\sc Composite} or {\sc Interpreter}
+patterns are normally accepted to provide a way to encode ASTs. Thus, 
+one possible interpretation is that \emph{according to definition of deep embedding
+  above, the OO approach classifies as a deep
+  embedding}.
 
@@
 \end{comment}
+
+In this paper, however, we prefer to interpret shallow embeddings
+differently. We argue that a straightforward OO approach is really a
+shallow embedding. The reason is that shallow embeddings are 
+implemented using \emph{procedural abstraction} and, as Cook~\cite{cook09abstraction} argued,
+procedural abstraction is the essence of Object-Oriented
+Programming. Thus the implementation of a shallow
+DSL in OOP languages should simply correspond to a standard
+object-oriented program.
+
+\begin{comment}
+If we accept Cook's view on OOP, 
+a natural way to distinguish implementations of EDSLs is 
+in terms of the data abstraction used to model the language
+constructs instead. As Reynold's~\cite{reynolds94proceduralabstraction} suggested there are two
+types of data abstraction: procedural abstraction and \emph{user-defined
+  types}. It is clear that shallow embeddings use \emph{procedural
+  abstraction}: the DSLs are modelled by interpretation
+functions. Thus, the other implementation option for EDSLs is to
+use \emph{user-defined types}. In Reynolds terminology user-defined
+types mean disjoint union types, which are nowadays commonly available
+in modern languages as \emph{algebraic datatypes}. Disjoint union
+types can also be emulated in OOP using the {\sc Visitor} pattern.
+
+A distinction based on data abstraction is more precise and provides a
+remedy for possible misinterpretation. An EDSL implemented with
+algebraic datatypes falls into the category of user-defined types
+(deep embedding), while a {\sc Composite}-based OO implementation falls
+under procedural abstraction (shallow embedding). 
+\end{comment}
+
+%At least some authors~\cite{} seem to implicitly
+%agree with the latter interpretation.
+
+The main goal of this paper is to argue that OOP languages have
+advantages for the implementation of shallow embeddings. This should
+not come as a surprise, as OOP languages have evolved over more than 50
+years to improve the use of procedural abstraction.
+
+In order to understand how OOP can help with shallow embeddings, lets
+first review some of the limitations commonly
+found in the literature:
+
+\begin{enumerate}
+
+\item {\bf Single Interpretation} An often stated limitation of
+  shallow embeddings is that they only support a single
+  interpretation.
+
+\item {\bf No Transformations} Another commonly stated limitation 
+of shallow embeddings is that they do not support transformations,
+preventing optimizations and other useful transformations.
+
+\end{enumerate}
+
+\noindent Both limitations are often used as motivators to switch to deep embeddings.
+
+We show that OOP abstractions, including \emph{(multiple)
+  inheritance}, \emph{subtyping} and \emph{type-refinement}, are
+helpful to address those problems. For the first problem, we can
+employ a recently proposed design pattern~\cite{eptrivially16}, which provides a simple
+solution to the \emph{Expression Problem}~\cite{expPb} in OOP languages. Thus
+using just standard OOP mechanims enables \emph{multiple modular
+  interpretations} to co-exist and be combined in shallow embeddings.
+For the second problem, we show that transformations can be encoded 
+with recursive objects. Interestingly enough, for transformations it
+is possible to port back the OO solution to
+Haskell.
+
+We make our arguments by taking a recent paper by Gibbons and Wu~\cite{gibbons2014folding},
+where procedural abstraction is used in Haskell to model a simple
+EDSL, and we recode that EDSL in Scala. Although from the
+\emph{syntactical} point of view there are minor inconveniences in
+the Scala version, from the \emph{semantic} and \emph{modularity} point of view the
+Scala version has clear advantages.
+
+\begin{comment}
+In summary, our contributions are:
+
+\begin{itemize}
+
+\item {}
+
+\item {\bf Multiple Modular Interpretations for Shallow Embeddings:} 
+  We show that with standard OOP mechanisms it is easy to support multiple modular
+  interpretations for shallow embeddings.
+
+\item {\bf Transformations for Shallow Embeddings:} We show that
+  transformations are encodable with recursive objects. Moreover, this technique
+  can be ported back into functional programming as well.
+
+\end{itemize}
+\end{comment}

### 2017-02-20 paper/paper.tex

@@
-\documentclass[10pt,preprint,numbers,nocopyrightspace]{sigplanconf}
+\documentclass{jfp1}
 
@@
 
-\usepackage{amsmath}
 \usepackage{color}
 \usepackage{listings}
-\usepackage{balance}
-\usepackage{amsmath}
-\usepackage{stmaryrd}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{fancyvrb}
 \usepackage{url}
-\usepackage{pstricks,pst-node,pst-tree}
-\usepackage{theorem}
-\usepackage{bbm}
-\usepackage{pgf}
-\usepackage{multirow}
-\usepackage{rotating}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{wrapfig}
 \usepackage{xspace}
-\usepackage{mdwlist}
-\usepackage{morefloats}
-\usepackage[T1]{fontenc}
-\usepackage[scaled=0.85]{beramono}
-\usepackage{mathptmx}
-\usepackage[normalem]{ulem}
-\usepackage{booktabs}
-\usepackage{tablefootnote}
-\usepackage{enumitem}
-\usepackage{stmaryrd}
 \usepackage{syntax}
+\usepackage{comment}
 
-\setlist[itemize]{noitemsep,nolistsep}
-\setlist[enumerate]{noitemsep,nolistsep}
-\setlength{\floatsep}{6.0pt plus 2.0pt minus 2.0pt}
-\setlength{\textfloatsep}{10.0pt plus 2.0pt minus 4.0pt}
-\setlength{\intextsep}{6.0pt plus 2.0pt minus 2.0pt}
-
-\newcommand{\name}{\textbf{@Family}\xspace}
 \newcommand{\dsl}{\emph{scans}\xspace}
@@
 
-\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
-\copyrightyear{20yy}
-\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
-\copyrightdoi{nnnnnnn.nnnnnnn}
-
 % Uncomment the publication rights you want to use.
@@
 
-\titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{Shallow EDSLs and Object-Oriented Programming}   % 'preprint' option specified.
+\title{Shallow EDSLs and Object-Oriented Programming}
+\subtitle{(Functional Pearl)}
 
-\title{Shallow EDSLs and Object-Oriented Programming}
-
-\authorinfo{Weixin Zhang \and Bruno C.d.S.Oliveira}
-           {The University of Hong Kong, Pokfulam, Hong Kong, China}
-           {\{wxzhang2,bruno\}@cs.hku.hk}
+\author[W. Zhang and B. C. d. S. Oliveira]
+        {WEIXIN ZHANG and BRUNO C. d. S. OLIVEIRA\\
+         The University of Hong Kong, Hong Kong\\
+         \email{bruno@cs.hku.hk}}
 %\authorinfo{Name}
@@
 \input{sections/Interpretations.tex}
-\input{sections/Transformations.tex}
+%\input{sections/Transformations.tex}
 %\input{sections/RelatedWorkShort.tex}

### 2017-02-20 paper/sections/Interpretations.tex

@@
 
+\subsection{Context-Sensitive Interpretations}
+Interpretations may rely on some mutable contexts.
+Consider an interpretation that simplifies the representation of a circuit.
+A circuit can be divided horizontally into layers.
+Each layer can be represented as a sequence of pairs $(i,j)$, denoting the connection from wire $i$ to wire $j$.
+For instance, circuit
+$$above\ (beside\ (fan\ 2)\ (fan\ 2))\ (stretch\ [2,2]\ (fan\ 2))$$
+can be represented as:
+$$[[(0,1), (2,3)], [(1,3)]]$$
+It has two layers: the first layer has connections from
+the first wire to the second, and from the second to the third; the second layer has
+only one connection from the first wire to the third one.
+
+\paragraph{Context-sensitive Interpretations in Haskell}
+The following Haskell code models the interpretation described above:
+\lstinputlisting[linerange=2-18]{./code/ContextSensitive.hs}%APPLY:TLAYOUT_HS
+\lstinline{tlayout} is firstly a dependent interpretation, relying on the
+previously defined \lstinline{width} interpretation.
+More importantly, it is a context-sensitive interpretation.
+The domain of \lstinline{tlayout} is not a value but a function \lstinline{(Int -> Int) -> Layout}.
+The argument \lstinline{f} is also a function that shifts the index of a wire.
+\lstinline{f} varies in recursive calls for accumulating transformations
+on wires and hence is context-sensitive.
+
+\paragraph{Context-sensitive Interpretations in Scala}
+Context-sensitive interpretations in our OO approach are unproblematic as well.
+Mutable contexts can be captured as arguments of methods:
+\begin{lstlisting}
+type Layout = List[List[Tuple2[Int,Int]]]
+
+def lzw[A](xs: List[A], ys: List[A])(f: (A, A) => A): List[A] = (xs, ys) match {
+  case (Nil,_) => ys
+  case (_,Nil) => xs
+  case (x::xs,y::ys) => f(x,y)::lzw(xs,ys)(f)
+}
+def partialSum(ns: List[Int]): List[Int] = ns.scanLeft(0)(_ + _).tail
+
+trait Circuit extends width.Circuit {
+  def tlayout(f: Int => Int): Layout
+}
+trait Id extends Circuit with width.Id {
+  def tlayout(f: Int => Int) = List()
+}
+trait Fan extends Circuit with width.Fan {
+  def tlayout(f: Int => Int) = List(for (i <- List.range(1,n)) yield (f(0),f(i)))
+}
+trait Above extends Circuit with width.Above {
+  val c1, c2: Circuit
+  def tlayout(f: Int => Int) = c1.tlayout(f) ++ c2.tlayout(f)
+}
+trait Beside extends Circuit with width.Beside {
+  val c1, c2: Circuit
+  def tlayout(f: Int => Int) =
+    lzw (c1.tlayout(f), c2.tlayout(f.andThen(c1.width + _))) (_ ++ _)
+}
+trait Stretch extends Circuit with width.Stretch {
+  val c: Circuit
+  def tlayout(f: Int => Int) = c.tlayout(f.andThen(partialSum(ns)(_) - 1))
+}
+\end{lstlisting}
 \subsection{Discussion}

### 2017-02-20 paper/sections/Overview.tex

@@
 Their types are given below:
-\lstinputlisting[language=haskell,linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\lstinputlisting[language=haskell,linerange=29-30]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 
@@
 
-\lstinputlisting[language=haskell,linerange=39-40]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
+\lstinputlisting[language=haskell,linerange=29-30]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
 \emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
@@
 \texttt{stretch} and \texttt{above}, for the semantic function:
-\lstinputlisting[language=haskell,linerange=44-45]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
+\lstinputlisting[language=haskell,linerange=34-35]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
 Similarly, defining two new interfaces

### 2017-02-20 paper/sections/RelatedWorkShort.tex

@@
 \paragraph{Deep and Shallow Embeddings}
-Shallow embedding yields flexible and concise EDSLs while deep embedding makes
-it easy to define optimizations.
-There are a lot of work~\cite{svenningsson2012combining,
-  Jovanovic:2014:YCD:2658761.2658771, scherr2014implicit} trying to blend these two
-approaches to enjoy benefits from both.
-They typically encode the surface language with a shallow embedding and
-then generate or translate to a deep embedded version for allowing optimizations.
-%Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at the same time so that clients can choose for a particular interpretation;
-The OO approach we present retains the simplicity of shallow embedding while
-makes it possible to implement optimizations.
-
-\paragraph{Modularity of EDSLs}
-Languages that solves the EP are capable of being host languages for
-implementing modular EDSLs. But many of these solutions require sophisticated
-type parameterization, advanced features or heavy encoding which impose
-complexity to the DSL implementers or even clients, preventing practical use.
-For example, solutions based on the \textsc{Visitor}
-pattern~\cite{oliveira09modular,hofer2010modular} %more references
-Relatively simple solution like Object Algebras~\cite{bruno12oa} may not support
-optimizations well.
-Our approach, however, is simple and supports transformations.

### 2017-02-20 paper/sections/Transformations.tex

@@
 The \emph{id} construct is an example of syntactic sugar:
-\vspace{-8pt}
-$$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
-The formula states that \emph{id n} can be rewritten as \emph{n} of \emph{fan 1}
+\vspace{-5pt}
+$$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{
+    times}}\ (fan\ 1)$$
+\noindent The formula states that \emph{id n} can be rewritten as \emph{n} of \emph{fan 1}
 combined with \emph{beside}.
 
+\vspace{-3pt}
 \paragraph{Desugaring in Scala}
@@
 %desugaring?}
-\paragraph{Desugaring in Haskell}
+\vspace{-3pt}\paragraph{Desugaring in Haskell}
 Interestingly we can port the implementation back to Haskell:

### 2017-02-26 src/scans/Scans.hs
Revision

@@ -47,7 +47,7 @@ identity4 n    =  (n,\f -> [])
 fan4 n         =  (n,\f -> [[(f 0,f j) | j <- [1..n-1]]])
 above4 c1 c2   =  (width c1,\f -> tlayout c1 f ++ tlayout c2 f)
 beside4 c1 c2  =  (width c1 + width c2,\f -> lzw (++) (tlayout c1 f) (tlayout c2 (f . (width c1+))))
-stretch4 ns c  =  (sum ns,\f -> tlayout c (f . pred . (vs!!)))
+stretch4 ns c  =  (sum ns,\f -> tlayout c (f . pred . (vs !!)))
   where vs = scanl1 (+) ns
 
 lzw                      ::  (a -> a -> a) -> [a] -> [a] -> [a]

### 2017-02-26 src/scans/Scans.scala

@@ -1,6 +1,6 @@
 package scans
 
-object Scans extends App {
+object Scans {
 trait Circuit1 {
   def width: Int
 }

### 2017-02-28 ICFP2017/Main.tex
Add old conclusion
Intro
Revision
Up to Sec 2

@@
 \newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
+%\newcommand{\an}[3]{}
 \newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}
+\newcommand{\bruno}[1]{\an{bruno}{blue}{#1}}
 
@@
 %advantages to encode shallow EDSLs.
-This paper argues that OOP abstractions
-(including \emph{inheritance}, \emph{subtyping} and
+This pearl argues that OOP abstractions
+(including \emph{inheritance}, \emph{subtyping}, and
 \emph{type-refinement})
@@
 that EDSL in Scala. From the \emph{semantic}
-and \emph{modularity} point of view the Scala version has clear advantages
+and \emph{modularity} point of view, the Scala version has clear advantages
 over the Haskell version.
@@
 
+\input{sections/Conclusion}
+
 % Bibliography

### 2017-02-28 ICFP2017/Main.tex
local

--- 2017-02-11 ICFP2017/Main.tex
@@
 abstraction is used in Haskell to model a simple shallow EDSL, and we recode
-that EDSL in Scala. From the \emph{semantic}
-and \emph{modularity} point of view the Scala version has clear advantages
-over the Haskell version.
+that EDSL in Scala. 
+To further illustrate the applicability of our OO approach, we conduct
+a case study on refactoring an existing deep embedded SQL DSL implementation to make it
+shallow, more modular and embedded.
 \end{abstract}

### 2017-02-28 ICFP2017/Main.tex
Abstract

--- 2017-02-28 ICFP2017/Main.tex
@@
 \newcommand{\interp}{\textsc{Interpreter}\xspace}
-\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
-%\newcommand{\an}[3]{}
+%\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
+\newcommand{\an}[3]{}
 \newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}
@@
 abstraction is used in Haskell to model a simple shallow EDSL, and we recode
-that EDSL in Scala. From the \emph{semantic}
-and \emph{modularity} point of view, the Scala version has clear advantages
-over the Haskell version.
+that EDSL in Scala. 
+To further illustrate the applicability of our OO approach, we conduct
+a case study on refactoring a deep external SQL implementation to make it
+more modular, shallow, and embedded.
 \end{abstract}

### 2017-02-28 ICFP2017/Main.tex
DSL

@@
 To further illustrate the applicability of our OO approach, we conduct
-a case study on refactoring a deep external SQL implementation to make it
+a case study on refactoring a deep external SQL DSL implementation to make it
 more modular, shallow, and embedded.

### 2017-02-28 ICFP2017/Main.tex
Conclusion

--- 2017-02-28 ICFP2017/Main.tex
@@
 that EDSL in Scala. 
-To further illustrate the applicability of our OO approach, we conduct
-a case study on refactoring a deep external SQL implementation to make it
+To further illustrate the applicability of our OOP approach, we conduct
+a case study on refactoring a deep external SQL DSL implementation to make it
 more modular, shallow, and embedded.

### 2017-04-18 ICFP2017/response.md
Add response & small revisions

We thank all the reviewers for their constructive comments.

Review A
===

- if you support multiple interpretations, then how do I write DSL terms, without committing to a particular interpretation of the DSL?

There are different options to build terms:

1) An easy way that offers some weak form of modularity but not independent from multiple interpretations is to provide smart constructors of same signatures in different versions of DSLs. For example,

```
object Circuit1 {
  def fan(x: Int) = new Fan1 {val n=x}
  def beside(x: Circuit1, y: Circuit1) = new Beside1 {val c1=x; val c2=y}
  ...
}

object Circuit2 {
  def fan(x: Int) = new Fan2 {val n=x}
  def beside(x: Circuit2, y: Circuit2) = new Beside2 {val c1=x; val c2=y}
  ...
}
```
By switching import statements from `Circuit1` to `Circuit2`, we can have a circuit that supports different interpretations without touching the construction code

2) Object Algebras are complementary to our own techniques.
By building on the strength of Object Algebras, we can have terms that are independent from interpretations.
Here is the Object Algebra interface for the circuit DSL:

```
trait CircuitFactory[C] {
  def fan(x: Int): C
  def beside(x: C, y: C): C
  ...
}
```

This interface can be used to construct interpretation-independent terms:

```
def mkCircuit[C](f: CircuitFactory[C]): C = f.beside(f.fan(2),f.fan(2))
```

Now, smart constructor definitions are moved to a concrete factory:

```
object Circuit1Factory extends CircuitFactory[Circuit1] {
  def fan(x: Int) = new Fan1 {val n=x}
  def beside(x: Circuit1, y: Circuit1) = new Beside1 {val c1=x; val c2=y}
  ...
}
object Circuit2Factory extends CircuitFactory[Circuit2] {
  def fan(x: Int) = new Fan2 {val n=x}
  def beside(x: Circuit1, y: Circuit1) = new Beside2 {val c1=x; val c2=y}
  ...
}
```

Similar concrete factories can be defined for `Circuit2`.
By supplying different concrete factories to `mkCircuit`, we obtain circuits that support different  interpretations.

Review B
===

Review C
===

- "Scala inheritance is used to save a tiny bit of redefinition"
inheritance allows reuse of existing interpretations
subtyping allows new interpretations

- This paper strangely (to me) spends 2 sentences acknowledging that the approach is a bad fit for DSLs that are used with program transformation -- as if that were not a crucial feature for many DSL domains, including the most realistic one treated in this paper!

- The OOP style demonstrated here forces a linearization of the method signatures -- each new widget inherits from a particular old one, even if the two widgets in question define two methods that aren't mutually recursive.

No. For example, `Circuit2` shown on page 6 can be defined separately:

```
trait Circuit2 {
  def depth: Int
}
trait Identity2 extends Circuit2 {
  def depth = 0
}
...
```

We can merge `Circuit1` and `Circuit2` via multiple trait inheritance:

```
trait Circuit extends Circuit1 with Circuit2
trait Identity extends Circuit12 with Idenity1 with Identity2
...
```

### 2017-05-02 JFP/Main.tex
JFP format

--- 2017-02-28 ICFP2017/Main.tex
@@
-\documentclass[acmlarge, anonymous, review]{acmart}\settopmatter{printfolios=true}
+\documentclass{jfp1}
 
@@
 \usepackage{syntax}
+\usepackage{textcomp}
 \newcommand{\dsl}{\textsc{Scans}\xspace}
@@
 
-\makeatletter\if@ACM@journal\makeatother
-%% Journal information (used by PACMPL format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{1}
-\acmArticle{1}
-\acmYear{2017}
-\acmMonth{1}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\else\makeatother
-%% Conference information (used by SIGPLAN proceedings format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
-\acmYear{2017}
-\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\fi
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission
-\setcopyright{none}             %% For review submission
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2017}           %% If different from \acmYear
-
 %% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
+\bibliographystyle{jfp}
 
@@
 \title{Shallow EDSLs and Object-Oriented Programming}
-\subtitle{Functional Pearl}
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
+
+\author[W. Zhang and B. Oliveira]
+        {WEIXIN ZHANG and BRUNO C. D. S. OLIVEIRA\\
+         The University of Hong Kong, Hong Kong}
+
+\maketitle[f]
 
@@
 
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
-
-\keywords{keyword1, keyword2, keyword3}
-
-\maketitle
-
 \input{sections/Introduction}

### 2017-07-09 JFP/Main.tex
Draft for interpretation-independent terms
Revision

@@
 \usepackage{textcomp}
+\usepackage{color}
 \newcommand{\dsl}{\textsc{Scans}\xspace}
@@
 \begin{abstract}
-Shallow Embedded Domain Specific Languages (EDSLs) use
+Shallow embedded domain-specific languages (EDSLs) use
 \emph{procedural abstraction} to directly encode a DSL into an existing host language. Procedural abstraction has
-been argued to be the essence of Object-Oriented Programming (OOP).
+been argued to be the essence of object-oriented programming (OOP).
 %Given
@@
 
+\input{sections/ObjectAlgebras}
+
 \input{sections/Casestudy}

### 2017-11-28 src/scans/Factory.scala
Table in text & Clean up refs

@@ -45,5 +45,5 @@ trait ExtendedFactory4 extends ExtendedFactory[Circuit4] with Factory4 {
 
 def c2[Circuit](f: ExtendedFactory[Circuit]) = f.rstretch(c(f),2,2,2,2)
 
-//println(new ExtendedTerm with ExtenededFactory4{}.c2.tlayout { x => x }) // List(List((1,3), (5,7)), List((3,7)), List((3,5)))
+println(c2(new ExtendedFactory4{}).tlayout { x => x }) // List(List((1,3), (5,7)), List((3,7)), List((3,5)))
 }

### 2017-11-28 JFP/Main.tex
A new section on AbstractFactory
Intro
Revised Intro
Section 3
Fix references
Table in text & Clean up refs
Side-by-side figures
Spellcheck

@@
 
+\usepackage{caption}
+\usepackage{wrapfig}
 \usepackage{xspace}
+\usepackage{xypic}
+%\usepackage{amsmath}
 \usepackage{url}
@@
 \usepackage{color}
+\usepackage{comment}
 \newcommand{\dsl}{\textsc{Scans}\xspace}
 \newcommand{\interp}{\textsc{Interpreter}\xspace}
-%\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
-\newcommand{\an}[3]{}
+\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
+%\newcommand{\an}[3]{}
 \newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}
 \newcommand{\bruno}[1]{\an{bruno}{blue}{#1}}
+
+\let\oldpar\paragraph
+\renewcommand{\paragraph}[1]{\vspace{-7pt}\oldpar{#1}}
+%\let\oldsec\section
+%\renewcommand{\section}[1]{\vspace{-7pt}\oldsec{#1}}
+\let\oldsubsec\subsection
+\renewcommand{\subsection}[1]{\vspace{-7pt}\oldsubsec{#1}}
 
@@
 % Title portion
-\title{Shallow EDSLs and Object-Oriented Programming}
+\title{Shallow EDSLs and Object-Oriented Programming: Beyond Simple Compositionality}
 
@@
 a case study on refactoring a deep external SQL DSL implementation to make it
-more modular, shallow, and embedded.
+more modular, shallow, and embedded without sacrificing the performance.%\bruno{Compositionality beyond Folds}
 \end{abstract}
@@
 
-\input{sections/ObjectAlgebras}
+\input{sections/Factory}
 
@@
 % Bibliography
+%\renewcommand{\section}[1]{\oldsec{#1}}
+
+\vspace{-7pt}
 \bibliography{Main}

### 2017-11-28 JFP/Main.tex
Section & paragraph style

@@
 a case study on refactoring a deep external SQL DSL implementation to make it
-more modular, shallow, and embedded without sacrificing the performance.%\bruno{Compositionality beyond Folds}
+more modular, shallow, and embedded without sacrificing performance.%\bruno{Compositionality beyond Folds}
 \end{abstract}

### 2017-11-28 src/scans/Circuit4.scala

@@ -1,29 +1,28 @@
 package scans
 
 object Circuit4 {
-type Layout = List[List[(Int,Int)]]
 trait Circuit4 extends Circuit1 {
-  def tlayout(f: Int => Int): Layout
+  def layout(f: Int => Int): List[List[(Int,Int)]]
 }
 trait Id4 extends Id1 with Circuit4 {
-  def tlayout(f: Int => Int) = List()
+  def layout(f: Int => Int) = List()
 }
 trait Fan4 extends Fan1 with Circuit4 {
-  def tlayout(f: Int => Int) = List(for (i <- List.range(1,n)) yield (f(0),f(i)))
+  def layout(f: Int => Int) = List(for (i <- List.range(1,n)) yield (f(0),f(i)))
 }
 trait Above4 extends Above1 with Circuit4 {
   val c1, c2: Circuit4
-  def tlayout(f: Int => Int) = c1.tlayout(f) ++ c2.tlayout(f)
+  def layout(f: Int => Int) = c1.layout(f) ++ c2.layout(f)
 }
 trait Beside4 extends Beside1 with Circuit4 {
   val c1, c2: Circuit4
-  def tlayout(f: Int => Int) = lzw (c1.tlayout(f), c2.tlayout(f.compose(c1.width + _))) (_ ++ _)
+  def layout(f: Int => Int) = lzw (c1.layout(f), c2.layout(f.compose(c1.width + _))) (_ ++ _)
 }
 trait Stretch4 extends Stretch1 with Circuit4 {
   val c: Circuit4
-  def tlayout(f: Int => Int) = {
+  def layout(f: Int => Int) = {
     val vs = ns.scanLeft(0)(_ + _).tail
-    c.tlayout(f.compose(vs(_) - 1))
+    c.layout(f.compose(vs(_) - 1))
   }
 }
 
@@ -34,9 +33,9 @@ def lzw[A](xs: List[A], ys: List[A])(f: (A, A) => A): List[A] = (xs, ys) match {
 }
 
 trait RStretch extends Stretch4 {
-  override def tlayout(f: Int => Int) = {
+  override def layout(f: Int => Int) = {
     val vs = ns.scanLeft(ns.last - 1)(_ + _).init
-    c.tlayout(f.compose(vs(_)))
+    c.layout(f.compose(vs(_)))
   }
 }
 }

### 2017-11-28 src/scans/Factory.scala

@@ -34,7 +34,7 @@ def c[Circuit](f: Factory[Circuit]) =
                         f.beside(f.beside(f.id(1),f.fan(2)),f.id(1))))
 
 println(c(new Factory1{}).width) // 4 
-println(c(new Factory4{}).tlayout { x => x }) // List(List((0,1), (2,3)), List((1,3)), List((1,2)))
+println(c(new Factory4{}).layout { x => x }) // List(List((0,1), (2,3)), List((1,3)), List((1,2)))
 
 trait ExtendedFactory[Circuit] extends Factory[Circuit] {
   def rstretch(x: Circuit, xs: Int*): Circuit
@@ -45,5 +45,5 @@ trait ExtendedFactory4 extends ExtendedFactory[Circuit4] with Factory4 {
 
 def c2[Circuit](f: ExtendedFactory[Circuit]) = f.rstretch(c(f),2,2,2,2)
 
-println(c2(new ExtendedFactory4{}).tlayout { x => x }) // List(List((1,3), (5,7)), List((3,7)), List((3,5)))
+println(c2(new ExtendedFactory4{}).layout { x => x }) // List(List((1,3), (5,7)), List((3,7)), List((3,5)))
 }

### 2017-11-28 src/scans/Scans.hs

@@ -44,9 +44,9 @@ type Layout    =  [[(Int, Int)]]
 type Circuit4  =  (Int,(Int -> Int) -> Layout)
 id4 n          =  (n,\f -> [])
 fan4 n         =  (n,\f -> [[(f 0,f j) | j <- [1..n-1]]])
-above4 c1 c2   =  (width c1,\f -> tlayout c1 f ++ tlayout c2 f)
-beside4 c1 c2  =  (width c1 + width c2,\f -> lzw (++) (tlayout c1 f) (tlayout c2 (f . (width c1+))))
-stretch4 ns c  =  (sum ns,\f -> tlayout c (f . pred . (vs !!)))
+above4 c1 c2   =  (width c1,\f -> layout c1 f ++ layout c2 f)
+beside4 c1 c2  =  (width c1 + width c2,\f -> lzw (++) (layout c1 f) (layout c2 (f . (width c1+))))
+stretch4 ns c  =  (sum ns,\f -> layout c (f . pred . (vs !!)))
   where vs = scanl1 (+) ns
 
 lzw                      ::  (a -> a -> a) -> [a] -> [a] -> [a]
@@ -54,7 +54,7 @@ lzw f [ ] ys             =  ys
 lzw f xs [ ]             =  xs
 lzw f (x : xs) (y : ys)  =  f x y : lzw f xs ys
 
-tlayout =  snd
+layout =  snd
 
 rstretch        ::  [Int] -> Circuit4 -> Circuit4
 rstretch  ns c  =   stretch4 (1 : init ns) c `beside4` id4 (last ns - 1)

### 2017-11-28 src/scans/Scans.scala

@@ -45,6 +45,6 @@ val c  = above(  beside(fan(2),fan(2)),
 println(c.width)
 println(c.depth)
 println(c.wellSized)
-println(c.tlayout(x => x))
+println(c.layout(x => x))
 }
 }

### 2017-12-01 JFP/Main.tex
local changes

@@
 been argued to be the essence of object-oriented programming (OOP).
-%Given
-%that OO languages have evolved over more than 50 years
-%to improve the use of procedural abstraction, they ought to have some
-%advantages to encode shallow EDSLs.
 This pearl argues that OOP abstractions
@@
 that EDSL in Scala. 
-To further illustrate the applicability of our OOP approach, we conduct
+To further illustrate the applicability of our functional OOP approach, we conduct
 a case study on refactoring a deep external SQL DSL implementation to make it
-more modular, shallow, and embedded without sacrificing performance.%\bruno{Compositionality beyond Folds}
+more modular, shallow, and embedded without sacrificing performance.
 \end{abstract}

### 2017-12-05 JFP/Main.tex
Improving Intro

@@
 Shallow embedded domain-specific languages (EDSLs) use
-\emph{procedural abstraction} to directly encode a DSL into an existing host language. Procedural abstraction has
+\emph{procedural abstraction} to directly encode a DSL into an
+existing host language. Procedural abstraction has 
 been argued to be the essence of object-oriented programming (OOP).
@@
 increase the modularity and reuse of shallow
-EDSLs when compared to classical procedural abstraction. We make this
+EDSLs when compared to classical procedural abstraction by enabling a
+simple way to express \emph{multiple, possibly dependent, interpretations}. We make this
 argument by taking a recent paper by Gibbons and Wu, where procedural
@@
 that EDSL in Scala. 
-To further illustrate the applicability of our functional OOP approach, we conduct
+We further  illustrate our functional OOP approach, with
 a case study on refactoring a deep external SQL DSL implementation to make it
-more modular, shallow, and embedded without sacrificing performance.
+more modular, shallow, and embedded.
 \end{abstract}

### 2018-09-07 src/scans/Typeclass.hs
Add a section on Finally Tagless

@@ -1,0 +1,73 @@
+{-# OPTIONS -XTypeSynonymInstances -XFlexibleInstances -XTypeOperators -XMultiParamTypeClasses -XFlexibleContexts -XOverlappingInstances #-}
+import Prelude hiding (id)
+
+class Circuit c where
+  id       ::  Int -> c
+  fan      ::  Int -> c
+  above    ::  c -> c -> c
+  beside   ::  c -> c -> c
+  stretch  ::  [Int] -> c -> c
+
+newtype Width = Width {width :: Int}
+
+instance Circuit Width where
+  id n          =  Width n
+  fan n         =  Width n
+  beside c1 c2  =  Width (width c1 + width c2)
+  above c1 c2   =  Width (width c1)
+  stretch ns c  =  Width (sum ns)
+
+newtype Depth = Depth {depth :: Int}
+
+instance Circuit Depth where
+  id n           =  Depth n
+  fan n          =  Depth n
+  beside c1 c2   =  Depth (depth c1 + depth c2)
+  above c1 c2    =  Depth (depth c1 `max` depth c2)
+  stretch ns c   =  Depth (depth c)
+
+newtype WellSized  = WellSized {wellSized :: Bool}
+
+type Compose i1 i2 = (i1, i2)
+
+class a :<: b where
+  inter :: b -> a
+
+instance a :<: a where
+  inter = \x -> x
+
+instance a :<: (Compose a b) where
+  inter = fst
+
+instance (c :<: b) => c :<: (Compose a b) where
+  inter = inter . snd
+
+
+ instance (Circuit width, Width :<: width) =>
+           Circuit (Compose WellSized width) where
+   id  n         =  (WellSized True, id n)
+   fan n         =  (WellSized True, fan n)
+   above c1 c2   =  (WellSized (gwellSized c1 && gwellSized c2 && gwidth c1 == gwidth c2)
+                    ,above (inter c1) (inter c2))
+   beside c1 c2  =  (WellSized (gwellSized c1 && gwellSized c2)
+                    ,beside (inter c1) (inter c2))
+   stretch ns c  =  (WellSized (gwellSized c && length ns == gwidth c)
+                    ,stretch ns (inter c))
+
+gwidth :: (Width :<: e) => e -> Int
+gwidth = width . inter
+
+gdepth :: (Depth :<: e) => e -> Int
+gdepth = depth . inter
+
[…]

### 2018-09-07 JFP/Main.tex
Revision

@@
-\documentclass{jfp1}
+\documentclass[english,submission]{programming}
 
+\usepackage{mathtools}
 \usepackage{caption}
@@
 \usepackage{xypic}
+\usepackage{hyperref}
 %\usepackage{amsmath}
@@
 
+\renewcommand*{\sectionautorefname}{Section}
+% for code that uses string literal in lhs2TeX
+\DeclareOldFontCommand{\tt}{\normalfont\ttfamily}{\texttt}
+
 \let\oldpar\paragraph
@@
 
-%% Bibliography style
-\bibliographystyle{jfp}
-
 % Document starts
@@
 % Title portion
-\title{Shallow EDSLs and Object-Oriented Programming: Beyond Simple Compositionality}
+\title{Shallow EDSLs and Object-Oriented Programming}
+\subtitle{Beyond Simple Compositionality}
 
-\author[W. Zhang and B. Oliveira]
-        {WEIXIN ZHANG and BRUNO C. D. S. OLIVEIRA\\
-         The University of Hong Kong, Hong Kong}
+\author[a]{Weixin Zhang}
+\author[a]{Bruno C. d. S. Oliveira}
+\affiliation[a]{The University of Hong Kong, Hong Kong, China}
+%\author[W. Zhang and B. Oliveira]
+%        {WEIXIN ZHANG and BRUNO C. D. S. OLIVEIRA\\
+%         The University of Hong Kong, Hong Kong}
 
-\maketitle[f]
+\paperdetails{
+  %% perspective options are: art, sciencetheoretical, scienceempirical, engineering.
+  %% Choose exactly the one that best describes this work. (see 2.1)
+  perspective=art,
+  %% State one or more areas, separated by a comma. (see 2.2)
+  %% Please see list of areas in http://programming-journal.org/cfp/
+  %% The list is open-ended, so use other areas if yours is/are not listed.
+  area={Domain-Specific Languages, Modularity and separation of concerns},
+  %% You may choose the license for your paper (see 3.)
+  %% License options include: cc-by (default), cc-by-nc
+  % license=cc-by,
+}
+
+\maketitle
 
@@
 that EDSL in Scala. 
-We further  illustrate our functional OOP approach, with
+We further illustrate our functional OOP approach with
 a case study on refactoring a deep external SQL DSL implementation to make it
@@
 
+\input{sections/TypeClass}
+
 \input{sections/Factory}
@@
 
-\vspace{-7pt}
 \bibliography{Main}

### 2018-09-07 src/scans/Circuit1.scala

@@ -1,8 +1,10 @@
 package scans
 
+// object interface
 trait Circuit1 {
   def width: Int
 }
+// concrete implementations
 trait Id1 extends Circuit1 {
   val n: Int
   def width = n

### 2018-09-07 src/scans/Circuit2.scala

@@ -1,5 +1,6 @@
 package scans
 
+// extended semantic domain
 trait Circuit2 extends Circuit1 {
   def depth: Int
 }
@@ -10,7 +11,7 @@ trait Fan2 extends Fan1 with Circuit2 {
   def depth = 1
 }
 trait Above2 extends Above1 with Circuit2 {
-  val c1, c2: Circuit2
+  val c1, c2: Circuit2 // type-refinement that allows depth invocations
   def depth = c1.depth + c2.depth
 }
 trait Beside2 extends Beside1 with Circuit2 {

### 2018-09-07 src/scans/Scans.hs

@@ -12,7 +12,9 @@ beside c1 c2   =  c1 + c2
 above c1 c2    =  c1
 stretch ns c   =  sum ns
 
-c  =  ( fan 2 `beside` fan 2) `above` stretch [2,2] (fan 2) `above` (id 1 `beside` fan 2 `beside` id 1)
+c  =  ( fan 2 `beside` fan 2)
+      `above` stretch [2,2] (fan 2)
+      `above` (id 1 `beside` fan 2 `beside` id 1)
 
 newtype Circuit1   =  Circuit1 {width1  ::  Int}
 id1 n              =  Circuit1 {width1  =   n}
@@ -56,3 +58,34 @@ layout =  snd
 
 rstretch        ::  [Int] -> Circuit4 -> Circuit4
 rstretch  ns c  =   stretch4 (1 : init ns) c `beside4` id4 (last ns - 1)
+
+class Scans circuit where
+  idF       ::  Int -> circuit
+  fanF      ::  Int -> circuit
+  aboveF    ::  circuit -> circuit -> circuit
+  besideF   ::  circuit -> circuit -> circuit
+  stretchF  ::  [Int] -> circuit -> circuit
+
+instance Scans Circuit1 where
+  idF = id1
+  fanF = fan1
+  besideF = beside1
+  aboveF = above1
+  stretchF = stretch1
+
+ {-
+newtype Depth = Depth {depth::Int}
+instance Scans Depth where
+  idF = id1
+  fanF = fan1
+  besideF = beside1
+  aboveF = above1
+  stretchF = stretch1
+-}
+
+-- generic term
+c  ::  Scans circuit => circuit
+c  =   ( fanF 2 `besideF` fanF 2) `aboveF`
+       stretchF [2,2] (fanF 2) `aboveF` (idF 1 `besideF` fanF 2 `besideF` idF 1)
+
+w2 = width1 (c2 :: Circuit1)

### 2018-09-07 src/scans/Typeclass.hs

@@ -43,8 +43,8 @@ instance (c :<: b) => c :<: (Compose a b) where
   inter = inter . snd
 
 
- instance (Circuit width, Width :<: width) =>
-           Circuit (Compose WellSized width) where
+instance (Circuit width, Width :<: width) =>
+          Circuit (Compose WellSized width) where
    id  n         =  (WellSized True, id n)
    fan n         =  (WellSized True, fan n)
    above c1 c2   =  (WellSized (gwellSized c1 && gwellSized c2 && gwidth c1 == gwidth c2)
@@ -53,6 +53,12 @@ instance (c :<: b) => c :<: (Compose a b) where
                     ,beside (inter c1) (inter c2))
    stretch ns c  =  (WellSized (gwellSized c && length ns == gwidth c)
                     ,stretch ns (inter c))
+
+class Circuit c => ExtendedCircuit c where
+  rstretch :: [Int] -> c -> c
+
+instance ExtendedCircuit Width where
+  rstretch = stretch
 
 gwidth :: (Width :<: e) => e -> Int
 gwidth = width . inter

### 2018-09-13 Programming19/Main.tex
Abstract

--- 2018-09-07 JFP/Main.tex
@@
 been argued to be the essence of object-oriented programming (OOP).
-This pearl argues that OOP abstractions
+This pearl argues that common OOP mechanisms
 (including \emph{inheritance}, \emph{subtyping}, and
@@
 abstraction is used in Haskell to model a simple shallow EDSL, and we recode
-that EDSL in Scala. 
-We further illustrate our functional OOP approach with
+that EDSL both in Scala and with an improved OO-inspired Haskell encoding. 
+We further illustrate our approach with
 a case study on refactoring a deep external SQL DSL implementation to make it

### 2018-09-18 Programming19/Main.tex
Intro
Abstract

@@
 \newcommand{\interp}{\textsc{Interpreter}\xspace}
-\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
+\newcommand{\an}[3]{{\color{#2} {#1}:#3}}
 %\newcommand{\an}[3]{}
@@
 \newcommand{\bruno}[1]{\an{bruno}{blue}{#1}}
+
+\DeclareOldFontCommand{\bf}{\normalfont\bfseries}{\mathbf}
 
@@
 \begin{abstract}
-Shallow embedded domain-specific languages (EDSLs) use
-\emph{procedural abstraction} to directly encode a DSL into an
-existing host language. Procedural abstraction has 
-been argued to be the essence of object-oriented programming (OOP).
+
+{\bf Context.} 
+Embedded Domain-Specific Languages (EDSLs) are a common and widely
+used approach to
+DSLs in various languages, including Haskell and Scala. There are two
+main implementation techniques for EDSLs: \emph{shallow embeddings}
+and \emph{deep embeddings}.
+
+{\bf Inquiry.} Shallow embeddings are quite simple, 
+but they have been criticized in the past for being quite limited in
+terms of modularity and reuse. In particular it is often argued that 
+shallow embeddings can only support a single DSL interpretation.
+
+{\bf Approach.}
+In this paper we argue that 
+shallow embedded domain-specific languages (EDSLs) and Object-Oriented
+Programming (OOP) are closely related. Gibbons and Wu already
+discussed the relationship between shallow EDSLs and procedural
+abstraction, while Cook discussed the connection between procedural
+abstraction and OOP. We make the transitive step in this paper by
+connecting shallow EDSLs directly to OOP via procedural abstraction. 
+The knowledge about this relationship enables us to
+improve on implementation techniques for EDSLs.
+
+{\bf Knowledge.}
 This pearl argues that common OOP mechanisms
@@
 EDSLs when compared to classical procedural abstraction by enabling a
-simple way to express \emph{multiple, possibly dependent, interpretations}. We make this
-argument by taking a recent paper by Gibbons and Wu, where procedural
+simple way to express \emph{multiple, possibly dependent,
+  interpretations}.
+
+{\bf Grounding}
+We make our
+arguments by taking a recent paper by Gibbons and Wu, where procedural
 abstraction is used in Haskell to model a simple shallow EDSL, and we recode
@@
 more modular, shallow, and embedded.
+
+{\bf Importance.} This work is important for two different
+reasons. Firstly, from an intelectual point of view, this work
+establishes the connection between shallow embeddings and 
+Object-Oriented Programming, which enables a better understanding
+of both concepts. Secondly, this work
+offers new programming techniques that can be used to improve 
+the modularity and reuse of shallow EDSLs.  
 \end{abstract}

### 2018-09-18 Programming19/Main.tex
Abstract

@@
 terms of modularity and reuse. In particular it is often argued that 
-shallow embeddings can only support a single DSL interpretation.
+supporting multiple DSL interpretations in shallow embeddings is difficult.
 
 {\bf Approach.}
-In this paper we argue that 
-shallow embedded domain-specific languages (EDSLs) and Object-Oriented
+This paper argues that shallow EDSLs and Object-Oriented
 Programming (OOP) are closely related. Gibbons and Wu already
@@
 {\bf Knowledge.}
-This pearl argues that common OOP mechanisms
+This paper argues that common OOP mechanisms
 (including \emph{inheritance}, \emph{subtyping}, and
@@
 {\bf Grounding}
-We make our
-arguments by taking a recent paper by Gibbons and Wu, where procedural
-abstraction is used in Haskell to model a simple shallow EDSL, and we recode
-that EDSL both in Scala and with an improved OO-inspired Haskell encoding. 
+We make our arguments by using Gibbons and Wu's examples, where procedural
+abstraction is used in Haskell to model a simple shallow EDSL. We recode that
+EDSL in Scala and with an improved OO-inspired Haskell encoding.
 We further illustrate our approach with
@@
 reasons. Firstly, from an intelectual point of view, this work
-establishes the connection between shallow embeddings and 
-Object-Oriented Programming, which enables a better understanding
-of both concepts. Secondly, this work
+establishes the connection between shallow embeddings and OOP, which enables a
+better understanding of both concepts. Secondly, this work
 offers new programming techniques that can be used to improve 

### 2018-09-24 src/scans/Typeclass.hs
Add references

@@ -34,7 +34,7 @@ class a :<: b where
   inter :: a -> b
 
 instance a :<: a where
-  inter = \x -> x
+  inter x = x
 
 instance (a,b) :<: a where
   inter = fst
@@ -43,7 +43,7 @@ instance (b :<: c) => (a,b) :<: c where
   inter = inter . snd
 
 
-instance (Circuit width, width :<: Width) => Circuit (WellSized, width) where
+instance (Circuit c, c :<: Width) => Circuit (WellSized, c) where
    id  n         =  (WellSized True, id n)
    fan n         =  (WellSized True, fan n)
    above c1 c2   =  (WellSized (gwellSized c1 && gwellSized c2 && gwidth c1 == gwidth c2)

### 2018-09-28 Programming19/Main.tex
Add references
Small fixes and Section name
Hack code indention

@@
 \newcommand{\interp}{\textsc{Interpreter}\xspace}
-\newcommand{\an}[3]{{\color{#2} {#1}:#3}}
-%\newcommand{\an}[3]{}
+\newcommand{\an}[3]{}
+%\newcommand{\an}[3]{{\color{#2} {#1}:#3}}
 \newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}
@@
 
-\renewcommand*{\sectionautorefname}{Section}
+\defcaptionname{english}{\sectionautorefname}{Section}
 % for code that uses string literal in lhs2TeX
@@
 \affiliation[a]{The University of Hong Kong, Hong Kong, China}
-%\author[W. Zhang and B. Oliveira]
-%        {WEIXIN ZHANG and BRUNO C. D. S. OLIVEIRA\\
-%         The University of Hong Kong, Hong Kong}
 
@@
 but they have been criticized in the past for being quite limited in
-terms of modularity and reuse. In particular it is often argued that 
+terms of modularity and reuse. In particular, it is often argued that
 supporting multiple DSL interpretations in shallow embeddings is difficult.
@@
 
-{\bf Grounding}
+{\bf Grounding.}
 We make our arguments by using Gibbons and Wu's examples, where procedural
@@
 
-{\bf Importance.} This work is important for two different
-reasons. Firstly, from an intelectual point of view, this work
+{\bf Importance.} This work is important for two reasons.
+Firstly, from an intellectual point of view, this work
 establishes the connection between shallow embeddings and OOP, which enables a

### 2018-09-29 Programming19/Main.tex
Add ccs

@@
 \input{Preamble}
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008.10011024</concept_id>
+<concept_desc>Software and its engineering~Language features</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011050.10011017</concept_id>
+<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
 
+\ccsdesc[500]{Software and its engineering~Language features}
+\ccsdesc[500]{Software and its engineering~Domain specific languages}
 % Document starts

### 2018-12-13 src/scans/Circuit2.scala
Code

@@ -23,8 +23,8 @@ trait Stretch2 extends Stretch1 with Circuit2 {
 }
 
 trait Circuit12 extends Circuit1 with Circuit2
-trait Id12 extends Id1 with Id2
-trait Fan12 extends Fan1 with Fan2
-trait Above12 extends Above1 with Above2 { val c1, c2: Circuit12 }
-trait Beside12 extends Beside1 with Beside2 { val c1, c2: Circuit12 }
-trait Stretch12 extends Stretch1 with Stretch2 { val c: Circuit12 }
+trait Id12 extends Circuit12 with Id1 with Id2
+trait Fan12 extends Circuit12 with Fan1 with Fan2
+trait Above12 extends Circuit12 with Above1 with Above2 { val c1, c2: Circuit12 }
+trait Beside12 extends Circuit12 with Beside1 with Beside2 { val c1, c2: Circuit12 }
+trait Stretch12 extends Circuit12 with Stretch1 with Stretch2 { val c: Circuit12 }

### 2018-12-17 Programming19/Main.tex
Acknowledgment
Code
Minor changes

@@
   %% The list is open-ended, so use other areas if yours is/are not listed.
-  area={Domain-Specific Languages, Modularity and separation of concerns},
+  area={Domain-Specific Languages, Modularity and Separation of Concerns},
   %% You may choose the license for your paper (see 3.)
@@
 better understanding of both concepts. Secondly, this work
-offers new programming techniques that can be used to improve 
+illustrates programming techniques that can be used to improve 
 the modularity and reuse of shallow EDSLs.  

### 2018-12-17 Programming19/Main.tex
Addressing problems
Table

@@
 %\renewcommand{\section}[1]{\oldsec{#1}}
+\acks We thank Willam R. Cook, Jeremy Gibbons, Ralf Hinze, Martin Erwig, and the
+anonymous reviewers of GPCE, ICFP, JFP, and Programming for their valuable
+comments that significantly improved this work. This work is based on the first
+author's master thesis~\cite{zhang2017extensible} and has been funded by Hong
+Kong Research Grant Council projects number 17210617 and 17258816.
 

### 2019-01-29 Programming19/Main.tex
Fixes for final version

@@
-\documentclass[english,submission]{programming}
+\documentclass[english]{programming}
 
 \usepackage{mathtools}
-\usepackage{caption}
-\usepackage{wrapfig}
-\usepackage{xspace}
 \usepackage{xypic}
-\usepackage{hyperref}
 %\usepackage{amsmath}
-\usepackage{url}
 \usepackage{syntax}
-\usepackage{textcomp}
-\usepackage{color}
-\usepackage{comment}
 \newcommand{\dsl}{\textsc{Scans}\xspace}
@@
 
-\DeclareOldFontCommand{\bf}{\normalfont\bfseries}{\mathbf}
+%\DeclareOldFontCommand{\bf}{\normalfont\bfseries}{\mathbf}
 
 \defcaptionname{english}{\sectionautorefname}{Section}
+\defcaptionname{english}{\figureautorefname}{Figure}
+\defcaptionname{english}{\tableautorefname}{Table}
 % for code that uses string literal in lhs2TeX
@@
 
-\let\oldpar\paragraph
-\renewcommand{\paragraph}[1]{\vspace{-7pt}\oldpar{#1}}
-%\let\oldsec\section
-%\renewcommand{\section}[1]{\vspace{-7pt}\oldsec{#1}}
-\let\oldsubsec\subsection
-\renewcommand{\subsection}[1]{\vspace{-7pt}\oldsubsec{#1}}
+\AtBeginDocument{\paperdetails{
+		submitted=2018-09-29,
+		published=2019-02-01,
+		year=2019,
+		volume=3,
+		issue=3,
+		articlenumber=10,
+}}
 
@@
 \author[a]{Weixin Zhang}
+\authorinfo[weixin]{is a PhD student at the University of Hong Kong. Contact him at wxzhang2@cs.hku.hk}
 \author[a]{Bruno C. d. S. Oliveira}
 \affiliation[a]{The University of Hong Kong, Hong Kong, China}
+
+\keywords{embedded domain-specific languages, shallow embedding, object-oriented programming}
 
@@
 
-{\bf Context.} 
+\textbf{Context.}
 Embedded Domain-Specific Languages (EDSLs) are a common and widely
@@
 
-{\bf Inquiry.} Shallow embeddings are quite simple, 
+\textbf{Inquiry.} Shallow embeddings are quite simple,
 but they have been criticized in the past for being quite limited in
@@
 
-{\bf Approach.}
+\textbf{Approach.}
 This paper argues that shallow EDSLs and Object-Oriented
@@
 
-{\bf Knowledge.}
+\textbf{Knowledge.}
 This paper argues that common OOP mechanisms
@@
 
-{\bf Grounding.}
+\textbf{Grounding.}
 We make our arguments by using Gibbons and Wu's examples, where procedural
@@
 
-{\bf Importance.} This work is important for two reasons.
+\textbf{Importance.} This work is important for two reasons.
 Firstly, from an intellectual point of view, this work

### 2019-01-29 src/scans/Scans.hs

@@ -35,15 +35,15 @@ beside2        ::  Circuit2 -> Circuit2 -> Circuit2
 above2         ::  Circuit2 -> Circuit2 -> Circuit2
 stretch2       ::  [Int] -> Circuit2 -> Circuit2
 
-data Circuit2  =  Circuit2 { width :: Int, depth :: Int }
-id2 n          =  Circuit2 (n,0)
+type Circuit2  =  (Int,Int)
+id2 n          =  (n,0)
 fan2 n         =  (n,1)
 above2 c1 c2   =  (width c1,depth c1 + depth c2)
 beside2 c1 c2  =  (width c1 + width c2, depth c1 `max` depth c2)
-  stretch2 ns c  =  (sum ns,depth c)
+stretch2 ns c  =  (sum ns,depth c)
 
---width  =  fst
---depth  =  snd
+width  =  fst
+depth  =  snd
 
 id3            ::  Int -> Circuit3
 fan3           ::  Int -> Circuit3

### 2019-01-30 Programming19/Main.tex
Add author info and fix margin problems

@@
 \author[a]{Weixin Zhang}
-\authorinfo[weixin]{is a PhD student at the University of Hong Kong. Contact him at wxzhang2@cs.hku.hk}
+\authorinfo[weixin]{is a PhD candidate at the University of Hong Kong. His
+  current research interests are centered around modularity, domain-specific languages
+  and generative programming. You can contact him at \email{wxzhang2@cs.hku.hk}
+  and find further information at \url{https://wxzh.github.io}.
+}
 \author[a]{Bruno C. d. S. Oliveira}
+\authorinfo[bruno]{is an assistant professor of the at the University of Hong Kong. His
+  research interests are centered around programming languages. His existing
+  research is mainly focused on type systems for modularity and the combination
+  of Object-Oriented and Functional Paradigms. You can contact him at
+  \email{bruno@cs.hku.hk} and find further information at \url{https://i.cs.hku.hk/~bruno}.}
 \affiliation[a]{The University of Hong Kong, Hong Kong, China}
@@
 We further illustrate our approach with
-a case study on refactoring a deep external SQL DSL implementation to make it
+a case study on refactoring a deep external SQL query processor to make it
 more modular, shallow, and embedded.
